### 2022-08-15 COLT/colt.tex
Updates from Overleaf

\documentclass[anon,12pt]{colt2022} % Anonymized submission
%\documentclass[final,12pt]{colt2022} % Include author names

% The following packages will be automatically loaded:
% amsmath, amssymb, natbib, graphicx, url, algorithm2e

\usepackage{amsfonts}

\def\R{{\mathbb{R}}}
\def\pr{{\textrm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\yh{{\widehat{y}}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\sign{{\rm sign}}
\def\PL{{\mbox{\rm PL}}}
\newcommand{\predict}{\mathbf{Predict}}
\newcommand{\learn}{\mathbf{Learn}}
\newcommand{\D}{{\cal D}}

\newcommand\bigsubset[1][1.19]{%
   \mathrel{\vcenter{\hbox{\scalebox{#1}{$\subset$}}}}}

\newcommand{\comment}[3]{{\color{#1} {\bf #2 :} #3}}
%\newcommand{\comment}[3]{}  % suppress comments
\newcommand{\sanjoy}[1]{\comment{red}{Sanjoy}{#1}}
\newcommand{\yoav}[1]{\comment{blue}{Yoav}{#1}}

\DeclareMathOperator*{\argmax}{arg\,max}

\title[Active learning using region-based sampling]{Active learning using region-based sampling}
\usepackage{times}
% Use \Name{Author Name} to specify the name.
% If the surname contains spaces, enclose the surname
% in braces, e.g. \Name{John {Smith Jones}} similarly
% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
% If the first letter in the forenames is a diacritic
% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

% Two authors with the same address
\coltauthor{\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
\Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
\addr UC San Diego}

% Three or more authors with the same address:
% \coltauthor{\Name{Author Name1} \Email{an1@sample.com}\\
%  \Name{Author Name2} \Email{an2@sample.com}\\
%  \Name{Author Name3} \Email{an3@sample.com}\\
%  \addr Address}

% Authors with different addresses:
%\coltauthor{%
% \Name{Author Name1} \Email{xyz@sample.com}\\
% \addr Address 1
% \AND
% \Name{Author Name2} \Email{xyz@sample.com}\\
% \addr Address 2%
%}

\begin{document}

\maketitle

\begin{abstract}%
Abstract goes here.
\end{abstract}

\begin{keywords}%
List of keywords
\end{keywords}

\section{Introduction}

\section{Setup}

We assume that examples are drawn from a fixed but unknown
distribution $\D$ over $\X \times \Y$, where $\X$ is a metric space and
$\Y = \{-1,+1\}$~\footnote{In a later section we define the notion of
  neighborhoods more generally to allow, for example, diadic
  partitions}
Given $x \in \X$, the label $y \in \Y$ is drawn
according to the conditional probability function
$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x).$$
The \emph{Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.

The task we wish to solve is \emph{finite population annotation}. We
are given a finite dataset $X \subset \X$ of $n$ points.
We are given the ability to query the label of any point $x \in X$. The
answer to the query: $a(x) \in \{-1,+1\}$ is a binary random variable
drawn according to $\eta(x)$. In particular, multiple queries on the
same point can recieve different answers.

Our goal is to design an active learning algorithm that converges to the Bayes
optimal $g^*(x)$ and does so with a much smaller number of queries
than a passive algorithm.

We want a procedure that chooses the next point to query. It will be applied repeatedly and then stopped at an arbitrary time, whereupon labels $y(x)$ must be provided for all $x \in X$, including those that were queried. We would like the inferred labels $y(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.

Specifically, we will be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
$$ \sum_{x \in X} {\bf 1}(y(x) \neq g^*(x),\, |\eta(x)| > \gamma) .$$
The procedure is allowed to fail with probability $\delta$, to account for sampling error.

\section{Active learning algorithm}

\subsection{Sampling regions}

The sampling is organized around a collection $\B$ of measurable subsets of $\X$. These are the atomic sets on which we assess label bias and we call them ``balls''.

\begin{enumerate}

\item[(a)] \emph{Points within a ball.} For any $B \in \B$, let $X_B$ denote the points in $X$ that can be used to assess the bias of $B$. This is simply $X_B = X \cap B$ if $\B$ is a predefined set of balls.

\item[(b)] \emph{Levels of sampling.} We group balls by the number of points that they contain: we put $B$ at {\it level} $\ell$ if
$$ \frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .$$
Here $\ell$ ranges from $0$ (consisting of balls that contain at least half the data points) to roughly $\lg n$ (containing a single point).

Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of very large balls and $\B_1, \B_2, \ldots$ consisting of successively smaller balls.

\item[(c)] \emph{The neighborhood of a point.} For any
  $x \in \X$, and level $\ell$.
  $N(\ell,x)$ is the collection of balls from $\B_\ell$ that
  contain $x$.
\item[(d)] We say that the neighborhood $N(x,\ell)$ is {\em covered}
  if there are $k$ query points in each $B \in N(x,\ell)$.

\end{enumerate}

\subsection{Bias estimates and provisional labels}

The active learning algorithm uses label-queries to assess the bias of balls $B \in \B$. In turn, these are used to estimate the labels of individual points. All these estimates are organized by sampling level.

\begin{enumerate}

\item[(a)] We take the \emph{bias of a ball} $B \in \B$ to be
$\eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \}$.
At any given time, the algorithm assigns $B$ a qualitative bias estimate,
$$ \yh(B) = 
\left\{
\begin{array}{cl}
+1 & \mbox{significant $+$ bias} \\
-1 & \mbox{significant $-$ bias} \\
0 & \mbox{no significant bias} \\
\bot & \mbox{not yet available}
\end{array}
\right.
$$
The option $\bot$ is used until sufficiently many points in $X_B$ have been labeled: the required number, $k$, is roughly $1/\gamma^2$ (recall that $\gamma$ is the smallest bias that needs to be detected). Once $k$ labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter. It will satisfy a basic correctness guarantee (Definition~\ref{def:accurate-bias-estimate}) based on large deviation theory.

\item[(b)] \emph{Possible labels for a point.} Pick any $x$ and any level $\ell$. Once bias-estimates are available for all balls $B \in \B_\ell(x)$, the possible labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,0,+1\}$, are defined as follows:
\begin{gather}
+1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in
  \B_\ell(x)$ with $\yh(B) = +1$} \notag \\
0 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = 0$} \notag \\
-1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = -1$}
\label{eq:PL}
\end{gather}
The label of point $x$ at level $\ell$ is a value $\yh_\ell(x) \in \{-1,+1,0,!,\bot\}$. It is initially $\bot$, but once bias-estimates are available for all $B \in \B_\ell(x)$, it is set to a value in $\{-1,+1,0,!\}$ and remains fixed thereafter:
\begin{equation}
\yh_\ell(x) = 
\left\{
  \begin{array}{cl}
+1   & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \notin PL_\ell(x)$} \\
-1   & \mbox{if $+1 \notin \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
!    & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
0    & \mbox{if $\PL_\ell(x)=\{0\}$} \\
\bot & \mbox{if $\PL_\ell(x)=\emptyset$}
\end{array}
\right.
\label{eq:provisional-label}
\end{equation}
A point $x$ for which $\yh_\ell(x)=!$ is said to have a ``conflict at
level $\ell$''.

\end{enumerate}

\subsection{The algorithm}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{5.9in}
\begin{itemize}
\item{\bf Initialize:} set $\ell=0$, cover all of the balls in $\B_0$.
\item   $\predict(x)$
  \begin{itemize}
  \item Predict with $\yh_\ell(x)$
  \end{itemize}
  \item
  $\learn$  Increment sampling level\\
  Repeat while there is an $x$ that has a conflict at level $0\leq i
  \leq \ell$ and $N(x,i+1)$ is not fully covered. 
  \begin{itemize}
  \item Cover $N(x,i+1)$ using $m$ queries
  \item Query $m$ points uniformly from the domain (background
    sampling)
  \item Update $\yh_{i+1}(y)$ For all $y$.
  \end{itemize}
  Let $\ell+1 \to \ell$
\end{itemize}
\end{minipage}}
\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
\label{alg:main}
\end{figure}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{5.9in}
\begin{itemize}
\item Initialize uncertainty regions at all levels:
\begin{itemize}
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}
\item Initialize all neighborhoods to ``not yet sampled'':
\begin{itemize}
\item For all $B \in \B$: \ $\yh(B) = \bot$
\end{itemize}
\item Initialize all point-labels, at all levels, to ``not yet available'':
\begin{itemize}
\item For all $x \in X$ and $\ell \geq 0$: \ $\yh_\ell(x) = \bot$
\end{itemize}
%
%\vspace{.1in}
\item Repeat (main loop):
\begin{itemize}
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}
\item Let $\ell$ be the smallest such level
\item Set query region:
$$ S =  \bigcup_{x \in U_\ell} \bigcup_{B \in \B_\ell(x)} X_B $$
\item {\tt Query}($S$) \ (this queries one point from $S$)
\end{itemize}
\item Background sampling: {\tt Query}($X$)
\item For any neighborhood $B$ that is now adequately sampled for the first time: 
\begin{itemize}
\item Compute bias estimate $\yh(B) \in \{-1,0,+1\}$
\end{itemize}
\item For any $\ell \geq 0$ and $x \in X$ for which $\B_\ell(x)$ is available for the first time:
\begin{itemize}
\item Set $\PL_\ell(x)$ and $\yh_{\ell}(x)$ as per equations (\ref{eq:PL}) and (\ref{eq:provisional-label})
\end{itemize}
\item Update uncertainty region at all levels:
\begin{itemize}
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
\end{itemize}
\end{itemize}

\end{itemize}

\end{minipage}}
\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
\label{alg:main}
\end{figure}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{5.9in}

\vspace{.1in}
\emph{Initialization:}
\begin{itemize}
\item Set $Q = \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
\end{itemize}

\vspace{.2in}
{\bf Query}($S, \ell$):
\begin{itemize}
\item Query all points in $S \setminus Q$ with $T_z \leq 2^\ell k/n$, and add to $Q$
\end{itemize}

\end{minipage}}
\caption{The sampling procedure. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried.} 
\label{alg:sampling}
\end{figure}

The algorithm of Figure~\ref{alg:main} can be run for any amount of time. Whenever it is stopped, the {\it final label} of any $x \in X$ is
\begin{equation}
\yh(x) = 
\left\{
\begin{array}{ll}
\yh_\ell(x) & \mbox{for the largest $\ell < L[x]$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
\mbox{as queried} & \mbox{if there is no such $\ell$ but $x \in Q$} \\
? & \mbox{otherwise}
\end{array}
\right.
\label{eq:final-label}
\end{equation}

\yoav{I don't think there is anything special about a point that has
  been queried, a single query is not enough, we need $k$ queries to
  determine the prediction.}

\section{Analysis of algorithm}

\subsection{Accuracy of bias estimates}

We'll start with a uniform guarantee on the bias estimates for all balls $B \in \B$.

Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution $\eta(x) = \E[Y | X=x]$. For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value in $B$, that is,
$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
The specific property we need is encapsulated in the following definition.
\begin{definition}
For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$(b_1,b_2)$-accurate} if the following holds:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
\end{itemize}
\label{def:accurate-bias-estimate}
\end{definition}

In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
\begin{equation}
c_1 = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
\label{eq:c1}
\end{equation}
\yoav{So $\B$ is defined relative to $X$?}

For any level $\ell$, define threshold $\tau_\ell$ as follows:
\begin{equation}
\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
\label{eq:threshold-ell}
\end{equation}
For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B)
= \{z \in X_B: T_z \leq \tau_\ell\}$. We base our bias-estimate for
$B$ on the labels of points in $\Gamma(B)$. The following is a
consequence of Lemma~\ref{lemma:large-deviation-bounds} in the
appendix.
\yoav{I guess $T_z \leq \tau_\ell$ refers to the poisson sampling
  trick, which is not discussed. I think that initially we should
  define an oracle that provides samples with the desired properties,
  and later we can (in an appendix?) describe how the oracle is implemented.}
\begin{theorem}
Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $(b_1,b_2)$-accurate.
\label{thm:accurate-bias-estimates}
\end{theorem}
In what follows, we will assume that this $(1-\delta)$-probability event holds.

\subsection{Mind changes}

As the level of sampling $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_\ell(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.

Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ at level $\ell$. We distinguish two critical levels: $L_0(x)$, a level beyond which the Bayes-optimal label $s = \mbox{sign}(\eta(x))$ is \emph{always} in $\PL_\ell(x)$; and $L_1(x)$, a level beyond which the opposite label, $-s$, is \emph{never} in $\PL_\ell(x)$.

%\begin{figure}
%\begin{center}
%\includegraphics[width=3.5in]{home-stretch.pdf}
%\end{center}
%\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively finer levels. Beyond level $L_s(x)$, this $\yh_\ell(x)$ will never be the wrong label, but might be $0$ or $!$. Beyond level $L_f(x)$, the correct label will be assigned.}
%\label{fig:two-levels}
%\end{figure}

\begin{definition}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We define $L_0(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there exists $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$. We define $L_1(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there is no $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
\label{defn:L01}
\end{definition}

The key properties of these critical levels are spelled out in the following lemma.
\begin{lemma}
Pick $x \in X$ with $\eta(x) \neq 0$. Let $g^*(x) \in \{+1,-1\}$ be its Bayes-optimal label. Then:
\begin{enumerate}
\item[(a)] For all $\ell \geq L_0(x)$, we must have $g^*(x) \in \PL_\ell(x)$.
\item[(b)] For all $\ell \geq L_1(x)$, we must have $-g^*(x) \not\in \PL_\ell(x)$.
\item[(c)] For all $\ell \geq \max(L_0(x), L_1(x))$, we must have $\yh_\ell(x) = g^*(x)$.
\end{enumerate}
\label{lemma:boundary}
\end{lemma}
\begin{proof}
This lemma follows from the $(b_1, b_2)$-accuracy of the bias estimates for balls. Pick $x \in X$ with $\eta(x) \neq 0$. If $\ell \geq L_0(x)$, there is necessarily a ball $B \in \B_\ell(x)$ with $g^*(x) \eta_X(B) > b_2$ and thus $\yh(B) = g^*(x)$; therefore $g^*(x) \in \PL_\ell(x)$. If $\ell \geq L_1(x)$, every $B \in B_\ell(x)$ has $g^*(x) \eta_X(B) \geq -b_1$ and thus $\yh(B) \in \{0,g^*(x)\}$; therefore $-g^*(x) \not\in \PL_\ell(x)$. Finally, when $\ell \geq \max(L_0(x), L_1(x))$, it follows from (a) and (b) that $\PL_\ell(x) = \{g^*(x)\}$ and thus $\yh_\ell(x) = g^*(x)$.
\end{proof}

\subsection{Bounding the amount of focused sampling}

We can now characterize the points that are queried as part of focused sampling.
\begin{lemma}
For any $\ell \geq 0$, define
\begin{equation}
\Delta_{\ell} \ \ \ = \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .
\label{eq:sampling-region}
\end{equation}
Then, all focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
\label{lemma:focused}
\end{lemma}
\begin{proof}
Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$. For any point $x$ with $L_1(x) < \ell$, we have from Lemma~\ref{lemma:boundary}(b) that $\PL_{\ell-1}(x)$ is either $\{\}$ or $\{g^*(x)\}$; either way, $x$ is not in the uncertainty set at level $\ell$. Thus $\overline{U}_\ell \subset \{x \in X: L_1(x) \geq \ell\}$.

The focused samples at level $\ell$ lie in
$$
\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau(B) \}
\ \ \ 
\bigsubset[1.44]
\ 
\bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \} .
$$
This is precisely $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
\end{proof}

A final property we will need for label complexity bounds is that various subsets of interest contain roughly the expected number of points at each level.
\begin{lemma}
Pick any positive integer $\ell_f$. With probability at least $1-2(\ell_f+1)e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \ell_f$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
\item[(b)] If $|\Delta_\ell| \geq k$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 |\Delta_\ell| \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution}
\end{lemma}
\begin{proof} 
Pick any subset $S \subset X$ and let $m = |S|$. Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$. The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m/3}$.

Both parts follow from this principle; and we take a union bound over all $\ell_o + 1$ levels.
\end{proof}

\subsection{Label complexity bounds}

\yoav{I think this is the main theorem, however, I find it very hard
  to parse. In particular, I don't see any reference to the fraction
  of examples that are in the focused group at each level. These
  fractions have to be small in order for our sampling to be effective.}

\begin{theorem}
Pick any $0 < \delta < 1$. Suppose that $k \geq (8c_1/(b_2-b_1))^2$, where $c_1$ is defined in (\ref{eq:c1}), and that the active learning algorithm is allowed to make $m$ queries. Let
$$ \ell_0 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
and let $\ell_1$ be the largest integer such that
$$ \sum_{\ell = \ell_0 + 1}^{\ell_1} |\Delta_{\ell}| \, \tau_\ell \ \leq \ \frac{m}{8} .$$
Then with probability at least $1-2\delta$, any point $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$ will have final label $\yh(x) = g^*(x)$.
\label{thm:label-complexity}
\end{theorem}
\begin{proof}
Under the setting of $k$, the high-probability events of Theorem~\ref{thm:accurate-bias-estimates} and Lemma~\ref{lemma:level-distribution} hold with probability at least $1-2\delta$. We will henceforth take these for granted.

Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}. We will analyze the effect of background sampling in phase one and focused sampling in phase two. We start with the former.

Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried. Now, for $\ell_0$ as defined, we have from (\ref{eq:threshold-ell}) that $\tau_{\ell_0} \leq m/8n$ and from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy
$$ T_x \leq \tau_{\ell_0} \leq \frac{m}{8n}.$$
Therefore all such points are queried in phase one, and thus all balls $B$ at level $\leq \ell_0$ are processed and get bias estimates $\yh(B)$.

It then follows from Lemma~\ref{lemma:boundary} that the following holds for any $x \in X$ with $L_0(x) \leq \ell_0$:
\begin{enumerate}
\item[(a)] For any $\ell \geq \ell_0$, if $\yh_\ell(x)$ is defined then it is either $g^*(x)$ or $!$.
\item[(b)] If $x$ ever leaves $U$ during phase two, then its final label as defined in (\ref{eq:final-label}) is $\yh(x) = g^*(x)$.
\item[(c)] If $x$ is processed at level $L_1(x)$, then it will leave $U$ for good.
\end{enumerate}

Now let's move on to phase two. Let $A = \{x \in X: L_0(x) \leq \ell_0, L_1(x) \leq \ell_1\}$. We will show that every point in $A$ must leave the uncertainty set at some time during phase two. From (b), we can conclude that all these points have their final labels set correctly, once and for all.

We will break the argument into three cases.

{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.

{\it Case 2:} A full $m/4$ focused queries are made in phase two, but at least one of them is at level $\ell_1 + 1$. In that round, $U$ cannot contain any $x \in A$, since by (c) all such points will leave the uncertainty set by level $\ell_1$.

{\it Case 3:} A full $m/4$ focused queries are made in phase two, all at levels $\leq \ell_1$. By the analysis of phase one, none of these queries can be at level $\leq \ell_0$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_0+1$ through $\ell_1$ inclusive is at most
$$\sum_{\ell=\ell_0+1}^{\ell_1} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
\ \leq \ 
\sum_{\ell=\ell_0+1}^{\ell_1} 2 |\Delta_{\ell}| \tau_\ell
\ \leq \ 
\frac{m}{4} ,
$$
where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus by the end of the phase two, every $x \in A$ will have been processed until it leaves $U$.
\end{proof}

\section{Label complexity bounds in various settings}

In order to apply Theorem~\ref{thm:label-complexity}, we need two things:
\begin{itemize}
\item An upper bound on the critical level $L_1(x)$ for each $x \in X$, which in turn (via Lemma~\ref{lemma:focused}) will yield an upper bound on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
\item An upper bound on $L_0(x)$ for points $x$ that we expect to be correctly labeled by the end.
\end{itemize}
We now illustrate how this works out in several canonical settings.

\subsection{One-dimensional data with Massart noise}

Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
\begin{itemize}
\item The Bayes-optimal labeling is determined by $p$ disjoint open intervals $I_1 = (\lambda_0, \lambda_1), I_2 = (\lambda_1, \lambda_2), \ldots, I_p = (\lambda_{p-1}, \lambda_p)$, where
$$ 0 = \lambda_0 < \lambda_1 < \lambda_2 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$$
Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
\item For each $j$, either $\eta(x) > c$ for all $x \in I_j$ or $\eta(x) < -c$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
\end{itemize}
Here $c > 0$ is some constant.

We will take $\B$ to be all open intervals of $[0,1]$, and $\B(x)$ to be intervals containing the point $x$. The accuracy parameters $b_1, b_2$ will be set so that $0 \leq b_1 < b_2 \leq c$.
\begin{lemma}
Pick any $x \in I_j$, for $1 \leq j \leq p$.
\begin{enumerate}
\item[(a)] $L_0(x) \leq \left\lceil \lg (n/n_j) - 1 \right\rceil$, where $n_j$ is the number of data points in $I_j$.
\item[(b)] $L_1(x) \leq \left\lceil \lg (n/r(x)) - 1 \right\rceil$, where $r(x)$ is the minimum number of data points between $x$ and a boundary point. 
\end{enumerate}
\label{lemma:critical-one-d-massart}
\end{lemma}
\begin{proof}
For (a), define $n_j = |X \cap I_j|$ and notice that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$. Moreover, $s(I_j) \cdot \eta_X(I_j) > c \geq b_2$. For levels $\ell' > \ell$, we can make a similar argument using a sub-interval $I \subset I_j$ that lies in $\B_{\ell'}(x)$.

For (b), we define $r(x)$ to be either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller. For $\ell \geq \lceil (\lg n/r(x)) - 1 \rceil$, any $B \in \B(x)$ will lie entirely within $I_j$ and must therefore have $s(I_j) \cdot \eta_X(B) > 0$.
\end{proof}

Using these bounds on the critical levels, we find that all but an $\epsilon$ fraction of $X$ will be correctly classified after making $m = O(\frac{(p-1)\log n}{c^2} \log \frac{p-1}{\epsilon})$ queries.
\begin{theorem}
In the one-dimensional scenario above, there are absolute constants $C_1, C_2, C_3$ for which the following holds. Pick any $0 < \delta < 1$ and take $k \geq (C_1/c^2) \ln (n/\delta)$. Then with probability at least $1-\delta$, after making $m$ queries, the active learning algorithm will correctly classify all but $n \cdot 2^{-C_2 m/(k(p-1))}$ points in any target interval $I_j$ with $n_j/n \geq C_3 k/m$.
\label{thm:one-d-massart}
\end{theorem}

\begin{proof}
Using Lemma~\ref{lemma:critical-one-d-massart}(b), we see that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
\begin{align*}
\Delta_\ell \ \ \ 
&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $2n/2^{\ell}$ points on either side of each boundary point, so $|\Delta_\ell| \leq 4(p-1)n/2^{\ell}$.

We then set $b_1 = 0$, $b_2 = c$ and apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
\begin{align*}
\ell_0
&= 
\lg \frac{m}{32k} \\
\ell_1
&=
\ell_0 + \frac{m}{128 k(p-1)},
\end{align*}
and $k \geq (192/c^2)(2 \ln n + \ln (2/\delta))$; in setting $k$, we have used $|\B| = {n \choose 2}$. The theorem statement follows by invoking Lemma~\ref{lemma:critical-one-d-massart}.
\end{proof}

\subsection{One-dimensional data with monotonic $\eta$}

For our next example, we once again take $X$ to consist of $n$ arbitrarily-placed points in $\X = [0,1]$. This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing. Let $\lambda \in (0,1)$ denote the point at which $\eta(\lambda) = 0$.

As before, $\B$ will consist of all open intervals, and $\B(x)$ be will all intervals containing $x$. Given a threshold $\gamma > 0$ such that we do not need to label points with $|\eta(x)| < \gamma$, we can set accuracy parameters $b_2 = \gamma, b_1 = 0$.

\begin{lemma}
Let $\lambda_L, \lambda_R$ be the points at which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$.
\begin{enumerate}
\item[(a)] Define $n^- = |[0,\lambda_L) \cap X|$ and $n^+ = |(\lambda_R,1] \cap X|$. Then
$$
L_0(x)
\leq
\left\{
\begin{array}{ll}
\lg \lceil (n/n^+) - 1 \rceil & \mbox{if $x > \lambda_R$} \\
\lg \lceil (n/n^-) -1 \rceil & \mbox{if $x < \lambda_L$}
\end{array}
\right.
$$
\item[(b)] For any $x \in X$, let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$. Then
$L_1(x) \leq \left\lceil \lg (n/r(x)) -1 \right\rceil$.
\end{enumerate}
\label{lemma:critical-one-d-monotonic}
\end{lemma}

\begin{proof}
For (a), take $x > \lambda_R$, for instance. The interval $B = (\lambda,1)$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta(B) > b_2$. For higher levels $\ell$, we can take a sub-interval of $(\lambda, 1)$ containing $x$.

For (b), we have $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$. For any $\ell \geq \lceil \lg (n/r(x)) - 1 \rceil$, no $B \in \B_\ell(x)$ extends to the other side of the boundary; thus $\eta_X(B) \geq 0$.
\end{proof}

In this setting, we are not required to label points in the interval $[\lambda_L, \lambda_R]$, and thus we don't need $L_0(x)$ values for them. However, we do need to bound $L_1(x)$ for \emph{all} points in order to control the size of the sampling region.

With Lemma~\ref{lemma:critical-one-d-monotonic}, the rest of the argument follows the pattern of the previous one-dimensional case.

\begin{theorem}
There are absolute constants $C_1, C_2, C_3$ for which the following holds. Pick any $0 < \delta < 1$ and take $k \geq (C_1/\lambda^2) \ln (n/\delta)$. Then with probability at least $1-\delta$, after making $m \geq C_2 k \max(n/n^+, n/n^-)$ queries:
\begin{enumerate}
\item[(a)] The active learning algorithm will correctly label all but $n \cdot 2^{-C_3 m/k}$ points in $[0,\lambda_L) \cup (\lambda_R,1]$. 
\item[(b)] If $n_o = \min(|X \cap [\lambda_L,\lambda)|, |X \cap (\lambda,\lambda_R]|) > 0$, then all points in $[0,\lambda_L) \cup (\lambda_R,1]$ will be correctly labeled after $C_3 k \lg (n/n_o)$ queries.
\end{enumerate}
\label{thm:one-d-monotonic}
\end{theorem}

\begin{proof}
From Lemma~\ref{lemma:critical-one-d-monotonic}(b), we have that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
\begin{align*}
\Delta_\ell \ \ \ 
&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $2n/2^{\ell}$ points on either side of the boundary, whereupon $|\Delta_\ell| \leq 4n/2^{\ell}$.

By Theorem~\ref{thm:label-complexity}, making $m$ queries will correctly classify all points with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
\begin{align*}
\ell_0
&= 
\lg \frac{m}{32k} \\
\ell_1
&=
\ell_0 + \frac{m}{128 k},
\end{align*}
and $k \geq (192/\lambda^2)(2 \ln n + \ln (2/\delta))$. If $m \geq 32k \cdot \max(n/n^+, n/n^-)$, this includes all points in $[0,\lambda_L) \cup (\lambda_R,1]$ with $r(x) \geq n \cdot 2^{-m/128k}$. 
\end{proof}

\subsection{Two-dimensional data with linear boundary}

\subsection{Higher dimension, with dyadic partitioning}

\subsection{Higher dimension, with smoothness conditions}

% Acknowledgments---Will not appear in anonymized version
\acks{We thank a bunch of people and funding agency.}

\bibliography{yourbibfile}

\vfill

\pagebreak

\appendix

\section{Technicalities}

\subsection{Large deviation bounds for the discrete setting}

\begin{lemma}
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
Now consider the following estimator $Z$ of this quantity:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta_o| \leq \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-discrete}
\end{lemma}
\begin{proof}
Let's start with (a). Define $c_1 = \sqrt{3 \ln (4/\delta)}$. We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c_1 \sqrt{2k}$ since this is the worst case. By assumption, $k \geq 2c_1^2$ and thus $qm \leq 2k$.

Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
\begin{align*}
\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
\end{align*}
Take $\epsilon = c_1/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c_1$, we see that with probability at least $1-\delta/2$, we get 
$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
The lower bound implies $N > qm(1-\epsilon) = qm - c_1\sqrt{qm} = k + c_1 \sqrt{2k} - c_1 \sqrt{qm} \geq k$.

For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$. The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$. Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$. Thus their sum $S$ has expectation
$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
and variance
$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
We can bound the concentration of $S$ around its expected value using Bernstein's inequality,
$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + 2t/3)} \right) .$$
Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.

Combining with the high-probability bound on $N$ above, we get
$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
whereupon (recalling $Z = S/N$)
$$ 
\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
Since $|\eta_o| \leq 1$,
$$ |Z - \eta_o | \leq \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
\leq 
4 \epsilon,$$
as claimed.  
\end{proof}

An immediate consequence is the following large deviation bound. Recall that for any ball $B \in \B_\ell$, its query-set $\Gamma(B)$ is $\{z \in X_B: T_z \leq \tau_\ell\}$, where $\tau_\ell$ is defined in (\ref{eq:threshold-ell}).
\begin{lemma}
Suppose that the constant $c_1$ is set as in (\ref{eq:c1}) and that $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$\left| \widehat{\eta}(B) - \eta_X(B) \right| \leq 4c_1/\sqrt{k}$.
\end{enumerate}
\label{lemma:large-deviation-bounds}
\end{lemma}
\begin{proof}
These follow by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
\end{proof}

\end{document}

### 2022-12-28 older stuff/Outline.tex

%%% Testing:basic functionality
\documentclass{article}
\title{Outline of paper}
\begin{document}
\maketitle
\section{Introduction}
\section{Setup}
\section{The algorithm}
\section{How to sample}
\section{Examples}
\begin{itemize}
    \item 1D
    \begin{itemize}
        \item Single threshold with jump in $\eta$ ($-c$ on one side, $+c$ on the other side)
        \item Single threshold with more general $\eta$, e.g. Tsybakov condition
        \item Multiple thresholds with $-c, +c$
    \end{itemize}
    \item 2D all Euclidean balls
    \begin{itemize}
        \item Linear separator with $-c, +c$, e.g. uniform distribution over $x$
        \item Separator of bounded curvature on both sides
    \end{itemize}
    \item Dyadic tree: instead of limiting attention to boxes that are cells of the tree, we look at all boxes of all sizes that supported on the underlying gridding of space. We get better bounds for two reasons:
    \begin{itemize}
        \item We look at large boxes, not just leaves of the tree.
        \item We use all boxes, not just those that are nested like the tree; this makes the boundary apparent earlier using !.
    \end{itemize}
    \item Arbitrary dimension, but Holder-smoothness and Tsybakov margin conditions (usual setting in nonparametric estimation). The rates obtainable using passive random sampling were determined by Audibert-Tsybakov. For active, can look at Hanneke. Will find out.
\end{itemize}

\end{document}

### 2022-12-28 older stuff/main.tex

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{xcolor}

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\yh{{\widehat{y}}}
\def\PL{{\rm PL}}
\def\sign{{\rm sign}}

\newtheorem{thm}{Theorem}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{claim}[thm]{Claim}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assump}{Assumption}
\newtheorem{open}{Open problem}
\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}

\newcommand{\sanjoy}[1]{{\color{red} {\bf Sanjoy:} #1}}
\newcommand{\yoav}[1]{{\color{blue} {\bf Yoav:} #1}}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{Active learning using the near-neighborhoods algorithm}

\begin{document}

\maketitle

\section{Setup}

\begin{itemize}
\item {\it Underlying model of data.}

\begin{itemize}
\item Instances lie in a metric space $(\X, d)$. 
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to an unknown conditional probability function $\Pr(Y|X)$, with
$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x) = 2 \pr(Y=1|X=x) - 1 .$$
\item The {\it Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
\end{itemize}

\item {\it Task: Finite population labeling.}

\begin{itemize}
\item We are given a finite dataset of $n$ points, and have the
  ability to query the labels of any of them. The response to the
  query $x$ is +1 or -1, chosen according the conditional probability
  function $\eta(x)$. When two or more queries are made on the
  instance $x$, the responses are chosen independently at random
  according to $\eta(x)$, in other words, the labels are not persistant.

\item We are given parameters $0 < \gamma, \delta < 1$.
\item We want a procedure that chooses the next point, or batch of points, to query. This process will be applied repeatedly and forever. At any point the procedure can be asked to make a prediction on any point in $X$. The prediction is $\yh(x) \in \{-1,0,+1\}$. We consider the prediction correct if $\yh(x)=+1$ and $\eta(x)=+1$, if $\yh(x)=-1$ and $\eta(x)=-1$ or $\yh(x)=0$ and $|\eta(x)| < \gamma$. 
\item 
Choosing the queries indepndently at random is guaranteed to reach the correct prediction after querying each example $O(1/\gamma^2)$ times for a total of $O(|X|/\gamma^2)$ queries. Our goal is to reach convergence after much fewer queries.
\item 
Our bounds are pointwise, i.e. we say that the query complexity of the point $x$ is $q(x)$ if for any number of queries larger than $q(x)$ the prediction on $x$ is correct. The hope is that for most $x$, $q(x) \ll |X|/\gamma^2)$.
\item 
Our bounds should be a function of the advantage, as defined in the adaptive KNN paper: $p \gamma^2$
\iffalse
We would like the inferred labels $\yh(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
\item Specifically, we will then be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
$$ \sum_{x \in X} {\bf 1}(\yh(x) \neq g^*(x),\ |\eta(x)| > \gamma) .$$
The procedure is allowed to fail with probability $\delta$, to account for sampling error.
\fi
\end{itemize}

\end{itemize}

\section{Active learning algorithm}

\subsection{Some details}

\begin{enumerate}

\item {\it Sampling regions.}
The sampling is organized around a collection $\B$ of measurable subsets of $\X$. These are the atomic sets on which we assess label bias and we call them ``balls''.

\begin{enumerate}

\item[(a)] {\it Points within a ball.} For any ball $B \in \B$, let $X_B$ denote the points in $X$ that can be used to assess the bias of $B$. This is simply $X_B = X \cap B$ unless $\B$ is constructed in a way that depends on $X$ and causes certain points to be excluded.\footnote{The construction of $\B$ may or may not depend on the unlabeled points $X$, and this distinction determines which points can be considered random draws from a particular $B \in \B$. For instance, consider these two alternatives for defining $\B$:
\begin{itemize}
\item $\B$ consists of a pre-defined set of balls.
\item $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
\end{itemize}
In the first case, all points $X \cap B$ are random draws from $B$. In the second case, this is also true except for the two points $x,x'$ that define the ball.}

\item[(b)] {\it Grouping balls by size.} We will group balls by the number of points that they contain: we put $B$ at {\it level} $\ell$ if
$$ \frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .$$
Here $\ell$ ranges from $0$ (consisting of balls that contain at least half the data points) to roughly $\lg n$ (containing a single point).

Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of very large balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls.

\item[(c)] {\it Balls used to evaluate a given point.}

For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that will be used in determining $x$'s label. There are many ways in which this could be set, for instance,
\begin{itemize}
\item all balls containing $x$, or
\item all balls centered at $x$.
\end{itemize}
As before, we can partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.

\end{enumerate}

\item {\it Bias-estimates for balls and for data points.} The active learning algorithm uses label-queries to assess the bias of balls $B \in \B$. In turn, these are used to estimate the labels of individual points.

\begin{enumerate}

\item[(a)] {\it Bias of a ball.} The bias of a ball $B \in \B$ is defined as
$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$

\item[(b)] {\it Qualitative estimates of bias.} Each ball $B$ is assigned qualitative bias estimate,
$$ \yh(B) = 
\left\{
\begin{array}{cl}
+1 & \mbox{significant $+$ bias} \\
-1 & \mbox{significant $-$ bias} \\
0 & \mbox{no significant bias} \\
\bot & \mbox{not yet available}
\end{array}
\right.
$$
The option $\bot$ is used until sufficiently many points in $X_B$ have been labeled: the required number, $k$, is roughly $1/\gamma^2$ (recall that $\gamma$ is the smallest bias that needs to be detected). Once $k$ labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.

\item[(c)] {\it Bias guarantees.} Our bias estimates will satisfy a guarantee of the following form: with probability $1-\delta$, for all $B \in \B$, once $\yh(B)$ is set to a value in $\{-1,0,+1\}$, 
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
\end{itemize}
for some constants $0 < b_1 < b_2 < 1$.

\item[(d)] {\it Possible labels for a point.} Pick any $x$ and any level $\ell$. Once bias-estimates are available for all balls $B \in \B_\ell(x)$, the possible labels for $x$ at level $\ell$ are $\PL_\ell(x) \subset \{-1,+1\}$, defined as follows:
\begin{itemize}
\item $\PL_\ell(x)$ contains $+1$ if and only if there exists $B \in \B_\ell(x)$ with $y(B) = +1$ such that there is no $B' \in \B_\ell(x)$ with $B' \subset B$ and $y(B') = -1$.
\item $\PL_\ell(x)$ contains $-1$ is defined symmetrically.
\end{itemize}

\item[(e)] {\it The label of a point at level $\ell$.} The label of point $x$ at level $\ell$ is a value $\yh_\ell(x) \in \{-1,+1,0,!,\bot\}$. It is initially $\bot$, but once bias-estimates are available for all $B \in \B_\ell(x)$, it is set to a value in $\{-1,+1,0,!\}$ and remains fixed thereafter:
\begin{equation}
\yh_\ell(x) = 
\left\{
\begin{array}{cl}
+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
\end{array}
\right.
\label{eq:provisional-label}
\end{equation}
Points labeled $!$ are ``known unknowns'': they are near the decision boundary and need further investigation. Points labeled $0$ show no significant bias.

\end{enumerate}

\end{enumerate}

\begin{figure}[ht!]
\framebox{
\begin{minipage}[t]{6.4in}
\vspace{.1in}
\emph{Variables:}
\begin{itemize}
\item $Q$ (queries made so far)
\end{itemize}

\vspace{.1in}
\emph{Initialization:}
\begin{itemize}
\item Set $Q := \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
\end{itemize}

\vspace{.2in}
{\bf Focused-query}($\ell, U$)

\vspace{.1in}
\begin{itemize}
\item Query region:
$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
\item Query the next unlabeled point in $S \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\vspace{.2in}
{\bf Background-query}

\vspace{.1in}
\begin{itemize}
\item Query the next unlabeled point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}
\yoav{I am not sure whether the set from which to pick the query
  should exclude $Q$: the queries that have been made o far. if a set
  consists of one point and that point has a weak bias, we need to
  sample the point several times.}
\end{minipage}}
\caption{The sampling procedures. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried. The focused querying procedure also uses threshold $\tau_\ell = 2^{\ell+2}k/n$.} 
\label{alg:sampling}
\end{figure}

\begin{figure}[ht!]
\framebox{
  \begin{minipage}[t]{6.4in}
    \vspace{.1in}
\emph{Variables:}
\begin{itemize}
\item $U_\ell$ The set of points level $\ell$ that are ``known unknowns''
\end{itemize}

\vspace{.2in}
{\bf Active Balls}
\begin{itemize}
\item Initialize uncertainty regions at all levels:
\begin{itemize}
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}

\vspace{.1in}
\item Repeat:
\begin{itemize}
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}
\item Let $\ell'$ be the smallest such level
\item {\tt Focused-query}($\ell', U_{\ell'}$)
\end{itemize}
\item {\tt Background-query}
\item Update $\yh_{\ell}$ values, $\ell \geq 0$, as per equation (\ref{eq:provisional-label})
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = ! \}$
%\yoav{The last run before convergence should contain only disagreements: $!$, not zeros $0$}
\end{itemize}

\end{itemize}

\end{minipage}}
\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
\label{alg:main}
\end{figure}

\section{Analysis of algorithm}

\subsection{Accuracy of bias estimates}

We'll start with a uniform guarantee on the bias estimates for all balls $B \in \B$.

Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
$$ \eta(x) = \E[Y | X=x].$$ 
For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value in $B$, that is,
$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
The specific property we need is encapsulated in the following definition.
\begin{defn}
For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$(b_1,b_2)$-accurate} if the following holds:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
\end{itemize}
\label{def:accurate-bias-estimate}
\end{defn}

In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
\begin{equation}
c_1 = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
\label{eq:c1}
\end{equation}

For any level $\ell$, define threshold $\tau_\ell$ as follows:
\begin{equation}
\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
\label{eq:threshold-ell}
\end{equation}
For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.

Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-discrete} in the appendix.

\begin{thm}
Suppose $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c_1}{\sqrt{k}}.$$
\end{enumerate}
\label{thm:large-deviation-bounds}
\end{thm}
\begin{proof}
These follow immediately by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
\end{proof}

The following corollary of Theorem~\ref{thm:large-deviation-bounds} is immediate.
\begin{cor}
Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\mbox{sign}(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $(b_1,b_2)$-accurate.
\label{cor:accurate-bias-estimates}
\end{cor}
In what follows, we will assume that this $(1-\delta)$-probability event holds.

\subsection{Mind changes}
\iffalse
\yoav{We say that a prediction {\bf has converged} at level $L_f$ if one of the following holds:
\begin{itemize}
\item For all $\ell\geq L_f$, $\yh_\ell(x) = +1 \mbox{ and } \eta(x) > 0$
\item For all $\ell\geq L_f$, $\yh_\ell(x) = -1 \mbox{ and } \eta(x) < 0$
\item For all $\ell\geq L_f$, $\yh_\ell(x) = 0  \mbox{ and } |\eta(x)| < \gamma$
\end{itemize}

}
\fi

As the level $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_{\ell}(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.

%\yoav{There seems to be a  gap: $\yh_\ell(x)$ is a random variable, but $L_f(x),L_s(x)$ are not.}

We distinguish two critical levels for each $x$: the level $L_s(x)$ beyond which this label no longer flips between $+1$ and $-1$ (although it may be $0$ or $!$), and the level $L_f(x)$ beyond which the label will be correct. See Figure~\ref{fig:two-levels} for a pictorial view.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{home-stretch.pdf}
\end{center}
\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively higher levels. Between levels $L_s(x)$ and $L_f(x)$ there are contradicting labels on $x$, i.e.  $\yh_\ell(x)=!$. Beyond level $L_f(x)$, the correct label will be assigned.}
\label{fig:two-levels}
\end{figure}

We begin with $L_s$.

\begin{defn}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We say $x$ is \emph{obscured} at level $\ell \geq 0$ if the following two conditions hold.
\begin{itemize}
\item No $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) > b_2$.
\item Some $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) < -b_1$.
\end{itemize}
Let $L_s(x)$ denote the smallest integer such that $x$ is not obscured at any level $\ell \geq L_s(x)$. If there is no such level, we define $L_s(x) = \infty$.
\label{defn:Ls}
\end{defn}
In words, for levels $\ell < L_s(x)$, there is no ball $B \in \B_\ell(x)$ which has a strong bias towards the correct (Bayes-optimal) label, and there are balls with at least a mild bias towards the wrong label.

Next, we define the level by which the label of a point ought to be apparent.
\begin{defn}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We say $x$ is \emph{resolved} at level $\ell \geq 0$ if the following two conditions hold for all $\ell' \geq \ell$.
\begin{itemize}
\item All $B \in \B_{\ell'}(x)$ have $s \cdot \eta_X(B) > -b_1$.
\item Some $B \in \B_{\ell'}(x)$ has $s \cdot \eta_X(B) > b_2$.
\end{itemize}
Let $R_\ell$ denote the points that are resolved by level $\ell$. Let $L_f(x)$ denote the smallest level at which $x$ is resolved, or $\infty$ is there is no such level.
\label{defn:Lf}
\end{defn}
In words, for levels $\ell \geq L_f(x)$, no ball $B \in \B_\ell(x)$ has a strong bias towards the wrong label, and at least one ball has a strong bias towards the correct label.

Thus $L_s(x) \leq L_f(x)$, and there might be a large gap between the two. Before the background sampling reaches level $L_s(x)$, it is quite possible that we will predict the wrong label for $x$. However, once all balls in $\B_{L_s(x)}(x)$ have been adequately sampled, point $x$ will either (i) be assigned label $0$ or $!$, and put in the uncertainty set, in which case it will benefit from focused sampling, or (ii) will get the correct label. Once the balls in $\B_{L_f(x)}(x)$ have been sampled, condition (ii) will hold.

We define the set $F_\ell$ or include all of the points $x \in X$ such that $L_s(x) \leq \ell \leq L_f(x)$. The probability of $F_\ell$ determines the sample complexity of the active learning algorithm

\subsection{A possible characterization of $L_f(x),L_s(x)$}

\renewcommand{\P}{{\cal P}}
We are given a distribution over $\P$ over $\X \times \Y$ and a point $x \in X$.
We sample a set of unlabeled examples $X \subset \X$, and then run our active learning algorithm.

Our goal is to upper bound the number of samples $|\X|$ and the number
of queries the algorithm needs before the prediction of the algorithm
on $x$ converges to the correct label (and will not change with additional samples).

As we are dealing with the true distribution, we redifine $\B_\ell(x)$
to be all balls containing $x$ whose probability is in the range
$\left[ \frac{1}{2^{\ell-1}}, \frac{1}{2^\ell}\right]$.

Given the distribution $\P$, an edge parameter $\gamma>0$, a
prediction point $x$ and a level $\ell$ we partition the balls in $\B_{\ell}(x)$ into three 
subsets:
$$\B_{\ell}^+ = \{ B \in \B_\ell(x) | \eta(B)  > \gamma\},\;\;
\B_{\ell}^- = \{ B \in \B_\ell(x) | \eta(B)  < -\gamma\},\;\;
\B_{\ell}^0 = \{ B \in \B_\ell(x) | -\gamma \leq \eta(B) \leq \gamma\}$$

let the ``correct label'' of x be $s$, i.e. $s = \sign(\eta(x))$.
We define the ``true'' values of  $L_f(x),L_s(x)$ as follows:
\begin{itemize}
\item $L_s(x)$, is the smallest integer $m$ such that for all $\ell\geq
  m$, $\B_\ell^s \neq \emptyset$ while $\B_\ell^{-s} = \emptyset$.
\item $L_f(x)$ is the smallest integer $n \leq L_s(x)$ such that for
  all $n\leq \ell < L_s(x)$, $\B_\ell^s \neq \emptyset$ and $\B_\ell^{-s} \neq \emptyset$.
\end{itemize}

In other words the level $L_s(x)$ is the minimal level below which the
prediction is always correct. $L_f(x)$ is the minimal level above $L_s(x)$
such that the prediction on $x$ for the levels between  $L_f(x)$ and
$L_s(x)$ is always $!$.

\subsection{Another definition of two critical levels}
\yoav{I prefer to work with this definition}

For each point $x \in X$, we define two critical levels $L_0(x)$ and $L_1(x)$.

Let $s = \mbox{sign}(\eta(x)) \in \{-1, +1\}$. Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ after balls in $\B_\ell(x)$ have been queried.

We will define $L_0(x)$ and $L_1(x)$ so that: $s$ is guaranteed to be in $\PL_\ell(x)$ for all $\ell \geq L_0(x)$ and $-s$ is guaranteed to \emph{not} be in $\PL_\ell(x)$ for all $\ell \geq L_1(x)$. More precisely:
\begin{itemize}
\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$,
\begin{itemize}
\item there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$, and
\item any ball $B' \in \B_{\ell'}(x)$ with $B' \subset B$ has $s \cdot \eta_X(B') > -b_1$.
\end{itemize}
\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, any ball $B \in \B_{\ell'}(x)$ either has $s \cdot \eta_X(B) > -b_1$ or contains some $B' \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B') > b_2$.
\item \yoav{I would leave the definition of $L_0$ unchanged, but define $L_1$ as follows} $L_1(x)$ is the smallest level $\ell$ such that for all $L_0(x) > \ell' \geq \ell$, 
there exists two balls $B^+,B^- \in \B_{\ell'}(x)$ such that $\eta_X(B^+) > b_2$ and $\eta_X(B^-) < -b_2$.
\end{itemize}

Without containment:
\begin{itemize}
\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$.
\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is no ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
\end{itemize}

\section{Characterizing $S_L,S_F$  for Euclidean balls}
Suppose the input
distribution is uniform on a ball $B$ of radius one  around the
origin. Denote the ``positive'' set to be $C$. Let $\gamma>0$ be the
edge such that
\[
  P(y=1|X=x) =\begin{cases}
    1/2+\gamma & \mbox{ if } x \in C \\
    1/2-\gamma & \mbox{ if } x \notin C
    \end{cases}
  \]

  $L_f(x)$ corresponds to half the distance between $x$ and the boundary

  $L_s(x)$ can be bounded independently of $x$ if $r$ is the largest
  radius such that both $C$ and $C^c$ can be expressed as a union of
  balls of radius $r$.

  We can define a bondary of bounded convexity in this way.

 For other definitions of ``ball'' we can use a similar
 characterization: $L_s(x)$ corresponds to the smallest ball
 containing both $x$ and a point with the opposite label. $L_f(x)$
 corresponds to the largest radius such that both $C$ and $C^c$ can be
 written as a union of balls.

\section{Examples}
\begin{itemize}
    \item 1D
    \begin{itemize}
        \item Single threshold with jump in $\eta$ ($-c$ on one side, $+c$ on the other side)
        \item Single threshold with more general $\eta$, e.g. Tsybakov condition
        \item Multiple thresholds with $-c, +c$
    \end{itemize}
    \item 2D all Euclidean balls
    \begin{itemize}
        \item Linear separator with $-c, +c$, e.g. uniform distribution over $x$\\
        I this case $L_s(x)$ is lower bounded the the probability of a
        ball whose radius is half the distance between $x$ and the
        linear separator. $L_f(x)$ is close to 0. Two large balls, one
        shifted towards the positive, the other shifted towards the
        negative, will suffice to create a disagreement.

        \item Separator of bounded curvature on both sides
    \end{itemize}
    \item Dyadic tree: instead of limiting attention to boxes that are cells of the tree, we look at all boxes of all sizes that supported on the underlying gridding of space. We get better bounds for two reasons:
    \begin{itemize}
        \item We look at large boxes, not just leaves of the tree.
        \item We use all boxes, not just those that are nested like the tree; this makes the boundary apparent earlier using !.
    \end{itemize}
    \item Arbitrary dimension, but Holder-smoothness and Tsybakov margin conditions (usual setting in nonparametric estimation). The rates obtainable using passive random sampling were determined by Audibert-Tsybakov. For active, can look at Hanneke. Will find out.
    \item A one-dimensional example, e.g. with data points in [0,1] and with
$\eta > 3/4$ or $\eta < 1/4$ everywhere. In this case, the balls could
be all intervals, or perhaps a finite collection of them based on some
gridding of the line. The Bayes-optimal classifier would have a finite
number of sign-changes, and we could quantify which of these we'd pick
up by each level of sampling.

\item A high-dimensional Euclidean setting in which we use balls
corresponding to the cells of a tree (e.g. RP tree).

\item A high-dimensional metric space setting in which we have a
collection of metric balls and we give label complexity bounds based
on smoothness conditions (Holder-smoothness of the $\eta$ function)
and boundary conditions (like Tsybakov margin condition).

\end{itemize}

\begin{figure}
\begin{center}
\includegraphics[width=6in]{IMG_0840.jpg}
\end{center}

\caption{\yoav{Picture should be changed to include circles containing $x$ but not centered at $x$.} This figure provides some examples for the concept of mind changes. Here we consider true probabilities, rather than estimates. The bounding rectangle defines the input space $X$. The blue region indicate $x$'s where the probability of +1 is larger than that of $-1$, the red region indicates the revers polarity. More specifically, we assume that in the blue region the probability of +1 is 0.6, while in the red region the probability of +1 is 0.4.
The black dots correspond to points on which we want to make a prediction. We consider all circles centered at each point. The {\em bias} of  circle is the probability of +1 conditioned on the point being in the circle. As labels are accumulated, the bias of smaller and smaller circles can be determined. For any point there is a large enough circle that contains the whole input space. We assume that the overall bias is negative, hence for a sufficiently large circle the bias is negative. This is indicated by the dotted circle which is always the largest circle. The solid line circles are the largest circles with the polarity indicated by the color of the line. The smallest circle corresponds to the last mind change which we also call the convergence circle.
A point is considered "easy" if its convergence circle is large, because that means that a small number of queries is required to find the correct label.
Following that definition we get that the order of points from the easiest to the hardest is {\bf a,b,c,e,d}. Comparing e and d we see that d has a smaller convergence circle, and one mind change, which (e) has a larger convergence circle, but has 4 mind changes.
}

\label{fig:mind-changes}
\end{figure}

\pagebreak

\section{Technicalities}

\subsection{Large deviation bounds}

\begin{lemma}
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
Now consider the following estimator $Z$ of this quantity:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta_o| \leq \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-discrete}
\end{lemma}
\begin{proof}
Let's start with (a). Define $c_1 = \sqrt{3 \ln (4/\delta)}$. We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c_1 \sqrt{2k}$ since this is the worst case. By assumption, $k \geq 2c_1^2$ and thus $qm \leq 2k$.

Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
\begin{align*}
\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
\end{align*}
Take $\epsilon = c_1/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c_1$, we see that with probability at least $1-\delta/2$, we get 
$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
The lower bound implies $N > qm(1-\epsilon) = qm - c_1\sqrt{qm} = k + c_1 \sqrt{2k} - c_1 \sqrt{qm} \geq k$.

For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$. The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$. Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$. Thus their sum $S$ has expectation
$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
and variance
$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + 2t/3)} \right) .$$
Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.

Combining with the high-probability bound on $N$ above, we get
$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
whereupon (recalling $Z = S/N$)
$$ 
\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
Since $|\eta_o| \leq 1$,
$$ |Z - \eta_o | \leq \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
\leq 
4 \epsilon,$$
as claimed.  
\end{proof}

\end{document}

### 2024-07-31 COLT/response.tex
Initial Overleaf Import

\documentclass[12pt]{colt2023} %% Anonymized submission
\usepackage{times}
\begin{document}
\title{}
This is a response to reviewers 2, 3, 4 (we haven't received Review 1).

\vspace{.1in}

\noindent
{\bf Reviewer 2:}

\noindent
\emph{Finiteness of the metric space?} We are given a finite set of points $X$ to begin with; the set of balls on these points, $\mathcal{B}$, is effectively finite.

\noindent
\emph{Connection to paper of Balsubramani et al?} We were indeed inspired by this earlier work; but that paper was not about active learning and doesn't directly yield a good active learning method. For instance, that paper assesses the status of a point $x$ using balls \emph{centered} at $x$, whereas for us it is crucial to use balls \emph{containing} $x$ but possibly centered elsewhere. Our three key challenges (section 1.2) are all specific to active learning and demand new approaches.

\vspace{.1in}
\noindent
{\bf Reviewer 3:}

\noindent
\emph{Questions about the algorithm:} A \emph{background query} is drawn at random from the entire data space while a \emph{focused query} is drawn from the current region of uncertainty (top of section 2.3). In Figure 3, each data point $x$ has an associated random variable $T_x$ drawn uniformly on $[0,1]$. The data points are ordered according to these values. The goal of this technical trick is to ensure that queries for any ball are drawn independently at random.

\noindent
\emph{Instantiations of Theorem 4:} Indeed, the theorem is pretty abstract. We give three concrete instantiations: two one-dimensional scenarios and a higher dimensional scenario with Massart noise and a curvature condition on the boundary (sec 4.3). It's also easy to give an instantiation under smoothness and Tsybakov noise (see below for Reviewer 4), which we plan to add in.

\noindent
\emph{Constants in Theorem 5:} Say the fraction of the data with $\eta$ value in $[0,0.4)$, $[0.4, 0.5)$, $(0.5, 0.6]$, and $(0.6,1]$ is $40\%, 10\%, 10\%, 40\%$ respectively. Then $n^- = n^+ = 0.4n$ and $r^- = r^+ = 0.1n$.

\noindent
\emph{Underlying assumptions?} Our \emph{algorithm} is pretty general, unlike many active learning procedures that make sense only under smoothness conditions, for instance. But our \emph{bounds} do need (smoothness or other) assumptions, as is inevitable in nonparametric estimation.

\noindent
\emph{More discussion:} We agree that the theorems and results need more discussion. We will figure out how to squeeze this in!

\vspace{.1in}
\noindent
{\bf Reviewer 4:}

\noindent
\emph{More discussion of results:} Some intuition behind the algorithm is in the introduction. The theorems are another matter, and certainly warrant more discussion. We will add this somehow!

\noindent
\emph{Theorem 9:} In Theorem 9, the rate of convergence is controlled by two parameters: $\sigma, \xi$ (the other constants are either inherent in the problem, like the number of points $n$ or data dimension $d$, or are absolute constants, like $C$). Both $\sigma$ and $\xi$ are exponents in analogs of the Tsybakov margin condition; the difference is that $\xi$ ignores all points whose $\eta$ value lies in $[1/2-\gamma, 1/2+\gamma]$, the range in which any prediction is acceptable.

We will also add another instantiation of Theorem 9, using a strong density assumption (as before), curvature condition on boundary (as before), Holder-smoothness (with exponent $\alpha$) and Tsybakov margin (with exponent $\beta$). In this case, it is easy to show that $\sigma = \xi = \alpha\beta$. We will add this to the main paper or the appendix, depending on space constraints.

\noindent
\emph{Related work:} A discussion of related work appears in Section 1.4, but this focuses on \emph{algorithmic strategies} rather than \emph{bounds}. One that is directly comparable, but is a special case, is Castro-Nowak (as mentioned at the bottom of page 12, we get similar bounds). But other work is hard to compare since the assumptions are different and specialized. For instance, Minsker has a special smoothness condition, Kpotufe et al have a probabilistic Lipschitz condition, etc. It is hard to bridge these gaps.

\end{document}

### 2024-07-31 AISTATS/aistats.tex
Initial Overleaf Import

\documentclass[twoside]{article}

\usepackage{aistats2024}

%\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
%\usepackage{natbib}
\usepackage{enumitem}

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\yh{{\widehat{y}}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\dist{{\rm dist}}
\def\sign{{\rm sign}}
\def\vol{{\rm vol}}
\def\PL{{\mbox{\rm PL}}}

\newtheorem{thm}{Theorem}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{claim}[thm]{Claim}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assump}{Assumption}
\newtheorem{open}{Open problem}
\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}

\newcommand{\comment}[2]{ {\bf #1 :} #2}
\newcommand{\yoav}[1]{\comment{Yoav}{#1}}
\newcommand{\sanjoy}[1]{\comment{blue}{Sanjoy}{#1}}

\DeclareMathOperator*{\argmax}{arg\,max}

%\setlength{\leftmargini}{0.5cm}
%\setlength{\leftmarginii}{0.25cm}
%\setlength{\leftmarginiii}{0.25cm}

\begin{document}

\twocolumn[

\aistatstitle{Active learning without smoothness}

\aistatsauthor{ Sanjoy Dasgupta \And Yoav Freund }

\aistatsaddress{ UC San Diego \And  UC San Diego} ]

\begin{abstract}%
We present a general-purpose active learning scheme for data in metric spaces. Unlike previous active learning algorithms it is effective even when there are discontinuities in the conditional distribution.

\iffalse
The algorithm maintains a collection of neighborhoods of different sizes and uses label queries to identify those that have a strong bias towards one particular label; when two such neighborhoods intersect and have different labels, the region of overlap is treated as a ``known unknown'' and is targeted in future active queries. We give label complexity bounds for this method that do not rely on assumptions about the data and we instantiate them in several cases of interest.
\fi
\end{abstract}

\section{Introduction}

In \emph{active learning}, the starting point is a data set whose labels are hidden but can be obtained individually for a price. The goal is to label the data set, or to find a good classifier, at low cost.

We consider a formulation in which we have a collection of $n$ points $X = \{x_1, \ldots, x_n\}$ that lie in a metric space $(\X, d)$. We can request the label of any of these points $x$, in which case we get a value $y \in \{-1, +1\}$ with conditional distribution
$$ \eta(x) \ = \ \E[y | x] \ = \ \pr(y=1|x) - \pr(y=-1|x) $$
for some unknown function $\eta: \X \to [-1,1]$. The \emph{Bayes-optimal} label for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and $+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.
%$$ g^*(x) = \left\{
%\begin{array}{cl}
%-1 & \mbox{if $\eta(x) < 0$} \\
%+1 & \mbox{if $\eta(x) > 0$} \\
%\mbox{either} & \mbox{if $\eta(x)= 0$}
%\end{array} \right.
%$$
We wish to find the Bayes-optimal labels for the given data $X$.

More precisely, at the outset we have: the data $X$; two parameters $0 < \gamma, \delta < 1$; and a \emph{query budget}, the number of label queries we can make. We want a procedure that chooses the next point, or batch of points, to query. This procedure will be applied repeatedly until the query budget is exhausted, whereupon labels $\yh(x)$ must be provided for all $x \in X$, including those that were queried. Ideally, these will equal the Bayes-optimal labels $g^*(x)$; however, we will only be judged on points $x \in X$ with non-negligible bias, $|\eta(x)| \geq \gamma$. That is, the number of mistakes is taken to be:
$$ \sum_{x \in X} {\bf 1} \big(\yh(x) \neq g^*(x),\ |\eta(x)| \geq \gamma \big) .$$
The overall procedure is allowed to fail with probability $\delta$, to account for sampling error.

\subsection{Approach and challenges}

In our setting there is no underlying assumption about $g^*$, for instance that it follows a linear model. Thus we adopt a nonparametric approach.

We begin with a large collection of \emph{neighborhoods}, $\B$, of varying sizes (Fig.~\ref{fig:challenges}(a)); we will refer to these as \emph{balls}. We use label queries to assign \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning with the largest balls. A bias of $+1$ means that the average $\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is significantly negative; and bias $0$ means the average $\eta$ is close to 0. As time goes on, we get biases for progressively smaller balls, as needed. The final label assigned to any point $x$ is based on the biases of balls containing $x$.

\begin{figure*}
\begin{center}
\includegraphics[width=5.2in]{ideas.pdf}
\end{center}
\caption{(a) A collection of balls $\B$. Ball $B$ has bias $\yh(B) \in \{-1,0,+1\}$. (b) If one ball has bias $+1$ and the other has bias $-1$, then points $x$ in the intersection are targeted for active querying. (c) The biases of progressively smaller balls containing $x$ might flip sign, leading to \emph{mind changes} about the label of $x$.}
\label{fig:challenges}
\end{figure*}

%, me and to use label queries to estimate the signs associated with these regions. Ultimately we hope to cover $\X$ by a patchwork of neighborhoods with $\eta(x)$ values of uniform sign. Without strong smoothness conditions, these neighborhoods could be of very different sizes. The most beneficial for active learning are neighborhoods that contain a lot of $X$, but to get good coverage we might have to include smaller neighborhoods. To address this heterogeneity, we group the regions $\B$ by size. We begin by estimating the signs of the largest of them, and then move on to progressively smaller neighborhoods as the need arises. When the label budget runs out, the process is stopped, and provisional labels are assigned to the individual points in $X$.

There are three key challenges to be addressed.

The first challenge is deciding \emph{where to query}. Initial random querying tells us the biases of the largest balls, which gives some information about the decision boundary. In what parts of $\X$ might it be helpful to know biases of smaller balls, warranting further label queries? Our approach is simple: if a ball $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the intersection region, $B \cap B'$, is a ``known unknown''. These are the regions targeted for active querying  (Fig.~\ref{fig:challenges}(b)).

This differs from the more conventional strategy of actively querying neighborhoods $B$ whose average $\eta$-value is close to zero. Such $B$ fall into two categories: (1) the $\eta$ values are close to zero throughout $B$, and (2) $B$ consists of two sub-regions, one strongly positive and the other strongly negative. In the first case, there is little merit in querying further. But in the second case, there is a lot to be gained. Our approach allows these two cases to be distinguished.

%To distinguish these two cases, we use a collection of neighborhoods that are \emph{overlapping}. For instance, we might take $\B$ to be \emph{all} balls in $\X$, which is effectively a finite collection once the given data points $X$ are taken into account. Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a neighborhood $B$ that is strongly positive as well as being in a neighborhood $B'$ that is strongly negative. Such points are ``known unknowns'', and these are the targets of our active querying.

A second challenge is that the sign of $\eta(\cdot)$ can change in an arbitrarily small neighborhood of any given point: the biases of balls containing $x$, say $B_1 \supset B_2 \supset \cdots$, might change sign a few times. At first, given only the bias of $B_1$, we might think $x$ has label $+1$. When we later estimate the bias of $B_2$, this could change to $-1$. And then $+1$, and so on. We call these \emph{mind changes} (Fig.~\ref{fig:challenges}(c)).

What this means is that the algorthm 
might correctly predict the label of $x$, but it does not know that, with more labels the bias might be flipped, which we call a "mind change". Thus in addition to focused (active) querying, we also do background sampling of the entire space to pick up on possible mind changes. For simplicity, we make one background query per focused query.

The third challenge is \emph{managing sampling} of overlapping balls. We are interested in detecting points, and thus neighborhoods, whose $\eta$-values are either $> \gamma$ or $<-\gamma$. This suggests querying $k \approx 1/\gamma^2$ points at random from each ball. Now, suppose we have queried this many points from ball $B$ and later want to query points from a different ball $B'$ that overlaps $B$. How can we reuse queries we have already made in $B \cap B'$?

\emph{Poisson sampling} provides a clean solution. Rather than choosing $k$ points at random from $B$, we pick each point in $B$ with probability (roughly) $k/|B \cap X|$, independently. We implement this by assigning each $x \in X$ a uniform-random value $T_x \in [0,1]$, at the outset. When sampling $B$, we choose to query $x$ if $T_x \leq k/|B \cap X|$. And when it comes time to sample a different $B'$ that also contains $x$, we choose $x$ if $T_x \leq k/|B' \cap X|$; if its label has already been obtained, we are able to reuse it. In this way, the random querying of overlapping balls is seamlessly managed.

These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$. By tackling all three of them, we are able to give a general-purpose active learning scheme.

%Our algorithm does not require any assumption on the input distribution. In the worst case the number of queries is twice the number of queries of a passive algorithm. We show that under several natural conditions the query complexity of our algorithm is logarithmic in the error.

The rest of the paper is organized as follows.
Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set in the metric space and allowing any $\eta$ function. We identify two \emph{critical levels} for any $x \in X$: two scales (of neighborhood sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).

In Section~\ref{sec:1D} we instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.

In Section~\ref{sec:continuous} we consider a statistical statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$. We define a notion of boundary set and use it to show that the algorithm has logarithmic label complexity(Theorem~\ref{thm:label-complexity-specific}) under assumption of bounded curvature.

\subsection{Related work}

Existing works on non-parametric active learning can mostly be grouped according to the overall principle they follow: either (1) they seek to obtain the Bayes-optimal labels of a few well-positioned points, and then propagate these to the rest of the space \cite{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or negative) of entire regions at a time \cite{DH08,M12}. In this paper, we follow the second strategy because the only reliable and general-purpose way to assess the bias of an individual point, $\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent smoothness assumptions, the sign can change abruptly in an arbitrarily small neighborhood around $x$. On the other hand, the bias of a region $B$---the sign of the average $\eta$ value in $B$---is easy to determine given a reasonable number of points from that region.

Early results of \cite{CN08} established upper and lower bounds on label complexity in situations where the Bayes-optimal boundary is of a simple form: a single threshold for one-dimensional data or a ``smooth boundary fragment'' in higher dimension.

An algorithm for active learning based on hierarchical sampling was given by \cite{DH08} and was analyzed under smoothness conditions by \cite{KUB15}. In these works, the idea is to begin with a hierarchical clustering of $X$, and to then use queries to discover a pruning of this tree whose leaf-clusters are almost-pure in their labels. The method is not well-suited to situations with significant noise levels. Another approach using dyadic partitions was given by \cite{M12} and analyzed under commonly-used smoothness, margin, and density assumptions---namely, that $\eta$ is Holder-smooth, the fraction of points with $|\eta(x)| \leq t$ is some polynomial in $t$, and the marginal density is close to uniform---along with an additional smoothness requirement on $\eta$.

A different strategy using nearest neighbors was explored by \cite{ASU18}. Their idea was to choose an appropriate scale $s$, find an $s$-cover of $X$, estimate the Bayes-optimal label for each point in this cover by querying its neighbors, and then use these cleanly-labeled points for 1-nearest neighbor classification. A somewhat more general approach was given by \cite{H17} and studied under the usual smoothness, margin, and density conditions, with resulting rates of convergence comparable to those found by \cite{M12}.

Finally, \cite{DNZ15} suggested a graph-based method for active learning based on adaptively looking for the cut in the graph corresponding to the correct decision boundary. Their assumptions are based on properties of this cut and are not easily comparable with earlier work.

The idea of predicting with balls of multiple sizes is a central theme of \cite{BDFM19}, which predicts at a point $x$ by choosing a ball centered at $x$ that is as small as possible while having a clear bias towards one label. That work does not consider active learning, which demands considerable further innovation. For instance, to assess the uncertainty at a point $x$, we need to look at balls containing $x$, not just those centered at $x$ (Fig.~\ref{fig:challenges}(b)).

\section{The active learning algorithm}
\label{sec:algorithm}

Before giving a high-level overview of our algorithm, we recall some basic notation. We have a collection of points $X$ lying in a metric space $(\X, d)$. The label of any $x \in X$ can be requested and the value returned will be either $-1$ or $+1$, according to the conditional probability function $\eta(x) = \E(y|x)$. We wish to assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.

\subsection{A collection of sampling regions}

In our active learning algorithm, sampling is organized around a collection $\B$ of subsets of $\X$. These are the atomic sets on which we assess label bias and we call them \emph{balls} or \emph{neighborhoods}. There are no formal requirements on $\B$---for instance, the so-called balls can be of arbitrary shape---but the intention is that each $x \in X$ is contained in balls of multiple sizes, including one that is so small as to exclude the rest of $X$.

For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the data points that lie in it.
%\footnote{Since $X_B$ is used to assess the bias of $B$, we need to exclude points that cannot be considered random samples from $B$. For instance, consider these two alternatives: (1) $\B$ consists of a pre-defined set of balls. (2) $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$. In the first case, all points $X \cap B$ are random draws from $B$. In the second case, we need to exclude the two points $x,x'$ that define the ball.} 
We group balls into levels by the number of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
\begin{equation}
\frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .
\label{eq:sampling-level}
\end{equation}
Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of highly-populated balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls. We will use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater, and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.

Balls in lower levels contain more points, and thus their biases (average $\eta$ values) are easier to estimate. Our sampling algorithm goes from easier to harder levels.

For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that contain $x$ and can thus be used in determining $x$'s label. We again partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$. 
%We will also refer to $\B_\ell(x)$ as the {\em level $\ell$ neighborhood of $x$}.

\subsection{Estimating the biases of balls}

The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
We estimate these biases using label queries and assign each ball a \emph{qualitative} bias estimate,
$$ \yh(B) = 
\left\{
\begin{array}{cl}
+1 & \mbox{significant $+$ bias} \\
-1 & \mbox{significant $-$ bias} \\
0 & \mbox{no significant bias} \\
\bot & \mbox{not yet available}
\end{array}
\right.
$$

The option $\bot$ is used until enough points in $X_B$ have been queried: the required number is $k = O(1/\gamma^2)$ since $\gamma$ is the smallest bias that needs to be detected. Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter. These bias estimates will with high probability be seen to satisfy the following guarantee.
\begin{defn}
For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > 0$
\item $\yh(B) = -1 \implies \eta_X(B) < 0$
\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
\end{itemize}
\label{def:accurate-bias-estimate}
\end{defn}

\subsection{Assigning labels to points}

The label of any point is inferred from the biases of balls containing it.

Pick any $x \in X$ and any level $\ell$. Once qualitative bias estimates $\yh(B)$ are available for all balls at levels $\leq \ell$ that contain $x$ (that is, $\B_{\leq \ell}(x)$), we use this information to infer a level-$\ell$ label for $x$, denoted $\yh_\ell(x)$.

To begin with, the set of \emph{possible} labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$, is defined thus:
\begin{itemize}[leftmargin=0.5cm]
\item $\PL_\ell(x)$ includes $+1$ if there exists a \emph{minimal} ball $B \in \B_{\leq \ell}(x)$ (with no other $B' \in \B_{\leq \ell}(x)$ strictly contained in it) for which $\yh(B) = +1$.
\item $\PL_\ell(x)$ includes $-1$ under a symmetrical condition.
\end{itemize}
This is made precise in Eqn~(\ref{eq:PL}) in Fig.~\ref{alg:main}. The label-estimate for $x$ at level $\ell$, denoted $\yh_\ell(x)$, is $+1$ if $\PL_\ell(x)=\{+1\}$, $-1$ if $\PL_\ell(x) = \{-1\}$, $0$ if $\PL_\ell(x) = \{\}$, and $!$ if $\PL_\ell(x) = \{-1,+1\}$ (Eqn~(\ref{eq:provisional-label})). 
%The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.

The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\cite{R11}. Our active learning algorithm makes its \emph{focused queries} at level $\ell+1$ in the vicinity of such points $x$, that is, in $\B_{\ell+1}(x)$.

% to be $-1$ if $\PL_\ell(x) = %\{-1\}$, or $+1$ if $\PL_\ell(x) = \{+1\}$, or $0$ if $\PL_\ell(x) = \{\}$, or $!$ %if $\PL_\ell(x) = \{-1,+1\}$, %and remains fixed thereafter. Points labeled $!$ are ``known unknowns'': they are %near the decision boundary and %need further investigation. Points labeled $0$ show no significant bias.

\subsection{The algorithm}

The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, makes two types of queries: \emph{focused queries} and \emph{background queries}. Background queries are random draws from $X$ and correspond to passive learning. Focused queries are made in the vicinity of ``uncertain'' points and correspond to active learning. The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$. Focused queries are drawn at random from balls in $\B_\ell$ that contain such points.

On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.

Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$. Then the next iteration begins.

The querying process can be stopped at any time, whereupon labels are assigned as follows:
\begin{equation}
\yh(x) = 
\left\{
\begin{array}{cl}
\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$,} \\
           & \mbox{if such an $\ell$ exists} \\
0 & \mbox{``don't know'', otherwise}
\end{array}
\right.
\label{eq:final-label}
\end{equation}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{3.1in}
\begin{center}
\begin{minipage}[t]{3in}
Initialize uncertainty regions at all levels:
\begin{itemize}[leftmargin=0.5cm]
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}
Initialize labels at all levels to ``unavailable'':
\begin{itemize}[leftmargin=0.5cm]
\item $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
\end{itemize}
Repeat:
\begin{itemize}[leftmargin=0.5cm]
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}[leftmargin=0.25cm]
\item Let $\ell'$ be the smallest such level
\item {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
\end{itemize}
\item {\tt Background-query} // Fig.~\ref{alg:sampling}
\item Update labels:
\begin{itemize}[leftmargin=0.25cm]
\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
\begin{itemize}[leftmargin=0.25cm]
\item Possible labels for $x$ at level $\ell$:
\begin{equation} 
\begin{split}
\PL_\ell(x) = & \{s \in \{-1, +1\}: \\
             & \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t.} \\
             & \mbox{$\yh(B) = s$ and no $B' \in \B_{\leq \ell}(x)$} \\ 
             & \mbox{has $X_{B'} \subsetneq X_B$.} \} 
\end{split}
\label{eq:PL}
\end{equation}
\item Set
\begin{equation} 
\yh_\ell(x) = 
\left\{
\begin{array}{cl}
+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
\end{array}
\right.
\label{eq:provisional-label}
\end{equation}
\end{itemize}
\end{itemize}
\item Update uncertainty regions:
\begin{itemize}[leftmargin=0.25cm]
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$:

$U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
\end{itemize}
\end{itemize}

\end{minipage}
\end{center}
\end{minipage}}
\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
\label{alg:main}
\end{figure}

\subsection{Sampling}

The uncertainty region at level $\ell$ consists of the known-unknowns from level $\ell-1$, that is, points $x \in X$ with $\yh_{\ell-1}(x) = \, !$. This region is revealed to us piecemeal, a few points at a time. To get good label complexity we need to actively query the part of it we know.

We manage this through \emph{Poisson sampling} (Fig.~\ref{alg:sampling}). To infer the bias of a ball $B$, we need labels for $k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$ have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$ fraction of the ball to be queried. Instead, we query every point in the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, \emph{independently}. The way we do this is to query the points $x$ in the ball that have $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.

Once a point is queried, it is never queried again.

\begin{figure}[t]
\framebox{
\begin{minipage}[t]{3.1in}

\vspace{.05in}
\emph{Initialization:}
\begin{itemize}[leftmargin=0.5cm]
\item Set $Q = \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
\end{itemize}

\vspace{.1in}
{\bf Background-query}

%\vspace{.1in}
\begin{itemize}[leftmargin=0.5cm]
\item Query the next unlabeled point in $X \setminus Q$, ordered by $T_x$ values, and add to $Q$
\end{itemize}

\vspace{.1in}
{\bf Focused-query}($\ell, U$)

%\vspace{.1in}
\begin{itemize}[leftmargin=0.5cm]
\item Define querying region:
$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
\item Query the next unlabeled point in $S \setminus Q$, ordered by $T_x$ values, and add to $Q$
\end{itemize}

\end{minipage}}
\caption{The two sampling procedures. Each $x \in X$ has an associated random variable $T_x$ drawn uniformly from $[0,1]$. The smaller $T_x$, the earlier $x$ is likely to be queried. \emph{Background queries} are drawn at random from $X$. \emph{Focused queries} are from the uncertainty region at a particular level $\ell$, and make use of level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$. Here $k = \tilde{O}(1/\gamma^2)$ is a global parameter.} 
\label{alg:sampling}
\end{figure}

\begin{figure}[t]
\framebox{
\begin{minipage}[t]{3.1in}
\vspace{.05in}
\begin{itemize}[leftmargin=0.25cm]
\item Initially $\yh(B) = \bot$
\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
\end{itemize}
\end{minipage}}
\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$.}
\label{alg:bias-estimate}
\end{figure}

\section{Analysis: finite population setting}
\label{sec:discrete}

We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on the manner in which $X$ is generated.

\subsection{Accuracy of bias estimates}

Fix the set of balls $\B$ and let $0 < \delta < 1$ be a
predefined confidence parameter. We start with a uniform guarantee on the
bias estimates for all balls $B \in \B$.

In Fig.~\ref{alg:sampling}, we see that the points queried in a ball $B$ at level $\ell$ are $\{z \in X_B: T_z \leq \tau_\ell\}$, which from the definition of $\tau_\ell$ has size $O(k)$. We use these labels to estimate the bias of $B$. Since we need to detect biases of magnitude greater than $\gamma$, we set $k \propto 1/\gamma^2$. 
%This intuition is borne out by the following result, proved in the appendix. 
\begin{lemma}
Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. 
Let $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}. Then with probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are $\gamma$-accurate in the sense of Definition~\ref{def:accurate-bias-estimate}.
\label{lemma:accurate-bias-estimates}
\end{lemma}
Henceforth assume $\yh(B)$ is $\gamma$-accurate for all  $B \in \B$.

\subsection{Critical levels}

Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes the possible labels for $x$ given the information from balls at levels $0$ through $\ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$. Despite the randomness in the querying process, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ will reliably contain the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ will reliably omit the wrong label.

\begin{defn}[Critical levels $L_1,L_2$]
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$ be its Bayes-optimal label. Define $L_1(x)$ to be the smallest level $\ell$ such that there exists $B_o \in \B_\ell(x)$ for which
\begin{itemize}[leftmargin=0.5cm]
\item $s(x) \cdot \eta_X(B_o) \geq \gamma$, and
\item $s(x) \cdot \eta_X(B) \geq \gamma$ for any $B \in \B(x)$ with $X_{B} \subset X_{B_o}$.
\end{itemize}
Define $L_2(x)$ to be the smallest level $\ell$ such that:
\begin{itemize}[leftmargin=0.5cm]
\item For all $B \in \B_{\geq \ell}(x)$, we have $s(x) \cdot \eta_X(B) \geq 0$.
\item For any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$, there exists $B' \in \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
\end{itemize}
Take $L_1(x)$ or $L_2(x)$ to be $\infty$ if no level meets the requirements. 
\label{defn:L12}
\end{defn}

Once $\ell$ reaches a critical level, $\PL_\ell(x)$ and thus $\yh_\ell(x)$ behave more predictably, as follows.
\begin{lemma}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) \in \{+1,-1\}$ denote its Bayes-optimal label. Then for any level $\ell$ and any time at which $\yh_\ell(x) \neq \bot$:
\begin{enumerate}
\item[(a)] If $\ell \geq L_1(x)$, then $s(x) \in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), !\}$. 
\item[(b)] If $\ell \geq L_2(x)$, then $-s(x) \not\in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), 0\}$.
\end{enumerate}
\label{lemma:boundary}
\end{lemma}

\subsection{Boundary sets and label complexity}

A common intuition about active learning is that successive queries gradually constrain the possible locations of the decision boundary. Let's consider the state of affairs when all balls at level $\leq \ell - 1$ have been sampled. The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points must have $L_2(x) \geq \ell$. Focused sampling at level $\ell$ is restricted to balls that contain these points. We will think of this region as the \emph{boundary set} at level $\ell$.
\begin{defn}[Boundary set $\bf \Delta_\ell$]
For any level $\ell$, define the \emph{boundary set} at level $\ell$ to be
\begin{equation}
\Delta_\ell 
= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B 
.
\label{eq:sampling-region}
\end{equation}
\end{defn}
%It can be shown that all focused samples at level $\ell$ must lie in this set.
\begin{lemma}
All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
\label{lemma:focused}
\end{lemma}

We can now give generic label complexity bounds in terms of $L_1$, $L_2$, and $\Delta_\ell$. These take two equivalent forms: a \emph{global} version that specifies what parts of $X$ are correctly labeled after $m$ queries and a \emph{local} version that specifies the number of queries after which a particular $x$ is correctly labeled. The global version (Theorem~\ref{thm:label-complexity-0}) is in the Appendix; here is the local.

\begin{thm}[Local label complexity]
Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. With probability at least $1-2\delta$, the following holds simultaneously for all $x \in X$. Let $L_1(x)$ and $L_2(x)$ be the critical levels for $x$, as in Definition~\ref{defn:L12}. If $L_2(x) \leq \lg (n/2k)$, let
\[
m_o(x) = 32k \cdot \max\bigg( 2^{L_1(x)}, \ \frac{1}{n} \sum_{\ell=L_1(x)+1}^{L_2(x)} |\Delta_\ell| \, 2^\ell\bigg) .
\]
Once the active learning algorithm has made $m_o(x)$ queries, $\yh(x)$ is fixed at the Bayes-optimal label $g^*(x)$.
\label{thm:label-complexity}
\end{thm}
Let's interpret this bound. For a specific $x \in X$, there is no telling how $\yh_\ell(x)$ will behave at levels $\ell < L_1(x)$. But for $\ell \geq L_1(x)$, either $\yh_\ell(x) = \, !$, in which case $x$ is in the uncertainty region, or the label of $x$ is correctly set once and for all, which occurs by level $L_2(x)$.

We will say that $x$ \emph{reaches level $\ell$} when bias estimates have been obtained for all balls in $\B_{\leq \ell}(x)$ and thus $\yh_\ell(x)$ is set. Recalling that the algorithm alternates between background and focused queries, we show that $m_o(x)/4$ \emph{background} queries suffice for $x$ to reach level $L_1(x)$ and an additional $m_o(x)/4$ \emph{focused} queries take it to level $L_2(x)$, by which time its label is correct.

In particular, $O(k \cdot 2^\ell)$ background queries give us the biases of all balls in $\B_{\leq \ell}$; this accounts for the first term in $m_o(x)$. Meanwhile, the number of focused queries at any level $\ell$ is $O(k \cdot 2^{\ell} \cdot |\Delta_\ell|/n)$: the second term.

%The argument for Theorem~\ref{thm:label-complexity} is roughly that during the first $m_o(x)/2$ queries, background sampling alone is enough to ensure that $\yh_\ell(x)$ is forever set to either the correct label or !. During the next $m_o(x)/2$ queries, focused sampling then correctly resolves the label.

A footnote: Our algorithm queries points at most once. However, in some applications, a point $x$ can be queried repeatedly, each time producing an independent draw from $\eta(x)$. If \emph{repeat queries} are permissible, $O(1/\gamma^2)$ copies should be made of each point in $X$ before the algorithm is applied; and in this case, points with $|\eta(x)| \geq \gamma$ will all be correctly labeled, eventually.

\section{One-dimensional examples}
\label{sec:1D}
In order to apply Theorem~\ref{thm:label-complexity}, we need upper bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each point $x \in X$, and upper bounds on the size of the sampling region $\Delta_\ell$ at each level $\ell$. We now illustrate how this works out in some canonical settings.

\begin{figure*}[h]
\begin{center}
\includegraphics[width=2.5in]{oned-monotonic.pdf}
\hspace{.5in}
\includegraphics[width=2.5in]{oned-massart.pdf}
\end{center}
\caption{(a) The conditional probability function $\eta$ is monotonically increasing; $\eta(x) = -\gamma, 0, \gamma$ at $x = \lambda_L, \lambda, \lambda_R$, respectively. (b) On each subinterval, $\eta$ is either $> \gamma$ or $<-\gamma$.}
\label{fig:oned}
\end{figure*}

\subsection{Example: monotonic $\eta$}

Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing (Fig.~\ref{fig:oned}(a)). Let $\lambda \in (0,1)$ be the point for which $\eta(\lambda) = 0$ and let $\lambda_L, \lambda_R$ be the points for which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$. Thus we are not required to label points in the interval $(\lambda_L, \lambda_R)$.

%\begin{figure}
%\begin{center}
%\includegraphics[width=3in]{oned-monotonic.pdf}
%\end{center}
%\caption{The conditional probability function $\eta$ is monotonically increasing. The values $x = %\lambda_L, \lambda, \lambda_R$ have $\eta(x) = -\gamma, 0, \gamma$ respectively.}
%\label{fig:oned-monotonic}
%\end{figure}

Let $\B$ consist of all closed intervals. Although there are infinitely many of these, we need only consider those whose endpoints lie in $X$, so effectively $|\B| = O(n^2)$.

The critical levels are easy to bound in this setting. Let $n^- = |[0,\lambda_L] \cap X|$  be the number of points to the left of $\lambda_L$ and $n^+ = |[\lambda_R,1] \cap X|$ the number of points to the right of $\lambda_R$. Then 
\begin{align*}
L_1(x)
&\leq
\left\{
\begin{array}{ll}
\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
\end{array}
\right.
\\
L_2(x)
&\leq
\lg (n/r(x))
\end{align*}
where $r(x)$ is the number of data points between $x$ and $\lambda$, counting $x$ as well.
%\begin{itemize}
%\item $L_1(x) \leq \lg (n/n^+)$ if $x \geq \lambda_R$ and $L_1(x) \leq \lg (n/n^-)$ if %$x \leq \lambda_L$, and
%\item $L_2(x) \leq \lg (n/r(x))$, 
%\end{itemize}

A counting argument then shows that the boundary region shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
\begin{thm}
Define $r^+ = \min \{r(x): x \in X \cap [\lambda_R,1]\}$ and $r^- = \min \{r(x): x \in X \cap [0,\lambda_L]\}$. Pick any $0 < \delta < 1$. Suppose we run the algorithm of Fig.~\ref{alg:main} with $k = O((\log (n/\delta))/\gamma^2)$ and that $\min(r^+, r^-) \geq 2k$.  Take any 
$$ m \geq 64k \cdot \max \left( \frac{n}{\min(n^+,n^-)}, \ 2 \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \right) .$$
With probability at least $1-\delta$, after $m$ queries the algorithm will correctly label all $x \in X$ with $|\eta(x)| \geq \gamma$.
\label{thm:oned-monotonic}
\end{thm}
Say that $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$. Then the label complexity is at most $O(\log^2 n)$. If in addition $r^+, r^- = \Theta(n)$, that is, a constant fraction of the points lie between $\lambda_L, \lambda_R$ and the decision boundary, the label complexity is $O(\log n)$.

\subsection{Example: Massart noise}

In Section~\ref{sec:oned-massart}, we develop another one-dimensional example, in which $\X$ consists of several pieces, such that on each piece $\eta$ is either entirely above $\gamma$ or entirely below $-\gamma$ (Fig.~\ref{fig:oned}(b)). Once again, a logarithmic label complexity is obtained.

\section{Analysis: distributional setting}
\label{sec:continuous}

We now turn to a setting where the points $X$ are sampled from a distribution $\mu$ on $\X \subset \R^d$. The algorithm is unchanged, but we seek to bound label complexity in terms of properties of $\mu$ and $\eta$.

For simplicity, we focus on the case where $\mu$ is absolutely continuous with respect to the Lebesgue measure on $\R^d$ and thus admits a density. We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.

In this setting, we specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under margin conditions (Theorem~\ref{thm:label-complexity-specific}). Let's begin with a concrete example of a scenario handled by the latter result.

\subsection{A probabilistic notion of distance}

Now we turn to a general analysis of the distributional setting. First, large deviation bounds show that with high probability $|X \cap B| \approx n \mu(B)$ for every $B$, provided $k = \Omega(d \log n)$. With this in place, the crucial step is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
\begin{defn}
For any $x \in \X$ and $S \subset \X$, define
$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
\label{defn:prob-dist}
\end{defn}
We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
\begin{lemma}
For any $x \in \X$ with $\eta(x) \neq 0$, let $s(x) = \sign(\eta(x))$. Then
$$ L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1.$$
\label{lemma:L2-bound}
\end{lemma}
We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
\begin{defn}
For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
\label{defn:boundary}
\end{defn}
Notice that this includes all $x$ with $\eta(x) = 0$.

The $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
\begin{defn}
For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
\label{defn:boundary2}
\end{defn}

We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
\begin{lemma}
For any $\ell \geq 0$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
\label{lemma:delta-continuous}
\end{lemma}

Theorem~\ref{thm:label-complexity} thus continues to hold with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.

\subsection{Example: Curvature and Massart noise}
\label{sec:Massart}

Here we consider data that satisfy two assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
\begin{itemize}[leftmargin=0.5cm]
\item (Strong density condition)
There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
$$ c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) ,$$
where $\vol(\cdot)$ is $d$-dimensional volume.
\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
\end{itemize}
The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on the separator. It is a common notion of curvature in computational geometry~\cite{F59,NSW06} and implies, for instance, that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
\begin{thm}
Under the two conditions above, there are constants $c_2, c_3$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$ queries, it assigns Bayes-optimal labels to all but $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
\label{thm:massart-dist}
\end{thm}
This error rate of $\tilde{O}(1/m^{1/(d-1)})$ is an improvement over the usual $1/m^{1/d}$ rate for random querying and has been obtained earlier in similar settings using specialized algorithms~\cite{CN08}.

\subsection{Bounds under distributional conditions}

We now give label complexity bounds under three assumptions. Let $\X_\gamma = \{x \in \X: |\eta(x)| \geq \gamma\}$; thus $X \cap \X_\gamma$ is the set of points whose assigned labels will be judged. We further divide this set by label: for $s \in \{-1,+1\}$, let $\X^s_\gamma = \{x \in \X: s \cdot \eta(x) \geq \gamma\}$.

The first assumption is on the curvature of the decision regions. It allows us to bound $L_1(x)$.
\begin{enumerate}
\item[(A1)] There is an absolute constant $p_o > 0$ for which the following holds: for any $x \in \X_\gamma$, there exists $B \in \B(x)$ such that $B \subset \X^{s(x)}_\gamma$ and $\mu(B) \geq p_o$.
\end{enumerate}
\begin{lemma}
Under (A1), every $x \in \X_\gamma$ has $L_1(x) \leq \lceil \lg (1/p_o) \rceil$.
\label{lemma:L1-bound}
\end{lemma} 
The second assumption is a variant of the Tsybakov margin condition and bounds $|\Delta_\ell|$.
\begin{enumerate}
\item[(A2)] There are absolute constants $C > 0$ and $0 < \sigma < 1$ for which the following holds: for any $p > 0$, we have $\mu(\partial_{p,p}) \leq C p^\sigma$.
\end{enumerate}
If, for instance, $\mu$ were the uniform distribution over $[0,1]^d$, we would expect $\sigma \sim 1/d$.

The third assumption bounds the fraction of points with large $L_2$ values. It asks, what fraction of the $p$-boundary touches $\X_\gamma$? We consider two options: under (A3), the fraction is zero for $p$ small enough; under (A3'), the fraction is $p^\xi$.
\begin{enumerate}
\item[(A3)] There is an absolute constant $p_1 > 0$ such that $\mu(\partial_p \cap \X_\gamma) = 0$ for any $p < p_1$.
\item[(A3')] There are constants $C' > 0$ and $0 < \xi < 1$ such that $\mu(\partial_{p} \cap \X_\gamma) \leq C' p^\xi$ for any $p > 0$.
\end{enumerate}

\begin{thm}
Assume (A1), (A2) and either (A3) or (A3'). There is an absolute constant $C''$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm of Fig.~\ref{alg:main} makes $m$ queries, for
$$ \frac{128k}{p_o} \ \leq \ m \ \leq \ \frac{64C \cdot 8^\sigma}{1-\sigma} \, n^{1-\sigma} k^\sigma,$$
then with probability at least $1-\delta$:
\begin{itemize}
\item Under (A3), all of $X \cap \X_\gamma$ get Bayes-optimal labels for $m > (512C/(1-\sigma)) \cdot (1/p_1)^{1-\sigma} \cdot k$.
\item Under (A3'),  Bayes-optimal labels get assigned to all but $ C'' (k/m)^{\xi/(1-\sigma)} + (2/n) \log (4/\delta)$
%$$ C'' \left(\frac{k}{m}\right)^{\xi/(1-\sigma)} + \frac{2 \log (4/\delta)}{n} $$
fraction of $X \cap \X_\gamma$.
\end{itemize}
\label{thm:label-complexity-specific}
\end{thm}
This theorem yields the result of Section~\ref{sec:Massart}. Another example is in Section~\ref{sec:Tsybakov}, where bounds are given under smoothness and Tsybakov margin conditions.

\newpage

%\input{active-knn.bbl}
\bibliography{refs}
\bibliographystyle{plain}

\newpage
\onecolumn

\appendix

\section{Technicalities: discrete setting}

\subsection{Large deviation bounds for the discrete setting}

\begin{lemma}
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
Now consider the following estimator $Z$ of this quantity:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta_o| < \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-discrete}
\end{lemma}
\begin{proof}
Let's start with (a). Define $c = \sqrt{3 \ln (4/\delta)}$. We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c \sqrt{2k}$ since this is the worst case. By assumption, $k \geq 2c^2$ and thus $qm \leq 2k$.

Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
\begin{align*}
\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
\end{align*}
Take $\epsilon = c/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c$, we see that with probability at least $1-\delta/2$, we get 
$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
The lower bound implies $N > qm(1-\epsilon) = qm - c\sqrt{qm} = k + c \sqrt{2k} - c \sqrt{qm} \geq k$.

For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$. The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$. Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$. Thus their sum $S$ has expectation
$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
and variance
$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + t/3)} \right) .$$
Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.

Combining with the high-probability bound on $N$ above, we get
$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
whereupon (recalling $Z = S/N$)
$$ 
\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
Since $|\eta_o| \leq 1$,
$$ |Z - \eta_o | < \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
\leq 
4 \epsilon,$$
as claimed.  
\end{proof}

\subsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}

We will use Lemma~\ref{lemma:large-deviation-discrete} to obtain a uniform guarantee on the bias estimates for all balls $B \in \B$.

Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
$$ \eta(x) = \E[y | x].$$ 
For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value of the points in $B$, that is,
$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$

For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$, where the sampling threshold $\tau_\ell$ for level $\ell$ is defined as:
\begin{equation}
\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
\end{equation}
for some $k$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.

For what follows, define
\begin{equation*}
c = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
\end{equation*}

\begin{lemma}
Suppose $k \geq 2c^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c}{\sqrt{k}}.$$
\end{enumerate}
\label{lemma:large-deviation-bounds}
\end{lemma}
\begin{proof}
Pick $B \in \B$. There are two cases to consider.

Case 1: $|X_B| \geq 2k$. The choice of $\tau_\ell$ then ensures $|X_B| \tau_\ell \geq 2k$, so that $|\widehat{\eta}(B) - \eta_X(B)|$ can be bounded by applying Lemma~\ref{lemma:large-deviation-discrete} to the points $X_B$ with sampling probability $q = \tau_\ell$.

Case 2: $k \leq |X_B| < 2k$. In this case, $B$ lies at a level $\ell$ for which $\tau_\ell = 1$. Lemma~\ref{lemma:large-deviation-discrete} does not apply directly, but its conclusion still holds. In particular, the query-set $\Gamma(B)$ is all of $X_B$, and the same Bernstein bound from the proof of Lemma~\ref{lemma:large-deviation-discrete} can be again be applied.

To complete the proof, we take a union bound over all $B \in \B$.
\end{proof}

The following corollary of Lemma~\ref{lemma:large-deviation-bounds} is immediate.
\begin{cor}
Suppose that $k \geq (8c/\gamma)^2$. For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\gamma$-accurate.
\label{cor:accurate-bias-estimates}
\end{cor}
\begin{proof}
By the choice of $k$, we have from Lemma~\ref{lemma:large-deviation-bounds} that 
$| \widehat{\eta}(B) - \eta_X(B) | < \gamma/2$
for all $B \in \B$.
\end{proof}

\subsection{Critical levels: Proof of Lemma~\ref{lemma:boundary}}

For part (a), note that some $B_o \in \B_{L_1(x)}(x)$ has significant bias (that is, bias $\geq \gamma$) towards the correct label $s(x)$, as does any ball $B \in \B(x)$ contained within it. For $\ell \geq L_1(x)$, the set $\{B \in \B_{\leq \ell}(x): X_{B} \subseteq X_{B_o}\}$ is nonempty (it contains $B_o$); pick any minimal ball $B$ within it. Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimates}) and thus $s(x) \in \PL_\ell(x)$.

For (b), take any $\ell \geq L_2(x)$. Consider any $B \in \B_{\leq \ell}(x)$ for which $s(x) \cdot \eta_X(B) < 0$. By definition of $L_2(x)$, this $B$ must lie in $B_{< L_2(x)}(x)$, and moreover there must exist $B' \in \B_{\leq L_2(x)}(x) \subset \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$. Thus, any minimal $B \in \B_{\leq \ell}(x)$ has bias $\geq 0$ towards the correct label $s(x)$, whereupon $\yh(B) \in \{0, s(x)\}$ by the $\gamma$-accuracy of bias estimates. Therefore, $-s(x) \not\in \PL_\ell(x)$.

\subsection{The region of focused sampling: Proof of Lemma~\ref{lemma:focused}}

Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$. From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty set at level $\ell$. In short,
\begin{equation}
\overline{U}_{\ell} \subset \{x \in X: L_2(x) \geq \ell\}.
\label{eq:uncertainty}
\end{equation}

We see from the {\tt Focused-query} subroutine (Figure~\ref{alg:sampling}) that all focused samples at level $\ell$ lie in
$$
\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
\subset 
\ 
\bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
=
\ 
\{z \in \Delta_\ell: T_z \leq \tau_\ell \}.
$$

\subsection{A generic label complexity bound}

We start by showing that various subsets of interest contain roughly the expected number of points at each level.
\begin{lemma}
With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/2k)$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| < 2 n \tau_\ell$.
\item[(b)] If $\Delta_\ell \neq \emptyset$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| < 2 |\Delta_\ell| \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution}
\end{lemma}
\begin{proof} 
Pick any subset $S \subset X$ and let $m = |S|$. Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$. The probability that it is greater than or equal to twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m \tau_{\ell}/3}$, which is $\leq e^{-k/3}$ as long as $m \tau_\ell \geq k$.

Both parts follow from this principle; and we take a union bound over all $\lg (n/k)$ levels. For (b), we need to check that $|\Delta_\ell| \tau_\ell \geq k$. To see this, observe from the definition (\ref{eq:sampling-region}) of $\Delta_\ell$ that if it is non-empty, then it contains $X_B$ for at least one ball $B \in \B_\ell$, and every such ball has at least $n/2^{\ell+1}$ points. Combining this with the definition $\tau_\ell = \min(2^{\ell+2}k/n, 1)$ yields $|\Delta_\ell| \tau_\ell \geq k$.
\end{proof}

Theorem~\ref{thm:label-complexity} is a restated version of the following.
\begin{thm}
Suppose the active learning algorithm makes $0 < m \leq n$ queries. Then all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
and $\ell_2$ is the largest integer $\leq \lg (n/2k)$ such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} |\Delta_{\ell}| \, \tau_{\ell} \ < \ \frac{m}{8} .$$
\label{thm:label-complexity-0}
\end{thm}

\begin{proof}
Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}. We will analyze the effect of background sampling in phase one and focused sampling in phase two. We start with the former.

Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.

Now, for $\ell_1$ as defined, we have that $\tau_{\ell_1} \leq m/8n$ and thus from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy $T_x \leq \tau_{\ell_1}$. Therefore all such points are queried in phase one, and all label-estimates $\{\yh_{\ell_1}(x): x \in X\}$ are set.

It then follows from Lemma~\ref{lemma:boundary}(a) that the following hold for any $x \in X$ with $L_1(x) \leq \ell_1$:
\begin{enumerate}
\item[(a)] By the end of phase one, $\yh_{\ell_1}(x) \in \{g^*(x), !\}$.
\item[(b)] For any $\ell > \ell_1$, when $\yh_\ell(x)$ becomes available, it lies in $\{g^*(x), !\}$.
\item[(c)] If $x$ ever leaves the combined uncertainty region $U = \cup_\ell U_\ell$ during phase two, then its final label as defined in (\ref{eq:final-label}) is henceforth always $\yh(x) = g^*(x)$.
\end{enumerate}

Now let's move on to phase two. Let $A = \{x \in X: L_1(x) \leq \ell_1, L_2(x) \leq \ell_2\}$. We will show that every point in $A$ must leave the uncertainty region $U$ at some time during phase two. From (c), we can conclude that all these points have their final labels set correctly, once and for all.

We will break the argument into two cases.

{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.

{\it Case 2:} A full $m/4$ focused queries are made in phase two. By the analysis of phase one, none of these queries can be at level $\leq \ell_1$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_1+1$ through $\ell_2$ inclusive is at most
$$\sum_{\ell=\ell_1+1}^{\ell_2} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
\ \leq \ 
\sum_{\ell=\ell_1+1}^{\ell_2} 2 |\Delta_{\ell}| \tau_\ell
\ < \ 
\frac{m}{4} ,
$$
where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus at least one query in phase two must be at level $\ell_2 +1$. When this query is made, every $U_\ell$ with $\ell \leq \ell_2$ must be empty and thus all of $A$ must have left the uncertainty region; recall from (\ref{eq:uncertainty}) that no $x \in A$ can be part of $U_\ell$ for $\ell > \ell_2 \geq L_2(x)$.

Thus every $x \in A$ must leave the uncertainty region at some point in phase two, and their final labels are subsequently set correctly.
\end{proof}

\subsection{One-dimensional monotonic $\eta$: Proof of Theorem~\ref{thm:oned-monotonic}}

We begin by bounding the critical levels $L_1$ and $L_2$ for points in $X$.
\begin{lemma}
Pick any $x \in [0, 1]$.
\begin{enumerate}
\item[(a)] Define $n^- = |[0,\lambda_L] \cap X|$ and $n^+ = |[\lambda_R,1] \cap X|$. Then
$$
L_1(x)
\leq
\left\{
\begin{array}{ll}
\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
\end{array}
\right.
$$
\item[(b)] Let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$, counting $x$ itself. That is, $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$. Then $L_2(x) \leq \lg (n/r(x))$.
\end{enumerate}
\label{lemma:oned-monotonic-L}
\end{lemma}
\begin{proof}
For (a), take any $x \geq \lambda_R$ (the other case is similar). The interval $B = [\lambda,1]$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta_X(B) \geq \gamma$. Furthermore, any $B' \subset B$ also has $\eta_X(B') \geq \gamma$.

For (b), take $x \geq \lambda_R$ and $\ell \geq \lg (n/r(x))$. Any $B \in \B_\ell(x)$ contains $< n/2^\ell \leq r(x)$ points and thus cannot possibly extend to the other side of the boundary. It follows that every $B \in \B_{\geq \ell}(x)$ has $\eta_X(B) \geq 0$. Moreover, any interval $B'$ that does extend to the other side of the boundary contains $B' \in \B_{\ell}(x)$ that is entirely on the same side as $x$.
\end{proof}

We can now bound the size of the query region at each level and find that it shrinks exponentially with $\ell$.
\begin{lemma}
For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq 4n/2^\ell$.
\label{lemma:oned-monotonic-query-region}
\end{lemma}

\begin{proof}
We have
\begin{align*}
\Delta_\ell 
&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $n/2^{\ell-1}$ points from $X$ on either side of $\lambda$. \end{proof}

We are now ready for the proof of Theorem~\ref{thm:oned-monotonic}.

Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}. Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.

First observe that for any $\ell_1 \leq \ell_2$, we have from Lemma~\ref{lemma:oned-monotonic-query-region} that
$$ \sum_{\ell=\ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell 
\ \leq \ \sum_{\ell=\ell_1+1}^{\ell_2} \frac{4n}{2^\ell} \cdot \frac{2^{\ell+2} k}{n} 
\ = \ 16k(\ell_2-\ell_1).$$
Now, let's define
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor, 
\ \ \ \ell_2 = \min \left( \ell_1 + \frac{m}{128k}, \ \ \lg \frac{n}{2k} \right).$$
Then for any $x \in [0, \lambda_L] \cup [\lambda_R,1]$, we have
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor 
\ \geq \ 
\lg \frac{m}{64k}
\ \geq \ 
\lg \frac{n}{\min(n^+,n^-)}
\ \geq \ 
L_1(x)$$
and, if $\min(r^+, r^-) \geq 2k$,
\begin{align*}
\ell_2 
\ = \ \ell_1 + \frac{m}{128k} 
&\geq \lg \frac{m}{64k} + \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \\
&\geq \ \lg \frac{m}{64k} + \lg \frac{64kn/m}{\min(r^+,r^-)} 
\ = \ \lg \frac{n}{\min(r^+,r^-)}
\ \geq \  
L_2(x).
\end{align*}
We get the algorithmic guarantee by applying Theorem~\ref{thm:label-complexity}.

\subsection{One-dimensional data with Massart noise}
\label{sec:oned-massart}

We now turn to another one-dimensional setting. Once again, $X$ consists of $n$ arbitrarily-placed points in $\X = [0,1]$. This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
\begin{itemize}
\item There are $p$ disjoint open intervals $I_1, \ldots, I_p$, such that $\X$ is (the closure of) their union, and
\item for each $j$, either $\eta(x) > \gamma$ for all $x \in I_j$ or $\eta(x) < -\gamma$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
\end{itemize}
Here $0 < \gamma < 1$ is some constant. See Figure~\ref{fig:oned}(b) for an illustrative example. For concreteness, the intervals $I_j$ can be written in the form $(\lambda_{j-1}, \lambda_j)$, where
$0 = \lambda_0 < \lambda_1 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$
Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.

We will take $\B$ to consist of all open intervals of $[0,1]$, with $\B(x)$ denoting intervals that contain point $x$.

We begin with bounds on the $L_1$ and $L_2$ levels for each point.
\begin{lemma}
Pick any $x \in \X$; suppose $x \in I_j$.
\begin{enumerate}
\item[(a)] Let $n_j = |X \cap I_j|$. Then $L_1(x) \leq \lg (n/n_j)$.
\item[(b)] Let $r(x)$ be the minimum number of data points that lie between $x$ and a boundary point, counting $x$ as well; this is either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller. Then $L_2(x) \leq \lg (n/r(x))$.
\end{enumerate}
\label{lemma:oned-massart-L}
\end{lemma}
\begin{proof}
For (a), notice first that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$. Moreover, $s(I_j) \cdot \eta_X(I_j) > \gamma$. Thus $I_j$ belongs to $\B_\ell(x)$ and is strongly biased towards the correct label. This strong bias also holds for any subset of $I_j$.

For (b), consider any $\ell \geq \lg (n/r(x))$. Any $B \in \B(x)$ with $\mbox{sign}(\eta_X(B)) \neq s(I_j)$ must contain either the entire left-interval $(\lambda_{j-1},x]$ or the entire right-interval $[x, \lambda_j)$, and thus has at least $r(x)$ points, which means that it is too large to be in $\B_\ell$. Thus all intervals $B \in \B_{\geq \ell}(x)$ have $s(I_j) \cdot \eta_X(B) > 0$. Also, for any interval $B \in \B_{<\ell}(x)$ there is some $B' \in \B_{\ell}(x)$ that is strictly contained within it.
\end{proof}

With $L_1(x)$ and $L_2(x)$ under control, it is easy to bound the size of the focused query region $\Delta_\ell$ at each level.
\begin{lemma}
For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq (p-1)n/2^{\ell-2}$.
\label{lemma:oned-massart-query-region}
\end{lemma}

\begin{proof}
We have
\begin{align*}
\Delta_\ell 
&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $n/2^{\ell-1}$ points from $X$ on either side of each boundary point $\lambda_j$.
\end{proof}

Notice that $|\Delta_\ell|$ shrinks exponentially with $\ell$. With $L_1$, $L_2$, and $|\Delta_\ell|$ values in place, Theorem~\ref{thm:label-complexity} can be applied directly to give the following label complexity bound.

\begin{thm}
Pick any $0 < \epsilon, \delta < 1$. Suppose we run the algorithm of Figure~\ref{alg:main} with $k = O((1/\gamma^2) \ln (n/\delta))$. With probability at least $1-\delta$, after making
$$ O \left( \frac{p-1}{\gamma^2} \ln \frac{p-1}{\epsilon} \ln \frac{n}{\delta} \right) $$
queries, the algorithm will assign the correct label $g^*(x)$ to at least $1-\epsilon$ fraction of $X$, except possibly the $2k$ points of either side of each boundary point.
\label{thm:oned-massart}
\end{thm}

\begin{proof}
Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}. Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.

Next, using Lemma~\ref{lemma:oned-massart-query-region}, we have that for any integers $0 \leq \ell_1 < \ell_2$,
$$ \sum_{\ell = \ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell
\ \leq \ 
\sum_{\ell = \ell_1+1}^{\ell_2} \frac{(p-1) n}{2^{\ell-2}} \cdot \frac{2^{\ell+2}k}{n}
\ = \ 
16(p-1)k (\ell_2 - \ell_1).
$$
We can then apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$, for
\begin{align*}
\ell_1
&= 
\left\lfloor \lg \frac{m}{32k} \right\rfloor \\
\ell_2
&=
\min \left( \ell_1 + \frac{m}{128 k(p-1)}, \ \lg \frac{n}{2k} \right)
\end{align*}
Using Lemma~\ref{lemma:oned-massart-L}, we have that in every target interval $I_j$ with $n_j/n = \Omega(k/m)$, all but $n \cdot 2^{-\ell_2-1}$ points will be correctly classified. For large enough $m$, this means that the fraction of misclassified or unclassified points in $X$ will be at most $\epsilon$ after $O(k(p-1) \log ((p-1)/\epsilon))$ queries, apart from the $2k$ points nearest the boundaries, which will remain unclassified.
\end{proof}

The $2k$ points nearest each boundary cannot be labeled by our algorithm with any certainty because they lie in intervals with a strongly positive bias as well as in intervals with a strongly negative bias. This qualification would be removed if were allowed to make multiple queries to each point, because in that case we would include $O(k)$ copies of each point, as explained earlier.

\section{Technicalities: distributional setting}

In the distributional setting, $X$ is drawn i.i.d.\ from a distribution $\mu$ on $\X \subset \R^d$.  We assume $\mu$ is absolutely continuous with respect to the Lebesgue measure on $\R^d$ and thus admits a density. Take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$, with $\B(x)$ being balls that contain $x$.

\subsection{Sampling level and probability mass}

To begin with, we relate the number of points in a ball $B \in \B$ (and thus the level of the ball) to its probability mass under the marginal distribution.
\begin{lemma}
With probability at least $1-\delta$, for all $B \in \B$ with $n \mu(B) \geq 12 \ln (2|\B|/\delta)$, we have
$$ \frac{\mu(B)}{2} \leq \frac{|X_B|}{n} \leq 2 \mu(B) .$$
\label{lemma:ball-size-bounds}
\end{lemma}
\begin{proof}
It is an immediate consequence of the multiplicative Chernoff bound that with probability at least $1-\delta$, for all $B \in \B$,
$$ \frac{|X_B|}{n} = \mu(B) \left(1 \pm \sqrt{\frac{3}{n \mu(B)} \ln \frac{2|\B|}{\delta}} \right) .$$
\end{proof}

Henceforth assume that this high-probability event holds. Next, we will see that balls of probability mass $p$ belong to level $\ell \approx \lg (1/p)$.
\begin{lemma}
For any level $\ell \geq 0$ and any $B \in \B_{\ell}$,
$$ \left\lceil \lg \frac{1}{\mu(B)} \right\rceil -2 \leq \ell \leq  \left\lceil \lg \frac{1}{\mu(B)} \right\rceil.$$
\label{lemma:mass-vs-level}
\end{lemma}
\begin{proof}
Recall the definition of level $\ell$:
$$ B \in \B_\ell 
\ \Longleftrightarrow \ \frac{n}{2^{\ell+1}} \leq |X_B| < \frac{n}{2^\ell} 
\ \Longleftrightarrow \ 2^\ell < \frac{n}{|X_B|} \leq 2^{\ell + 1}.$$
By Lemma~\ref{lemma:ball-size-bounds}, 
$$ \frac{1}{2 \mu(B)} \leq \frac{n}{|X_B|} \leq \frac{2}{\mu(B)}.$$
Thus we must have $2^{\ell} < 2/\mu(B)$ and $1/(2 \mu(B)) \leq 2^{\ell+1}$. These translate into the stated bounds on $\ell$.
\end{proof}

\subsection{Bounding $L_2$ using probabilistic distance: Proof of Lemma~\ref{lemma:L2-bound}}

Let $p = \dist(x, \X^{-s(x)})$. If $p = 0$, the statement is vacuous, so assume $p > 0$.

Consider $\ell = \lceil \lg (1/p) \rceil + 1$. For any $B \in \B_{\geq \ell}(x)$, we have $\mu(B) \leq 2|X_B|/n < 2/2^\ell \leq p$, using Lemma~\ref{lemma:ball-size-bounds}, the definition of sampling levels, and the definition of $\ell$, in that order. It follows that $B$ does not intersect $\X^{-s(x)}$, whereupon $s(x) \cdot \eta_X(B) \geq 0$.

Next, pick any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$. Thus $B$ must intersect $\X^{-s(x)}$ and has probability mass $\geq p$. We will show that there exists $B' \in \B_{\leq \ell}(x)$ such that $B' \subset B$ and $B'$ does not intersect $\X^{-s(x)}$; whereupon $s(x) \cdot \eta_X(B') \geq 0$. Indeed, take $B' \in \B(x)$ to be a subset of $B$ of $\mu$-mass $p-\epsilon$ for some very small $\epsilon$; we can do this because of the absolute continuity of $\mu$. Then $B'$ does not touch $\X^{-s(x)}$ and by Lemma~\ref{lemma:mass-vs-level}, for small enough $\epsilon$, it lies at level $\leq \ell$.

\subsection{Uncertainty region in the distributional setting: Proof of Lemma~\ref{lemma:delta-continuous}}

Recall from (\ref{eq:sampling-region}) that
$$ \Delta_\ell = \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .$$
Consider any $z \in \Delta_\ell$. Then there exists $x \in \X$ with $L_2(x) \geq \ell$ and $B \in \B_\ell(x)$ such that $z \in B$. Now, $B \in \B_\ell(x)$ implies $|X_B|/n < 1/2^\ell$ and so (by Lemma~\ref{lemma:ball-size-bounds}) $\mu(B) < 2/2^\ell$. We will show that $x \in \partial_{4/2^\ell}$ and thus $z \in \partial_{4/2^\ell, 2/2^{\ell}}$.

There are two cases for $x$. If $\eta(x) = 0$ then we immediately have $x \in \partial_{4/2^\ell}$. Otherwise, $\eta(x) \neq 0$. In this case, since $L_2(x) \geq \ell$, we can apply Lemma~\ref{lemma:L2-bound} to get 
$$ \ell 
\ \leq \ \left\lceil \lg \frac{1}{\dist(x,\X^{-s(x)})} \right\rceil + 1 
\ \leq \ 
\lg \frac{1}{\dist(x,\X^{-s(x)})} + 2 .
$$
Thus $\dist(x, \X^{-s(x)}) \leq 1/2^{\ell-2}$ and $x \in \partial_{4/2^\ell}$.

\subsection{A generic label complexity bound in the distributional setting}

We will need to relate the size of the query region $\Delta_\ell$ to the probability mass of the corresponding second-order boundary. For this, we provide an analog of Lemma~\ref{lemma:level-distribution} for the distributional setting.
\begin{lemma}
With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/2k)$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
\item[(b)] $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 n \mu(\partial_{4/2^\ell, 2/2^\ell}) \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution-dist}
\end{lemma}
\begin{proof} 
Part (a) is as in Lemma~\ref{lemma:level-distribution}.

For (b), pick any subset $S \subset \X$. If $X$ consists of $n$ independent draws from $\mu$, then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(n, \mu(S) \tau_\ell)$ distribution with expectation $n \mu(S) \tau_\ell$. The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-n \mu(S) \tau_\ell/3}$. We will apply this to the various sets $S = \partial_{4/2^\ell, 2/2^\ell}$ and take a union bound over them. In each application, we will also see that $n \mu(S) \tau_\ell \geq k$.

Pick any $\ell \leq \lg (n/2k)$. If $\partial_{4/2^\ell, 2/2^\ell} = \emptyset$, then the statement in (b) is trivially true given Lemma~\ref{lemma:delta-continuous}. So assume this is not the case. Writing $p = 4/2^\ell$, we need to check that $\mu(\partial_{p, p/2}) \tau_\ell \geq k/n$, or equivalently, $\mu(\partial_{p, p/2}) \geq \max(1/2^{\ell+2}, k/n)$. Now, $\partial_{p,p/2} \neq \emptyset \Longrightarrow \partial_p \neq \emptyset$. Pick any $x \in \partial_p$. By absolute continuity of $\mu$, we can grow a ball around $x$ of probability mass arbitrarily close to $p/2$, so that this ball is contained within $\partial_{p,p/2}$. Thus $\mu(\partial_{p,p/2}) \geq p/2-\epsilon$ for any $\epsilon > 0$. The required conditions then follow from the value of $p$.
\end{proof}

Theorem~\ref{thm:label-complexity} now takes on the following form. Since the set of balls in $\R^d$ has VC dimension $d+1$, we can take $|\B|$ to be $O(n^{d+1})$.
\begin{thm}
Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$ and that the active learning algorithm makes $0 < m \leq n$ queries. Then with probability at least $1-3\delta$, all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where 
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
and $\ell_2$ is the largest integer $\leq \lg (n/2k)$ such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell}) \ < \ \frac{m}{32k} .$$
\label{thm:label-complexity-dist}
\end{thm}

\begin{proof}
The proof is identical to that of Theorem~\ref{thm:label-complexity-0}; the only change is to use Lemma~\ref{lemma:level-distribution-dist}(b) in place of Lemma~\ref{lemma:level-distribution}(b).
\end{proof}

\subsection{Bounding $L_1$ by curvature: Proof of Lemma~\ref{lemma:L1-bound}}

Pick any $x \in X_\gamma$; apply (A1) to get $B \in \B(x)$ for which $\mu(B) \geq p_o$ and $B \subset \X^{s(x)}_\gamma$. By Lemma~\ref{lemma:mass-vs-level}, $B \in \B_\ell$ for $\ell \leq \lceil \lg (1/p_o) \rceil$. Now, $s(x) \cdot \eta_X(B) \geq \gamma$; moreover, for any $B' \in \B(x)$ with $X_{B'} \subset X_B$ we have $X_{B'} \subset \X^{s(x)}_\gamma$ and thus $s(x) \cdot \eta_X(B') \geq \gamma$ as well.

\subsection{Label complexity under three assumptions: Proof of Theorem~\ref{thm:label-complexity-specific}}

In this case, $\B$ is infinite, but by standard VC-dimension arguments there are only $O(n^{d+1})$ balls with distinct sets $X_B$. This governs the setting of $k$.

First, define $\ell_1 = \lfloor \lg (m/32k) \rfloor$ and observe that by Lemma~\ref{lemma:L1-bound}, all points in $\X_\gamma$ have 
$$ L_1(x) 
\ \leq \ 
\left\lceil \lg \frac{1}{p_o} \right\rceil
\ \leq \ 
\lg \frac{2}{p_o} 
\ \leq \ 
\lg \frac{m}{64k}
\ \leq \ 
\ell_1 .
$$
Next, pick  
$$ \ell_2 = \left\lfloor \frac{1}{1-\sigma} \left( \lg \frac{m}{32k} + \lg \frac{1-\sigma}{C \cdot 4^{1+\sigma}} \right) \right\rfloor.$$
The upper bound on $m$ ensures that this is at most $\lg (n/2k)$. Then, using (A2),
$$
\sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell})
\ \leq \ 
C \sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \left( \frac{4}{2^\ell} \right)^\sigma 
\ \leq \ 
C \cdot 4^\sigma \cdot \frac{2}{2^{1-\sigma}-1} \cdot 2^{(1-\sigma)\ell_2}
\ \leq \ 
\frac{m}{32k} .
$$
Applying Theorem~\ref{thm:label-complexity-dist}, we then see that all points in $\X_\gamma$ with $L_2(x) \leq \ell_2$ will be correctly classified. Any remaining point $x \in \X_\gamma$ has $L_2(x) > \ell_2$ and thus (by Lemma~\ref{lemma:L2-bound})
$$ \left\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \right\rceil + 1 > \ell_2 
\ \Longrightarrow \ 
\dist(x, \X^{-s(x)}) < \frac{4}{2^{\ell_2}} \leq  \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} . 
$$
Call this quantity $p$; thus any such $x$ lies in $\partial_p$.

Under (A3), $\partial_p \cap \X_\gamma$ has zero probability mass for $p < p_1$, that is, if 
$$ \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} < p_1
\ \Longleftrightarrow \ 
m > \frac{512C}{1-\sigma} \cdot \frac{1}{p_1^{1-\sigma}} \cdot k .
$$

Under (A3'), $\mu(\partial_p \cap \X_\gamma) \leq C' p^{\xi}$; we can then apply a Bernstein bound to assert that with probability at least $1-\delta$, 
$$ |X \cap (\partial_p \cap \X_\gamma)| \ \leq \ \frac{3}{2} C' n p^{\xi} + 2 \log \frac{1}{\delta},$$
from which the bound in the theorem follows by defining $C''$ appropriately.

\subsection{Label complexity under curvature and Massart noise: Proof of Theorem~\ref{thm:massart-dist}}

Let (C1) denote the strong density condition and (C2) the Massart noise and boundary condition. Theorem~\ref{thm:massart-dist} follows immediately from the more general result of Theorem~\ref{thm:label-complexity-specific}, once we establish how (C1) and (C2) relate to assumptions (A1), (A2), and (A3').

\begin{lemma}
Conditions (C1) and (C2) yield assumption (A1), with $p_o = r_o^d \cdot c_o \cdot v_d$, where $v_d$ is the volume of the unit ball in $\R^d$.
\label{lemma:A1}
\end{lemma}

\begin{proof}
Pick any $x \in \X_\gamma^+$ (the negative case is similar), and let $r = \inf_{z \in \X^0} \|x - z\|$. If $r \geq r_o$, then $B = B(x,r_o)$ is entirely in $\X_\gamma^+$. Otherwise, the reach condition (C2) implies the existence of a ball $B \subset \X_\gamma^+$ that contains $x$ and has radius $r_o$. Either way, $\mu(B) \geq c_o \vol(B) = c_o v_d r_o^d$ by (C1). 
\end{proof}

\begin{lemma}
Conditions (C1) and (C2) yield assumptions (A2) and (A3') with $\sigma = \xi = 1/d$. 
\end{lemma}

\begin{proof}
Under (C1), any ball of probability mass $\leq p$ has volume $\leq p/c_o$ and radius $\leq (p/(c_o v_d))^{1/d}$. Let's call this latter quantity $r$. Thus, any point in $\partial_p$ lies within distance $2r$ of the boundary, while a point in $\partial_{p,p}$ lies within distance $4r$.

To bound the volume of $\partial_{p,p}$, we can associate each point in this region with its nearest neighbor in $\X^0$; by condition (C1), this projection map is uniquely defined for $r < r_o/4$. The volume of the region is thus $O(r)$ and under (C1), has probability mass $O(r) = O(p^{1/d})$.
\end{proof}

\subsection{Label complexity under smoothness and Tsybakov noise}
\label{sec:Tsybakov}

Now we specialize Theorem~\ref{thm:label-complexity-specific} to another situation, where the data distribution satisfies Holder smoothness and Tsybakov margin conditions.

Recall that $\dist(x,S)$ denotes the probability-distance from $x$ and set $S$. We will overload notation so that for $x,x' \in \X$,
$$ \dist(x,x') = \dist(x, \{x'\}) = \inf\{\mu(B): B \in \B(x) \cap \B(x')\},$$
that is, the probability mass of the smallest ball containing both $x$ and $x'$. We will impose the following conditions on the data.
\begin{enumerate}
\item[(C1)] [Strong density condition] This was introduced in the earlier (Massart) example.
\item[(C2')] [Holder-smoothness of conditional probability function] There exist constants $L, \alpha$ such that
$$ |\eta(x) - \eta(x')| \ \leq \ L \cdot \dist(x,x')^\alpha$$
for all $x,x' \in \X$.
\item[(C3')] [Tsybakov margin condition] There exists constants $M, \beta$ such that
$$ \mu(\{x \in \X: |\eta(x)| \leq \tau\}) \leq M \tau^\beta$$
for all $\tau \in (0,1)$.
\item[(C4')] [Bounded curvature] The boundaries $\{x \in \X: \eta(x) = \gamma\}$ and $\{x \in \X: \eta(x) = -\gamma\}$ are $(d-1)$-dimensional Riemannian manifolds of reach $r_o > 0$.
\end{enumerate}

Lemma~\ref{lemma:A1} continues to hold, with condition (C4') doing the job of (C2). This yields assumption (A1). For the remaining assumptions, we first obtain a consequence of the Holder condition.

\begin{lemma}
Under (C2'), for any $p, q > 0$,
\begin{enumerate}
\item[(a)] $x \in \partial_p \implies |\eta(x)| \leq L p^\alpha$.
\item[(b)] $x \in \partial_{p,q} \implies |\eta(x)| \leq L(p^\alpha + q^\alpha)$.
\end{enumerate}
\label{lemma:holder-consequence}
\end{lemma}
\begin{proof}
Pick any $x \in \partial_p$. Let $s(x) = \mbox{sign}(\eta(x))$. By definition of the $p$-boundary set, for any $\epsilon > 0$, there exists $x' \in \X^{-s(x)}$ such that $\dist(x,x') < p+\epsilon$. By the Holder condition, $|\eta(x) - \eta(x')| < L(p+\epsilon)^\alpha$ and thus $|\eta(x)| < L(p+\epsilon)^\alpha$. Since this holds for any $\epsilon > 0$, we get part (a).

For (b), pick $x \in \partial_{p,q}$ and $\epsilon > 0$. Then there exists $x' \in \partial_p$ with $\dist(x,x') < q+\epsilon$. As before, we use the Holder condition to conclude that $|\eta(x)| < |\eta(x')| + L(q+\epsilon)^\alpha$ and then invoke (a).
\end{proof}

\begin{lemma}
Conditions (C2') and (C3') yield assumptions (A2) and (A3) with $C = (2L)^\beta M$, $\sigma = \alpha \beta$, and $p_1 = (\gamma/L)^{1/\alpha}$.
\label{lemma:boundary-constants} 
\end{lemma}

\begin{proof}
By Lemma~\ref{lemma:holder-consequence}, $\partial_{p,p} \subset \{x \in \X: |\eta(x)| \leq 2Lp^\alpha\}$; the probability mass of this set can be bounded by (C3').

Also by Lemma~\ref{lemma:holder-consequence}, $\partial_p$ is entirely contained in $\{x \in \X: |\eta(x)| \leq L p^\alpha\}$. For $p < p_1$, this does not intersect $\X_\gamma$.
\end{proof}

\begin{thm}
Assume conditions (C1) and (C2')--(C4'). There are constants $c_2, c_3$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. Suppose the algorithm of Fig.~\ref{alg:main} makes $m$ queries, where
$$ \frac{c_2k}{p_1^{1-\sigma}} \ \leq \ m \ \leq \ c_3 n^{1-\sigma} ,$$
with $p_1, \sigma$ as given in Lemma~\ref{lemma:boundary-constants}. Then with probability at least $1-\delta$, all of $X \cap \X_\gamma$ is assigned Bayes-optimal labels.
\label{thm:smoothness-dist}
\end{thm}

In this setting, regular $k$-nearest neighbor classification would also correctly classify all of $X \cap X_\gamma$ if given $O(k/p_1)$ random labeled points. The benefit of active learning is to thus reduce the label complexity to $O(k/p_1^{1-\alpha\beta})$.
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2024-07-31 COLT/COLT-old/active-knn-old.tex
Initial Overleaf Import

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\yh{{\widehat{y}}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\sign{{\rm sign}}
\def\PL{{\mbox{\rm PL}}}

\newtheorem{thm}{Theorem}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{claim}[thm]{Claim}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assump}{Assumption}
\newtheorem{open}{Open problem}
\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}

\newcommand{\comment}[3]{{\color{#1} {\bf #2 :} #3}}
%\newcommand{\comment}[3]{}  % suppress comments
\newcommand{\yoav}[1]{\comment{cyan}{Yoav}{#1}}
\newcommand{\sanjoy}[1]{\comment{blue}{Sanjoy}{#1}}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{Active learning using region-based sampling}

\begin{document}

\maketitle

\section{Setup}

\begin{itemize}
\item {\it Underlying model of data.}

\begin{itemize}
\item Instances lie in a metric space $(\X, d)$. 
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to an unknown conditional probability function $\Pr(Y|X)$, with
$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x) = 2 \pr(Y=1|X=x) - 1 .$$
\item The {\it Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
\end{itemize}

\item {\it Task: Finite population labeling.}

\begin{itemize}
\item We are given a finite dataset $X \subset \X$ of $n$ points, and have the ability to query the labels of any of them. We are also given parameters $0 < \gamma, \delta < 1$.
\item We want a procedure that chooses the next point, or batch of points, to query. This process will be applied repeatedly and then stopped at an arbitrary time, whereupon labels $\yh(x)$ must be provided for all $x \in X$ (including those that were queried).
\item We would like the inferred labels $\yh(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
\item Specifically, we are judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
$$ \sum_{x \in X} {\bf 1}(\yh(x) \neq g^*(x),\ |\eta(x)| > \gamma) .$$
The procedure is allowed to fail with probability $\delta$, to account for sampling error.
\end{itemize}

\end{itemize}

\section{Active learning algorithm}

\subsection{Some details}

\begin{enumerate}

\item {\it Sampling regions.}
The sampling is organized around a collection $\B$ of measurable subsets of $\X$. These are the atomic sets on which we assess label bias and we call them ``balls''.

\begin{enumerate}

\item[(a)] {\it Points within a ball.} For any ball $B \in \B$, let $X_B = X \cap B$ denote the points in $X$ that will be used to assess the bias of $B$.\footnote{Sometimes, $\B$ might be constructed in a way that depends on $X$ and causes certain points to be excluded. For instance, consider these two alternatives for defining $\B$:
\begin{itemize}
\item $\B$ consists of a pre-defined set of balls.
\item $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
\end{itemize}
In the first case, all points $X \cap B$ are random draws from $B$. In the second case, this is also true except for the two points $x,x'$ that define the ball.}

\item[(b)] {\it Grouping balls by size.} We will group balls by the number of points that they contain: we put $B$ at {\it level} $\ell$ if
$$ \frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .$$
Here $\ell$ ranges from $0$ (consisting of balls that contain at least half the data points) to roughly $\lg n$ (containing a single point).

Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of very large balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls.

\item[(c)] {\it Balls used to evaluate a given point.}

For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that will be used in determining $x$'s label. We will take these to be all balls that contain $x$. As before, we can partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.

\end{enumerate}

\item {\it Bias-estimates for balls and for data points.} The active learning algorithm uses label-queries to assess the bias of balls $B \in \B$. In turn, these are used to estimate the labels of individual points.

\begin{enumerate}

\item[(a)] {\it Bias of a ball.} The bias of a ball $B \in \B$ is defined as
$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$

\item[(b)] {\it Qualitative estimates of bias.} Each ball $B$ is assigned qualitative bias estimate,
$$ \yh(B) = 
\left\{
\begin{array}{cl}
+1 & \mbox{significant $+$ bias} \\
-1 & \mbox{significant $-$ bias} \\
0 & \mbox{no significant bias} \\
\bot & \mbox{not yet available}
\end{array}
\right.
$$
The option $\bot$ is used until sufficiently many points in $X_B$ have been labeled: the required number, $k$, is roughly $1/\gamma^2$ (recall that $\gamma$ is the smallest bias that needs to be detected). Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.

\item[(c)] {\it Bias guarantees.} Our bias estimates will satisfy a guarantee of the following form: with probability $1-\delta$, for all $B \in \B$, once $\yh(B)$ is set to a value in $\{-1,0,+1\}$, 
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
\end{itemize}
for some constants $0 < b_1 < b_2 < 1$.

\item[(d)] {\it supported labels for a point.} Pick an instance $x$ and a level $\ell$. We say that a ball $B \in \B_\ell(x)$ {\em supports} the label $s \in \{-1,+1\}$ if $\yh(B) = s$. The set of {\em supported labels} for $x$ at level $\ell$ is defined as:
\begin{equation}
\PL_\ell(x) = \{s \in \{-1, +1\}: \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = s$} \} 
\label{eq:PL}
\end{equation}

\item[(e)] {\it The label of a point at level $\ell$.} The label of point $x$ at level $\ell$ is a value $\yh_\ell(x) \in \{-1,+1,0,!,\bot\}$. It is initially $\bot$ (meaning ``not yet available''), but once bias-estimates are available for all $B \in \B_\ell(x)$, it is set to a value in $\{-1,+1,0,!\}$ and remains fixed thereafter:
\begin{equation}
\yh_\ell(x) = 
\left\{
\begin{array}{cl}
+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
\end{array}
\right.
\label{eq:provisional-label}
\end{equation}
\end{enumerate}
\end{enumerate}

It is worth noting that, while non-parametric methods, of the type described here are guaranteed to converge to the Bayes optimal classifier, the algorithm never knows whether the label on any point $x \in \X$ has converged. We say that our algorithm had a {\em mind change} on $x$ if $\yh_\ell(x)=+1$ and $\yh_{\ell+1}=-1$ or vice versa. It is not hard to construct cases where the number of mind changes for a point $x$ is arbitrarily large.

In other words the algorithm never knows what is the Bayes optimal label of any point. We divide the points into those labeled $!$ which are the {\em known unknowns} and all of the other points which are the {\em unknown unknowns}.
The basic idea of our algorithm is to focus half of its queries on the known unknowns, and the other half  querying all instances.

\subsection{The algorithm}

The overall active learning algorithm is in Figure~\ref{alg:main}. It makes calls to subroutines for background and focused sampling, shown in Figure~\ref{alg:sampling}. The querying process can be stopped at any time, whereupon labels will be assigned as follows:
\begin{equation}
\yh(x) = 
\left\{
\begin{array}{ll}
\yh_\ell(x) & \mbox{for the largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
0 & \mbox{(don't know), otherwise}
\end{array}
\right.
\label{eq:final-label}
\end{equation}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.4in}
\begin{itemize}
\item Initialize uncertainty regions at all levels:
\begin{itemize}
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}
\item Initialize labels at all levels to ``unavailable'':
\begin{itemize}
\item $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
\end{itemize}
\item Repeat:
\begin{itemize}
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}
\item Let $\ell'$ be the smallest such level
\item {\tt Focused-query}($\ell', U_{\ell'}$)
\end{itemize}
\item {\tt Background-query}
\item Update labels:
\begin{itemize}
\item Get bias-estimate $\yh(B) \in \{-1,0,+1\}$ for all $B \in \B$ that are now adequately sampled
\item For each $x \in X$ and level $\ell \geq 0$ for which all of $\B_\ell(x)$ is now sampled:
\begin{itemize}
\item Determine the possible labels for $x$ at level $\ell$:
$$\PL_\ell(x) = \{s \in \{-1, +1\}: \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = s$} \}$$
\item Set
$$
\yh_\ell(x) = 
\left\{
\begin{array}{cl}
+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
\end{array}
\right.
$$
\end{itemize}
\end{itemize}
\item Update uncertainty regions:
\begin{itemize}
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
\end{itemize}
\end{itemize}

\end{itemize}

\end{minipage}}
\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query. The subroutines {\tt Focused-query} and {\tt Background-query} appear in Figure~\ref{alg:sampling}.}
\label{alg:main}
\end{figure}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.4in}

\vspace{.1in}
\emph{Initialization:}
\begin{itemize}
\item Set $Q = \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
\end{itemize}

\vspace{.2in}
{\bf Focused-query}($\ell, U$)

\vspace{.1in}
\begin{itemize}
\item Define querying region:
$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
\item Query the next unlabeled point in $S \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\vspace{.2in}
{\bf Background-query}

\vspace{.1in}
\begin{itemize}
\item Query the next unlabeled point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\end{minipage}}
\caption{The sampling procedures. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried. The focused querying procedure uses level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$.} 
\label{alg:sampling}
\end{figure}

\section{Analysis of algorithm: finite population setting}

We now analyze the active learning procedure for the finite population annotation problem introduced earlier.

\subsection{Accuracy of bias estimates}

We'll start with a uniform guarantee on the bias estimates for all balls $B \in \B$.

Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
$$ \eta(x) = \E[Y | X=x].$$ 
For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value of the points in $B$, that is,
$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
The specific property we need is encapsulated in the following definition.

\yoav{I think instead of this definition it would be  better to point to item (c) "Bias estimatess ..."}
\begin{defn}
For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$(b_1,b_2)$-accurate} if the following holds:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
\end{itemize}
\label{def:accurate-bias-estimate}
\end{defn}
For reference, we will take $b_1 = \gamma/2$ and $b_2 = \gamma$, where $\gamma$ is the minimum bias we need to detect, as per the problem specification.

In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
\begin{equation}
c_1 = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
\label{eq:c1}
\end{equation}

For any level $\ell$, define threshold $\tau_\ell$ as follows:
\begin{equation}
\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
\label{eq:threshold-ell}
\end{equation}
\yoav{What is $k$? Oh, I think it is the minimal number of queries to identify a bias $\gamma$,... not sure where it is defined.}

For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.

Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-discrete} in the appendix.
\begin{thm}[checked]
Suppose $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c_1}{\sqrt{k}}.$$
\end{enumerate}
\label{thm:large-deviation-bounds}
\end{thm}
\begin{proof}
These follow immediately by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. In particular, note that the choice of $\tau_\ell$ ensures $|X_B| \tau_\ell \geq 2k$.
\end{proof}

The following corollary of Theorem~\ref{thm:large-deviation-bounds} is immediate.
\begin{cor}[checked]
Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $(b_1,b_2)$-accurate.
\label{cor:accurate-bias-estimates}
\end{cor}
\begin{proof}
By the choice of $k$, we have from Theorem~\ref{thm:large-deviation-bounds} that 
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| < \frac{b_2 - b_1}{2} $$
for all $B \in \B$.
\end{proof}

In what follows, we will assume that this $(1-\delta)$-probability event holds.

\subsection{Mind changes}

The label assigned to a data point $x$ at level $\ell$, which we denote $\yh_\ell(x)$, might change as $\ell$ increases; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.

We distinguish two critical levels for each $x$: the level $L_s(x)$ by which $\yh_\ell(x)$ has ``stabilized'' in the sense that it no longer flips between $+1$ and $-1$ (although it may be $!$), and the level $L_f(x)$ by which $\yh_\ell(x)$ is correct and thus ``finalized''. See Figure~\ref{fig:two-levels} for a pictorial view.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{home-stretch.pdf}
\end{center}
\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively finer levels. Beyond level $L_s(x)$, this $\yh_\ell(x)$ will never be the wrong label, but might be $!$. Beyond level $L_f(x)$, the correct label will be assigned.}
\label{fig:two-levels}
\end{figure}

\begin{defn}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $g^*(x) = \sign(\eta(x))$ be its Bayes-optimal label. We define $L_s(x)$ to be the smallest integer such that for all $\ell \geq L_s(x)$, the following condition holds:
\begin{enumerate}
\item[1.] Some $B \in \B_\ell(x)$ has $g^*(x) \cdot \eta_X(B) \geq b_2$.
\end{enumerate}
(If there is no such level, we define $L_s(x) = \infty$.) We define $L_f(x)$ to be the smallest integer such that for all $\ell \geq L_f(x)$, the above condition holds, and additionally:
\begin{enumerate}
\item[2.] Every $B \in \B_\ell(x)$ has $g^*(x) \cdot \eta_X(B) \geq -b_1$.
\end{enumerate} 
(Again, take $L_f(x) = \infty$ if there is no such level.) 
\label{defn:Lsf}
\end{defn}
Condition 1 says that there is a good neighborhood $B \in \B_\ell(x)$, one with a strong bias towards the correct label of $x$. Condition 2 says that there are no bad neighborhoods, that are weakly biased towards the wrong label.

Thus $L_s(x) \leq L_f(x)$, and there might be a large gap between the two. Before the background sampling reaches level $L_s(x)$, it is quite possible that we will predict the wrong label for $x$. However, once all balls in $\B_{L_s(x)}(x)$ have been adequately sampled, point $x$ will either (i) be assigned label $!$ and put in the uncertainty set, in which case it will benefit from focused sampling, or (ii) will get the correct label. Once the balls in $\B_{L_f(x)}(x)$ have been sampled, condition (ii) will hold.

\begin{lemma}[checked]
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $g^*(x) \in \{+1,-1\}$ denote its Bayes-optimal label. Then for any level $\ell$ and any time at which $\yh_\ell(x) \neq \bot$:
\begin{enumerate}
\item[(a)] If $\ell \geq L_s(x)$, then $\yh_\ell(x) \in \{g^*(x), !\}$.
\item[(b)] If $\ell \geq L_f(x)$, then $\yh_\ell(x) = g^*(x)$.
\end{enumerate}
\label{lemma:boundary}
\end{lemma}
\begin{proof}
For part (a), note that if $\ell \geq L_s(x)$, then some $B \in \B_\ell(x)$ has significant bias towards the correct label $g^*(x)$, in which case (by Corollary~\ref{cor:accurate-bias-estimates}) $\yh(B) = g^*(x)$. Thus $\yh_\ell(x)$ must either be $!$ (if there is some $B' \in \B_\ell(x)$ with $\yh(B') = -g^*(x)$) or $g^*(x)$.

For (b), both conditions (1) and (2) of Definition~\ref{defn:Lsf} hold, so there is at least one $B \in \B_\ell(x)$ with $\yh(B) = g^*(x)$ and no $B$ with $\yh(B) = -g^*(x)$. Thus $\yh_\ell(x) = g^*(x)$.
\end{proof}

\subsection{Bounding the amount of focused sampling}

From Lemma~\ref{lemma:boundary}(b), we see that the label of a point $x$ will be apparent by the time it has been sampled at level $L_f(x)$, and thus $x$ will not appear in the uncertainty set for subsequent levels. That is, 
$$U_{\ell} \subset \{x \in X: L_f(x) \geq \ell\}.$$ 
This in turns gives a simple characterization of the points that are queried as part of focused sampling.
\begin{lemma}[checked]
All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$, where
\begin{equation}
\Delta_\ell 
= \bigcup_{x \in X: L_f(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B 
.
\label{eq:sampling-region}
\end{equation}
\label{lemma:focused}
\end{lemma}
\begin{proof}
Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$. As is apparent from the {\tt Focused-query} subroutine (Figure~\ref{alg:sampling}), all focused samples at level $\ell$ lie in
$$
\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
\subset 
\ 
\bigcup_{x \in X: L_f(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
=
\ 
\{z \in \Delta_\ell: T_z \leq \tau_\ell \},
$$
where the first containment is a consequence of Lemma~\ref{lemma:boundary}.
\end{proof}

A final property we will need for label complexity bounds is that various subsets of interest contain roughly the expected number of points at each level.
\begin{lemma}
Pick any $0 < \ell_o \leq \lg n$. With probability at least $1-2(\ell_o+1)e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \ell_o$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| < 2 n \tau_\ell$.
\item[(b)] If $|\Delta_\ell| \tau_\ell \geq k$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| < 2 |\Delta_\ell| \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution}
\end{lemma}
\begin{proof} 
Pick any subset $S \subset X$ and let $m = |S|$. Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$. The probability that it is greater than or equal to twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m \tau_{\ell}/3}$.

Both parts follow from this principle; and we take a union bound over all $\ell_o + 1$ levels.
\end{proof}

Henceforth we assume that these two bounds hold.

\subsection{Label complexity bounds}

\begin{thm}
Suppose the active learning algorithm is allowed to make $0 < m \leq n$ queries. Define
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor .$$
and let $\ell_2$ be the largest integer such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} |\Delta_{\ell}| \, \tau_\ell \ \leq \ \frac{m}{8} .$$
Then any point $x \in X$ with $L_s(x) \leq \ell_1$ and $L_f(x) \leq \ell_2$ will have final label $\yh(x) = g^*(x)$.
\label{thm:label-complexity}
\end{thm}
\begin{proof}
Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}. We will analyze the effect of background sampling in phase one and focused sampling in phase two. We start with the former.

Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.

Now, for $\ell_1$ as defined, we have from (\ref{eq:threshold-ell}) that $\tau_{\ell_1} \leq m/8n$ and thus from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy $T_x \leq \tau_{\ell_1}$. Therefore all such points are queried in phase one, and all label-estimates $\{\yh_{\ell_1}(x): x \in X\}$ are set.

It then follows from Lemma~\ref{lemma:boundary} that the following hold for any $x \in X$ with $L_s(x) \leq \ell_1$:
\begin{enumerate}
\item[(a)] By the end of phase one, $\yh_{\ell_1}(x) \in \{g^*(x), !\}$.
\item[(b)] For any $\ell > \ell_1$, when $\yh_\ell(x)$ becomes available, it lies in $\{g^*(x), !\}$.
\item[(c)] If $x$ ever leaves the combined uncertainty region $U = \cup_\ell U_\ell$ during phase two, then its final label as defined in (\ref{eq:final-label}) is henceforth always $\yh(x) = g^*(x)$.
\end{enumerate}

Now let's move on to phase two. Let $A = \{x \in X: L_s(x) \leq \ell_1, L_f(x) \leq \ell_2\}$. We will show that every point in $A$ must leave the uncertainty region $U$ at some time during phase two. From (c), we can conclude that all these points have their final labels set correctly, once and for all.

We will break the argument into two cases.

{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.

{\it Case 2:} A full $m/4$ focused queries are made in phase two. By the analysis of phase one, none of these queries can be at level $\leq \ell_1$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_1+1$ through $\ell_2$ inclusive is at most
$$\sum_{\ell=\ell_1+1}^{\ell_2} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
\ < \ 
\sum_{\ell=\ell_1+1}^{\ell_2} 2 |\Delta_{\ell}| \tau_\ell
\ \leq \ 
\frac{m}{4} ,
$$
where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus at least one query in phase two must be at level $\ell_2 +1$. When this query is made, every $U_\ell$ with $\ell \leq \ell_2$ must be empty and thus all of $A$ must have left the uncertainty region (no $x \in A$ can be part of $U_\ell$ for $\ell > \ell_2 \geq L_f(x)$).

Thus every $x \in A$ must leave the uncertainty region at some point in phase two, and their final labels are subsequently set correctly.
\end{proof}

\subsection{Example: One-dimensional data with Massart noise}

In order to apply Theorem~\ref{thm:label-complexity}, we need
\begin{itemize}
\item upper bounds on the critical levels $L_s(x)$ and $L_f(x)$ for each point $x \in X$, and
\item upper bounds on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
\end{itemize}
We now illustrate how this works out in a canonical setting.

Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
\begin{itemize}
\item There are $p$ disjoint open intervals $I_1, \ldots, I_p$, such that $\X$ is (the closure of) their union, and
\item for each $j$, either $\eta(x) > \gamma$ for all $x \in I_j$ or $\eta(x) < -\gamma$ for all $x \in I_j$. In the first case, we write $g^*(I_j) = +1$ and in the second case, $g^*(I_j) = -1$.
\end{itemize}
Here $0 < \gamma < 1$ is some constant. See Figure~\ref{fig:oned-massart} for an illustrative example.

\begin{figure}
\begin{center}
\includegraphics[width=3in]{oned-massart.pdf}
\end{center}
\caption{In this example the conditional probability function $\eta$ is defined by four intervals; on each, $\eta$ is either $> \gamma$ or $<-\gamma$.}
\label{fig:oned-massart}
\end{figure}

We will take $\B$ to consist of all open intervals of $[0,1]$, with $\B(x)$ denoting intervals that contain point $x$. We will choose accuracy parameters $b_1, b_2$ so that $0 < b_1 < b_2 \leq \gamma$.

The intervals $I_j$ can be written in the form $(\lambda_{j-1}, \lambda_j)$, where
$$ 0 = \lambda_0 < \lambda_1 < \lambda_2 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$$
Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
\begin{lemma}
Pick any $x \in \X$; suppose $x \in I_j$.
\begin{enumerate}
\item[(a)] Let $n_j = |X \cap I_j|$. Then $L_s(x) \leq \lg (n/n_j)$.
\item[(b)] Let $r(x)$ be the minimum number of data points that lie between $x$ and a boundary point, counting $x$ as well; this is either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller. Then $L_f(x) \leq \lg (n/r(x))$.
\end{enumerate}
\label{lemma:oned-massart-Lsf}
\end{lemma}
\begin{proof}
For (a), notice first that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$. Moreover, $g^*(I_j) \cdot \eta_X(I_j) > \gamma \geq b_2$. Thus $I_j$ belongs to $\B_\ell(x)$ and is strongly biased towards the correct label. For levels $> \ell$, we can show a similar property for suitable sub-intervals of $I_j$.

For (b), consider any $\ell \geq \lceil (\lg n/r(x)) - 1 \rceil$. First of all, $\B_\ell(x)$ includes some sub-interval $B \subset I_j$ which is strongly biased towards the correct label: $g^*(I_j) \cdot \eta_X(B) > \gamma \geq b_2$. At the same time, $\B_\ell(x)$ does not contain any sets that are weakly biased towards the wrong label. To see this, notice that any $B \in \B(x)$ with $\mbox{sign}(\eta_X(B)) \neq g^*(I_j)$ must contain either the entire left-interval $(\lambda_{j-1},x]$ or the entire right-interval $[x, \lambda_j)$, and thus has at least $r(x) + 1$ points, which means that it is too large to be in $\B_\ell$.
\end{proof}

Based on these $L_s(x)$ and $L_f(x)$, it is easy to bound the size of the focused query region $\Delta_\ell$ at each level.
\begin{lemma}
For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq (p-1)n/2^{\ell-2}$.
\label{lemma:oned-massart-query-region}
\end{lemma}

\begin{proof}
We have
\begin{align*}
\Delta_\ell 
&= \bigcup_{x \in X: L_f(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $n/2^{\ell-1}$ points from $X$ on either side of each boundary point $\lambda_j$.
\end{proof}

\begin{thm}
Pick any $0 < \epsilon, \delta < 1$. Suppose we run the algorithm of Figure~\ref{alg:main} with $b_1 = \gamma/2$ and $b_2 = \gamma$. With probability at least $1-\delta$, after making
$$ O \left( \frac{p-1}{\gamma^2} \ln \frac{p-1}{\epsilon} \ln \frac{n}{\delta} \right) $$
queries, the algorithm will assign the correct label $g^*(x)$ to at least $1-\epsilon$ fraction of $X$.
\label{thm:oned-massart}
\end{thm}
\begin{proof}
Under this choice of $b_1$ and $b_2$, we will set $k$ to $O((1/\gamma^2) \ln (n/\delta))$, as per Corollary~\ref{cor:accurate-bias-estimates}. Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.

Next, using Lemma~\ref{lemma:oned-massart-query-region}, we have that for any integers $0 \leq \ell_1 < \ell_2$,
$$ \sum_{\ell = \ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell
\ \leq \ 
\sum_{\ell = \ell_1+1}^{\ell_2} \frac{(p-1) n}{2^{\ell-2}} \cdot \frac{2^{\ell+2}k}{n}
\ = \ 
16(p-1)k (\ell_2 - \ell_1).
$$
We can then apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_s(x) \leq \ell_1$ and $L_f(x) \leq \ell_2$, for
\begin{align*}
\ell_1
&= 
\left\lfloor \lg \frac{m}{32k} \right\rceil \\
\ell_2
&=
\ell_1 + \frac{m}{128 k(p-1)}
\end{align*}
Using Lemma~\ref{lemma:oned-massart-Lsf}, we have that in every target interval $I_j$ with $n_j/n = \Omega(k/m)$, all but $n \cdot 2^{-\Omega(m/k(p-1))}$ points will be correctly classified. For large enough $m$, this means that the fraction of misclassified or unclassified points in $X$ will be at most $\epsilon$ after $O(k(p-1) \log ((p-1)/\epsilon))$ queries.
\end{proof}

\subsection{Example: One-dimensional data with monotonic $\eta$}

We now turn to another one-dimensional setting. Once again, $X$ consists of $n$ arbitrarily-placed points in $\X = [0,1]$. This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing. Let $\lambda \in (0,1)$ denote the point for which $\eta(\lambda) = 0$.

As before, we will take $\B$ to consist of all open intervals, and $\B(x)$ to be intervals containing $x$. Given a threshold $\gamma > 0$ such that we do not need to label points with $|\eta(x)| < \gamma$, we can set accuracy parameters $b_2 = \gamma, b_1 = \gamma/2$.

Let $\lambda_L, \lambda_R$ be the points for which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$. Thus we are not required to label points in the interval $[\lambda_L, \lambda_R]$. See Figure~\ref{fig:oned-monotonic} for an example.

\begin{figure}
\begin{center}
\includegraphics[width=3in]{oned-monotonic.pdf}
\end{center}
\caption{The conditional probability function $\eta$ is monotonically increasing. The values $x = \lambda_L, \lambda, \lambda_R$ have $\eta(x) = -\gamma, 0, \gamma$ respectively.}
\label{fig:oned-monotonic}
\end{figure}

\begin{lemma}
Pick any $x \in [0, \lambda_L] \cup [\lambda_R, 1]$.
\begin{enumerate}
\item[(a)] Define $n^- = |[0,\lambda_L) \cap X|$ and $n^+ = |(\lambda_R,1] \cap X|$. Then
$$
L_s(x)
\leq
\left\{
\begin{array}{ll}
\lg n/n^+ & \mbox{if $x > \lambda_R$} \\
\lg n/n^- & \mbox{if $x < \lambda_L$}
\end{array}
\right.
$$
\item[(b)] Let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$, counting $x$ itself. That is, $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$. Then
$$ L_f(x) \leq \max\left( L_s(x), \left\lceil \lg \frac{n}{r(x)} -1 \right\rceil \right).$$
\end{enumerate}
\label{lemma:oned-monotonic-Lsf}
\end{lemma}
\begin{proof}
For (a), take any $x > \lambda_R$ (the other case is similar). The interval $B = (\lambda,1)$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta_X(B) > b_2$. For larger $\ell$, we get a similar result with a suitable sub-interval of $(\lambda,1)$.

For (b), note that for $\ell \geq \lceil \lg (n/r(x)) - 1 \rceil$, there is no $B \in \B_\ell(x)$ that extends to the other side of the boundary. Moreover, for $\ell \geq L_s(x)$, there is at least one interval in $\B_\ell(x)$ with bias of the correct sign and magnitude at least $b_2$.
\end{proof}

\section{Label complexity bounds in the distributional setting}

We now turn to a setting where the points $X$ are sampled from a distribution $\mu$ on $\X$. The algorithm is unchanged, but we seek to bound label complexity in terms of properties of the joint distribution specified by $\mu$ and $\eta$.

\subsection{Sampling level and probability mass}

To begin with, we relate the number of points in a ball $B \in \B$ (and thus the level of the ball) to its probability mass under the marginal distribution.
\begin{lemma}
With probability at least $1-\delta$, for all $B \in \B$ with $n \mu(B) \geq 12 \ln (2|\B|/\delta)$, we have
$$ \frac{\mu(B)}{2} \leq \frac{|X_B|}{n} \leq 2 \mu(B) .$$
\label{lemma:ball-size-bounds}
\end{lemma}
\begin{proof}
It is an immediate consequence of the multiplicative Chernoff bound that with probability at least $1-\delta$, for all $B \in \B$,
$$ \frac{|X_B|}{n} = \mu(B) \left(1 \pm \sqrt{\frac{3}{n \mu(B)} \ln \frac{2|\B|}{\delta}} \right) .$$
\end{proof}

Henceforth assume that this high-probability event holds. Now, if $B \in \B_\ell$, then
$$ \mu(B) \leq \frac{2 |X_B|}{n} < \frac{2}{2^\ell} $$
and thus $\ell < \lg (2/\mu(B))$. A similar chain of inequalities in the other direction yields $\ell \geq \lg 1/(4 \mu(B))$. We thus see that balls of probability mass $p$ belong to level $\ell \approx \lg (1/p)$.

\subsection{Adjustments to the finite population analysis}

In the distributional setting, our notion of the \emph{average bias} of a ball $B \in \B$ changes from the quantity $\eta_X(B)$ defined earlier (namely, the average $\eta$-value of the finite set of points $X \cap B$) to
$$ \eta(B) = \frac{1}{\mu(B)} \int \eta \, d\mu ,$$
which is defined as long as $\mu(B) > 0$. Correspondingly, we relate the accuracy of bias estimates $\yh(B)$ to $\eta(B)$ rather than $\eta_X(B)$.
\begin{defn}
For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{distributionally $(b_1,b_2)$-accurate} if the following holds:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta(B) > b_1$
\item $\yh(B) = -1 \implies \eta(B) < -b_1$
\item $\yh(B) = 0 \implies |\eta(B)| < b_2$
\end{itemize}
\label{def:accurate-bias-estimate-dist}
\end{defn}

We can show that Theorem~\ref{thm:large-deviation-bounds} and Corollary~\ref{cor:accurate-bias-estimates} continue to hold with $\eta(B)$ in place of $\eta_X(B)$. We state the result here for convenience.
\begin{thm}
Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are distributionally $(b_1,b_2)$-accurate.
\label{thm:accurate-bias-estimates-dist}
\end{thm}

We need to make a similar small adjustment to the definitions of $L_s$ and $L_f$, merely swapping $\eta_X(B)$ for $\eta(B)$. Under the modified definitions, Lemma~\ref{lemma:focused} continues to hold; we restate it here for reference.
\begin{lemma}
All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$, where
$$ \Delta_\ell = \bigcup_{x \in X: L_f(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .$$
\label{lemma:focused-dist}
\end{lemma}

In a similar vein, we have an analog of Lemma~\ref{lemma:level-distribution} for the distributional setting.
\begin{lemma}
With probability at least $1-2(\ell_o+1)e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \ell_o$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
\item[(b)] If $\mu(\Delta_\ell) \tau_\ell \geq k/n$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 n \mu(\Delta_\ell) \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution-dist}
\end{lemma}
\begin{proof} 
  Part (a) is as in Lemma~\ref{lemma:level-distribution-dist}. For (b), pick any subset $S \subset \X$. If $X$ consists of $n$ independent draws from $\mu$, then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(n, \mu(S) \tau_\ell)$ distribution with expectation $n \mu(S) \tau_\ell$. The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-n \mu(S) \tau_\ell/3}$.
\end{proof}

Theorem~\ref{thm:label-complexity} continues to hold, in the following form. The only change to the proof is to use Lemma~\ref{lemma:level-distribution-dist}(b) in place of Lemma~\ref{lemma:level-distribution}(b).
\begin{thm}
Suppose the active learning algorithm is allowed to make $0 < m \leq n$ queries. Define
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor .$$
and let $\ell_2$ be the largest integer such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} 2^\ell \mu(\Delta_{\ell}) \ \leq \ \frac{m}{32k} .$$
Then any point $x \in X$ with $L_s(x) \leq \ell_1$ and $L_f(x) \leq \ell_2$ will have a final label $\yh(x)$ that is correctly set to $g^*(x)$.
\label{thm:label-complexity-dist}
\end{thm}

\pagebreak

\appendix

\section{Technicalities}

\subsection{Large deviation bounds for the discrete setting}

\begin{lemma}[rechecked]
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
Now consider the following estimator $Z$ of this quantity:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta_o| < \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-discrete}
\end{lemma}
\begin{proof}
Let's start with (a). Define $c_1 = \sqrt{3 \ln (4/\delta)}$. We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c_1 \sqrt{2k}$ since this is the worst case. By assumption, $k \geq 2c_1^2$ and thus $qm \leq 2k$.

Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
\begin{align*}
\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
\end{align*}
Take $\epsilon = c_1/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c_1$, we see that with probability at least $1-\delta/2$, we get 
$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
The lower bound implies $N > qm(1-\epsilon) = qm - c_1\sqrt{qm} = k + c_1 \sqrt{2k} - c_1 \sqrt{qm} \geq k$.

For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$. The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$. Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$. Thus their sum $S$ has expectation
$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
and variance
$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + t/3)} \right) .$$
Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.

Combining with the high-probability bound on $N$ above, we get
$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
whereupon (recalling $Z = S/N$)
$$ 
\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
Since $|\eta_o| \leq 1$,
$$ |Z - \eta_o | < \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
\leq 
4 \epsilon,$$
as claimed.  
\end{proof}

\subsection{Large deviation bounds for the continuous setting}

Let $\mu$ be a distribution on the instance space $\X$.

\begin{lemma}[checked]
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Pick a set $B \subset \X$ with $\mu(B) > 0$. Suppose $m$ points $x_1, \ldots, x_m$ are drawn from $\mu|_B$, the restriction of $\mu$ to $B$, and their labels $y_1, \ldots, y_m \in \{-1,+1\}$ are generated according to $\eta$, that is, $\E[Y_i|x_i] = \eta(x_i)$.

Consider the following estimator $Z$ of $\eta(B)$:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item The labels of the selected points are averaged, yielding a value $Z$ (which defaults to 0 if no points were selected).
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta(B)| < \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-continuous}
\end{lemma}

\begin{proof}
This exactly follows the proof of Lemma~\ref{lemma:large-deviation-discrete}, with $\eta(B)$ in place of $\eta_o$.
\end{proof}

Next, we show that the $\eta_X(B)$ are, in turn, close to the corresponding $\eta(B)$.
\begin{lemma}
Fix any confidence parameter $0 < \delta < 1$. Then with probability at least $1-\delta$, the following holds for all $B \in \B$:
$$ \left| \eta(B) - \eta_X(B) \right| \leq \sqrt{\frac{2}{|X_B|} \ln \frac{2|\B|}{\delta}} .$$
\label{lemma:discrete-to-continuous}
\end{lemma}
\begin{proof}
Pick any $B \in \B$. The points in $X_B$ are random draws from $\mu|_B$. Each such point $Z$ has $\eta(Z) \in [-1,1]$ with $\E[\eta(Z)] = \eta(B)$. By a Hoeffding bound,
$$ \pr(|\eta_X(B) - \eta(B)| \geq t) \leq 2e^{-t^2|X_B|/2} $$
for any $t > 0$. The result follows by taking 
$$t = \sqrt{\frac{2}{|X_B|} \ln \frac{2|\B|}{\delta}}$$ 
and applying a union bound over all $B \in \B$.
\end{proof}

\end{document}

### 2024-07-31 COLT/colt.tex
Initial Overleaf Import

@@
-\documentclass[anon,12pt]{colt2022} % Anonymized submission
-%\documentclass[final,12pt]{colt2022} % Include author names
-
-% The following packages will be automatically loaded:
-% amsmath, amssymb, natbib, graphicx, url, algorithm2e
-
-\usepackage{amsfonts}
+\documentclass[anon,12pt]{colt2024}
+
+\usepackage{enumitem}
+\usepackage{multicol}
 
 \def\R{{\mathbb{R}}}
-\def\pr{{\textrm Pr}}
+\def\pr{{\rm Pr}}
 \def\E{{\mathbb E}}
@@
 \def\supp{{\rm supp}}
+\def\dist{{\rm dist}}
 \def\sign{{\rm sign}}
+\def\vol{{\rm vol}}
 \def\PL{{\mbox{\rm PL}}}
-\newcommand{\predict}{\mathbf{Predict}}
-\newcommand{\learn}{\mathbf{Learn}}
-\newcommand{\D}{{\cal D}}
-
-\newcommand\bigsubset[1][1.19]{%
-   \mathrel{\vcenter{\hbox{\scalebox{#1}{$\subset$}}}}}
-
-\newcommand{\comment}[3]{{\color{#1} {\bf #2 :} #3}}
-%\newcommand{\comment}[3]{}  % suppress comments
-\newcommand{\sanjoy}[1]{\comment{red}{Sanjoy}{#1}}
-\newcommand{\yoav}[1]{\comment{blue}{Yoav}{#1}}
+
+\newtheorem{thm}{Theorem}
+%\newtheorem{lemma}[thm]{Lemma}
+\newtheorem{cor}[thm]{Corollary}
+%\newtheorem{claim}[thm]{Claim}
+\newtheorem{defn}[thm]{Definition}
+\newtheorem{assump}{Assumption}
+%\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}
+
+\newcommand{\comment}[2]{ {\bf #1 :} #2}
+\newcommand{\yoav}[1]{\comment{Yoav}{\em #1}}
+\newcommand{\sanjoy}[1]{\comment{blue}{Sanjoy}{#1}}
 
@@
 
+%\setlength{\leftmargini}{0.5cm}
+%\setlength{\leftmarginii}{0.25cm}
+%\setlength{\leftmarginiii}{0.25cm}
+
 \title[Active learning using region-based sampling]{Active learning using region-based sampling}
+
 \usepackage{times}
-% Use \Name{Author Name} to specify the name.
-% If the surname contains spaces, enclose the surname
-% in braces, e.g. \Name{John {Smith Jones}} similarly
-% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
-% If the first letter in the forenames is a diacritic
-% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}
-
-% Two authors with the same address
-\coltauthor{\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
+
+\coltauthor{
+\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
 \Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
-\addr UC San Diego}
-
-% Three or more authors with the same address:
-% \coltauthor{\Name{Author Name1} \Email{an1@sample.com}\\
-%  \Name{Author Name2} \Email{an2@sample.com}\\
-%  \Name{Author Name3} \Email{an3@sample.com}\\
-%  \addr Address}
-
-% Authors with different addresses:
-%\coltauthor{%
-% \Name{Author Name1} \Email{xyz@sample.com}\\
-% \addr Address 1
-% \AND
-% \Name{Author Name2} \Email{xyz@sample.com}\\
-% \addr Address 2%
-%}
+\addr University of California, San Diego}
 
@@
 \begin{abstract}%
-Abstract goes here.
+We present a general-purpose active learning scheme that maintains a collection of neighborhoods of different sizes and uses label queries to identify those with a strong bias towards one particular label.
+When two such neighborhoods intersect and have different labels, the region of overlap is treated as a ``known unknown'' and is targeted in future active queries.
+We give label complexity bounds for this method that do not rely on assumptions about the data, and we instantiate them in several cases of interest.
 \end{abstract}
 
-\begin{keywords}%
-List of keywords
+\begin{keywords}
+Active learning, nonparametric estimation.
 \end{keywords}
@@
 
-\section{Setup}
-
-We assume that examples are drawn from a fixed but unknown
-distribution $\D$ over $\X \times \Y$, where $\X$ is a metric space and
-$\Y = \{-1,+1\}$~\footnote{In a later section we define the notion of
-  neighborhoods more generally to allow, for example, diadic
-  partitions}
-Given $x \in \X$, the label $y \in \Y$ is drawn
-according to the conditional probability function
-$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x).$$
-The \emph{Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
-
-The task we wish to solve is \emph{finite population annotation}. We
-are given a finite dataset $X \subset \X$ of $n$ points.
-We are given the ability to query the label of any point $x \in X$. The
-answer to the query: $a(x) \in \{-1,+1\}$ is a binary random variable
-drawn according to $\eta(x)$. In particular, multiple queries on the
-same point can recieve different answers.
-
-Our goal is to design an active learning algorithm that converges to the Bayes
-optimal $g^*(x)$ and does so with a much smaller number of queries
-than a passive algorithm.
-
-We want a procedure that chooses the next point to query.
-It will be applied repeatedly and then stopped at an arbitrary time, whereupon labels $y(x)$ must be provided for all $x \in X$, including those that were queried.
-We would like the inferred labels $y(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
-
-Specifically, we will be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
-$$ \sum_{x \in X} {\bf 1}(y(x) \neq g^*(x),\, |\eta(x)| > \gamma) .$$
-The procedure is allowed to fail with probability $\delta$, to account for sampling error.
-
-\section{Active learning algorithm}
-
-\subsection{Sampling regions}
-
-The sampling is organized around a collection $\B$ of measurable subsets of $\X$. These are the atomic sets on which we assess label bias and we call them ``balls''.
-
-\begin{enumerate}
-
-\item[(a)] \emph{Points within a ball.} For any $B \in \B$, let $X_B$ denote the points in $X$ that can be used to assess the bias of $B$.
-This is simply $X_B = X \cap B$ if $\B$ is a predefined set of balls.
-
-\item[(b)] \emph{Levels of sampling.} We group balls by the number of points that they contain: we put $B$ at {\it level} $\ell$ if
-$$ \frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .$$
-Here $\ell$ ranges from $0$ (consisting of balls that contain at least half the data points) to roughly $\lg n$ (containing a single point).
-
+In \emph{active learning}, the starting point is a data set whose labels are hidden but can be obtained individually for a price.
+The goal is to label the data set, or to find a good classifier, at low cost.
+
+We consider a formulation in which we have an arbitrary collection of points $X = \{x_1, \ldots, x_n\}$ from some space $\X$.
+We can request the label of any of these points $x$, in which case we get a value $y \in \{-1, +1\}$ with conditional distribution
+$ \eta(x) = \E[y | x] = \pr(y=1|x) - \pr(y=-1|x) $
+for some unknown function $\eta: \X \to [-1,1]$.
+The \emph{Bayes-optimal} label for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and $+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.
+%$$ g^*(x) = \left\{
+%\begin{array}{cl}
+%-1 & \mbox{if $\eta(x) < 0$} \\
+%+1 & \mbox{if $\eta(x) > 0$} \\
+%\mbox{either} & \mbox{if $\eta(x)= 0$}
+%\end{array} \right.
+%$$
+We wish to find the Bayes-optimal labels for the given data points $X$.
+
+More precisely, at the outset we have the data $X$ and parameters $0 < \gamma, \delta < 1$.
+We want a procedure that chooses the next point, or batch of points, to query.
+This procedure will be applied repeatedly and can be stopped at any time, whereupon labels $\yh(x)$ must be provided for all $x \in X$, including those that were queried.
+Ideally, these will be the Bayes-optimal labels $g^*(x)$; however, we will only be judged on points with non-negligible bias, $|\eta(x)| \geq \gamma$.
+The number of mistakes is:
+$$ \sum_{x \in X} {\bf 1} \big(\yh(x) \neq g^*(x),\ |\eta(x)| \geq \gamma \big) .$$
+The overall procedure is allowed to fail with probability $\delta$, to account for sampling error.
+
+\subsection{Approach and challenges}
+
+In our setting there is no underlying assumption about $g^*$, for instance that it follows a linear model. Thus we adopt a nonparametric approach.
+
+We begin with a large collection of \emph{neighborhoods}, $\B$, of varying sizes (Fig.~\ref{fig:challenges}(a)); we will refer to these as \emph{balls}.
+We use label queries to assign \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning with the largest balls.
+A bias of $+1$ means that the average $\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is significantly negative; and bias $0$ means the average $\eta$ is close to 0.
+As time goes on, we get biases for progressively smaller balls, as needed.
+The final label assigned to any point $x$ is based on the biases of balls containing $x$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{ideas.pdf}
+\end{center}
+\caption{(a) A collection of balls $\B$.
+Ball $B$ has bias $\yh(B) \in \{-1,0,+1\}$.
+(b) If one ball has bias $+1$ and the other $-1$, points in their intersection are targeted for active querying.
+(c) Biases of progressively smaller balls containing $x$ might flip sign, leading to \emph{mind changes} about the label of $x$.}
+\label{fig:challenges}
+\end{figure}
+
+%, me and to use label queries to estimate the signs associated with these regions.
+Ultimately we hope to cover $\X$ by a patchwork of neighborhoods with $\eta(x)$ values of uniform sign.
+Without strong smoothness conditions, these neighborhoods could be of very different sizes.
+The most beneficial for active learning are neighborhoods that contain a lot of $X$, but to get good coverage we might have to include smaller neighborhoods.
+To address this heterogeneity, we group the regions $\B$ by size.
+We begin by estimating the signs of the largest of them, and then move on to progressively smaller neighborhoods as the need arises.
+When the label budget runs out, the process is stopped, and provisional labels are assigned to the individual points in $X$.
+
+There are three key challenges to be addressed.
+The first is deciding \emph{where to query}.
+Initial random querying tells us the biases of the largest balls, which gives some information about the decision boundary.
+In what parts of $\X$ might it be helpful to know biases of smaller balls, warranting further label queries?
+Our approach is simple: if a ball $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the intersection $B \cap B'$ is a ``known unknown'' and is targeted for active querying  (Fig.~\ref{fig:challenges}(b)).
+
+This differs from the more conventional strategy of actively querying neighborhoods $B$ whose average $\eta$-value is close to zero.
+Such $B$ fall into two categories: (1) the $\eta$ values are close to zero throughout $B$, and (2) $B$ consists of two sub-regions, one strongly positive and the other strongly negative.
+In the first case, there is little merit in querying further.
+But in the second case, there is a lot to be gained.
+Our approach allows these two cases to be distinguished.
+
+%To distinguish these two cases, we use a collection of neighborhoods that are \emph{overlapping}.
+For instance, we might take $\B$ to be \emph{all} balls in $\X$, which is effectively a finite collection once the given data points $X$ are taken into account.
+Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a neighborhood $B$ that is strongly positive as well as being in a neighborhood $B'$ that is strongly negative.
+Such points are ``known unknowns'', and these are the targets of our active querying.
+
+A second challenge is that the sign of $\eta(\cdot)$ can change in an arbitrarily small neighborhood of any given point: the biases of balls containing $x$, say $B_1 \supset B_2 \supset \cdots$, might change sign a few times.
+At first, given only the bias of $B_1$, we might think $x$ has label $+1$.
+When we later estimate the bias of $B_2$, we could change our mind to $-1$.
+And then $+1$, and so on (Fig.~\ref{fig:challenges}(c)).
+
+What this means is that we can \emph{never} be sure of having correctly determined the label of $x$.
+Thus in addition to focused (active) querying, we also do background sampling of the entire space to pick up on possible mind changes.
+For simplicity, we make one background query per focused query.
+
+The third challenge is \emph{managing sampling} of overlapping balls.
+We are interested in detecting points, and thus neighborhoods, whose $\eta$-values are either $> \gamma$ or $<-\gamma$.
+This suggests querying $k \approx 1/\gamma^2$ points at random from each ball.
+Now, suppose we have queried this many points from ball $B$ and later want to query points from a different ball $B'$ that overlaps $B$.
+How can we reuse queries already made in $B \cap B'$?
+
+\emph{Poisson sampling} provides a clean solution.
+Rather than choosing $k$ points at random from $B$, we pick each point in $B$ with probability (roughly) $k/|B \cap X|$, independently.
+We implement this by assigning each $x \in X$ a uniform-random value $T_x \in [0,1]$, at the outset.
+When sampling $B$, we choose to query $x$ if $T_x \leq k/|B \cap X|$.
+And when it comes time to sample a different $B'$ that also contains $x$, we choose $x$ if $T_x \leq k/|B' \cap X|$; if its label has already been obtained, we are able to reuse it.
+In this way, the random querying of overlapping balls is seamlessly managed.
+
+These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$.
+By tackling all three of them, we are able to give a general-purpose active learning scheme.
+
+\subsection{Results}
+
+%Our algorithm does not require any assumption on the input distribution.
+In the worst case the number of queries is twice the number of queries of a passive algorithm.
+We show that under several natural conditions the query complexity of our algorithm is logarithmic in the error.
+
+Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function.
+We identify two \emph{critical levels} for any $x \in X$: two scales (of neighborhood sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).
+We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.
+
+Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$.
+We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.
+
+\subsection{Related work}
+
+Existing works on nonparametric active learning can mostly be grouped by overall principle: either (1) they seek to obtain the Bayes-optimal labels of a few well-positioned points, and then propagate these to the rest of the space \citep{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or negative) of entire regions at a time \citep{DH08,M12}.
+In this paper, we follow the second strategy because the only reliable and general-purpose way to assess the bias of an individual point, $\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent smoothness assumptions, the sign can change abruptly in an arbitrarily small neighborhood around $x$.
+On the other hand, the bias of a region $B$---the sign of the average $\eta$ value in $B$---is easy to determine given a reasonable number of points from that region.
+
+Early results of \cite{CN08} established upper and lower bounds on label complexity in situations where the Bayes-optimal boundary is of a simple form: a single threshold for one-dimensional data or a ``smooth boundary fragment'' in higher dimension.
+
+An algorithm for active learning based on hierarchical sampling was given by \cite{DH08} and was analyzed under smoothness conditions by \cite{KUB15}.
+The idea is to begin with a hierarchical clustering of $X$, and to then use queries to discover a pruning of this tree whose leaf-clusters are almost-pure in their labels.
+The method is not well-suited to situations with significant noise levels.
+Another approach using dyadic partitions was given by \cite{M12} and analyzed under commonly-used smoothness, margin, and density assumptions---namely, that $\eta$ is Holder-smooth, the fraction of points with $|\eta(x)| \leq t$ is some polynomial in $t$, and the marginal density is close to uniform---along with an additional smoothness requirement on $\eta$.
+
+A different strategy using nearest neighbors was explored by \cite{ASU18}.
+Their idea was to choose an appropriate scale $s$, find an $s$-cover of $X$, estimate the Bayes-optimal label for each point in this cover by querying its neighbors, and then use these cleanly-labeled points for 1-nearest neighbor classification.
+A somewhat more general approach was given by \cite{H17} and studied under the usual smoothness, margin, and density conditions, with resulting rates of convergence comparable to those found by \cite{M12}.
+
+Finally, \cite{DNZ15} suggested a graph-based method for active learning based on adaptively looking for the cut in the graph corresponding to the correct decision boundary.
+Their assumptions are based on properties of this cut and are not easily comparable with earlier work.
+
+The idea of predicting with balls of multiple sizes is a central theme of \cite{BDFM19}, which predicts at a point $x$ by choosing a ball centered at $x$ that is as small as possible while having a clear bias towards one label.
+That work does not consider active learning, which demands considerable further innovation.
+For instance, to assess the uncertainty at a point $x$, we need to look at balls containing $x$, not just those centered at $x$ (Fig.~\ref{fig:challenges}(b)).
+
+\section{The active learning algorithm}
+\label{sec:algorithm}
+
+Recall that we have a collection of points $X$ from a space $\X$.
+The label of any $x \in X$ can be requested and the value returned will be either $-1$ or $+1$, according to the conditional probability function $\eta(x) = \E[y|x]$.
+We wish to assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.
+
+\subsection{A collection of sampling regions}
+
+In our active learning algorithm, sampling is organized around a collection $\B$ of subsets of $\X$.
+These are the atomic sets on which we assess label bias and we call them \emph{balls} or \emph{neighborhoods}.
+There are no formal requirements on $\B$---for instance, the so-called balls can be of arbitrary shape---but the intention is that each $x \in X$ is contained in balls of multiple sizes, including one that is so small as to exclude the rest of $X$.
+
+For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the data points that lie in it.
+%\footnote{Since $X_B$ is used to assess the bias of $B$, we need to exclude points that cannot be considered random samples from $B$.
+For instance, consider these two alternatives: (1) $\B$ consists of a pre-defined set of balls.
+(2) $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
+In the first case, all points $X \cap B$ are random draws from $B$.
+In the second case, we need to exclude the two points $x,x'$ that define the ball.}
+We group balls into levels by the number of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
+\begin{equation}
+\frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .
+\label{eq:sampling-level}
+\end{equation}
 Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$.
-Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of very large balls and $\B_1, \B_2, \ldots$ consisting of successively smaller balls.
-
-\item[(c)] \emph{The neighborhood of a point.} For any
-  $x \in \X$, and level $\ell$.
-  $N(\ell,x)$ is the collection of balls from $\B_\ell$ that
-  contain $x$.
-\item[(d)] We say that the neighborhood $N(x,\ell)$ is {\em covered}
-  if there are $k$ query points in each $B \in N(x,\ell)$.
-
-\end{enumerate}
-
-\subsection{Bias estimates and provisional labels}
-
-The active learning algorithm uses label-queries to assess the bias of balls $B \in \B$.
-In turn, these are used to estimate the labels of individual points.
-All these estimates are organized by sampling level.
-
-\begin{enumerate}
-
-\item[(a)] We take the \emph{bias of a ball} $B \in \B$ to be
-$\eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \}$.
-At any given time, the algorithm assigns $B$ a qualitative bias estimate,
+Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of highly-populated balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls.
+We will use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater, and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.
+
+Balls in lower levels contain more points, and thus their biases (average $\eta$ values) are easier to estimate. Our sampling algorithm goes from easier to harder levels.
+
+For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that contain $x$ and can thus be used in determining $x$'s label.
+We again partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.
+%We will also refer to $\B_\ell(x)$ as the {\em level $\ell$ neighborhood of $x$}.
+
+\subsection{Estimating the biases of balls}
+
+The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
+$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
+We estimate these biases using label queries and assign each ball a \emph{qualitative} bias estimate,
 $$ \yh(B) = 
@@
 $$
-The option $\bot$ is used until sufficiently many points in $X_B$ have been labeled: the required number, $k$, is roughly $1/\gamma^2$ (recall that $\gamma$ is the smallest bias that needs to be detected).
-Once $k$ labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.
-It will satisfy a basic correctness guarantee (Definition~\ref{def:accurate-bias-estimate}) based on large deviation theory.
-
-\item[(b)] \emph{Possible labels for a point.} Pick any $x$ and any level $\ell$.
-Once bias-estimates are available for all balls $B \in \B_\ell(x)$, the possible labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,0,+1\}$, are defined as follows:
-\begin{gather}
-+1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in
-  \B_\ell(x)$ with $\yh(B) = +1$} \notag \\
-0 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = 0$} \notag \\
--1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = -1$}
-\label{eq:PL}
-\end{gather}
-The label of point $x$ at level $\ell$ is a value $\yh_\ell(x) \in \{-1,+1,0,!,\bot\}$.
-It is initially $\bot$, but once bias-estimates are available for all $B \in \B_\ell(x)$, it is set to a value in $\{-1,+1,0,!\}$ and remains fixed thereafter:
-\begin{equation}
-\yh_\ell(x) = 
-\left\{
-  \begin{array}{cl}
-+1   & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \notin PL_\ell(x)$} \\
--1   & \mbox{if $+1 \notin \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
-!    & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
-0    & \mbox{if $\PL_\ell(x)=\{0\}$} \\
-\bot & \mbox{if $\PL_\ell(x)=\emptyset$}
-\end{array}
-\right.
-\label{eq:provisional-label}
-\end{equation}
-A point $x$ for which $\yh_\ell(x)=!$ is said to have a ``conflict at
-level $\ell$''.
-
-\end{enumerate}
+
+The option $\bot$ is used until enough points in $X_B$ have been queried: the required number is $k = O(1/\gamma^2)$ since $\gamma$ is the smallest bias that needs to be detected.
+Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.
+These bias estimates will with high probability be seen to satisfy the following guarantee.
+\begin{defn}
+For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
+\begin{itemize}
+\item $\yh(B) = +1 \implies \eta_X(B) > 0$
+\item $\yh(B) = -1 \implies \eta_X(B) < 0$
+\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
+\end{itemize}
+\label{def:accurate-bias-estimate}
+\end{defn}
+
+\subsection{Assigning labels to points}
+
+The label of any point is inferred from the biases of balls containing it.
+
+Pick any $x \in X$ and any level $\ell$.
+Once qualitative bias estimates $\yh(B)$ are available for all balls at levels $\leq \ell$ that contain $x$ (that is, $\B_{\leq \ell}(x)$), we use these to infer a level-$\ell$ label for $x$.
+To begin with, the set of \emph{possible} labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$, is defined thus:
+\begin{itemize}[leftmargin=0.5cm]
+\item $\PL_\ell(x)$ includes $+1$ if there exists a \emph{minimal} ball $B \in \B_{\leq \ell}(x)$ (with no other $B' \in \B_{\leq \ell}(x)$ strictly contained in it) for which $\yh(B) = +1$.
+\item $\PL_\ell(x)$ includes $-1$ under a symmetrical condition.
+\end{itemize}
+This is made precise in Eqn~(\ref{eq:PL}) in Fig.~\ref{alg:main}.
+The label-estimate for $x$ at level $\ell$, denoted $\yh_\ell(x)$, is $+1$ if $\PL_\ell(x)=\{+1\}$, $-1$ if $\PL_\ell(x) = \{-1\}$, $0$ if $\PL_\ell(x) = \{\}$, and $!$ if $\PL_\ell(x) = \{-1,+1\}$ (Eqn~(\ref{eq:provisional-label})).
+%The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.
+
+The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\citep{R11}.
+Our active learning algorithm makes its \emph{focused queries} at level $\ell+1$ in the vicinity of such points $x$, that is, in $\B_{\ell+1}(x)$.
+
+% to be $-1$ if $\PL_\ell(x) = %\{-1\}$, or $+1$ if $\PL_\ell(x) = \{+1\}$, or $0$ if $\PL_\ell(x) = \{\}$, or $!$ %if $\PL_\ell(x) = \{-1,+1\}$, %and remains fixed thereafter.
+Points labeled $!$ are ``known unknowns'': they are %near the decision boundary and %need further investigation.
+Points labeled $0$ show no significant bias.
 
@@
 
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
-\begin{itemize}
-\item{\bf Initialize:} set $\ell=0$, cover all of the balls in $\B_0$.
-\item   $\predict(x)$
-  \begin{itemize}
-  \item Predict with $\yh_\ell(x)$
-  \end{itemize}
-  \item
-  $\learn$  Increment sampling level\\
-  Repeat while there is an $x$ that has a conflict at level $0\leq i
-  \leq \ell$ and $N(x,i+1)$ is not fully covered. 
-  \begin{itemize}
-  \item Cover $N(x,i+1)$ using $m$ queries
-  \item Query $m$ points uniformly from the domain (background
-    sampling)
-  \item Update $\yh_{i+1}(y)$ For all $y$.
-  \end{itemize}
-  Let $\ell+1 \to \ell$
-\end{itemize}
-\end{minipage}}
-\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
-\label{alg:main}
-\end{figure}
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
-\begin{itemize}
-\item Initialize uncertainty regions at all levels:
-\begin{itemize}
-\item $U_0 = X$
-\item $U_\ell = \emptyset$ for $\ell \geq 1$
-\end{itemize}
-\item Initialize all neighborhoods to ``not yet sampled'':
-\begin{itemize}
-\item For all $B \in \B$: \ $\yh(B) = \bot$
-\end{itemize}
-\item Initialize all point-labels, at all levels, to ``not yet available'':
-\begin{itemize}
-\item For all $x \in X$ and $\ell \geq 0$: \ $\yh_\ell(x) = \bot$
-\end{itemize}
-%
-%\vspace{.1in}
-\item Repeat (main loop):
-\begin{itemize}
-\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
-\begin{itemize}
-\item Let $\ell$ be the smallest such level
-\item Set query region:
-$$ S =  \bigcup_{x \in U_\ell} \bigcup_{B \in \B_\ell(x)} X_B $$
-\item {\tt Query}($S$) \ (this queries one point from $S$)
-\end{itemize}
-\item Background sampling: {\tt Query}($X$)
-\item For any neighborhood $B$ that is now adequately sampled for the first time: 
-\begin{itemize}
-\item Compute bias estimate $\yh(B) \in \{-1,0,+1\}$
-\end{itemize}
-\item For any $\ell \geq 0$ and $x \in X$ for which $\B_\ell(x)$ is available for the first time:
-\begin{itemize}
-\item Set $\PL_\ell(x)$ and $\yh_{\ell}(x)$ as per equations (\ref{eq:PL}) and (\ref{eq:provisional-label})
-\end{itemize}
-\item Update uncertainty region at all levels:
-\begin{itemize}
-\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
-\end{itemize}
-\end{itemize}
-
-\end{itemize}
-
-\end{minipage}}
-\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
-\label{alg:main}
-\end{figure}
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
-
-\vspace{.1in}
-\emph{Initialization:}
-\begin{itemize}
-\item Set $Q = \emptyset$ (points queried so far)
-\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
-\end{itemize}
-
-\vspace{.2in}
-{\bf Query}($S, \ell$):
-\begin{itemize}
-\item Query all points in $S \setminus Q$ with $T_z \leq 2^\ell k/n$, and add to $Q$
-\end{itemize}
-
-\end{minipage}}
-\caption{The sampling procedure. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried.} 
-\label{alg:sampling}
-\end{figure}
-
-The algorithm of Figure~\ref{alg:main} can be run for any amount of time. Whenever it is stopped, the {\it final label} of any $x \in X$ is
+The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, has two types of query.
+\emph{Background queries} are random draws from $X$ and correspond to passive learning.
+\emph{Focused queries} are made in the vicinity of ``uncertain'' points and correspond to active learning.
+The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$.
+Focused queries are drawn at random from balls in $\B_\ell$ containing such points.
+
+On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
+
+Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$.
+Then the next iteration begins.
+
+The querying process can be stopped at any time, whereupon labels are assigned as follows:
 \begin{equation}
@@
 \left\{
-\begin{array}{ll}
-\yh_\ell(x) & \mbox{for the largest $\ell < L[x]$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
-\mbox{as queried} & \mbox{if there is no such $\ell$ but $x \in Q$} \\
-? & \mbox{otherwise}
+\begin{array}{cl}
+\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
+0 & \mbox{``don't know'', otherwise}
 \end{array}
@@
 
-\yoav{I don't think there is anything special about a point that has
-  been queried, a single query is not enough, we need $k$ queries to
-  determine the prediction.}
-
-\section{Analysis of algorithm}
-
-\subsection{Accuracy of bias estimates}
-
-We'll start with a uniform guarantee on the bias estimates for all balls $B \in \B$.
-
-Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution $\eta(x) = \E[Y | X=x]$.
-For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value in $B$, that is,
-$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
-The specific property we need is encapsulated in the following definition.
-\begin{definition}
-For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$(b_1,b_2)$-accurate} if the following holds:
-\begin{itemize}
-\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
-\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
-\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
-\end{itemize}
-\label{def:accurate-bias-estimate}
-\end{definition}
-
-In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
-\begin{equation}
-c_1 = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
-\label{eq:c1}
-\end{equation}
-\yoav{So $\B$ is defined relative to $X$?}
-
-For any level $\ell$, define threshold $\tau_\ell$ as follows:
-\begin{equation}
-\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
-\label{eq:threshold-ell}
-\end{equation}
-For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B)
-= \{z \in X_B: T_z \leq \tau_\ell\}$. We base our bias-estimate for
-$B$ on the labels of points in $\Gamma(B)$. The following is a
-consequence of Lemma~\ref{lemma:large-deviation-bounds} in the
-appendix.
-\yoav{I guess $T_z \leq \tau_\ell$ refers to the poisson sampling
-  trick, which is not discussed. I think that initially we should
-  define an oracle that provides samples with the desired properties,
-  and later we can (in an appendix?) describe how the oracle is implemented.}
-\begin{theorem}
-Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
-For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
+\begin{figure}[t]
+\framebox{
+\begin{minipage}[t]{6in}
+\begin{center}
+\begin{minipage}[t]{6in}
+\vspace{.05in}
+
+Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
+
+\vspace{.1in}
+
+Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
+
+\vspace{.1in}
+
+Repeat:
+\begin{itemize}[leftmargin=0.5cm]
+\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
+\begin{itemize}[leftmargin=0.25cm]
+\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
+\end{itemize}
+\item {\tt Background-query} // Fig.~\ref{alg:sampling}
+\item Update labels:
+\begin{itemize}[leftmargin=0.25cm]
+\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
+\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
+\begin{align}
+\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
+             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
+\yh_\ell(x) &= 
+\left\{
+\begin{array}{cl}
++1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
+-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
+0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
+! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
+\end{array}
+\right.
+\label{eq:provisional-label}
+\end{align}
+
+\end{itemize}
+\item Update uncertainty regions:
+\begin{itemize}[leftmargin=0.25cm]
+\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
+\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
+\end{itemize}
+\end{itemize}
+
+\end{minipage}
+\end{center}
+\end{minipage}}
+\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
+\label{alg:main}
+\end{figure}
+
+\subsection{Sampling}
+
+The uncertainty region at level $\ell$ consists of the known-unknowns from level $\ell-1$, that is, points $x \in X$ with $\yh_{\ell-1}(x) = \, !$.
+This region is revealed to us piecemeal, a few points at a time.
+To get good label complexity we need to actively query the part of it we know.
+
+We manage this through \emph{Poisson sampling} (Fig.~\ref{alg:sampling}).
+To infer the bias of a ball $B$, we need labels for $k = \tilde{O}(1/\gamma^2)$ random points in it.
+Balls at level $\ell$ have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$ fraction of the ball to be queried.
+Instead, we query every point in the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, \emph{independently}.
+Specifically, we query points $x$ in the ball with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
+
+Once a point is queried, it is never queried again.
+
+\begin{figure}[t]
+\framebox{
+\begin{minipage}[t]{6in}
+
+\vspace{.05in}
+\emph{Initialization:}
+\begin{itemize}[leftmargin=0.5cm]
+\item Set $Q = \emptyset$ (points queried so far)
+\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Background-query}
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Focused-query}($\ell, U$)
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Define querying region:
+$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
+\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\end{minipage}}
+\caption{The two sampling procedures.
+Each $x \in X$ has an associated r.v.
+$T_x$ drawn uniformly from $[0,1]$.
+The smaller $T_x$, the earlier $x$ is likely to be queried.
+\emph{Background queries} are drawn at random from $X$.
+\emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.}
+\label{alg:sampling}
+\end{figure}
+
+\begin{figure}
+\framebox{
+\begin{minipage}[t]{6in}
+\vspace{.05in}
+\begin{itemize}[leftmargin=0.25cm]
+\item Initially $\yh(B) = \bot$
+\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
 $$ \yh(B)
@@
 \begin{array}{ll}
-\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
+\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
 0 & \mbox{otherwise}
@@
 $$
-With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $(b_1,b_2)$-accurate.
-\label{thm:accurate-bias-estimates}
-\end{theorem}
-In what follows, we will assume that this $(1-\delta)$-probability event holds.
-
-\subsection{Mind changes}
-
-As the level of sampling $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_\ell(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.
-
-Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ at level $\ell$.
-We distinguish two critical levels: $L_0(x)$, a level beyond which the Bayes-optimal label $s = \mbox{sign}(\eta(x))$ is \emph{always} in $\PL_\ell(x)$; and $L_1(x)$, a level beyond which the opposite label, $-s$, is \emph{never} in $\PL_\ell(x)$.
+\end{itemize}
+\end{minipage}}
+\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
+\label{alg:bias-estimate}
+\end{figure}
+
+\section{Analysis: finite population setting}
+\label{sec:discrete}
+
+We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on the manner in which $X$ is generated.
+
+\subsection{Accuracy of bias estimates}
+
+Fix the set of balls $\B$ and let $0 < \delta < 1$ be a
+predefined confidence parameter. We start with a uniform guarantee on the
+bias estimates for all balls $B \in \B$.
+
+In Fig.~\ref{alg:sampling}, we see that the points queried in a ball $B$ at level $\ell$ are $\{z \in X_B: T_z \leq \tau_\ell\}$, which from the definition of $\tau_\ell$ has size $O(k)$.
+We use these labels to estimate the bias of $B$.
+Since we need to detect biases of magnitude greater than $\gamma$, we set $k \propto 1/\gamma^2$.
+%This intuition is borne out by the following result, proved in the appendix. 
+\begin{lemma}
+Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. 
+Let $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}.
+Then with probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are $\gamma$-accurate in the sense of Definition~\ref{def:accurate-bias-estimate}.
+\label{lemma:accurate-bias-estimates}
+\end{lemma}
+Henceforth assume $\yh(B)$ is $\gamma$-accurate for all  $B \in \B$.
+
+\subsection{Critical levels}
+
+Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes the possible labels for $x$ given information from balls at levels $\leq \ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$.
+Despite randomness in querying, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ reliably contains the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ reliably omits the wrong label.
+
+\begin{defn}[Critical levels $L_1,L_2$]
+Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$. Define $L_1(x)$ to be the smallest level $\ell$ such that there exists $B_o \in \B_\ell(x)$ for which
+\begin{itemize}[leftmargin=0.5cm]
+\item $s(x) \cdot \eta_X(B_o) \geq \gamma$, and
+\item $s(x) \cdot \eta_X(B) \geq \gamma$ for any $B \in \B(x)$ with $X_{B} \subset X_{B_o}$.
+\end{itemize}
+Define $L_2(x)$ to be the smallest level $\ell$ such that:
+\begin{itemize}[leftmargin=0.5cm]
+\item For all $B \in \B_{\geq \ell}(x)$, we have $s(x) \cdot \eta_X(B) \geq 0$.
+\item For any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$, there exists $B' \in \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
+\end{itemize}
+Take $L_1(x)$ or $L_2(x)$ to be $\infty$ if no level meets the requirements, or if $\eta(x) = 0$. 
+\label{defn:L12}
+\end{defn}
+
+Once $\ell$ reaches a critical level, $\PL_\ell(x)$ and thus $\yh_\ell(x)$ behave more predictably, as follows.
+\begin{lemma}
+Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) \in \{+1,-1\}$ denote its Bayes-optimal label. Then for any level $\ell$ and any time at which $\yh_\ell(x) \neq \bot$:
+\begin{enumerate}
+\item[(a)] If $\ell \geq L_1(x)$, then $s(x) \in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), !\}$. 
+\item[(b)] If $\ell \geq L_2(x)$, then $-s(x) \not\in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), 0\}$.
+\end{enumerate}
+\label{lemma:boundary}
+\end{lemma}
+
+\subsection{Boundary sets and label complexity}
+
+A common intuition about active learning is that successive queries gradually constrain the possible locations of the decision boundary.
+Let's consider the state of affairs when all balls at level $\leq \ell - 1$ have been sampled.
+The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points must have $L_2(x) \geq \ell$.
+Focused sampling at level $\ell$ is restricted to balls that contain these points.
+We will think of this region as the \emph{boundary set} at level $\ell$.
+\begin{defn}[Boundary set $\bf \Delta_\ell$]
+For any level $\ell$, define the \emph{boundary set} at level $\ell$ to be
+\begin{equation}
+\Delta_\ell 
+= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B 
+.
+\label{eq:sampling-region}
+\end{equation}
+\end{defn}
+%It can be shown that all focused samples at level $\ell$ must lie in this set.
+\begin{lemma}
+All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
+\label{lemma:focused}
+\end{lemma}
+
+We can now give generic label complexity bounds in terms of $L_1$, $L_2$, and $\Delta_\ell$.
+These take two equivalent forms: a \emph{global} version that specifies what parts of $X$ are correctly labeled after $m$ queries and a \emph{local} version that specifies the number of queries after which a particular $x$ is correctly labeled.
+The global version (Theorem~\ref{thm:label-complexity-0}) is in the Appendix; here is the local version.
+
+\begin{thm}[Local label complexity]
+Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$.
+With probability at least $1-2\delta$, the following holds simultaneously for all $x \in X$.
+Let $L_1(x)$ and $L_2(x)$ be the critical levels for $x$, as in Definition~\ref{defn:L12}.
+If $L_2(x) \leq \lg (n/2k)$, let
+\[
+m_o(x) = 32k \cdot \max\bigg( 2^{L_1(x)}, \ \frac{1}{n} \sum_{\ell=L_1(x)+1}^{L_2(x)} |\Delta_\ell| \, 2^\ell\bigg) .
+\]
+Once the active learner has made $m_o(x)$ queries, $\yh(x)$ is fixed at the Bayes-optimal label $g^*(x)$.
+\label{thm:label-complexity}
+\end{thm}
+Let's interpret this bound.
+For a specific $x \in X$, there is no telling how $\yh_\ell(x)$ will behave at levels $\ell < L_1(x)$.
+But for $\ell \geq L_1(x)$, either $\yh_\ell(x) = \, !$, in which case $x$ is in the uncertainty region, or the label of $x$ is correctly set once and for all, which occurs by level $L_2(x)$.
+
+We will say that $x$ \emph{reaches level $\ell$} when bias estimates have been obtained for all balls in $\B_{\leq \ell}(x)$ and thus $\yh_\ell(x)$ is set.
+Recalling that the algorithm alternates between background and focused queries, we show that $m_o(x)/4$ \emph{background} queries suffice for $x$ to reach level $L_1(x)$ and an additional $m_o(x)/4$ \emph{focused} queries take it to level $L_2(x)$, by which time its label is correct.
+
+In particular, $O(k \cdot 2^\ell)$ background queries give us the biases of all balls in $\B_{\leq \ell}$; this is the first term in $m_o(x)$.
+And there are $O(k \cdot 2^\ell \cdot |\Delta_\ell|/n)$ focused queries at any level $\ell$: the second term.
+
+A footnote: Our algorithm queries points at most once.
+However, in some applications, a point $x$ can be queried repeatedly, each time producing an independent draw from $\eta(x)$.
+If \emph{repeat queries} are permissible, $O(1/\gamma^2)$ copies should be made of each point in $X$ before the algorithm is applied; and in this case, points with $|\eta(x)| \geq \gamma$ will all be correctly labeled, eventually.
+
+\section{One-dimensional examples}
+
+To apply Theorem~\ref{thm:label-complexity}, we need bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each $x \in X$, and on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
+We now derive these in some canonical settings.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=2.5in]{oned-monotonic.pdf}
+\hspace{.5in}
+\includegraphics[width=2.5in]{oned-massart.pdf}
+\end{center}
+\caption{(a) The conditional probability function $\eta$ is monotonically increasing; $\eta(x) = -\gamma, 0, \gamma$ at $x = \lambda_L, \lambda, \lambda_R$, respectively.
+(b) On each subinterval, $\eta$ is either $> \gamma$ or $<-\gamma$.}
+\label{fig:oned}
+\end{figure}
+
+\subsection{Example: monotonic $\eta$}
+
+Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing (Fig.~\ref{fig:oned}(a)).
+Let $\lambda \in (0,1)$ be the point for which $\eta(\lambda) = 0$ and let $\lambda_L, \lambda_R$ be the points for which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$.
+Thus we are not required to label points in the interval $(\lambda_L, \lambda_R)$.
 
@@
 %\begin{center}
-%\includegraphics[width=3.5in]{home-stretch.pdf}
+%\includegraphics[width=3in]{oned-monotonic.pdf}
 %\end{center}
-%\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively finer levels.
-Beyond level $L_s(x)$, this $\yh_\ell(x)$ will never be the wrong label, but might be $0$ or $!$.
-Beyond level $L_f(x)$, the correct label will be assigned.}
-%\label{fig:two-levels}
+%\caption{The conditional probability function $\eta$ is monotonically increasing. The values $x = %\lambda_L, \lambda, \lambda_R$ have $\eta(x) = -\gamma, 0, \gamma$ respectively.}
+%\label{fig:oned-monotonic}
 %\end{figure}
 
-\begin{definition}
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label.
-We define $L_0(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there exists $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$.
-We define $L_1(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there is no $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
-\label{defn:L01}
-\end{definition}
-
-The key properties of these critical levels are spelled out in the following lemma.
-\begin{lemma}
-Pick $x \in X$ with $\eta(x) \neq 0$. Let $g^*(x) \in \{+1,-1\}$ be its Bayes-optimal label. Then:
-\begin{enumerate}
-\item[(a)] For all $\ell \geq L_0(x)$, we must have $g^*(x) \in \PL_\ell(x)$.
-\item[(b)] For all $\ell \geq L_1(x)$, we must have $-g^*(x) \not\in \PL_\ell(x)$.
-\item[(c)] For all $\ell \geq \max(L_0(x), L_1(x))$, we must have $\yh_\ell(x) = g^*(x)$.
-\end{enumerate}
-\label{lemma:boundary}
-\end{lemma}
-\begin{proof}
-This lemma follows from the $(b_1, b_2)$-accuracy of the bias estimates for balls.
-Pick $x \in X$ with $\eta(x) \neq 0$.
-If $\ell \geq L_0(x)$, there is necessarily a ball $B \in \B_\ell(x)$ with $g^*(x) \eta_X(B) > b_2$ and thus $\yh(B) = g^*(x)$; therefore $g^*(x) \in \PL_\ell(x)$.
-If $\ell \geq L_1(x)$, every $B \in B_\ell(x)$ has $g^*(x) \eta_X(B) \geq -b_1$ and thus $\yh(B) \in \{0,g^*(x)\}$; therefore $-g^*(x) \not\in \PL_\ell(x)$.
-Finally, when $\ell \geq \max(L_0(x), L_1(x))$, it follows from (a) and (b) that $\PL_\ell(x) = \{g^*(x)\}$ and thus $\yh_\ell(x) = g^*(x)$.
-\end{proof}
-
-\subsection{Bounding the amount of focused sampling}
-
-We can now characterize the points that are queried as part of focused sampling.
-\begin{lemma}
-For any $\ell \geq 0$, define
-\begin{equation}
-\Delta_{\ell} \ \ \ = \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .
-\label{eq:sampling-region}
-\end{equation}
-Then, all focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
-\label{lemma:focused}
-\end{lemma}
-\begin{proof}
-Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$.
-For any point $x$ with $L_1(x) < \ell$, we have from Lemma~\ref{lemma:boundary}(b) that $\PL_{\ell-1}(x)$ is either $\{\}$ or $\{g^*(x)\}$; either way, $x$ is not in the uncertainty set at level $\ell$.
-Thus $\overline{U}_\ell \subset \{x \in X: L_1(x) \geq \ell\}$.
-
-The focused samples at level $\ell$ lie in
-$$
-\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau(B) \}
-\ \ \ 
-\bigsubset[1.44]
-\ 
-\bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \} .
-$$
-This is precisely $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
-\end{proof}
-
-A final property we will need for label complexity bounds is that various subsets of interest contain roughly the expected number of points at each level.
-\begin{lemma}
-Pick any positive integer $\ell_f$. With probability at least $1-2(\ell_f+1)e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \ell_f$.
-\begin{enumerate}
-\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
-\item[(b)] If $|\Delta_\ell| \geq k$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 |\Delta_\ell| \tau_\ell$.
-\end{enumerate}
-\label{lemma:level-distribution}
-\end{lemma}
-\begin{proof} 
-Pick any subset $S \subset X$ and let $m = |S|$.
-Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$.
-The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m/3}$.
-
-Both parts follow from this principle; and we take a union bound over all $\ell_o + 1$ levels.
-\end{proof}
-
-\subsection{Label complexity bounds}
-
-\yoav{I think this is the main theorem, however, I find it very hard
-  to parse. In particular, I don't see any reference to the fraction
-  of examples that are in the focused group at each level. These
-  fractions have to be small in order for our sampling to be effective.}
-
-\begin{theorem}
-Pick any $0 < \delta < 1$. Suppose that $k \geq (8c_1/(b_2-b_1))^2$, where $c_1$ is defined in (\ref{eq:c1}), and that the active learning algorithm is allowed to make $m$ queries. Let
-$$ \ell_0 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
-and let $\ell_1$ be the largest integer such that
-$$ \sum_{\ell = \ell_0 + 1}^{\ell_1} |\Delta_{\ell}| \, \tau_\ell \ \leq \ \frac{m}{8} .$$
-Then with probability at least $1-2\delta$, any point $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$ will have final label $\yh(x) = g^*(x)$.
-\label{thm:label-complexity}
-\end{theorem}
-\begin{proof}
-Under the setting of $k$, the high-probability events of Theorem~\ref{thm:accurate-bias-estimates} and Lemma~\ref{lemma:level-distribution} hold with probability at least $1-2\delta$.
-We will henceforth take these for granted.
-
-Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}.
-We will analyze the effect of background sampling in phase one and focused sampling in phase two.
-We start with the former.
-
-Of the $m/2$ queries in phase one, at least $m/4$ will be background samples.
-Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.
-Now, for $\ell_0$ as defined, we have from (\ref{eq:threshold-ell}) that $\tau_{\ell_0} \leq m/8n$ and from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy
-$$ T_x \leq \tau_{\ell_0} \leq \frac{m}{8n}.$$
-Therefore all such points are queried in phase one, and thus all balls $B$ at level $\leq \ell_0$ are processed and get bias estimates $\yh(B)$.
-
-It then follows from Lemma~\ref{lemma:boundary} that the following holds for any $x \in X$ with $L_0(x) \leq \ell_0$:
-\begin{enumerate}
-\item[(a)] For any $\ell \geq \ell_0$, if $\yh_\ell(x)$ is defined then it is either $g^*(x)$ or $!$.
-\item[(b)] If $x$ ever leaves $U$ during phase two, then its final label as defined in (\ref{eq:final-label}) is $\yh(x) = g^*(x)$.
-\item[(c)] If $x$ is processed at level $L_1(x)$, then it will leave $U$ for good.
-\end{enumerate}
-
-Now let's move on to phase two.
-Let $A = \{x \in X: L_0(x) \leq \ell_0, L_1(x) \leq \ell_1\}$.
-We will show that every point in $A$ must leave the uncertainty set at some time during phase two.
-From (b), we can conclude that all these points have their final labels set correctly, once and for all.
-
-We will break the argument into three cases.
-
-{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.
-
-{\it Case 2:} A full $m/4$ focused queries are made in phase two, but at least one of them is at level $\ell_1 + 1$.
-In that round, $U$ cannot contain any $x \in A$, since by (c) all such points will leave the uncertainty set by level $\ell_1$.
-
-{\it Case 3:} A full $m/4$ focused queries are made in phase two, all at levels $\leq \ell_1$.
-By the analysis of phase one, none of these queries can be at level $\leq \ell_0$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_0+1$ through $\ell_1$ inclusive is at most
-$$\sum_{\ell=\ell_0+1}^{\ell_1} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
-\ \leq \ 
-\sum_{\ell=\ell_0+1}^{\ell_1} 2 |\Delta_{\ell}| \tau_\ell
-\ \leq \ 
-\frac{m}{4} ,
-$$
-where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus by the end of the phase two, every $x \in A$ will have been processed until it leaves $U$.
-\end{proof}
-
-\section{Label complexity bounds in various settings}
-
-In order to apply Theorem~\ref{thm:label-complexity}, we need two things:
-\begin{itemize}
-\item An upper bound on the critical level $L_1(x)$ for each $x \in X$, which in turn (via Lemma~\ref{lemma:focused}) will yield an upper bound on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
-\item An upper bound on $L_0(x)$ for points $x$ that we expect to be correctly labeled by the end.
-\end{itemize}
-We now illustrate how this works out in several canonical settings.
-
-\subsection{One-dimensional data with Massart noise}
-
-Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
-\begin{itemize}
-\item The Bayes-optimal labeling is determined by $p$ disjoint open intervals $I_1 = (\lambda_0, \lambda_1), I_2 = (\lambda_1, \lambda_2), \ldots, I_p = (\lambda_{p-1}, \lambda_p)$, where
-$$ 0 = \lambda_0 < \lambda_1 < \lambda_2 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$$
-Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
-\item For each $j$, either $\eta(x) > c$ for all $x \in I_j$ or $\eta(x) < -c$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
-\end{itemize}
-Here $c > 0$ is some constant.
-
-We will take $\B$ to be all open intervals of $[0,1]$, and $\B(x)$ to be intervals containing the point $x$. The accuracy parameters $b_1, b_2$ will be set so that $0 \leq b_1 < b_2 \leq c$.
-\begin{lemma}
-Pick any $x \in I_j$, for $1 \leq j \leq p$.
-\begin{enumerate}
-\item[(a)] $L_0(x) \leq \left\lceil \lg (n/n_j) - 1 \right\rceil$, where $n_j$ is the number of data points in $I_j$.
-\item[(b)] $L_1(x) \leq \left\lceil \lg (n/r(x)) - 1 \right\rceil$, where $r(x)$ is the minimum number of data points between $x$ and a boundary point. 
-\end{enumerate}
-\label{lemma:critical-one-d-massart}
-\end{lemma}
-\begin{proof}
-For (a), define $n_j = |X \cap I_j|$ and notice that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$.
-Moreover, $s(I_j) \cdot \eta_X(I_j) > c \geq b_2$.
-For levels $\ell' > \ell$, we can make a similar argument using a sub-interval $I \subset I_j$ that lies in $\B_{\ell'}(x)$.
-
-For (b), we define $r(x)$ to be either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller.
-For $\ell \geq \lceil (\lg n/r(x)) - 1 \rceil$, any $B \in \B(x)$ will lie entirely within $I_j$ and must therefore have $s(I_j) \cdot \eta_X(B) > 0$.
-\end{proof}
-
-Using these bounds on the critical levels, we find that all but an $\epsilon$ fraction of $X$ will be correctly classified after making $m = O(\frac{(p-1)\log n}{c^2} \log \frac{p-1}{\epsilon})$ queries.
-\begin{theorem}
-In the one-dimensional scenario above, there are absolute constants $C_1, C_2, C_3$ for which the following holds.
-Pick any $0 < \delta < 1$ and take $k \geq (C_1/c^2) \ln (n/\delta)$.
-Then with probability at least $1-\delta$, after making $m$ queries, the active learning algorithm will correctly classify all but $n \cdot 2^{-C_2 m/(k(p-1))}$ points in any target interval $I_j$ with $n_j/n \geq C_3 k/m$.
-\label{thm:one-d-massart}
-\end{theorem}
-
-\begin{proof}
-Using Lemma~\ref{lemma:critical-one-d-massart}(b), we see that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
-\begin{align*}
-\Delta_\ell \ \ \ 
-&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $2n/2^{\ell}$ points on either side of each boundary point, so $|\Delta_\ell| \leq 4(p-1)n/2^{\ell}$.
-
-We then set $b_1 = 0$, $b_2 = c$ and apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
-\begin{align*}
-\ell_0
-&= 
-\lg \frac{m}{32k} \\
-\ell_1
-&=
-\ell_0 + \frac{m}{128 k(p-1)},
-\end{align*}
-and $k \geq (192/c^2)(2 \ln n + \ln (2/\delta))$; in setting $k$, we have used $|\B| = {n \choose 2}$. The theorem statement follows by invoking Lemma~\ref{lemma:critical-one-d-massart}.
-\end{proof}
-
-\subsection{One-dimensional data with monotonic $\eta$}
-
-For our next example, we once again take $X$ to consist of $n$ arbitrarily-placed points in $\X = [0,1]$.
-This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing.
-Let $\lambda \in (0,1)$ denote the point at which $\eta(\lambda) = 0$.
-
-As before, $\B$ will consist of all open intervals, and $\B(x)$ be will all intervals containing $x$.
-Given a threshold $\gamma > 0$ such that we do not need to label points with $|\eta(x)| < \gamma$, we can set accuracy parameters $b_2 = \gamma, b_1 = 0$.
-
-\begin{lemma}
-Let $\lambda_L, \lambda_R$ be the points at which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$.
-\begin{enumerate}
-\item[(a)] Define $n^- = |[0,\lambda_L) \cap X|$ and $n^+ = |(\lambda_R,1] \cap X|$. Then
-$$
-L_0(x)
+Let $\B$ consist of all closed intervals. Although there are infinitely many of these, we need only consider those whose endpoints lie in $X$, so effectively $|\B| = O(n^2)$.
+
+The critical levels are easy to bound in this setting.
+Let $n^- = |[0,\lambda_L] \cap X|$  be the number of points to the left of $\lambda_L$ and $n^+ = |[\lambda_R,1] \cap X|$ the number of points to the right of $\lambda_R$.
+Then
+%\begin{align*}
+%L_1(x)
+%&\leq
+%\left\{
+%\begin{array}{ll}
+%\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
+%\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
+%\end{array}
+%\right.
+%\\
+%L_2(x)
+%&\leq
+%\lg (n/r(x))
+%\end{align*}
+$$
+L_1(x)
 \leq
@@
 \begin{array}{ll}
-\lg \lceil (n/n^+) - 1 \rceil & \mbox{if $x > \lambda_R$} \\
-\lg \lceil (n/n^-) -1 \rceil & \mbox{if $x < \lambda_L$}
+\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
+\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
 \end{array}
 \right.
-$$
-\item[(b)] For any $x \in X$, let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$. Then
-$L_1(x) \leq \left\lceil \lg (n/r(x)) -1 \right\rceil$.
-\end{enumerate}
-\label{lemma:critical-one-d-monotonic}
-\end{lemma}
-
-\begin{proof}
-For (a), take $x > \lambda_R$, for instance.
-The interval $B = (\lambda,1)$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta(B) > b_2$.
-For higher levels $\ell$, we can take a sub-interval of $(\lambda, 1)$ containing $x$.
-
-For (b), we have $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$.
-For any $\ell \geq \lceil \lg (n/r(x)) - 1 \rceil$, no $B \in \B_\ell(x)$ extends to the other side of the boundary; thus $\eta_X(B) \geq 0$.
-\end{proof}
-
-In this setting, we are not required to label points in the interval $[\lambda_L, \lambda_R]$, and thus we don't need $L_0(x)$ values for them.
-However, we do need to bound $L_1(x)$ for \emph{all} points in order to control the size of the sampling region.
-
-With Lemma~\ref{lemma:critical-one-d-monotonic}, the rest of the argument follows the pattern of the previous one-dimensional case.
-
-\begin{theorem}
-There are absolute constants $C_1, C_2, C_3$ for which the following holds.
-Pick any $0 < \delta < 1$ and take $k \geq (C_1/\lambda^2) \ln (n/\delta)$.
-Then with probability at least $1-\delta$, after making $m \geq C_2 k \max(n/n^+, n/n^-)$ queries:
-\begin{enumerate}
-\item[(a)] The active learning algorithm will correctly label all but $n \cdot 2^{-C_3 m/k}$ points in $[0,\lambda_L) \cup (\lambda_R,1]$. 
-\item[(b)] If $n_o = \min(|X \cap [\lambda_L,\lambda)|, |X \cap (\lambda,\lambda_R]|) > 0$, then all points in $[0,\lambda_L) \cup (\lambda_R,1]$ will be correctly labeled after $C_3 k \lg (n/n_o)$ queries.
-\end{enumerate}
-\label{thm:one-d-monotonic}
-\end{theorem}
-
-\begin{proof}
-From Lemma~\ref{lemma:critical-one-d-monotonic}(b), we have that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
-\begin{align*}
-\Delta_\ell \ \ \ 
-&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $2n/2^{\ell}$ points on either side of the boundary, whereupon $|\Delta_\ell| \leq 4n/2^{\ell}$.
-
-By Theorem~\ref{thm:label-complexity}, making $m$ queries will correctly classify all points with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
-\begin{align*}
-\ell_0
-&= 
-\lg \frac{m}{32k} \\
-\ell_1
-&=
-\ell_0 + \frac{m}{128 k},
-\end{align*}
-and $k \geq (192/\lambda^2)(2 \ln n + \ln (2/\delta))$. If $m \geq 32k \cdot \max(n/n^+, n/n^-)$, this includes all points in $[0,\lambda_L) \cup (\lambda_R,1]$ with $r(x) \geq n \cdot 2^{-m/128k}$. 
-\end{proof}
-
-\subsection{Two-dimensional data with linear boundary}
-
-\subsection{Higher dimension, with dyadic partitioning}
-
-\subsection{Higher dimension, with smoothness conditions}
-
-% Acknowledgments---Will not appear in anonymized version
-\acks{We thank a bunch of people and funding agency.}
-
-\bibliography{yourbibfile}
-
-\vfill
-
-\pagebreak
-
+\mbox{\ \ \ \ and \ \ \ }
+L_2(x)
+\leq
+\lg (n/r(x))
+$$
+where $r(x)$ is the number of data points between $x$ and $\lambda$, counting $x$ as well.
+%\begin{itemize}
+%\item $L_1(x) \leq \lg (n/n^+)$ if $x \geq \lambda_R$ and $L_1(x) \leq \lg (n/n^-)$ if %$x \leq \lambda_L$, and
+%\item $L_2(x) \leq \lg (n/r(x))$, 
+%\end{itemize}
+
+A counting argument then shows that the boundary region shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
+\begin{thm}
+Define $r^+ = \min \{r(x): x \in X \cap [\lambda_R,1]\}$ and $r^- = \min \{r(x): x \in X \cap [0,\lambda_L]\}$.
+Pick any $0 < \delta < 1$.
+Suppose we run the algorithm of Fig.~\ref{alg:main} with $k = O((\log (n/\delta))/\gamma^2)$ and that $\min(r^+, r^-) \geq 2k$.
+Take any
+$$ m \geq 64k \cdot \max \left( \frac{n}{\min(n^+,n^-)}, \ 2 \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \right) .$$
+With probability $\geq 1-\delta$, after $m$ queries the algorithm correctly labels all $x \in X$ with $|\eta(x)| \geq \gamma$.
+\label{thm:oned-monotonic}
+\end{thm}
+If $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$, then the label complexity is $O(\log^2 n)$.
+If in addition $r^+, r^- = \Theta(n)$, that is, a constant fraction of the points lie between $\lambda_L, \lambda_R$ and the decision boundary, the label complexity is $O(\log n)$.
+
+\subsection{Example: Massart noise}
+
+In Section~\ref{sec:oned-massart}, we develop another one-dimensional example, in which $\X$ consists of several pieces, such that on each piece $\eta$ is either entirely above $\gamma$ or entirely below $-\gamma$ (Fig.~\ref{fig:oned}(b)).
+Once again, a logarithmic label complexity is obtained.
+
+\section{Analysis: distributional setting}
+\label{sec:continuous}
+
+We now turn to a setting where the points $X$ are drawn from some distribution $\mu$ on a topological space $\X$.
+For the same algorithm, we bound label complexity in terms of properties of $\mu$ and $\eta$.
+
+\subsection{A sufficiently-rich collection of neighborhoods}
+
+Let $\B$ be a finite collection of subsets of $\X$; we spell out a basic property we need it to possess.
+
+\begin{defn}
+We say $\B$ is \emph{$k$-conforming} with respect to $X$ if 
+\begin{itemize}
+\item ({\bf Richness}) For any $x \in X$, there exists $B \in \B_0(x)$.
+Moreover, for any level $0 \leq \ell \leq \lg (n/k)$ and any $B \in \B_\ell(x)$, there exists $B' \in \B_{\ell+1}(x)$ such that $X_{B'} \subset X_B$.
+\item ({\bf Fidelity}) $\mu(B)/2 \leq |X_B|/n \leq 2\mu(B)$ for all $B \in \B$ with $|X_B| \geq k$.
+\end{itemize}
+\label{defn:conformity}
+\end{defn}
+When $\B$ consists of Euclidean balls in $\R^d$, made finite by restricting to $X$, the first property is immediate and the second holds with high probability (Lemma~\ref{lemma:ball-size-bounds}).
+We also envisage scenarios where $\B$ is constructed using $X$, for instance from the cells of multiple spatial partition trees.
+
+Henceforth assume $\B$ satisfies this property.  We ignore $B$ with $|X_B| < k$, thereby restricting attention to levels $0 \leq \ell < \lg (n/k)$.
+
+\subsection{A probabilistic notion of distance}
+
+A crucial step in the analysis is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
+\begin{defn}
+For any $x \in \X$ and $S \subset \X$, define
+$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
+\label{defn:prob-dist}
+\end{defn}
+We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
+\begin{lemma}
+For any $x \in \X$, let $s(x) = \sign(\eta(x))$. If $\dist(x, \X^{-s(x)}) \geq 2k/n$, then
+$L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1$.
+\label{lemma:L2-bound}
+\end{lemma}
+We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
+\begin{defn}
+For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
+\label{defn:boundary}
+\end{defn}
+This includes all $x$ with $\eta(x) = 0$. Next, the $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
+\begin{defn}
+For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
+\label{defn:boundary2}
+\end{defn}
+
+We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
+\begin{lemma}
+For any $0 \leq \ell \leq \lg (2n/k)$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
+\label{lemma:delta-continuous}
+\end{lemma}
+Theorem~\ref{thm:label-complexity} thus still holds with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.
+
+Using these notions, we are able to specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under suitable margin conditions; this is given in Theorem~\ref{thm:label-complexity-specific} in the appendix.
+Here we just give a concrete scenario handled by the latter result; another such scenario, with smoothness and Tsybakov noise conditions, appears in the appendix.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=6in]{COLT/curvature.pdf}
+\end{center}
+\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
+The orange and light blue indicate the boundary sets for negative and positive respectively.
+{\em Right:} examples of the $L_1,L_2$ balls for different points.
+The green $L_1$ circles are the largest monochrome circle that contains the point.
+The purple $L_2$ circles are the union of all circles of half the radius or smaller.
+All of these circles are monochrome.}
+\label{fig:curvature}
+\end{figure}
+
+\subsection{Example: Curvature and Massart noise}
+\label{sec:Massart}
+
+By way of example, consider a case where $\X \subset \R^d$ and $\mu$ admits a density.
+We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.
+By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\citep{BC03}.
+
+We make two key assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
+\begin{itemize}[leftmargin=0.5cm]
+\item (Strong density condition)
+There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
+we have $c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) $, where $\vol(\cdot)$ is $d$-dimensional volume.
+\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
+\end{itemize}
+The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$.
+It is a common notion of curvature~\citep{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
+A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
+\begin{thm}
+Under the two conditions above, there are constants $c_2, c_3$ for which the following holds.
+Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
+If the algorithm of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$ queries, it assigns Bayes-optimal labels to all but $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
+\label{thm:massart-dist}
+\end{thm}
+Here the constant $c_2$ includes a term $(1/r_o)^d$.
+The $\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over the usual $1/m^{1/d}$ rate for random querying and was also observed earlier, in a more restricted setting, using specialized algorithms~\citep{CN08}.
+
+\newpage
+
+%\input{active-knn.bbl}
+\bibliography{refs}
+%\bibliographystyle{plain}
+
+\newpage
 \appendix
 
-\section{Technicalities}
+\section{Technicalities: discrete setting}
 
@@
 \item[(a)] $N \geq k$, and 
-\item[(b)] $| Z - \eta_o| \leq \sqrt{(48/k) \ln (4/\delta)}$.
+\item[(b)] $| Z - \eta_o| < \sqrt{(48/k) \ln (4/\delta)}$.
 \end{enumerate}
@@
 Let's start with (a).
-Define $c_1 = \sqrt{3 \ln (4/\delta)}$.
-We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c_1 \sqrt{2k}$ since this is the worst case.
-By assumption, $k \geq 2c_1^2$ and thus $qm \leq 2k$.
+Define $c = \sqrt{3 \ln (4/\delta)}$.
+We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c \sqrt{2k}$ since this is the worst case.
+By assumption, $k \geq 2c^2$ and thus $qm \leq 2k$.
 
@@
 \end{align*}
-Take $\epsilon = c_1/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c_1$, we see that with probability at least $1-\delta/2$, we get 
+Take $\epsilon = c/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c$, we see that with probability at least $1-\delta/2$, we get 
 $$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
-The lower bound implies $N > qm(1-\epsilon) = qm - c_1\sqrt{qm} = k + c_1 \sqrt{2k} - c_1 \sqrt{qm} \geq k$.
+The lower bound implies $N > qm(1-\epsilon) = qm - c\sqrt{qm} = k + c \sqrt{2k} - c \sqrt{qm} \geq k$.
 
@@
 $$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
-We can bound the concentration of $S$ around its expected value using Bernstein's inequality,
-$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + 2t/3)} \right) .$$
+We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
+$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + t/3)} \right) .$$
 Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.
@@
 Since $|\eta_o| \leq 1$,
-$$ |Z - \eta_o | \leq \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
+$$ |Z - \eta_o | < \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
 \leq 
@@
 
-An immediate consequence is the following large deviation bound.
-Recall that for any ball $B \in \B_\ell$, its query-set $\Gamma(B)$ is $\{z \in X_B: T_z \leq \tau_\ell\}$, where $\tau_\ell$ is defined in (\ref{eq:threshold-ell}).
-\begin{lemma}
-Suppose that the constant $c_1$ is set as in (\ref{eq:c1}) and that $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
+\subsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}
+
+We will use Lemma~\ref{lemma:large-deviation-discrete} to obtain a uniform guarantee on the bias estimates for all balls $B \in \B$.
+
+Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
+$$ \eta(x) = \E[y | x].$$ 
+For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value of the points in $B$, that is,
+$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
+
+For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$, where the sampling threshold $\tau_\ell$ for level $\ell$ is defined as:
+\begin{equation}
+\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
+\end{equation}
+for some $k$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.
+
+For what follows, define
+\begin{equation*}
+c = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
+\end{equation*}
+
+\begin{lemma}
+Suppose $k \geq 2c^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
 \begin{enumerate}
@@
 \item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
-$\left| \widehat{\eta}(B) - \eta_X(B) \right| \leq 4c_1/\sqrt{k}$.
+$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c}{\sqrt{k}}.$$
 \end{enumerate}
@@
 \begin{proof}
-These follow by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
-\end{proof}
-
+Pick $B \in \B$. There are two cases to consider.
+
+Case 1: $|X_B| \geq 2k$.
+The choice of $\tau_\ell$ then ensures $|X_B| \tau_\ell \geq 2k$, so that $|\widehat{\eta}(B) - \eta_X(B)|$ can be bounded by applying Lemma~\ref{lemma:large-deviation-discrete} to the points $X_B$ with sampling probability $q = \tau_\ell$.
+
+Case 2: $k \leq |X_B| < 2k$.
+In this case, $B$ lies at a level $\ell$ for which $\tau_\ell = 1$.
+Lemma~\ref{lemma:large-deviation-discrete} does not apply directly, but its conclusion still holds.
+In particular, the query-set $\Gamma(B)$ is all of $X_B$, and the same Bernstein bound from the proof of Lemma~\ref{lemma:large-deviation-discrete} can be again be applied.
+
+To complete the proof, we take a union bound over all $B \in \B$.
+\end{proof}
+
+The following corollary of Lemma~\ref{lemma:large-deviation-bounds} is immediate.
+\begin{cor}
+Suppose that $k \geq (8c/\gamma)^2$. For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
+$$ \yh(B)
+= 
+\left\{
+\begin{array}{ll}
+\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+0 & \mbox{otherwise}
+\end{array}
+\right.
+$$
+With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\gamma$-accurate.
+\label{cor:accurate-bias-estimates}
+\end{cor}
+\begin{proof}
+By the choice of $k$, we have from Lemma~\ref{lemma:large-deviation-bounds} that 
+$| \widehat{\eta}(B) - \eta_X(B) | < \gamma/2$
+for all $B \in \B$.
+\end{proof}
+
+\subsection{Critical levels: Proof of Lemma~\ref{lemma:boundary}}
+
+For part (a), note that some $B_o \in \B_{L_1(x)}(x)$ has significant bias (that is, bias $\geq \gamma$) towards the correct label $s(x)$, as does any ball $B \in \B(x)$ contained within it.
+For $\ell \geq L_1(x)$, the set $\{B \in \B_{\leq \ell}(x): X_{B} \subseteq X_{B_o}\}$ is nonempty (it contains $B_o$); pick any minimal ball $B$ within it.
+Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimates}) and thus $s(x) \in \PL_\ell(x)$.
+
+For (b), take any $\ell \geq L_2(x)$.
+Consider any $B \in \B_{\leq \ell}(x)$ for which $s(x) \cdot \eta_X(B) < 0$.
+By definition of $L_2(x)$, this $B$ must lie in $B_{< L_2(x)}(x)$, and moreover there must exist $B' \in \B_{\leq L_2(x)}(x) \subset \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
+Thus, any minimal $B \in \B_{\leq \ell}(x)$ has bias $\geq 0$ towards the correct label $s(x)$, whereupon $\yh(B) \in \{0, s(x)\}$ by the $\gamma$-accuracy of bias estimates.
+Therefore, $-s(x) \not\in \PL_\ell(x)$.
+
+\subsection{The region of focused sampling: Proof of Lemma~\ref{lemma:focused}}
+
+Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$.
+From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty set at level $\ell$.
+In short,
+\begin{equation}
+\overline{U}_{\ell} \subset \{x \in X: L_2(x) \geq \ell\}.
+\label{eq:uncertainty}
+\end{equation}
+
+We see from the {\tt Focused-query} subroutine (Figure~\ref{alg:sampling}) that all focused samples at level $\ell$ lie in
+$$
+\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
+\ 
+\subset 
+\ 
+\bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
+\ 
+=
+\ 
+\{z \in \Delta_\ell: T_z \leq \tau_\ell \}.
+$$
+
+\subsection{A generic label complexity bound}
+
+We start by showing that various subsets of interest contain roughly the expected number of points at each level.
+\begin{lemma}
+With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/2k)$.
+\begin{enumerate}
+\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| < 2 n \tau_\ell$.
+\item[(b)] If $\Delta_\ell \neq \emptyset$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| < 2 |\Delta_\ell| \tau_\ell$.
+\end{enumerate}
+\label{lemma:level-distribution}
+\end{lemma}
+\begin{proof} 
+Pick any subset $S \subset X$ and let $m = |S|$.
+Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$.
+The probability that it is greater than or equal to twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m \tau_{\ell}/3}$, which is $\leq e^{-k/3}$ as long as $m \tau_\ell \geq k$.
+
+Both parts follow from this principle; and we take a union bound over all $\lg (n/k)$ levels.
+For (b), we need to check that $|\Delta_\ell| \tau_\ell \geq k$.
+To see this, observe from the definition (\ref{eq:sampling-region}) of $\Delta_\ell$ that if it is non-empty, then it contains $X_B$ for at least one ball $B \in \B_\ell$, and every such ball has at least $n/2^{\ell+1}$ points.
+Combining this with the definition $\tau_\ell = \min(2^{\ell+2}k/n, 1)$ yields $|\Delta_\ell| \tau_\ell \geq k$.
+\end{proof}
+
+Theorem~\ref{thm:label-complexity} is a restated version of the following.
+\begin{thm}
+Suppose the active learning algorithm makes $0 < m \leq n$ queries. Then all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
+$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
+and $\ell_2$ is the largest integer $\leq \lg (n/2k)$ such that
+$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} |\Delta_{\ell}| \, \tau_{\ell} \ < \ \frac{m}{8} .$$
+\label{thm:label-complexity-0}
+\end{thm}
+
+\begin{proof}
+Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}.
+We will analyze the effect of background sampling in phase one and focused sampling in phase two.
+We start with the former.
+
+Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.
+
+Now, for $\ell_1$ as defined, we have that $\tau_{\ell_1} \leq m/8n$ and thus from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy $T_x \leq \tau_{\ell_1}$.
+Therefore all such points are queried in phase one, and all label-estimates $\{\yh_{\ell_1}(x): x \in X\}$ are set.
+
+It then follows from Lemma~\ref{lemma:boundary}(a) that the following hold for any $x \in X$ with $L_1(x) \leq \ell_1$:
+\begin{enumerate}
+\item[(a)] By the end of phase one, $\yh_{\ell_1}(x) \in \{g^*(x), !\}$.
+\item[(b)] For any $\ell > \ell_1$, when $\yh_\ell(x)$ becomes available, it lies in $\{g^*(x), !\}$.
+\item[(c)] If $x$ ever leaves the combined uncertainty region $U = \cup_\ell U_\ell$ during phase two, then its final label as defined in (\ref{eq:final-label}) is henceforth always $\yh(x) = g^*(x)$.
+\end{enumerate}
+
+Now let's move on to phase two.
+Let $A = \{x \in X: L_1(x) \leq \ell_1, L_2(x) \leq \ell_2\}$.
+We will show that every point in $A$ must leave the uncertainty region $U$ at some time during phase two.
+From (c), we can conclude that all these points have their final labels set correctly, once and for all.
+
+We will break the argument into two cases.
+
+{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.
+
+{\it Case 2:} A full $m/4$ focused queries are made in phase two.
+By the analysis of phase one, none of these queries can be at level $\leq \ell_1$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_1+1$ through $\ell_2$ inclusive is at most
+$$\sum_{\ell=\ell_1+1}^{\ell_2} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
+\ \leq \ 
+\sum_{\ell=\ell_1+1}^{\ell_2} 2 |\Delta_{\ell}| \tau_\ell
+\ < \ 
+\frac{m}{4} ,
+$$
+where the first inequality is from Lemma~\ref{lemma:level-distribution}(b).
+Thus at least one query in phase two must be at level $\ell_2 +1$.
+When this query is made, every $U_\ell$ with $\ell \leq \ell_2$ must be empty and thus all of $A$ must have left the uncertainty region; recall from (\ref{eq:uncertainty}) that no $x \in A$ can be part of $U_\ell$ for $\ell > \ell_2 \geq L_2(x)$.
+
+Thus every $x \in A$ must leave the uncertainty region at some point in phase two, and their final labels are subsequently set correctly.
+\end{proof}
+
+\subsection{One-dimensional monotonic $\eta$: Proof of Theorem~\ref{thm:oned-monotonic}}
+
+We begin by bounding the critical levels $L_1$ and $L_2$ for points in $X$.
+\begin{lemma}
+Pick any $x \in [0, 1]$.
+\begin{enumerate}
+\item[(a)] Define $n^- = |[0,\lambda_L] \cap X|$ and $n^+ = |[\lambda_R,1] \cap X|$. Then
+$$
+L_1(x)
+\leq
+\left\{
+\begin{array}{ll}
+\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
+\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
+\end{array}
+\right.
+$$
+\item[(b)] Let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$, counting $x$ itself.
+That is, $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$.
+Then $L_2(x) \leq \lg (n/r(x))$.
+\end{enumerate}
+\label{lemma:oned-monotonic-L}
+\end{lemma}
+\begin{proof}
+For (a), take any $x \geq \lambda_R$ (the other case is similar).
+The interval $B = [\lambda,1]$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta_X(B) \geq \gamma$.
+Furthermore, any $B' \subset B$ also has $\eta_X(B') \geq \gamma$.
+
+For (b), take $x \geq \lambda_R$ and $\ell \geq \lg (n/r(x))$.
+Any $B \in \B_\ell(x)$ contains $< n/2^\ell \leq r(x)$ points and thus cannot possibly extend to the other side of the boundary.
+It follows that every $B \in \B_{\geq \ell}(x)$ has $\eta_X(B) \geq 0$.
+Moreover, any interval $B'$ that does extend to the other side of the boundary contains $B' \in \B_{\ell}(x)$ that is entirely on the same side as $x$.
+\end{proof}
+
+We can now bound the size of the query region at each level and find that it shrinks exponentially with $\ell$.
+\begin{lemma}
+For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq 4n/2^\ell$.
+\label{lemma:oned-monotonic-query-region}
+\end{lemma}
+
+\begin{proof}
+We have
+\begin{align*}
+\Delta_\ell 
+&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
+&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
+.
+\end{align*}
+This includes at most $n/2^{\ell-1}$ points from $X$ on either side of $\lambda$. \end{proof}
+
+We are now ready for the proof of Theorem~\ref{thm:oned-monotonic}.
+
+Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}.
+Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.
+
+First observe that for any $\ell_1 \leq \ell_2$, we have from Lemma~\ref{lemma:oned-monotonic-query-region} that
+$$ \sum_{\ell=\ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell 
+\ \leq \ \sum_{\ell=\ell_1+1}^{\ell_2} \frac{4n}{2^\ell} \cdot \frac{2^{\ell+2} k}{n} 
+\ = \ 16k(\ell_2-\ell_1).$$
+Now, let's define
+$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor, 
+\ \ \ \ell_2 = \min \left( \ell_1 + \frac{m}{128k}, \ \ \lg \frac{n}{2k} \right).$$
+Then for any $x \in [0, \lambda_L] \cup [\lambda_R,1]$, we have
+$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor 
+\ \geq \ 
+\lg \frac{m}{64k}
+\ \geq \ 
+\lg \frac{n}{\min(n^+,n^-)}
+\ \geq \ 
+L_1(x)$$
+and, if $\min(r^+, r^-) \geq 2k$,
+\begin{align*}
+\ell_2 
+\ = \ \ell_1 + \frac{m}{128k} 
+&\geq \lg \frac{m}{64k} + \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \\
+&\geq \ \lg \frac{m}{64k} + \lg \frac{64kn/m}{\min(r^+,r^-)} 
+\ = \ \lg \frac{n}{\min(r^+,r^-)}
+\ \geq \  
+L_2(x).
+\end{align*}
+We get the algorithmic guarantee by applying Theorem~\ref{thm:label-complexity}.
+
+\subsection{One-dimensional data with Massart noise}
+\label{sec:oned-massart}
+
+We now turn to another one-dimensional setting.
+Once again, $X$ consists of $n$ arbitrarily-placed points in $\X = [0,1]$.
+This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
+\begin{itemize}
+\item There are $p$ disjoint open intervals $I_1, \ldots, I_p$, such that $\X$ is (the closure of) their union, and
+\item for each $j$, either $\eta(x) > \gamma$ for all $x \in I_j$ or $\eta(x) < -\gamma$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
+\end{itemize}
+Here $0 < \gamma < 1$ is some constant. See Figure~\ref{fig:oned}(b) for an illustrative example. For concreteness, the intervals $I_j$ can be written in the form $(\lambda_{j-1}, \lambda_j)$, where
+$0 = \lambda_0 < \lambda_1 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$
+Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
+
+We will take $\B$ to consist of all open intervals of $[0,1]$, with $\B(x)$ denoting intervals that contain point $x$.
+
+We begin with bounds on the $L_1$ and $L_2$ levels for each point.
+\begin{lemma}
+Pick any $x \in \X$; suppose $x \in I_j$.
+\begin{enumerate}
+\item[(a)] Let $n_j = |X \cap I_j|$. Then $L_1(x) \leq \lg (n/n_j)$.
+\item[(b)] Let $r(x)$ be the minimum number of data points that lie between $x$ and a boundary point, counting $x$ as well; this is either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller.
+Then $L_2(x) \leq \lg (n/r(x))$.
+\end{enumerate}
+\label{lemma:oned-massart-L}
+\end{lemma}
+\begin{proof}
+For (a), notice first that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$.
+Moreover, $s(I_j) \cdot \eta_X(I_j) > \gamma$.
+Thus $I_j$ belongs to $\B_\ell(x)$ and is strongly biased towards the correct label.
+This strong bias also holds for any subset of $I_j$.
+
+For (b), consider any $\ell \geq \lg (n/r(x))$.
+Any $B \in \B(x)$ with $\mbox{sign}(\eta_X(B)) \neq s(I_j)$ must contain either the entire left-interval $(\lambda_{j-1},x]$ or the entire right-interval $[x, \lambda_j)$, and thus has at least $r(x)$ points, which means that it is too large to be in $\B_\ell$.
+Thus all intervals $B \in \B_{\geq \ell}(x)$ have $s(I_j) \cdot \eta_X(B) > 0$.
+Also, for any interval $B \in \B_{<\ell}(x)$ there is some $B' \in \B_{\ell}(x)$ that is strictly contained within it.
+\end{proof}
+
+With $L_1(x)$ and $L_2(x)$ under control, it is easy to bound the size of the focused query region $\Delta_\ell$ at each level.
+\begin{lemma}
+For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq (p-1)n/2^{\ell-2}$.
+\label{lemma:oned-massart-query-region}
+\end{lemma}
+
+\begin{proof}
+We have
+\begin{align*}
+\Delta_\ell 
+&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
+&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
+.
+\end{align*}
+This includes at most $n/2^{\ell-1}$ points from $X$ on either side of each boundary point $\lambda_j$.
+\end{proof}
+
+Notice that $|\Delta_\ell|$ shrinks exponentially with $\ell$.
+With $L_1$, $L_2$, and $|\Delta_\ell|$ values in place, Theorem~\ref{thm:label-complexity} can be applied directly to give the following label complexity bound.
+
+\begin{thm}
+Pick any $0 < \epsilon, \delta < 1$. Suppose we run the algorithm of Figure~\ref{alg:main} with $k = O((1/\gamma^2) \ln (n/\delta))$. With probability at least $1-\delta$, after making
+$$ O \left( \frac{p-1}{\gamma^2} \ln \frac{p-1}{\epsilon} \ln \frac{n}{\delta} \right) $$
+queries, the algorithm will assign the correct label $g^*(x)$ to at least $1-\epsilon$ fraction of $X$, except possibly the $2k$ points of either side of each boundary point.
+\label{thm:oned-massart}
+\end{thm}
+
+\begin{proof}
+Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}.
+Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.
+
+Next, using Lemma~\ref{lemma:oned-massart-query-region}, we have that for any integers $0 \leq \ell_1 < \ell_2$,
+$$ \sum_{\ell = \ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell
+\ \leq \ 
+\sum_{\ell = \ell_1+1}^{\ell_2} \frac{(p-1) n}{2^{\ell-2}} \cdot \frac{2^{\ell+2}k}{n}
+\ = \ 
+16(p-1)k (\ell_2 - \ell_1).
+$$
+We can then apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$, for
+\begin{align*}
+\ell_1
+&= 
+\left\lfloor \lg \frac{m}{32k} \right\rfloor \\
+\ell_2
+&=
+\min \left( \ell_1 + \frac{m}{128 k(p-1)}, \ \lg \frac{n}{2k} \right)
+\end{align*}
+Using Lemma~\ref{lemma:oned-massart-L}, we have that in every target interval $I_j$ with $n_j/n = \Omega(k/m)$, all but $n \cdot 2^{-\ell_2-1}$ points will be correctly classified.
+For large enough $m$, this means that the fraction of misclassified or unclassified points in $X$ will be at most $\epsilon$ after $O(k(p-1) \log ((p-1)/\epsilon))$ queries, apart from the $2k$ points nearest the boundaries, which will remain unclassified.
+\end{proof}
+
+The $2k$ points nearest each boundary cannot be labeled by our algorithm with any certainty because they lie in intervals with a strongly positive bias as well as in intervals with a strongly negative bias.
+This qualification would be removed if were allowed to make multiple queries to each point, because in that case we would include $O(k)$ copies of each point, as explained earlier.
+
+\section{Technicalities: distributional setting}
+
+In the distributional setting, $X$ is drawn i.i.d.\ from a distribution $\mu$ on a topological space $\X$. We have a collection of regions $\B$.
+
+\subsection{Sampling level and probability mass}
+
+We start by showing how the fidelity of $\B$ (Definition~\ref{defn:conformity}) can be established.
+\begin{lemma}
+With probability at least $1-\delta$,
+\begin{enumerate}
+\item[(a)] For all $B \in \B$ with $n \mu(B) \geq 12 \ln (2|\B|/\delta)$, we have
+$$ \frac{\mu(B)}{2} \leq \frac{|X_B|}{n} \leq 2 \mu(B) .$$
+\item[(b)] For all $B \in \B$ with $n \mu(B) < 12 \ln (2|\B|/\delta)$, we have $|X_B| < 24 \ln (2 |\B|/\delta)$.
+\end{enumerate}
+\label{lemma:ball-size-bounds}
+\end{lemma}
+\begin{proof}
+It is an immediate consequence of the multiplicative Chernoff bound that with probability at least $1-\delta$, for all $B \in \B$,
+$$ \frac{|X_B|}{n} = \mu(B) \left(1 \pm \sqrt{\frac{3}{n \mu(B)} \ln \frac{2|\B|}{\delta}} \right) .$$
+\end{proof}
+
+In particular, taking $k > 24 \ln (2 |\B|/\delta)$ ensures that condition (a) hold will hold for all $B \in \B$ with $|X_B| \geq k$.
+
+Next, we will see that balls of probability mass $p$ belong to level $\ell \approx \lg (1/p)$.
+\begin{lemma}
+For any level $\ell \geq 0$ and any $B \in \B_{\ell}$ with $|X_B| \geq k$,
+$$ \left\lceil \lg \frac{1}{\mu(B)} \right\rceil -2 \leq \ell \leq  \left\lceil \lg \frac{1}{\mu(B)} \right\rceil.$$
+\label{lemma:mass-vs-level}
+\end{lemma}
+\begin{proof}
+Recall the definition of level $\ell$:
+$$ B \in \B_\ell 
+\ \Longleftrightarrow \ \frac{n}{2^{\ell+1}} \leq |X_B| < \frac{n}{2^\ell} 
+\ \Longleftrightarrow \ 2^\ell < \frac{n}{|X_B|} \leq 2^{\ell + 1}.$$
+By Lemma~\ref{lemma:ball-size-bounds}, 
+$$ \frac{1}{2 \mu(B)} \leq \frac{n}{|X_B|} \leq \frac{2}{\mu(B)}.$$
+Thus $2^{\ell} < 2/\mu(B)$ and $1/(2 \mu(B)) \leq 2^{\ell+1}$. These translate into the stated bounds on $\ell$.
+\end{proof}
+
+\subsection{Bounding $L_2$ using probabilistic distance: Proof of Lemma~\ref{lemma:L2-bound}}
+
+Let $p = \dist(x, \X^{-s(x)}) \geq 2k/n$ and let $\ell = \lceil \lg (1/p) \rceil + 1 \leq \lg (2n/k)$.
+
+For any $B \in \B_{\geq \ell}(x)$, we have $\mu(B) \leq 2|X_B|/n < 2/2^\ell \leq p$, using fidelity (Definition~\ref{defn:conformity}), the definition of sampling levels, and the definition of $\ell$, in that order.
+It follows that $B$ does not intersect $\X^{-s(x)}$, whereupon $s(x) \cdot \eta_X(B) \geq 0$.
+
+Next, pick any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$; thus $B \in \B_{< \ell}(x)$.
+By the richness of $\B$, there exists $B' \in \B_\ell(x)$ with $X_{B'} \subset X_B$; and so $s(x) \cdot \eta_X(B') \geq 0$.
+% Thus $B$ must intersect $\X^{-s(x)}$ and has probability mass $\geq p$.
+We will show that there exists $B' \in \B_{\leq \ell}(x)$ such that $B' \subset B$ and $B'$ does not intersect $\X^{-s(x)}$; whereupon $s(x) \cdot \eta_X(B') \geq 0$.
+Indeed, take $B' \in \B(x)$ to be a subset of $B$ of $\mu$-mass $p-\epsilon$ for some very small $\epsilon$; we can do this because of the absolute continuity of $\mu$.
+Then $B'$ does not touch $\X^{-s(x)}$ and by Lemma~\ref{lemma:mass-vs-level}, for small enough $\epsilon$, it lies at level $\leq \ell$.
+
+\subsection{Uncertainty region in the distributional setting: Proof of Lemma~\ref{lemma:delta-continuous}}
+
+Recall from (\ref{eq:sampling-region}) that
+$$ \Delta_\ell = \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .$$
+Consider any $z \in \Delta_\ell$.
+Then there exists $x \in \X$ with $L_2(x) \geq \ell$ and $B \in \B_\ell(x)$ such that $z \in B$.
+Now, $B \in \B_\ell(x)$ implies $|X_B|/n < 1/2^\ell$ and so (by the fidelity property) $\mu(B) < 2/2^\ell$.
+We will show that $x \in \partial_{4/2^\ell}$ and thus $z \in \partial_{4/2^\ell, 2/2^{\ell}}$.
+
+There are two cases for $x$. If $\dist(x, \X^{-s(x)}) < 2k/n$ then we immediately have $x \in \partial_{4/2^\ell}$ since $4/2^\ell \geq 2k/n$.
+
+On the other hand, if $\dist(x, \X^{-s(x)}) \geq 2k/n$ then, since $L_2(x) \geq \ell$, we can apply Lemma~\ref{lemma:L2-bound} to get 
+$$ \ell 
+\ \leq \ \left\lceil \lg \frac{1}{\dist(x,\X^{-s(x)})} \right\rceil + 1 
+\ \leq \ 
+\lg \frac{1}{\dist(x,\X^{-s(x)})} + 2 .
+$$
+Thus $\dist(x, \X^{-s(x)}) \leq 1/2^{\ell-2}$ and $x \in \partial_{4/2^\ell}$.
+
+\subsection{A generic label complexity bound in the distributional setting}
+
+We will need to relate the size of the query region $\Delta_\ell$ to the probability mass of the corresponding second-order boundary.
+For this, we provide an analog of Lemma~\ref{lemma:level-distribution} for the distributional setting.
+\begin{lemma}
+With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/4k)$.
+\begin{enumerate}
+\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
+\item[(b)] $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 n \mu(\partial_{4/2^\ell, 2/2^\ell}) \tau_\ell$.
+\end{enumerate}
+\label{lemma:level-distribution-dist}
+\end{lemma}
+\begin{proof} 
+Part (a) is as in Lemma~\ref{lemma:level-distribution}.
+
+For (b), pick any subset $S \subset \X$.
+If $X$ consists of $n$ independent draws from $\mu$, then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(n, \mu(S) \tau_\ell)$ distribution with expectation $n \mu(S) \tau_\ell$.
+The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-n \mu(S) \tau_\ell/3}$.
+We will apply this to the various sets $S = \partial_{4/2^\ell, 2/2^\ell}$ and take a union bound over them.
+In each application, we will also see that $n \mu(S) \tau_\ell \geq k$.
+
+Pick any $\ell \leq \lg (n/4k)$.
+If $\partial_{4/2^\ell, 2/2^\ell} = \emptyset$, then the statement in (b) is trivially true given Lemma~\ref{lemma:delta-continuous}.
+So assume this is not the case.
+Writing $p = 4/2^\ell$, we need to check that $\mu(\partial_{p, p/2}) \tau_\ell \geq k/n$, or equivalently, $\mu(\partial_{p, p/2}) \geq \max(1/2^{\ell+2}, k/n) = 1/2^{\ell+2}$.
+Now, $\partial_{p,p/2} \neq \emptyset \Longrightarrow \partial_p \neq \emptyset$.
+Pick any $x \in \partial_p$.
+We have by the richness of $\B$ (Definition~\ref{defn:conformity}) that there exists $B \in \B_\ell(x)$.
+By the definition of level, $p/8 \leq |X_B|/n < p/4$ and thus, by the fidelity property, $p/16 \leq \mu(B) < p/2$.
+It follows that $B \subset \partial_{p,p/2}$, whereupon and $\mu(\partial_{p,p/2}) \geq \mu(B) \geq p/16 = 1/2^{\ell+2}$.
+%By absolute continuity of $\mu$, we can grow a ball around $x$ of probability mass arbitrarily close to $p/2$, so that this ball is contained within $\partial_{p,p/2}$.
+Thus $\mu(\partial_{p,p/2}) \geq p/2-\epsilon$ for any $\epsilon > 0$.
+The required conditions then follow from the value of $p$.
+\end{proof}
+
+\begin{thm}
+Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$ and that the active learning algorithm makes $0 < m \leq n$ queries.
+Then with probability at least $1-3\delta$, all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
+$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
+and $\ell_2$ is the largest integer $\leq \lg (n/4k)$ such that
+$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell}) \ < \ \frac{m}{32k} .$$
+\label{thm:label-complexity-dist}
+\end{thm}
+
+\begin{proof}
+The proof is identical to that of Theorem~\ref{thm:label-complexity-0}; the only change is to use Lemma~\ref{lemma:level-distribution-dist}(b) in place of Lemma~\ref{lemma:level-distribution}(b).
+\end{proof}
+
+\subsection{Rates of convergence under distributional conditions}
+
+We now give label complexity bounds under three assumptions.
+Let $\X_\gamma = \{x \in \X: |\eta(x)| \geq \gamma\}$; thus $X \cap \X_\gamma$ is the set of points whose assigned labels will be judged.
+We further divide this set by label: for $s \in \{-1,+1\}$, let $\X^s_\gamma = \{x \in \X: s \cdot \eta(x) \geq \gamma\}$.
+
+The first assumption is on the curvature of the decision regions. It allows us to bound $L_1(x)$.
+\begin{enumerate}
+\item[(A1)] There is an absolute constant $p_o > 0$ for which the following holds: for any $x \in \X_\gamma$, there exists $B \in \B(x)$ such that $B \subset \X^{s(x)}_\gamma$ and $\mu(B) \geq p_o$.
+\end{enumerate}
+\begin{lemma}
+Under (A1), every $x \in \X_\gamma$ has $L_1(x) \leq \lceil \lg (1/p_o) \rceil$, provided $n > k/2p_o$.
+\label{lemma:L1-bound}
+\end{lemma} 
+\begin{proof}
+Pick any $x \in X_\gamma$; apply (A1) to get $B \in \B(x)$ for which $\mu(B) \geq p_o$ and $B \subset \X^{s(x)}_\gamma$.
+By Lemma~\ref{lemma:mass-vs-level}, $B \in \B_\ell$ for $\ell \leq \lceil \lg (1/p_o) \rceil$.
+Now, $s(x) \cdot \eta_X(B) \geq \gamma$; moreover, for any $B' \in \B(x)$ with $X_{B'} \subset X_B$ we have $X_{B'} \subset \X^{s(x)}_\gamma$ and thus $s(x) \cdot \eta_X(B') \geq \gamma$ as well.
+\end{proof}
+
+The second assumption is a variant of the Tsybakov margin condition and bounds $|\Delta_\ell|$.
+\begin{enumerate}
+\item[(A2)] There are absolute constants $C > 0$ and $0 < \sigma < 1$ for which the following holds: for any $p > 0$, we have $\mu(\partial_{p,p}) \leq C p^\sigma$.
+\end{enumerate}
+If, for instance, $\mu$ were the uniform distribution over $[0,1]^d$, we would expect $\sigma \sim 1/d$.
+
+The third assumption bounds the fraction of points with large $L_2$ values.
+It asks, what fraction of the $p$-boundary touches $\X_\gamma$?
+We consider two options: under (A3), the fraction is zero for $p$ small enough; under (A3'), the fraction is $p^\xi$.
+\begin{enumerate}
+\item[(A3)] There is an absolute constant $p_1 > 0$ such that $\mu(\partial_p \cap \X_\gamma) = 0$ for any $p < p_1$.
+\item[(A3')] There are constants $C' > 0$ and $0 < \xi < 1$ such that $\mu(\partial_{p} \cap \X_\gamma) \leq C' p^\xi$ for any $p > 0$.
+\end{enumerate}
+
+Under these assumptions, we get rates of convergence as follows.
+\begin{thm}
+Assume (A1), (A2) and either (A3) or (A3').
+There is an absolute constant $C''$ for which the following holds.
+Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
+If the algorithm of Fig.~\ref{alg:main} makes $m$ queries, for
+$$ \frac{128k}{p_o} \ \leq \ m \ \leq \ \frac{64C \cdot 8^\sigma}{1-\sigma} \, n^{1-\sigma} k^\sigma,$$
+then with probability at least $1-\delta$:
+\begin{itemize}
+\item Under (A3), all of $X \cap \X_\gamma$ get Bayes-optimal labels for $m > (512C/(1-\sigma)) \cdot (1/p_1)^{1-\sigma} \cdot k$.
+\item Under (A3'),  Bayes-optimal labels get assigned to all but $ C'' (k/m)^{\xi/(1-\sigma)} + (2/n) \log (4/\delta)$
+%$$ C'' \left(\frac{k}{m}\right)^{\xi/(1-\sigma)} + \frac{2 \log (4/\delta)}{n} $$
+fraction of $X \cap \X_\gamma$.
+\end{itemize}
+\label{thm:label-complexity-specific}
+\end{thm}
+This theorem yields the result of Section~\ref{sec:Massart}. Another example is in Section~\ref{sec:Tsybakov}, where bounds are given under smoothness and Tsybakov margin conditions.
+
+\begin{proof}
+In this case, $\B$ is infinite, but by standard VC-dimension arguments there are only $O(n^{d+1})$ balls with distinct sets $X_B$. This governs the setting of $k$.
+
+First, define $\ell_1 = \lfloor \lg (m/32k) \rfloor$ and observe that by Lemma~\ref{lemma:L1-bound}, all points in $\X_\gamma$ have 
+$$ L_1(x) 
+\ \leq \ 
+\left\lceil \lg \frac{1}{p_o} \right\rceil
+\ \leq \ 
+\lg \frac{2}{p_o} 
+\ \leq \ 
+\lg \frac{m}{64k}
+\ \leq \ 
+\ell_1 .
+$$
+Next, pick  
+$$ \ell_2 = \left\lfloor \frac{1}{1-\sigma} \left( \lg \frac{m}{32k} + \lg \frac{1-\sigma}{C \cdot 4^{1+\sigma}} \right) \right\rfloor.$$
+The upper bound on $m$ ensures that this is at most $\lg (n/2k)$. Then, using (A2),
+$$
+\sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell})
+\ \leq \ 
+C \sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \left( \frac{4}{2^\ell} \right)^\sigma 
+\ \leq \ 
+C \cdot 4^\sigma \cdot \frac{2}{2^{1-\sigma}-1} \cdot 2^{(1-\sigma)\ell_2}
+\ \leq \ 
+\frac{m}{32k} .
+$$
+Applying Theorem~\ref{thm:label-complexity-dist}, we then see that all points in $\X_\gamma$ with $L_2(x) \leq \ell_2$ will be correctly classified.
+Any remaining point $x \in \X_\gamma$ has $L_2(x) > \ell_2$ and thus (by Lemma~\ref{lemma:L2-bound})
+$$ \left\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \right\rceil + 1 > \ell_2 
+\ \Longrightarrow \ 
+\dist(x, \X^{-s(x)}) < \frac{4}{2^{\ell_2}} \leq  \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} . 
+$$
+Call this quantity $p$; thus any such $x$ lies in $\partial_p$.
+
+Under (A3), $\partial_p \cap \X_\gamma$ has zero probability mass for $p < p_1$, that is, if 
+$$ \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} < p_1
+\ \Longleftrightarrow \ 
+m > \frac{512C}{1-\sigma} \cdot \frac{1}{p_1^{1-\sigma}} \cdot k .
+$$
+
+Under (A3'), $\mu(\partial_p \cap \X_\gamma) \leq C' p^{\xi}$; we can then apply a Bernstein bound to assert that with probability at least $1-\delta$, 
+$$ |X \cap (\partial_p \cap \X_\gamma)| \ \leq \ \frac{3}{2} C' n p^{\xi} + 2 \log \frac{1}{\delta},$$
+from which the bound in the theorem follows by defining $C''$ appropriately.
+\end{proof}
+
+\subsection{Label complexity under curvature and Massart noise: Proof of Theorem~\ref{thm:massart-dist}}
+
+Let (C1) denote the strong density condition and (C2) the Massart noise and boundary condition.
+Theorem~\ref{thm:massart-dist} follows immediately from the more general result of Theorem~\ref{thm:label-complexity-specific}, once we establish how (C1) and (C2) relate to assumptions (A1), (A2), and (A3').
+
+\begin{lemma}
+Conditions (C1) and (C2) yield assumption (A1), with $p_o = r_o^d \cdot c_o \cdot v_d$, where $v_d$ is the volume of the unit ball in $\R^d$.
+\label{lemma:A1}
+\end{lemma}
+
+\begin{proof}
+Pick any $x \in \X_\gamma^+$ (the negative case is similar), and let $r = \inf_{z \in \X^0} \|x - z\|$.
+If $r \geq r_o$, then $B = B(x,r_o)$ is entirely in $\X_\gamma^+$.
+Otherwise, the reach condition (C2) implies the existence of a ball $B \subset \X_\gamma^+$ that contains $x$ and has radius $r_o$.
+Either way, $\mu(B) \geq c_o \vol(B) = c_o v_d r_o^d$ by (C1).
+\end{proof}
+
+\begin{lemma}
+Conditions (C1) and (C2) yield assumptions (A2) and (A3') with $\sigma = \xi = 1/d$. 
+\end{lemma}
+
+\begin{proof}
+Under (C1), any ball of probability mass $\leq p$ has volume $\leq p/c_o$ and radius $\leq (p/(c_o v_d))^{1/d}$.
+Let's call this latter quantity $r$.
+Thus, any point in $\partial_p$ lies within distance $2r$ of the boundary, while a point in $\partial_{p,p}$ lies within distance $4r$.
+
+To bound the volume of $\partial_{p,p}$, we can associate each point in this region with its nearest neighbor in $\X^0$; by condition (C1), this projection map is uniquely defined for $r < r_o/4$.
+The volume of the region is thus $O(r)$ and under (C1), has probability mass $O(r) = O(p^{1/d})$.
+\end{proof}
+
+\subsection{Label complexity under smoothness and Tsybakov noise}
+\label{sec:Tsybakov}
+
+Now we specialize Theorem~\ref{thm:label-complexity-specific} to another situation, where the data distribution satisfies Holder smoothness and Tsybakov margin conditions.
+
+Recall that $\dist(x,S)$ denotes the probability-distance from $x$ and set $S$. We will overload notation so that for $x,x' \in \X$,
+$$ \dist(x,x') = \dist(x, \{x'\}) = \inf\{\mu(B): B \in \B(x) \cap \B(x')\},$$
+that is, the probability mass of the smallest ball containing both $x$ and $x'$. We will impose the following conditions on the data.
+\begin{enumerate}
+\item[(C1)] [Strong density condition] This was introduced in the earlier (Massart) example.
+\item[(C2')] [Holder-smoothness of conditional probability function] There exist constants $L, \alpha$ such that
+$$ |\eta(x) - \eta(x')| \ \leq \ L \cdot \dist(x,x')^\alpha$$
+for all $x,x' \in \X$.
+\item[(C3')] [Tsybakov margin condition] There exists constants $M, \beta$ such that
+$$ \mu(\{x \in \X: |\eta(x)| \leq \tau\}) \leq M \tau^\beta$$
+for all $\tau \in (0,1)$.
+\item[(C4')] [Bounded curvature] The boundaries $\{x \in \X: \eta(x) = \gamma\}$ and $\{x \in \X: \eta(x) = -\gamma\}$ are $(d-1)$-dimensional Riemannian manifolds of reach $r_o > 0$.
+\end{enumerate}
+
+Lemma~\ref{lemma:A1} continues to hold, with condition (C4') doing the job of (C2). This yields assumption (A1). For the remaining assumptions, we first obtain a consequence of the Holder condition.
+
+\begin{lemma}
+Under (C2'), for any $p, q > 0$,
+\begin{enumerate}
+\item[(a)] $x \in \partial_p \implies |\eta(x)| \leq L p^\alpha$.
+\item[(b)] $x \in \partial_{p,q} \implies |\eta(x)| \leq L(p^\alpha + q^\alpha)$.
+\end{enumerate}
+\label{lemma:holder-consequence}
+\end{lemma}
+\begin{proof}
+Pick any $x \in \partial_p$.
+Let $s(x) = \mbox{sign}(\eta(x))$.
+By definition of the $p$-boundary set, for any $\epsilon > 0$, there exists $x' \in \X^{-s(x)}$ such that $\dist(x,x') < p+\epsilon$.
+By the Holder condition, $|\eta(x) - \eta(x')| < L(p+\epsilon)^\alpha$ and thus $|\eta(x)| < L(p+\epsilon)^\alpha$.
+Since this holds for any $\epsilon > 0$, we get part (a).
+
+For (b), pick $x \in \partial_{p,q}$ and $\epsilon > 0$.
+Then there exists $x' \in \partial_p$ with $\dist(x,x') < q+\epsilon$.
+As before, we use the Holder condition to conclude that $|\eta(x)| < |\eta(x')| + L(q+\epsilon)^\alpha$ and then invoke (a).
+\end{proof}
+
+\begin{lemma}
+Conditions (C2') and (C3') yield assumptions (A2) and (A3) with $C = (2L)^\beta M$, $\sigma = \alpha \beta$, and $p_1 = (\gamma/L)^{1/\alpha}$.
+\label{lemma:boundary-constants} 
+\end{lemma}
+
+\begin{proof}
+By Lemma~\ref{lemma:holder-consequence}, $\partial_{p,p} \subset \{x \in \X: |\eta(x)| \leq 2Lp^\alpha\}$; the probability mass of this set can be bounded by (C3').
+
+Also by Lemma~\ref{lemma:holder-consequence}, $\partial_p$ is entirely contained in $\{x \in \X: |\eta(x)| \leq L p^\alpha\}$. For $p < p_1$, this does not intersect $\X_\gamma$.
+\end{proof}
+
+\begin{thm}
+Assume conditions (C1) and (C2')--(C4').
+There are constants $c_2, c_3$ for which the following holds.
+Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
+Suppose the algorithm of Fig.~\ref{alg:main} makes $m$ queries, where
+$$ \frac{c_2k}{p_1^{1-\sigma}} \ \leq \ m \ \leq \ c_3 n^{1-\sigma} ,$$
+with $p_1, \sigma$ as given in Lemma~\ref{lemma:boundary-constants}. Then with probability at least $1-\delta$, all of $X \cap \X_\gamma$ is assigned Bayes-optimal labels.
+\label{thm:smoothness-dist}
+\end{thm}
+
+In this setting, regular $k$-nearest neighbor classification would also correctly classify all of $X \cap X_\gamma$ if given $O(k/p_1)$ random labeled points.
+The benefit of active learning is to thus reduce the label complexity to $O(k/p_1^{1-\alpha\beta})$.
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2024-07-31 old/active-knn-old1.tex
Initial Overleaf Import

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{natbib}

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}

\newtheorem{thm}{Theorem}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{claim}[thm]{Claim}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assump}{Assumption}
\newtheorem{open}{Open problem}
\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}

\DeclareMathOperator*{\argmax}{arg\,max}

\title{An active $k$-nearest neighbor algorithm}

\begin{document}

\maketitle

\section{Setup and goals}

Basic setup:
\begin{itemize}
\item Instance space $\X \subset \R^d$. Data are generated from a distribution $\mu$ on $\X$.
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
\item Let $g^*$ denote the Bayes-optimal classifier on $\X$:
$$ g^*(x) = \mbox{sign}(2 \eta(x) - 1) .$$
\end{itemize}

\noindent
Goal: \emph{pool-based active learning}
\begin{itemize}
\item We have a finite dataset $X \subset \R^d$ of $n$ points, and we are able to query the labels of any of them.
\item Given parameters $0 < \epsilon, \delta < 1$, our goal is to query the labels of some points $Q \subset X$, and infer the labels of the remainder, so that with probability at least $1-\delta$, at most $\epsilon n$ of the inferred labels disagree with $g^*$.
\end{itemize}

\noindent
Alternative goal: \emph{stream-based active learning}
\begin{itemize}
\item Data arrive continuously, one point at a time. For each point, we decide whether or not to query its label.
\item The classifier $g_t$ at time $t$ is (some kind of) $k$-nearest neighbor classifier based on the points labeled so far. We require $g_t \rightarrow g^*$ and we consider the rate of convergence in terms of the number of labeled points.
\end{itemize}

\section{Notation}

For any $x \in X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$.

For any $p, \gamma> 0$, let $\X_{p, \gamma}$ consist of all points $x \in \supp(\mu)$ such that:
\begin{itemize}
\item either $\eta(B(x,r)) \geq 1/2 + \gamma$ for all $0 \leq r \leq r_p(x)$ 
\item or $\eta(B(x,r)) \leq 1/2 - \gamma$ for all $0 \leq r \leq r_p(x)$.
\end{itemize}

\section{Algorithm for pool-based active learning}

Global variables:
\begin{itemize}
\item $U \subset X$, a subset of points deemed ``uncertain''
\item $Q \subset X$, points whose labels have been queried
\item $R \subset Q$, points in $Q$ that are random draws from the entire space $X$
\item $I \subset X$, points whose labels have been inferred, possibly incorrectly
\end{itemize}

We will use large deviation bounds for the class $\B$ of balls centered at points of $X$. This is a class of size at most ${n \choose 2}$. The bounds will assert the following: suppose $m$ points are drawn i.i.d. from $\mu|_S$, the restriction of $\mu$ to a subset $S \subset \X$, and labeled according to $\eta$. Then with probability at least $1-\delta$, the following properties hold for all $B \in \B$:
\begin{enumerate}
\item[(P1)] If $\mu|_S(B) \geq (k/m) \ln (n/\delta)$, then $B$ gets at least $k+1$ points.
\item[(P2)] If $B$ gets at least $k$ points, then the average of the labels in $B$ is equal to $\eta(B) \pm k^{-1/2}$.
\end{enumerate}
We will use several rounds of sampling. Let $G_t$ denote the good event that these conditions hold for the points drawn in round $t$.

\subsection{A single round of sampling}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.3in}
{\bf function Extend}($k$, $p$, $\delta$):
\begin{itemize}
\item Setup:
\begin{itemize}
\item Define the region of uncertainty:
$$ U = X \setminus (Q \cup I) $$
\item Determine the sampling region:
$$ S =  \left( \bigcup_{x \in U} B(x, r_p(x)) \right) \cap X $$
\item Choose number of samples:
$$ m = \frac{|S|}{|X|} \cdot \frac{k}{p} \cdot \ln \frac{n}{\delta} $$
\end{itemize}

\item Querying:
\begin{itemize}
\item Let $M$ consist of $m$ points chosen uniformly at random from $S$
\item Let $M'$ consist of $m$ points chosen uniformly at random from $X$
\item For all $x \in (M \cup M') \setminus Q$: 
\begin{itemize}
\item Query $x$'s label
\item Add $x$ to $Q$
\end{itemize}
\item $R = R \cup M'$
\end{itemize}

\item Update:
\begin{itemize}
\item For all $x \in U$:
\begin{itemize}
\item Find the $k$ nearest neighbors of $x$ in $M$ (can probably do $M \cup R$)
\item If their labels have significant bias, in the sense that their average has absolute value $\geq k^{-1/2}$, label $x$ accordingly and add it to $I$
\end{itemize}
\item For all $x \in I$:
\begin{itemize}
\item Find the $k$ nearest neighbors of $x$ in $R$
\item If their labels have significant bias and disagree with $x$'s current label, move all of $B(x,r_p(x)) \cap (I \setminus \{x\})$ out of $I$ (not sure exactly what to do here)
\end{itemize}
\end{itemize}
\end{itemize}
\end{minipage}}
\caption{Extend: the key subroutine for pool-based active learning}
\label{fig:extend}
\end{figure}

\begin{lemma}
Consider round $t$ of sampling. Suppose event $G_t$ holds. For any $x$, let $\widetilde{r}(x)$ denote the smallest radius $r$ such that $B(x,r)$ contains $k+1$ points of $M$. Then for all $x \in U$,
\begin{enumerate}
\item[(a)] $\widetilde{r}(x) \leq r_p(x)$
\item[(b)] the bias of the $k$ nearest neighbors of $x$ in $M$ agrees with $\eta(B(x,\widetilde{r}(x))) \pm 1/k^{1/2}$.
\end{enumerate}
\label{lemma:bias-bound}
\end{lemma}
\begin{proof}
Let's begin with (a). Choose any $x \in U$ and write $B = B(x,r_p(x))$. The points in $M$ can be considered to be $m$ i.i.d. draws from $\mu|_S$. Ignoring the small discrepancy between $\mu(S)$ and $|S|/|X|$, we have 
$$ \mu|_S(B) = \frac{\mu(B)}{\mu(S)} \geq \frac{p}{\mu(S)} \geq \frac{k}{m} \ln \frac{n}{\delta}$$
and thus by (P1) $M$ includes at least $k+1$ points of $B$.

Part (b) then follows directly by appealing to (P2).
\end{proof}

\begin{thm}
Suppose that in round $t$, the starting uncertainty region is $U$. Assume event $G_t$ occurs. Then for all $x \in U$ that aren't queried, the following two properties hold.
\begin{enumerate}
\item[(a)] If $x \in \X_{p, 2k^{-1/2}}$, then by the end of the round, its label is inferred to be $g^*(x)$.
\item[(b)] If the label of $x$ is inferred but is not equal to $g^*(x)$, then $x \not\in \X_{p, k^{-1/2}}$.
\end{enumerate}
\end{thm}
\begin{proof}
Pick any $x \in U$ whose label is not queried by the end of round $t$. By Lemma~\ref{lemma:bias-bound}, its $k$ nearest neighbors in $M$ lie within distance $\widetilde{r}(x) \leq r_p(x)$, and its $k$ nearest neighbors have average label in the range $\eta(B(x, \widetilde{r}(x))) \pm k^{-1/2}$.

For part (a), observe that by the definition of $\X_{p,\gamma}$, this average label has sign $g^*(x)$ and absolute value $> k^{-1/2}$. Thus the bias is significant and the label of $x$ is correctly inferred.

For part (b), we observe that if $x \in \X_{p, k^{-1/2}}$, then the average label of $x$'s $k$ nearest neighbors would have the correct sign, $g^*(x)$.
\end{proof}

\subsection{A simple active learning loop}

Here's a simple idea: keep $k$ fixed, with the goal of trying to correctly classify all $x \in \X_{p,\gamma}$ with $p > 1/n$ and $\gamma > 1/\sqrt{k}$. We won't quite be able to do this because of mind-changes.

\begin{itemize}
\item Given: $X$, $\gamma$, $\delta$
\item $k = 1/\gamma^2$
\item $Q = I = \emptyset$
\item for $t = 0, 1, \ldots, \log n$:
\begin{itemize}
\item $p_t = 1/c^t$
\item $\delta_t = \delta/(1 + \log n)$
\item Extend($k$, $p_t$, $\delta_t$)
\end{itemize}
\end{itemize}

Let's start by analyzing this in the case where there are no mind-changes.

\subsection{Analysis: no mind changes}

Fix $\gamma = k^{-1/2}$. For $\tilde{p} \sim 1/n$, define $\widetilde{\X}^+ = \X_{\tilde{p}, \gamma}$ and likewise $\widetilde{\X}^-$. Then $\widetilde{\X} = \widetilde{\X}^+ \cup \widetilde{\X}^-$ is exactly the set of points that could reasonably be expected to be correctly classified by a $k$-NN classifier with access to all the labels in $X$. We'd like to do almost as well after querying just a fraction of the labels.

\begin{enumerate}
\item {\it No mind changes.} We will assume no mind-changes: that is, for any $r>0$,
\begin{itemize}
\item $x \in \widetilde{\X}^+ \implies \eta(B(x,r)) > 1/2 - \gamma$
\item $x \in \widetilde{\X}^- \implies \eta(B(x,r)) < 1/2 + \gamma$
\end{itemize}
This means that no point in $\widetilde{\X}$ will ever be misclassified.

\item {\it After round $t$.} At the end of round $t$, every point in $\X_{p_t, \gamma}$ will either have been queried or will have had its label correctly inferred. What remains, at most, is then $\partial_{p_t, \gamma}$, where we define
$$ \partial_{p, \gamma} = \X \setminus \X_{p, \gamma} .$$
The sampling region in the next round is then $S_{t+1} \subset R_{t+1} \cap X$, where
$$ R_{t+1} = \bigcup_{x \in \partial_{p_t, \gamma}} B(x, r_{p_{t+1}}(x))  .$$

\item {\it Smoothness condition.} Recall our nearest neighbor-inspired variant of the Holder condition: $\eta$ is $(\alpha, L)$-smooth if for all $x \in \supp(\mu)$ and all $r > 0$,
$$ |\eta(B(x,r)) - \eta(x)| \leq L \mu(B^o(x,r))^\alpha .$$
In [CD14], it is shown that under this condition, for any $p, \gamma$, we have
$$ \partial_{p,\gamma} \cap \supp(\mu) \subset \{x: |\eta(x) - 1/2| \leq \gamma + Lp^\alpha\} .$$

Here's a stronger condition: for all $x \in \supp(\mu)$ and all $r > 0$,
$$ |\eta(x') - \eta(x)| \leq L \mu(B^o(x,r))^\alpha \ \ \mbox{for all $x' \in B(x,r)$}.$$
Under this condition, it easily follows that
$$ \bigcup_{x \in \partial_{p_1, \gamma}} B(x, r_{p_2}(x)) \ \subset \ \{x: |\eta(x) - 1/2| \leq \gamma + L(p_1^\alpha + p_2^\alpha)\} .$$
\item {\it Margin condition.} For $\beta > 0$, we say $(\mu, \eta)$ satisfies the $\beta$-margin condition if there exists a constant $C > 0$ such that for any $t$,
$$ \mu(\{x: |\eta(x) - 1/2| \leq t\}) \leq C t^\beta .$$

\item {\it Bounding the size of the sampling region.} Under the strong smoothness condition, we have
$$ R_{t+1} \subset  \{x: |\eta(x) - 1/2| 
\ \leq \ 
\gamma + L(p_t^\alpha + (p_t/c)^\alpha)\} .$$
Define $L' = L(1 + c^{-\alpha})$. Under the margin condition, we then get
$$ \mu(R_{t+1}) 
\ \leq \ 
C (\gamma + L' p_t^\alpha)^\beta 
\ \leq \ 
2C \cdot \max \left(\frac{1}{k^{1/2}}, \frac{L'}{c^{\alpha t}} \right)^\beta.
$$
Initially (when $t$ is small), the dominant term is $(1/c^{\alpha \beta})^t$. Once $t \geq (1/(2 \alpha)) \log_c k$, the dominant term becomes $1/k^{\beta/2}$.

\item {\it Label complexity analysis under optimal choice of $k$.} Under the smoothness and margin condition, the optimum setting of $k$ is $n^{2\alpha/(2\alpha+1)}$.

\end{enumerate}

\begin{center}
\begin{huge}
Older material
\end{huge}
\end{center}

\pagebreak

\section{Setup}

Basic Transductive setup:
\begin{itemize}
\item Intance set $\X =\{x_1,\dots,x_n\}$, $x_i \in {\bf X}$, where
  ${\bf X}$ is a metric space. . The set
  $\X$ is fixed and known to the learner. The probability of each $x
  \in \R$ is $1/n$.
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is associated with
  a conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
\item Let $g^*$ denote the Bayes-optimal classifier on $\X$:
  $$ g^*(x) = \mbox{sign}(2 \eta(x) - 1) .$$
  \item Label Queries: the algorithm can select any example $x \in
    \X$, as a response it gets a label in $\Y$ chosen acccording to
    $\eta(x)$.    
  \end{itemize}

\section{Notation}
\begin{itemize}
\item {\bf Balls:} For any $x \in \X$ an $r\geq0$,
  $B(x,r)$ denotes the subset of $\X$ inside ball of radius $r$ centered at $x$.
  \item {\bf Probability of a ball} The probability of a ball is the
    fraction of points in $\X$ that fall inside $B(x,r)$.
    $$ \mu(B(x,r)) = \frac{ |B(x,r)|}{n}$$
  \item {\bf conditional probability given a ball:}
    $$ \eta(B(x,r)) = \frac{\sum_{x \in B(x,r)}\eta(z)}{|B(x,r)|}$$
  \item {\bf measuring radius in terms of probability}
For any $x \in X$ and any $p > 0$, let $r_p(x)$ denote the smallest
radius $r$ such that $\mu(B(x, r)) \geq p$.
\end{itemize}

\newcommand{\NNB}{{\bf NNB}}
\newcommand{\ANNB}{{\bf ANNB}}
\newcommand{\ONNB}{{\bf oracle-NNB}}

\section{The near-neighborhoods algorithm}
The near neighborhood (\NNB) algorithm is a variant of AKNN where the balls
that are considered for the prediction of a point $x$ are {\em all} balls
the contain $x$ while in AKNN only balls that are centered at $x$ are
considered.

Suppose that we are given $n$ labeled examples chosen uniformly at
random from $\X$ and labeled according to $\eta(x)$.

Given that set of labels each ball is as $-,?,+$ according to the
condition used in AKNN. We say that a ball labeled $\pm$ is {\em
  determined}, a ball labeled $?$ is {\em undetermined}.

We define the partial order on balls that is defined by containment.

The labels of the balls induce a labeling on the points in $\X$. Each
point has one of {\em four} labels: $-1,+1,?,!$. Given a point $x \in
\X$ we define the {\em minimal set of determined balls for $x$},
denoted $\B(x)$, to be the set of determined balls that contain $x$
that is minimal (in terms of the containment partial order). We can
construct this set by starting with the set of all determined balls containing
$x$ and eliminating any ball that has a subset-ball that is also in
the set.

Given the set of balls $\B(x)$ the label associated with $x$
according to the following rules:
\begin{itemize}
\item If $\B(x)$ is empty the label is undetermined, or $?$.
\item If all of the balls in $\B(x)$ are labeled $+$, the label is $+1$.
\item If all of the balls in $\B(x)$ are labeled $-$, the label is $-1$
\item If $\B(x)$ contains both labels, then the label is {\em
    over-determined} or $!$. We will also refer to over-determined
  examples as the ``known unknown''.
\end{itemize}

We define the numerical prediction $\hat{y}$ to be $0$  if the label
is undetermined or overdetermined. We set it equal to the label otherwise

We define the error of the prediction as the produce
$-\hat{y}{y}$. Note that the error can be $-1,0,+1$.

\NNB\ is a passive learning algorithm, in other words the queriy
points are selected IID from $\X$.

\begin{claim}
  The algorithm \NNB\ converges to Bayes rule as the number of
  labeled examples increases to $\infty$.
  \end{claim}

\section{Enforcing locality}
For technical reasons, we consider balls that are smaller and smaller as the algorithm progresses. Let
$1=p_1 >p_2>\cdots >p_k=1/n$ be the radius probabilities. Then in iteration $i$ we restrict the set of balls to be such that $\mu(B(x,r)) \leq p_i$.

\section{Active \NNB}

The main idea behind active \NNB\, or \ANNB\, is to concentrate the
queries on the known unknown. However, this is not done to the
exclusion of the other parts of the space, as any example can have an
unbounded number of mind changes. In other words, an example whose label at some point is $+1$
can be labeled $-1$ later on and vice versa. (These are the ``unknown
unknowns'').

\ANNB\ operates in phases. At the start of phase $t$ a concentration
area $C_t \subset \X$ is computed. Then, $2n$ queries are are
made. There are $n$ {\em passive} queries and $n$ {\em active} queries.
The passive queries are selected uniformly at random from $\X$, the
active quries are selected uniformly at random from $C_t$.

After the answers to the queries are recieved, they are used to update
the counter $k_+,k_-$ in the balls. Passive points are used in each
ball that contains them. Active points are a bit more subtle, they are
used only in balls that are conained in $C_t$. In other words if a
ball contains an element $x$ that is {\em not} in $C_t$ then this ball
cannot use {\em any} of the active samples for epoch $t$.

Based on the additional queries, the label $\pm +1,?$ for each ball is
updated, and based on that the label of each $x \in \X$ is updated.

Finally, $C_{t+1}$ is calculated as follows. Let $U_t$ be the set $x
\in \X$ such that the label of $x$ is overdetermined, or ``!''.  $C_{t+1}$
is defined as the union of the minimal sets of determined balls for $x
\in U_t$, In other words:
\[
  C_{t+1} = \bigcup \left\{\B(x,r) | x \in U_t, \mu(\B(x,r)) \leq p_t \right\}
  \]

  \section{A comparator algorithm}
  In order to bound the query complexity of \ANNB\ we consider a
  variant of \ANNB\ that is aided by and oracle. Intuitively, the
  oracle removes the need for a passive sample by informing the
  algorithm that the prediction on a point $x$ is ``final''. I.e. it
  will not change as the number of queries increases to infinity.

  For this algorithm, which we denote by \ONNB\, we can more easily
  give bounds. As this framework is local, our goal is to provide
  point-wise bounds.

  \ONNB\ differs from \ANNB\ in two important ways
  \begin{enumerate}
  \item It makes only active queries, not passive queries.
  \item The set $U_t$ contains, in addition to overdetermined
    points. The points where the label is $+1$ or $-1$ but the oracle
    informs the algorithm that there will be mind cahnges on $x$ in
    the future. 
  \end{enumerate}

  Intuitively, \ONNB\ contines to collect information on $x$ until the
  label of $x$ converges, at which point it does not need more
  samples for $\B(x)$ so it stop collecting this information.

  \section{Pointwise analysis}

  Fix a point $x \in \X$ and compare the behaviour of \NNB\ and \ONNB\
  on $x$.

  Let the support set for $x$ be th set of queries that influence the predicted label for $x$. In other words removing the examples not in the support does not effect change the way the algorithm unfolds wrt that point.

  {\bf Can we prove that } The size of the support for any example using \ONNB\ is at least half of the size of the support using \ANNB\

\pagebreak

\section{Setup}

\begin{enumerate}
\item Setting: {\it Binary classification.}

\begin{itemize}
\item Instance space $\X \subset \R^d$. Data points $x$ are generated from a distribution $\mu$ on $\X$.
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
\item The Bayes-optimal classifier on $\X$ is $g^*(x) = \mbox{sign}(2 \eta(x) - 1)$.
\end{itemize}

\item Goal: {\it Pool-based active learning.}

\begin{itemize}
\item We have a finite dataset $X \subset \R^d$ of $n$ points, and we are able to query the labels of any of them.
\item Given parameters $0 < \epsilon, \delta < 1$, our goal is to query the labels of some of the points, and infer the labels of the remainder, so that with probability at least $1-\delta$, at most $\epsilon n$ of the inferred labels disagree with the Bayes-optimal $g^*$.
\end{itemize}

\end{enumerate}

\section{An active learning algorithm}

\begin{enumerate}

\item {\it Algorithm overview.}

There are multiple rounds of labeling, and during each, query points are chosen at random from the current ``uncertainty region''.

\item {\it Sampling balls.}

There is a set of ``balls'' $\B$, balls of different sizes that can get arbitrarily small. These are the balls on which we assess label bias.

We query points that are chosen at random, under two types of sampling. 
\begin{itemize}
\item Background random sampling of the entire data set $X$.
\item Focused random sampling of a subset of $X$.
\end{itemize}
For each $B \in \B$, we know which of the labels seen so far can be considered random samples from $B$. On that basis, we assign a {\it provisional label}
$$ \yh(B) = 
\left\{
\begin{array}{ll}
+ & \mbox{significant $+$ bias} \\
- & \mbox{significant $-$ bias} \\
? & \mbox{otherwise}
\end{array}
\right.
$$

\item {\it $\Delta$-accurate bias estimates.}

The bias estimator that we use returns $\{+,-,?\}$ for any ball $B$ and has the following property that holds (with high probability) uniformly over all balls $B$ under consideration:
\begin{itemize}
\item If $\eta(B) \geq 1/2 + 2 \Delta$, it returns $+$.
\item If $\eta(B) \leq 1/2$, it does not return $+$.
\item Likewise with $-$.
\end{itemize}
Roughly speaking, this requires having $1/\Delta^2$ random labeled points in each ball.

Over time, as more random samples appear in any given ball $B$, it is theoretically possible that this provisional label would change from $+$ to $-$ or vice versa. In such cases, our large deviation bounds allow us to retain either label.

\item {\it Probability radius.}

For any $x \in \X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$.

\item {\it Balls used to evaluate a given point.}

For any $x \in \X$, let $\B_x \subset \B$ denote the collection of balls that may be used in determining $x$'s label. This could be, for instance,
\begin{itemize}
\item[(a)] all balls containing $x$, or
\item[(b)] all balls centered at $x$.
\end{itemize}

\item {\it Regions at a given probability scale.}

For any point $x \in \X$ and value $p > 0$, let $\B_{x,p}$ be a subset of $\B_x$ that is considered to be at ``probability-scale $p$''. For instance, we could define
$$ \B_{x,p} = \{B \in \B_x: B \subseteq B(x,r_p(x))\},$$
but we leave it flexible for now.

\noindent
{\it Assumption:} For any $x \in \X$ and $p > 0$, there exists some $B \in \B_{x,p}$ with $\mu(B) \geq c_o p$. Here $c_o > 0$ is an absolute constant.

\item {\it Provisional labels for data points.}

At any given time, at scale $p$, we will assign $x$ a provisional label
$$ \yh_p(x) = 
\left\{
\begin{array}{ll}
? & \mbox{if all $B \in \B_{x,p}$ have provisional label $\yh(B) = \ ?$} \\
+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
! & \mbox{if there is both a $+$ ball and a $-$ ball}
\end{array}
\right.
$$
Intuitively, the points we will deem uncertain are those that are labeled $!$ or $?$.

\item {\it The effective decision boundary.}

The true decision boundary is the set $\{x: \eta(x) = 1/2\}$, but this doesn't tell us which points we can reasonably hope to correctly label with a finite sample.

For any $p, \gamma > 0$, let $\X_{p, \gamma}^+$ consist of all points $x \in \supp(\mu)$ such that:
\begin{itemize}
\item $\eta(x) \geq 1/2 + \gamma$
\item $\eta(B) \geq 1/2$ for all $B \in \B_{x,p}$
\item $\eta(B) \geq 1/2 + \gamma$ for some $B \in \B_{x,p}$ with $\mu(B) \geq c_o p$
\end{itemize}
We can define $\X_{p,\gamma}^-$ similarly, and $\X_{p,\gamma} = \X_{p,\gamma}^+ \cup \X_{p,\gamma}^-$. The {\it effective boundary} then consists of all remaining points,
$$ \partial_{p,\gamma} = \X \setminus \X_{p,\gamma} .$$

\end{enumerate}

{\bf Yoav:} Might be possible to prove an upper bound on the number of queries if the probability of the boundary $\partial_{p,\gamma}$ is small.

\subsection{One round of sampling}

The algorithm proceeds in rounds, each with an equal amount of {\it background} sampling and {\it focused} sampling. At the beginning of round $t$:
\begin{itemize}
\item $Q_t \subseteq X$ is the set of points queried so far.

Each point $x \in Q_t$ has accompanying information that tells us, for any $B \in \B$, whether $x$ can be considered a random sample from $B$.

\item $I_t \subseteq X$ is the set of points whose labels were inferred.

\item $U_t \subseteq X$ is the set of points deemed ``uncertain'' according to some criterion.
\end{itemize}
Based on these, the focused sampling region for round $t$ is
$$ S_t = \bigcup_{x \in U_t} \bigcup_{B \in \B_{x, p_t}} B .$$
Here $(p_t)$ is a predefined sequence with $p_t \downarrow 0$.

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.3in}
{\bf function Extend}($k$, $p$, $\delta$):

\vspace{.1in}
Global variables: $Q, I, U$.

\vspace{.1in}
Setup:
\begin{itemize}
\item Sampling region:
$$ S =  \left( \bigcup_{x \in U} \bigcup_{B \in \B_{x,p}} B \right) \cap X $$
\item Choose number of samples:
  $$ m = \frac{|S|}{|X|} \cdot \frac{k + c_1 \sqrt{k}}{c_o p} $$
\end{itemize}

\vspace{.1in}
Querying:
\begin{itemize}
\item Let $M$ consist of $m$ points chosen uniformly at random from $S$
\item Query the labels of all $x \in M \setminus Q$ and add these points to $Q$
\end{itemize}

\vspace{.1in}
Infer labels using $M$:
\begin{itemize}
\item For all $B \in \{B \in \B_{x,p}: x \in U\}$: Use $M \cap B$ to set $\yh(B)$
\item For all $x \in U \setminus Q$:
\begin{itemize}
\item Use $\{\yh(B): B \in \B_{x,p}\}$ to set $\yh_{p}(x)$
\item If $\yh_{p}(x) \in \{+,-\}$ and $\mbox{\tt oracle-check}(x,p)$: Add $x$ to $I$ with the given label
\end{itemize}
\end{itemize}

\vspace{.1in}
Update uncertainty region: 
\begin{itemize}
\item $U = U \setminus (Q \cup I)$
\end{itemize}

\end{minipage}}
\caption{Subroutine for active learning given {\tt oracle-check}.}
\label{fig:extend}
\end{figure}

\subsection{Large deviation bounds}

We will use large deviation bounds for the class $\B$ of regions. 
\begin{lemma}
Choose any $0 < \delta < 1$ and define $C_\delta = \ln (4 |\B|/\delta)$. There are universal constants $c',c''>0$ for which the following is true. If $m$ points are drawn i.i.d. from $\mu|_S$, the restriction of $\mu$ to a subset $S \subset \X$, and labeled according to $\eta$, then with probability at least $1-\delta$, the following properties hold for all $B \in \B$:
\begin{enumerate}
\item[(a)] If $k \geq C_\delta$ and $\mu|_S(B) \geq (k + c' \sqrt{k C_\delta})/m$, then $B$ gets at least $k$ points.
\item[(b)] If $B \subseteq S$ gets at least $k$ points, then the empirical bias of the labels in $B$ is equal to $\eta(B) \pm c'' \sqrt{C_\delta/k}$.
\end{enumerate}
\label{lemma:bias-estimates}
\end{lemma}
\begin{proof}
These properties follow immediately from Lemmas~\ref{lemma:points-in-balls-finite} and \ref{lemma:bias}.
\end{proof}

Let $G_t$ denote the good event that these conditions hold for the points drawn in round $t$ of sampling, and let $G_{\leq t} = G_1 \cap G_2 \cap \cdots \cap G_t$.

In what follows, we will take $\delta > 0$ to be a constant and will assume that $k \geq C_\delta$. We also define $c_1 = c' \sqrt{C_\delta}$ and $c_2 = c'' \sqrt{C_\delta}$. Finally, let
$$ \Delta_k = \frac{c_2}{k^{1/2}}.$$

Our bias-estimates are computed as follows. On any round of the algorithm, let $S \subseteq \X$ denote the sampling region and $M$ the randomly chosen samples in this region. For any region $B \subseteq S$ under consideration, 
$$ \yh(B) = 
\left\{
\begin{array}{ll}
+ & \mbox{if $|B \cap M| \geq k$ and empirical bias $\geq 1/2 + \Delta_k$} \\
- & \mbox{if $|B \cap M| \geq k$ and empirical bias $\leq 1/2 - \Delta_k$} \\
? & \mbox{otherwise}
\end{array}
\right.
$$
The key property we will need for these estimates is that they are $\Delta_k$-accurate in the following sense.
\begin{defn}
Pick any $\Delta > 0$. We say bias-estimates $\yh(B) \in \{+,-,?\}$ are $\Delta$-accurate if they satisfy the following properties:
\begin{itemize}
\item If $\eta(B) \geq 1/2 + 2\Delta$, then $\yh(B) = +$.
\item If $\eta(B) \leq 1/2 - 2\Delta$, then $\yh(B) = -$.
\item If $\yh(B) = +$ then $\eta(B) > 1/2$.
\item If $\yh(B) = -$ then $\eta(B) < 1/2$.
\end{itemize}
\label{defn:delta-accurate}
\end{defn}

\begin{lemma}
Consider round $t$ of sampling, with uncertainty set $U$ and parameters $k,p$. Suppose event $G_t$ holds. Then for all $x \in U$, the bias-estimates $\{\yh(B): B \in \B_{x,p}\}$ are $\Delta_k$-accurate.
\label{lemma:accurate-bias}
\end{lemma}
\begin{proof}
This follows immediately from Lemma~\ref{lemma:bias-estimates}.
\end{proof}

\begin{thm}
Suppose that in round $t$, the starting uncertainty region is $U$. Assume event $G_{\leq t}$ occurs. Then for all $x \in U$ that aren't queried, the following two properties hold.
\begin{enumerate}
\item[(a)] If $x \in \X_{p,2\Delta_k}$, then by the end of the round, its label is inferred to be $\yh_{p}(x) = g^*(x)$.
\item[(b)] If $\yh_{p}(x) \in \{+,-\}$ but is not equal to $g^*(x)$, then $x \not\in \X_{p,0}$.
\end{enumerate}
\end{thm}
\begin{proof}
Pick any $x \in U$ whose label isn't queried by the end of round $t$. Without loss of generality $\eta(x) \geq 1/2$.

For part (a), if $x \in \X_{p,2\Delta_k}$, then every $B \in \B_{x,p}$ has $\eta(B) \geq 1/2$ and thus, by Lemma~\ref{lemma:accurate-bias}, gets $\yh(B) \in \{+,?\}$ in any round of sampling. Moreover, by assumption there is some $B' \in \B_{x,p}$ for which $\mu(B') \geq c_o p$ and $\eta(B') \geq 1/2 + 2\Delta_k$. Letting $S$ denote the sampling region in round $t$ and ignoring the small discrepancy between $\mu(S)$ and $|S|/|X|$,
$$ \mu|_S(B') = \frac{\mu(B')}{\mu(S)} \geq \frac{c_o p}{\mu(S)} \geq \frac{k + c_1 \sqrt{k}}{m}$$
and thus by Lemma~\ref{lemma:bias-estimates}(a) we have $|M \cap B'| \geq k$. Thus $\yh(B)$ would be assigned $+$ in round $t$, if not earlier, whereupon $\yh_p(x) = +$.

For part (b), suppose $\yh_{p}(x) = -$. Then some $B \in \B_{x,p}$ has $\yh(B) = -$, implying by Lemma~\ref{lemma:accurate-bias} that $\eta(B) < 1/2$. This can only be happen if $x \not\in \X_{p,0}$. 
\end{proof}

The Algorithm in Figure~\ref{fig:extend} makes use of an external subroutine called {\tt oracle-check}. This has the following semantics: $\mbox{\tt oracle-check}(x,p)$ returns {\tt true} if $x \in \X_{p,\gamma}$ and {\tt false} otherwise.

\begin{cor}
\item Pick any round $t$, and say parameters $k,p$ are used in that round. Suppose that $G_{\leq t}$ holds and that $\gamma \leq 2\Delta_k$. Let $U$ denote the uncertainty set at the start of the round. During round $t$, all points in $U \cap \X_{p,2\Delta_k}$ will be placed in $I$. Moreover, assuming the correctness of the {\tt oracle-check} subroutine, any point placed in $I$ will be labeled correctly.
\end{cor}

\subsection{An active learning loop (using the oracle)}

Here's a simple idea: keep $k$ fixed, with the goal of trying to correctly classify all $x \in \X_{p,\gamma}$ with $p > 1/n$ and $\gamma = \Delta_k$.

\begin{itemize}
\item Given: $X$, $\gamma$, $\delta$
\item Choose $k$ so that $\Delta_k = \gamma$ (i.e., $k = 4c_2^2/\gamma^2$)
\item $Q = I = \emptyset$
\item for $t = 0, 1, \ldots, T$:
\begin{itemize}
\item $p_t = 1/c^t$ for some $c > 1$
\item $\delta = \delta_o/(T + 1)$
\item Extend($k$, $p_t$, $\delta$)
\end{itemize}
\end{itemize}

Pieces of the analysis:
\begin{enumerate}
\item {\it One useful instantiation of $\B_x$.}

Suppose we define $\B_x$ to be all balls centered at $x$, and $\B_{x,p}$ to be all such balls with radius at most $r_p(x)$. In our earlier work we found that in this setting, in order to get good labels for $\X_{p,\gamma}$, 
$$ n \sim \frac{1}{p \gamma^2}$$
labels are necessary and sufficient. This is our point of reference.

\item {\it Invariants after round $t$.}

At the end of round $t$, every point in $\X_{p_t,\gamma}$ will either have been queried or will have had its label correctly inferred. What remains, at most, is then $\partial_{p_t,\gamma}$. The sampling region in the next round is then a subset of $R_{t+1} = \partial_{p_t, p_{t+1}, \gamma}$, where we define
$$ \partial_{p,q,\gamma} = \bigcup_{x \in \partial_{p,\gamma}} \bigcup_{B \in \B_{x,q}} B .$$

\item {\it Label complexity: high level insights.}

The total number of queries is upper-bounded by, roughly,
$$ \sum_{t=0}^T \mu(R_t) \frac{k_t}{p_t} .$$
If we keep $k$ fixed at $1/\gamma^2$ and use geometrically decreasing $p_t$, this is at most $1/(p_T \gamma^2)$, even if $\mu(R_t)$ remains stuck at 1. This is exactly the label complexity of passive learning for $\X_{p_T, \gamma}$. So the benefit of active sampling comes from the shrinking of $\mu(R_t)$.

\end{enumerate}

\subsection{Label complexity under smoothness assumptions}

\begin{enumerate}

\item {\it A further assumption on $\B_{x,p}$.}

In this section we will impose the further condition that $B \in \B_{x,p} \implies B \subseteq B(x, r_p(x))$. As a consequence,
$$ \partial_{p,q,\gamma} \subseteq \bigcup_{x \in \partial_{p,\gamma}} B(x, r_q(x)) .$$

\item {\it Smoothness and margin conditions.}

\begin{itemize}
\item Notion of margin: for any $t \in (0, 1/2)$, define the {\it $\tau$-width margin} to be
$$ M(\tau) = \{x \in \X: |\eta(x) - 1/2| \leq \tau\} .$$
\item Smoothness: we define a notion of $(\alpha,L)$-smoothness condition under which, for any $p,q > 0$,
$$ \bigcup_{x \in \partial_p} B(x, r_q(x)) \subseteq M(\gamma + L(p^\alpha + q^\alpha)) .$$
\item Tsybakov margin condition: there is $C > 0$ and $\beta > 0$ for which
$$ \mu(M(\tau)) \leq C \tau^\beta.$$
\end{itemize}

\item {\it Bounding the size of the sampling region.}

Under the smoothness condition, we have by Lemma~\ref{lemma:boundary-plus} that
$$ R_{t+1} \subseteq M(\gamma + L(p_t^\alpha + (p_t/c)^\alpha)) .$$
Define $L' = L(1 + c^{-\alpha})$. Under the margin condition, we then get
$$ \mu(R_{t+1}) 
\ \leq \ 
C (\gamma + L' p_t^\alpha)^\beta 
\ \leq \ 
2C \cdot \max \left(\frac{1}{k^{1/2}}, \frac{L'}{c^{\alpha t}} \right)^\beta.
$$
Define $t_o = (1/(2 \alpha)) \log_c k$. 
\begin{itemize}
\item When $t \leq t_o$, the dominant term is $(1/c^{\alpha \beta})^t$, which is exponentially shrinking, as we'd like.
\item When $t > t_o$, the dominant term becomes $1/k^{\beta/2}$, which is constant.
\end{itemize}
During the second regime, there is no benefit to active learning. So we should stop when $t = t_o$.

\item {\it Label complexity: more detail.}

Use $p_t = 1/2^t$ and stop after $T = (1/(2 \alpha)) \lg k$ rounds. Then the label complexity is, roughly,
$$ \sum_{t=0}^T \mu(R_t) \frac{k}{p_t} = k \sum_{t=0}^T \frac{2^t}{2^{\alpha\beta t}} = k \sum_{t = 0}^T \left( \frac{1}{2^{\alpha\beta - 1}}\right)^t .$$
Consider three cases for the geometric series $(1/2^{\alpha\beta - 1})^t$.
\begin{itemize}
\item $\alpha \beta > 1$. Then the geometric series has ratio $< 1$ and the label complexity is $k$.
\item $\alpha \beta = 1$. The geometric series has ratio 1 and the label complexity is $k T \sim k \lg k$.
\item $\alpha \beta < 1$. The geometric series has ratio $> 1$ and the label complexity is $k 2^{(1-\alpha\beta)T} = k^{(2\alpha + 1 - \alpha\beta)/(2 \alpha)}$.
\end{itemize}
This correctly classifies $\X_{p_T, \gamma}$, for which passive sampling would need roughly 
$$ \frac{1}{p_T^2 \gamma^2} = k2^T = k^{(2\alpha + 1)/(2 \alpha)}$$
samples. If we think of this as $n$, then we should set $k = n^{2\alpha/(2\alpha+1)}$.
Putting it all together, we get the following label complexity bounds (compared to $n$):
$$ 
\begin{array}{ll}
n^{2\alpha/(2\alpha+1)} & \mbox{if $\alpha \beta > 1$} \\
n^{2\alpha/(2\alpha+1)} \log n & \mbox{if $\alpha \beta = 1$} \\
n^{(2\alpha + 1 - \alpha\beta)/(2\alpha+1)} & \mbox{if $\alpha \beta < 1$}
\end{array}
$$

\end{enumerate}

\section{Deterministic characterizations}

The goal of this section is to characterize the conditional distribution of the labels given a point using quantities that control the random quantities define earlier. The general plan is to use these quantities in guarantees regarding the query complexity of the algorithm. Note that the sample space $\X$ is fixed and know. The unknown quaantity is $\eta(x)$

\begin{enumerate}

\item {\it Sampling regions.}

There is a set of ``balls'' $\B$, regions of different sizes that can be as small as a single point in $\X$. These are the atomic regions on which we assess label bias.

\item {\it bias of balls} For each $B \in \B$ we define the bias $B$ as 
$$ \eta(B) = \frac{1}{|B|} \sum_{x \in B}\eta(x) $$
\item {\it thresholding}
We define a provisional label for a ball:
$$\yh(B) =  
\begin{cases}
+ & \mbox{ if } \eta(B)>\gamma \\
? & \mbox{ if } |\eta(B)| \leq \gamma \\
- & \mbox{ if } \eta(B)<-\gamma 
\end{cases}
$$

\item {\it Probability radius.}

For any $x \in \X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$. 
In other words $\mu(B) = |B| / |\X|$

\item {\it Regions at a given probability scale.}

For any point $x \in \X$ and value $p > 0$, let $\B_{x,p}$ be a subset of $\B_x$ that is considered to be at ``probability-scale $p$''. For instance, we could define
$$ \B_{x,p} = \{B | x \in B, p\leq \mu(B) \leq 2p\}$$

\item {\it Provisional labels for data points.}

given a scale $p$ and a conditional probability threshold $\gamma$, we  assign $x \in \X$ a provisional label
$$ \yh_{p,\gamma}(x) = 
\left\{
\begin{array}{ll}
? & \mbox{if all $B \in \B_{x,p}$ have provisional label $\yh(B) = \ ?$} \\
+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
! & \mbox{if there is both a $+$ ball and a $-$ ball}
\end{array}
\right.
$$

\item{\it mind changes in $[0,1]$}
We fix $\gamma>0$ and $x$, $s(x,p) = \yh_{p,\gamma}(x)$ defines a maps $p \in [0,1]$ to $(!,-,+,?)$.
$s$ partitions the segment $[0,1]$ to sub-segments with constant labels. The segment that contains $p=0$ is the convergence, or, no-more-mind-changes segments. The segment above it is the final $!$ segment. My hypothesis is that our algorithm relies on the uniform sampling half for all segments but the last 2, and it uses the active region sampling for the final $!$ sampling until it reaches convergence. (it does not know that it reached convergence, therefor it has to continue uniform sampling.)

\item {\it The effective decision boundary.}
The true decision boundary is the set $\{x: \eta(x) = 1/2\}$, but this doesn't tell us which points we can reasonably hope to correctly label with a finite sample.

The effective decision boundary for a given level $\gamma$ and a radius $p$ is the subset of $x \in \X$ for which 
$p$ is not in the final convergence segment.

\end{enumerate}

\section{Another way to write the algorithm}
\label{sec:alg-2}

\subsection{Key ideas}

\begin{enumerate}

\item {\it Pool of unlabeled points.}

We begin with a pool $X$ of $n$ points drawn at random from an unknown underlying distribution $\mu$ on the instance space $\X$. We are allowed to query the label of any $x \in X$; this label follows the conditional probability distribution
$$ \pr(Y = 1|X = x) = \eta(x)$$
where $\eta$ is also unknown.

\item {\it Poisson sampling.}

We adopt a slightly unusual sampling process.
\begin{itemize}
\item Typical sampling scheme: Choose $m'$ points at random from a given sampling region $S$ of size $m$.
\item What we do: Pick each $x \in S$ with probability $m'/m$, independently.
\end{itemize}
This permits a cleaner implementation and analysis, as we now outline.

Every point $x \in X$ has its own random value $T_x$ chosen independently from the uniform distribution over $[0,1]$. If we merely wanted to query uniform-random points from $X$, we could achieve this by choosing points in increasing order of their $T_x$ values. This can be visualized as having a ``knob'' $\tau$ that begins at zero and is gradually increased; at any given time, the queried points are then precisely those with $T_x \leq \tau$.

For active learning, will have a combination of
\begin{itemize}
\item Background random sampling of the entire data set $X$.
\item Focused random sampling of subsets of $X$.
\end{itemize}
We will do this by having several knobs: a background sampling knob that works as described above, and other, higher-valued, knobs that apply to different subsets of $X$ where we want denser sampling.

\item {\it Sampling regions.}

There is a set of ``balls'' $\B$, regions of different sizes that can get arbitrarily small. These are the regions on which we assess label bias.

\begin{enumerate}

\item[(a)] {\it Random samples within a region.} The construction of $\B$ may or may not depend on the unlabeled points $X$, and this distinction determines which points can be considered {\it random draws} from a particular $B \in \B$.

For instance, consider these two alternatives for defining $\B$:
\begin{itemize}
\item $\B$ consists of a pre-defined set of regions.
\item $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
\end{itemize}
In the first case, all points $X \cap B$ are random draws from $\mu|_B$, the restriction of $\mu$ to region $B$. In the second case, this is also true except for the two points $x,x'$ that define the ball.

Let $X_B$ denote the points in $X$ that can be considered random draws from $\mu|_B$; this will typically be almost all of $B \cap X$.

\item[(b)] {\it Query-set for a region.} For each region $B$, we define its {\it query-set} as the $O(k)$ points in it with lowest $T_x$ values. More precisely, we have a threshold $\tau(B) \in [0,1]$  associated with $B$, and the query-set for $B$ consists of 
$$\{x \in X_B: T_x \leq \tau(B)\} .$$ 
The threshold $\tau(B)$ will be roughly $O(k/|X_B|)$.

\item[(c)] {\it Processed regions.} When the labels of $B$'s query-set have all been queried, we say {\it $B$ has been processed}. We then use the labels of these points to assign a label to the region as a whole:
$$ \yh(B) = 
\left\{
\begin{array}{ll}
+ & \mbox{significant $+$ bias} \\
- & \mbox{significant $-$ bias} \\
? & \mbox{otherwise}
\end{array}
\right.
$$

\item[(d)] {\it Regions at different sampling scales.} Regions $B$ that are large---in the sense of containing many points---have low thresholds $\tau(B)$ and are thus likely to get processed earlier, modulo differences in the density of focused sampling across $X$.

We will group regions $B$ according to intervals of $[0,1]$ in which their thresholds $\tau(B)$ lie. Define a sequence 
$$ 0 < \tau_0 < \tau_1 < \ldots < \tau_{\ell_o} = 1, $$
for some positive integer $\ell_o$. Each $0 \leq \ell \leq \ell_o$ denotes a ``level of sampling''. A region $B$ belongs to the smallest level $\ell$ such that $\tau(B) \leq \tau_\ell$; let $\B_\ell$ be the collection of all such regions.

Thus $\B_0, \B_1, \ldots$ is a partition of $\B$ by sampling level, with $\B_0$ consisting of very large regions and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller regions that may not need to be considered if they are far from the decision boundary.

\end{enumerate}

\item {\it Balls used to evaluate a given point.}

For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that may be used in determining $x$'s label. This could be, for instance,
\begin{itemize}
\item all balls containing $x$, or
\item all balls centered at $x$.
\end{itemize}
As before, we can partition these regions by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.

\begin{enumerate}
\item[(a)] {\it Provisional label at a given level.} The provisional label of $x$ at a given sampling level $\ell$ is a function of the labels $\yh(B)$ of regions $B \in \B_\ell(x)$. It is defined as follows:
\begin{equation}
\yh_\ell(x) = 
\left\{
\begin{array}{ll}
? & \mbox{if all $B \in \B_\ell(x)$ have $\yh(B) = \ ?$} \\
+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
! & \mbox{if there is both a $+$ ball and a $-$ ball}
\end{array}
\right.
\label{eq:provisional-label}
\end{equation}
Points labeled $!$ are ``known unknowns'': they are near the decision boundary and need further investigation. Points labeled $?$ are also uncertain, and will be treated as such.

\item[(b)] {\it Point processed at level $\ell$.} When $\yh_\ell(x)$ becomes known, we say ``$x$ has been processed at level $\ell$''. In the case $\yh_\ell(x) = \ !$, this could happen before some of the individual regions $B \in \B_\ell(x)$ have been processed.

\item[(c)] Let $L(x) \in \{0,1,\ldots,\ell_o+1\}$ be the first level at which $x$ has {\it not} yet been processed. Initially, all $L(x) = 0$.

At any given time, the density of sampling will vary across $X$. It might be high at points known to be near the decision boundary, low near points believed to be far from the boundary, and intermediate at points which were once believed to be far from the boundary but have recently be found to possibly be close to the boundary.

\end{enumerate}

\item {\it Uncertainty region.} At any given time, the focused sampling takes place in the neighborhood of an ``uncertainty region'' $U \subset X$. We can partition this region by level: $U_\ell = \{x \in U: L(x) = \ell\}$.

Initially, $U = X$.

\end{enumerate}

\subsection{One round of sampling}

The algorithm proceeds in rounds, each with an equal amount of {\it background} sampling and {\it focused} sampling. At the beginning of a round:
\begin{itemize}
\item $Q \subseteq X$ is the set of points queried so far.

\item $U \subseteq X \setminus Q$ is the ``uncertain'' region. We can partition it by level: $U_\ell = \{x \in U: L(x) = \ell\}$.
\end{itemize}
The focused sampling for the round occurs around the {\it lowest-numbered non-empty uncertainty set} $U_\ell$.

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.4in}
{\bf function One-Round}

\vspace{.1in}
Global variables: 
\begin{itemize}
\item $Q$ - points queried so far.
\item $U$ - uncertain region. Can be partitioned by level: $U_\ell = \{x \in U: L(x) = \ell\}$
\end{itemize}

\vspace{.1in}
{\it Focused sampling.}
\begin{itemize}
\item Find the smallest $0 \leq \ell \leq \ell_o$ such that $U_\ell \neq \emptyset$. If there is no such $\ell$, move to the next step.
\item Query region:
$$ S =  \bigcup_{x \in U_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau(B)\} $$
\item Query the next point in $S \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\vspace{.1in}
{\it Background sampling.}
\begin{itemize}
\item Query the next point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\vspace{.1in}
{\it Processing points and regions.}
\begin{itemize}
\item For any region $B$ that is now processed: compute $\yh(B)$
\item For any point $x$ that is now processed at level $L(x)$:
\begin{itemize}
\item Determine $\yh_\ell(x)$ using (\ref{eq:provisional-label})
\item If $\yh_\ell(x) \in \{+,-\}$: set $U = U \setminus \{x\}$
\item If $\yh_\ell(x) \in \{!,?\}$: set $U = U \cup \{x\}$
\item $L(x) = L(x) + 1$
\end{itemize}
\end{itemize}

\end{minipage}}
\caption{One round of active learning.}
\label{fig:one-round}
\end{figure}

\subsection{Accuracy of bias estimates}

Our algorithm will rely upon a uniform bound on the bias estimates of all $B \in \B$, under Poisson sampling of the data.

Recall that the pool of data points $X$ is sampled from an underlying distribution $\mu$, and that each such point $x$ is labeled according to the conditional probability $\pr(Y = 1|X = x) = \eta(x)$. Let $\eta(B)$ be the average $\eta$-value in $B$, that is,
$$ \eta(B) = \frac{1}{\mu(B)} \int_B \eta(x) \ d\mu .$$
The specific property we need is encapsulated in the following definition.
\begin{defn}
For any region $B \in \B$ and $\Delta > 0$, we say bias estimate $\yh(B) \in \{+,-,?\}$ is {\it $\Delta$-accurate} if the following holds:
\begin{itemize}
\item If $\eta(B) \geq 1/2 + 2 \Delta$, then $\yh(B) = +$.
\item If $\yh(B) = +$ then $\eta(B) > 1/2$.
\item If $\eta(B) \leq 1/2 - 2 \Delta$, then $\yh(B) = -$.
\item If $\yh(B) = -$ then $\eta(B) < 1/2$.
\end{itemize}
\label{def:accurate-bias-estimate}
\end{defn}

For any region $B \in \B$, let $X_B$ be the points in $X$ that are random draws from $\mu|_B$. The number of such points, $n(B) = |X_B|$, determines the ``level'' in which $B$ lies. We now make this precise.

In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
\begin{equation}
c_1 = \sqrt{2 \ln \frac{2|\B|}{\delta}} .
\label{eq:c1}
\end{equation}

\begin{defn}
For any region $B \in \B$, set 
\begin{equation}
\tau(B) = \frac{k + c_1 \sqrt{2k}}{n(B)} .
\label{eq:ball-threshold}
\end{equation}
We define the {\it query-set} for $B$ to be 
$$ \{x \in X_B: T_x \leq \tau(B)\}.$$
\label{def:query-set}
\end{defn}

Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-single-ball} in the appendix.
\begin{thm}
With probability at least $1-\delta$, the following is true for all $B \in \B$:
\begin{enumerate}
\item[(a)] The query-set for $B$ has size at least $k$.
\item[(b)] The average label on the query-set, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta(B) \right| \leq \frac{c_1}{2\sqrt{k}}.$$
\end{enumerate}
\label{thm:large-deviation-bounds}
\end{thm}
\begin{proof}
These follow immediately from Lemma~\ref{lemma:large-deviation-single-ball}. Note that $X_B$ consists of $n(B)$ random draws from $\mu|_B$; picking those with $T_x \leq \tau(B)$ is tantamount to picking each of these points with probability $\tau(B)$.
\end{proof}

\begin{cor}
For $c_1$ as in (\ref{eq:c1}), define
$$ \Delta = \frac{c_1}{2 \sqrt{k}} .$$
For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set, as above, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
+ & \mbox{if $\widehat{\eta}(B) > 1/2 + \Delta$} \\
- & \mbox{if $\widehat{\eta}(B) < 1/2 - \Delta$} \\
? & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\Delta$-accurate.
\label{cor:accurate-bias-estimates}
\end{cor}
In what follows, we will assume that this $(1-\delta)$-probability event holds.

\subsection{Algorithmic guarantees}

Pick any $x \in \X$ with $\eta(x) > 1/2$. We say $x$ has {\it significant bias at level $\ell$} if the following is true:
\begin{itemize}
\item $\eta(B) > 1/2$ for all $B \in \B_\ell(x)$
\item $\eta(B) \geq 1/2 + 2\Delta$ for some $B \in \B_\ell(x)$
\end{itemize}
We say $x$ is {\it resolved at level $\ell$} if it has significant bias at level $\ell$ and moreover $\eta(B) > 1/2$ for all $B \in \B_{\ell'}(x), \ell' > \ell$.

We can make a similar definition for $x$ with $\eta(x) < 1/2$.

\begin{thm}
Suppose all bias estimates $\yh(B)$ are $\Delta$-accurate as defined above. Suppose $x$ has $\eta(x) > 1/2$ and is resolved at level $\ell$. Then:
\begin{enumerate}
\item[(a)] $\yh_\ell(x) = +$ and 
\item[(b)] for any level $\ell' > \ell$, we have $\yh_{\ell'}(x) \in \{+, ?\}$ and $x \not\in U_{\ell'}$.
\end{enumerate}
Likewise if $\eta(x) < 1/2$.
\end{thm}

\begin{proof}
Suppose $\eta(x) > 1/2$ and $x$ is resolved at level $\ell$.

For any $\ell' \geq \ell$ and $B \in \B_{\ell'}(x)$, we have $\eta(B) > 1/2$ and thus $\yh(B) \in \{+,?\}$. Moreover, since $x$ has significant bias at level $\ell$, there is some $B_o \in \B_\ell(x)$ with $\eta(B_o) > 1/2 + 2\Delta$ and thus $\yh(B_o) = +$.

These imply that $\yh_\ell(x) = +$ and $\yh_{\ell'}(x) \in \{+,?\}$ for all $\ell' > \ell$. Thus, if $x \in U_\ell$, it will be removed from $U_\ell$ as soon as it is processed at level $\ell$. It will never be added to $U$ thereafter.
\end{proof}

\pagebreak
\appendix

\section{Technicalities}

\subsection{Large deviation bounds}

We need large deviation bounds that will guarantee two properties: (i) regions of interest receive at least $k$ samples, and (ii) the estimated label bias within these regions is close to the true bias.

For both results, we have an instance space $\X$ with underlying distribution $\mu$, and label space $\Y = \{-1,+1\}$ with conditional probability distribution $\eta$.
The first bound is comes from [BBL05].
\begin{lemma}
Let $\B$ consist of finitely many subsets of $\X$. Pick any $0 < \delta < 1$. Suppose $n$ points $S$ are drawn i.i.d.\ from a distribution $\mu$ on $\X$. Then with probability at least $1-\delta$ over the random sample, for all $B \in \B$, we have
$$ \mu(B) \geq \frac{k}{n} + \frac{5 \sqrt{k \ln (4|\B|/\delta)}}{n} \ \implies \ |S \cap B| \geq k,$$
provided $k \geq \ln (4|\B|/\delta)$.
\label{lemma:points-in-balls-finite}
\end{lemma}
\begin{proof}
From [BBL05, Thm 5.1], we have that with probability $> 1-\delta$, for all $B \in \B$, 
$$ \frac{|S \cap B|}{n} \geq \mu(B) - 2 \sqrt{\frac{\mu(B) \ln (4|\B|/\delta)}{n}} .$$ 
If $k \geq \ln (4|\B|/\delta)$ and $\mu(B)$ is written in the form
$$ \mu(B) = \frac{k + \alpha\sqrt{k \ln (4|\B|/\delta)}}{n} $$
for some $\alpha > 0$, then by algebraic manipulation we get $|S \cap B| \geq k$ as long as $\alpha \geq 5$.
\end{proof}

If $\B$ is a class of subsets of $\X$ of VC dimension $d_o$, then the same bound holds with $|\B|$ replaced by $d_o \ln n$.

\iffalse
Here is a weaker bound that is a bit cleaner.
\begin{lemma}[CD10, Lemma 7]
There is a universal constant $c_1$ such that the following holds. Let $\B$ be a class of measurable subsets of $\X$ of VC dimension $d_o$. Pick any $0 < \delta < 1$. Then with probability at least $1-\delta$, if a set of $n$ points $S$ is drawn i.i.d.\ from $\mu$, then for any integer $k$,
$$ \mu(B) \geq \frac{k}{n} + \frac{c_1}{n} \max \left( k, d_o \log \frac{n}{\delta} \right)
\ \implies \ 
|S \cap B| \geq k.$$ 
\label{lemma:points-in-balls-vc}
\end{lemma}
\fi

The next lemma, bounding empirical biases, is from [BDFM19].
\begin{lemma}[BDFM19, Lemma 7]
There is a universal constant $c_2$ for which the following holds. Let $\mathcal{C}$ be a class of subsets of $\X$ with VC dimension $d_o$. Suppose $n$ points $(x_1, y_1), \ldots, (x_n, y_n)$ are drawn i.i.d.\ from the underlying distribution on $\X \times \Y$ given by $\mu$ and $\eta$, and are used to determine empirical biases
$$ \eta_n(C) = \frac{|\{i \in [n]: x_i \in C, y_i = 1\}|}{n}.$$
Pick any $0 < \delta < 1$. Then with probability at least $1-\delta$ over the sample, for all $C \in \mathcal{C}$,
$$ \left| \eta_n(C) - \eta(C) \right| \ \leq \ \Delta(n, \#_n(C), \delta) $$
where $\#_n(C) = |\{i: x_i \in C\}|$ is the number of points in $C$ and 
\begin{equation}
\Delta(n,k,\delta) = c_2 \sqrt{\frac{d_o \log n + \log (1/\delta)}{k}} .
\label{eq:delta-defn}
\end{equation}
\label{lemma:bias}
\end{lemma}

\subsection{Smoothness condition}

The work of [CD14] introduced a nearest neighbor-inspired variant of the Holder condition on $\eta$: for some constants $\alpha, L > 0$, for all $x \in \supp(\mu)$ and all $r > 0$,
$$ |\eta(B(x,r)) - \eta(x)| \leq L \mu(B^o(x,r))^\alpha .$$
It was shown in [CD14, Lemma 18] that under this condition, for any $p, \gamma$,
$$ \partial_{p,\gamma} \cap \supp(\mu) \subset M(\gamma + Lp^\alpha) ,$$
where $M(\tau)$ is a shorthand for $\{x \in \X: |\eta(x) - 1/2| \leq \tau\}$.

In this paper, we consider a stronger condition.
\begin{defn}
We say $\eta$ is $(\alpha, L)$-smooth if for all $x \in \supp(\mu)$ and all $p > 0$,
$$ |\eta(x') - \eta(x)| \leq L p^\alpha \ \ \mbox{for all $x' \in B(x,r_p(x))$}.$$
\end{defn}

\begin{lemma}
Suppose $\eta$ is $(\alpha, L)$-smooth. Then for any $p,q > 0$,
$$ \bigcup_{x \in \partial_{p, \gamma}} B(x, r_q(x)) \ \subset \ M(\gamma + L(p^\alpha + q^\alpha)).$$
\label{lemma:boundary-plus}
\end{lemma}
\begin{proof}
Pick any $x \in \partial_{p, \gamma}$ and $x' \in B(x,r_q(x))$. Then
$$
|\eta(x') - 1/2| 
\ \leq \ 
|\eta(x) - 1/2| + |\eta(x') - \eta(x)|
\ \leq \ 
(\gamma + Lp^\alpha) + Lq^\alpha,
$$
using the smoothness of $\eta$ along with Lemma 18 of [CD14].
\end{proof}

A more common assumption is that $\X = \R^d$ and $\eta$ is $\alpha_H$-Holder continuous for some $\alpha_H > 0$, that is,
$$ |\eta(x) - \eta(x')| \leq C \|x-x'\|^{\alpha_H}$$
for some $C > 0$. Typically such analysis also requires $\mu$ to be bounded below. Here we relate this to our assumption.
\begin{lemma}
Suppose that $\X \subseteq \R^d$, that $\eta$ is $\alpha_H$-Holder continuous, and that for some $\mu_{\scriptsize\rm min} > 0$ it is the case that 
$$\mu(B(x,r)) \geq \min(\mu_{\scriptsize\rm min} r^d,1)$$ 
for all $x \in \X$ and $r > 0$. Then there is a constant $L > 0$ such that $\eta$ is $(\alpha_H/d, L)$-smooth.
\end{lemma}
\begin{proof}
Suppose $\eta$ satisfies the $\alpha_H$-Holder condition: for some $C > 0$,
$$ |\eta(x) - \eta(x')| \leq C \|x-x'\|^{\alpha_H}.$$
Now pick any $x \in \X$, any $p > 0$, and a point $x' \in B(x, r_p(x))$. Then
$$ |\eta(x) - \eta(x')| 
\leq C \|x-x'\|^{\alpha_H}
\leq C \left( \frac{\mu(B(x, \|x-x'\|))}{\mu_{\scriptsize\rm min}} \right)^{\alpha_H/d}
\leq C \left( \frac{p}{\mu_{\scriptsize\rm min}} \right)^{\alpha_H/d}.
$$
The result follows by taking $L = C \mu_{\scriptsize\rm min}^{-\alpha_H/d}$.
\end{proof}

\end{document}

### 2024-07-31 old/active-knn.tex
Initial Overleaf Import

--- 2024-07-31 old/active-knn-old1.tex
@@
 
-\section{Setup and goals}
+\section{Setup}
 
-Basic setup:
+Basic Transductive setup:
 \begin{itemize}
-\item Instance space $\X \subset \R^d$. Data are generated from a distribution $\mu$ on $\X$.
-\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
+\item Intance set $\X =\{x_1,\dots,x_n\}$, $x_i in \R^d$. The set $\X$
+  is fixed and known to the learner. The probability of each $x \in
+  \R$ is $1/n$. 
+\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is associated with
+  a conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
 \item Let $g^*$ denote the Bayes-optimal classifier on $\X$:
   $$ g^*(x) = \mbox{sign}(2 \eta(x) - 1) .$$
-\end{itemize}
-
-\noindent
-Goal: \emph{pool-based active learning}
-\begin{itemize}
-\item We have a finite dataset $X \subset \R^d$ of $n$ points, and we are able to query the labels of any of them.
-\item Given parameters $0 < \epsilon, \delta < 1$, our goal is to query the labels of some points $Q \subset X$, and infer the labels of the remainder, so that with probability at least $1-\delta$, at most $\epsilon n$ of the inferred labels disagree with $g^*$.
-\end{itemize}
-
-\noindent
-Alternative goal: \emph{stream-based active learning}
-\begin{itemize}
-\item Data arrive continuously, one point at a time. For each point, we decide whether or not to query its label.
-\item The classifier $g_t$ at time $t$ is (some kind of) $k$-nearest neighbor classifier based on the points labeled so far.
-We require $g_t \rightarrow g^*$ and we consider the rate of convergence in terms of the number of labeled points.
+  \item Label Queries: the algorithm can select any example $x \in
+    \X$, as a response it gets a label in $\Y$ chosesn acccording to
+    $\eta(x)$.    
   \end{itemize}
@@
 \section{Notation}
-
+\begin{itemize}
+  \item {\bf Balls:} $B(x,r)$ denotes a ball of radius $r$ centered at $x$.
+  \item {\bf Probability of a ball} The probability of a ball is the
+    fraction of points in $\X$ that fall inside $B(x,r)$.
+    $$ \mu(B(x,r)) = \frac{ |B(x,r) \cap \X|}{n} $$
+  \item {\bf measuring radius in terms of probability}
 For any $x \in X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$. 
+\end{itemize}
 
@@
 
-\begin{center}
-\begin{huge}
-Older material
-\end{huge}
-\end{center}
-
-\pagebreak
-
-\section{Setup}
-
-Basic Transductive setup:
-\begin{itemize}
-\item Intance set $\X =\{x_1,\dots,x_n\}$, $x_i \in {\bf X}$, where
-  ${\bf X}$ is a metric space. . The set
-  $\X$ is fixed and known to the learner. The probability of each $x
-  \in \R$ is $1/n$.
-\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is associated with
-  a conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
-\item Let $g^*$ denote the Bayes-optimal classifier on $\X$:
-  $$ g^*(x) = \mbox{sign}(2 \eta(x) - 1) .$$
-  \item Label Queries: the algorithm can select any example $x \in
-    \X$, as a response it gets a label in $\Y$ chosen acccording to
-    $\eta(x)$.    
-  \end{itemize}
-
-\section{Notation}
-\begin{itemize}
-\item {\bf Balls:} For any $x \in \X$ an $r\geq0$,
-  $B(x,r)$ denotes the subset of $\X$ inside ball of radius $r$ centered at $x$.
-  \item {\bf Probability of a ball} The probability of a ball is the
-    fraction of points in $\X$ that fall inside $B(x,r)$.
-    $$ \mu(B(x,r)) = \frac{ |B(x,r)|}{n}$$
-  \item {\bf conditional probability given a ball:}
-    $$ \eta(B(x,r)) = \frac{\sum_{x \in B(x,r)}\eta(z)}{|B(x,r)|}$$
-  \item {\bf measuring radius in terms of probability}
-For any $x \in X$ and any $p > 0$, let $r_p(x)$ denote the smallest
-radius $r$ such that $\mu(B(x, r)) \geq p$.
-\end{itemize}
-
-\newcommand{\NNB}{{\bf NNB}}
-\newcommand{\ANNB}{{\bf ANNB}}
-\newcommand{\ONNB}{{\bf oracle-NNB}}
-
-\section{The near-neighborhoods algorithm}
-The near neighborhood (\NNB) algorithm is a variant of AKNN where the balls
-that are considered for the prediction of a point $x$ are {\em all} balls
-the contain $x$ while in AKNN only balls that are centered at $x$ are
-considered.
-
-Suppose that we are given $n$ labeled examples chosen uniformly at
-random from $\X$ and labeled according to $\eta(x)$.
-
-Given that set of labels each ball is as $-,?,+$ according to the
-condition used in AKNN. We say that a ball labeled $\pm$ is {\em
-  determined}, a ball labeled $?$ is {\em undetermined}.
-
-We define the partial order on balls that is defined by containment.
-
-The labels of the balls induce a labeling on the points in $\X$. Each
-point has one of {\em four} labels: $-1,+1,?,!$. Given a point $x \in
-\X$ we define the {\em minimal set of determined balls for $x$},
-denoted $\B(x)$, to be the set of determined balls that contain $x$
-that is minimal (in terms of the containment partial order). We can
-construct this set by starting with the set of all determined balls containing
-$x$ and eliminating any ball that has a subset-ball that is also in
-the set.
-
-Given the set of balls $\B(x)$ the label associated with $x$
-according to the following rules:
-\begin{itemize}
-\item If $\B(x)$ is empty the label is undetermined, or $?$.
-\item If all of the balls in $\B(x)$ are labeled $+$, the label is $+1$.
-\item If all of the balls in $\B(x)$ are labeled $-$, the label is $-1$
-\item If $\B(x)$ contains both labels, then the label is {\em
-    over-determined} or $!$. We will also refer to over-determined
-  examples as the ``known unknown''.
-\end{itemize}
-
-We define the numerical prediction $\hat{y}$ to be $0$  if the label
-is undetermined or overdetermined. We set it equal to the label otherwise
-
-We define the error of the prediction as the produce
-$-\hat{y}{y}$. Note that the error can be $-1,0,+1$.
-
-\NNB\ is a passive learning algorithm, in other words the queriy
-points are selected IID from $\X$.
-
-\begin{claim}
-  The algorithm \NNB\ converges to Bayes rule as the number of
-  labeled examples increases to $\infty$.
-  \end{claim}
-
-\section{Enforcing locality}
-For technical reasons, we consider balls that are smaller and smaller as the algorithm progresses. Let
-$1=p_1 >p_2>\cdots >p_k=1/n$ be the radius probabilities. Then in iteration $i$ we restrict the set of balls to be such that $\mu(B(x,r)) \leq p_i$.
-
-\section{Active \NNB}
-
-The main idea behind active \NNB\, or \ANNB\, is to concentrate the
-queries on the known unknown. However, this is not done to the
-exclusion of the other parts of the space, as any example can have an
-unbounded number of mind changes. In other words, an example whose label at some point is $+1$
-can be labeled $-1$ later on and vice versa. (These are the ``unknown
-unknowns'').
-
-\ANNB\ operates in phases. At the start of phase $t$ a concentration
-area $C_t \subset \X$ is computed. Then, $2n$ queries are are
-made. There are $n$ {\em passive} queries and $n$ {\em active} queries.
-The passive queries are selected uniformly at random from $\X$, the
-active quries are selected uniformly at random from $C_t$.
-
-After the answers to the queries are recieved, they are used to update
-the counter $k_+,k_-$ in the balls. Passive points are used in each
-ball that contains them. Active points are a bit more subtle, they are
-used only in balls that are conained in $C_t$. In other words if a
-ball contains an element $x$ that is {\em not} in $C_t$ then this ball
-cannot use {\em any} of the active samples for epoch $t$.
-
-Based on the additional queries, the label $\pm +1,?$ for each ball is
-updated, and based on that the label of each $x \in \X$ is updated.
-
-Finally, $C_{t+1}$ is calculated as follows. Let $U_t$ be the set $x
-\in \X$ such that the label of $x$ is overdetermined, or ``!''.  $C_{t+1}$
-is defined as the union of the minimal sets of determined balls for $x
-\in U_t$, In other words:
-\[
-  C_{t+1} = \bigcup \left\{\B(x,r) | x \in U_t, \mu(\B(x,r)) \leq p_t \right\}
-  \]
-
-  \section{A comparator algorithm}
-  In order to bound the query complexity of \ANNB\ we consider a
-  variant of \ANNB\ that is aided by and oracle. Intuitively, the
-  oracle removes the need for a passive sample by informing the
-  algorithm that the prediction on a point $x$ is ``final''. I.e. it
-  will not change as the number of queries increases to infinity.
-
-  For this algorithm, which we denote by \ONNB\, we can more easily
-  give bounds. As this framework is local, our goal is to provide
-  point-wise bounds.
-
-  \ONNB\ differs from \ANNB\ in two important ways
-  \begin{enumerate}
-  \item It makes only active queries, not passive queries.
-  \item The set $U_t$ contains, in addition to overdetermined
-    points. The points where the label is $+1$ or $-1$ but the oracle
-    informs the algorithm that there will be mind cahnges on $x$ in
-    the future. 
-  \end{enumerate}
-
-  Intuitively, \ONNB\ contines to collect information on $x$ until the
-  label of $x$ converges, at which point it does not need more
-  samples for $\B(x)$ so it stop collecting this information.
-
-  \section{Pointwise analysis}
-
-  Fix a point $x \in \X$ and compare the behaviour of \NNB\ and \ONNB\
-  on $x$.
-
-  Let the support set for $x$ be th set of queries that influence the predicted label for $x$.
-In other words removing the examples not in the support does not effect change the way the algorithm unfolds wrt that point.
-
-  {\bf Can we prove that } The size of the support for any example using \ONNB\ is at least half of the size of the support using \ANNB\
-
-\pagebreak
-
-\section{Setup}
-
-\begin{enumerate}
-\item Setting: {\it Binary classification.}
-
-\begin{itemize}
-\item Instance space $\X \subset \R^d$. Data points $x$ are generated from a distribution $\mu$ on $\X$.
-\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$.
-\item The Bayes-optimal classifier on $\X$ is $g^*(x) = \mbox{sign}(2 \eta(x) - 1)$.
-\end{itemize}
-
-\item Goal: {\it Pool-based active learning.}
-
-\begin{itemize}
-\item We have a finite dataset $X \subset \R^d$ of $n$ points, and we are able to query the labels of any of them.
-\item Given parameters $0 < \epsilon, \delta < 1$, our goal is to query the labels of some of the points, and infer the labels of the remainder, so that with probability at least $1-\delta$, at most $\epsilon n$ of the inferred labels disagree with the Bayes-optimal $g^*$.
-\end{itemize}
-
-\end{enumerate}
-
-\section{An active learning algorithm}
-
-\begin{enumerate}
-
-\item {\it Algorithm overview.}
-
-There are multiple rounds of labeling, and during each, query points are chosen at random from the current ``uncertainty region''.
-
-\item {\it Sampling balls.}
-
-There is a set of ``balls'' $\B$, balls of different sizes that can get arbitrarily small. These are the balls on which we assess label bias.
-
-We query points that are chosen at random, under two types of sampling. 
-\begin{itemize}
-\item Background random sampling of the entire data set $X$.
-\item Focused random sampling of a subset of $X$.
-\end{itemize}
-For each $B \in \B$, we know which of the labels seen so far can be considered random samples from $B$. On that basis, we assign a {\it provisional label}
-$$ \yh(B) = 
-\left\{
-\begin{array}{ll}
-+ & \mbox{significant $+$ bias} \\
-- & \mbox{significant $-$ bias} \\
-? & \mbox{otherwise}
-\end{array}
-\right.
-$$
-
-\item {\it $\Delta$-accurate bias estimates.}
-
-The bias estimator that we use returns $\{+,-,?\}$ for any ball $B$ and has the following property that holds (with high probability) uniformly over all balls $B$ under consideration:
-\begin{itemize}
-\item If $\eta(B) \geq 1/2 + 2 \Delta$, it returns $+$.
-\item If $\eta(B) \leq 1/2$, it does not return $+$.
-\item Likewise with $-$.
-\end{itemize}
-Roughly speaking, this requires having $1/\Delta^2$ random labeled points in each ball.
-
-Over time, as more random samples appear in any given ball $B$, it is theoretically possible that this provisional label would change from $+$ to $-$ or vice versa.
-In such cases, our large deviation bounds allow us to retain either label.
-
-\item {\it Probability radius.}
-
-For any $x \in \X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$.
-
-\item {\it Balls used to evaluate a given point.}
-
-For any $x \in \X$, let $\B_x \subset \B$ denote the collection of balls that may be used in determining $x$'s label. This could be, for instance,
-\begin{itemize}
-\item[(a)] all balls containing $x$, or
-\item[(b)] all balls centered at $x$.
-\end{itemize}
-
-\item {\it Regions at a given probability scale.}
-
-For any point $x \in \X$ and value $p > 0$, let $\B_{x,p}$ be a subset of $\B_x$ that is considered to be at ``probability-scale $p$''. For instance, we could define
-$$ \B_{x,p} = \{B \in \B_x: B \subseteq B(x,r_p(x))\},$$
-but we leave it flexible for now.
-
-\noindent
-{\it Assumption:} For any $x \in \X$ and $p > 0$, there exists some $B \in \B_{x,p}$ with $\mu(B) \geq c_o p$. Here $c_o > 0$ is an absolute constant.
-
-\item {\it Provisional labels for data points.}
-
-At any given time, at scale $p$, we will assign $x$ a provisional label
-$$ \yh_p(x) = 
-\left\{
-\begin{array}{ll}
-? & \mbox{if all $B \in \B_{x,p}$ have provisional label $\yh(B) = \ ?$} \\
-+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
-- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
-! & \mbox{if there is both a $+$ ball and a $-$ ball}
-\end{array}
-\right.
-$$
-Intuitively, the points we will deem uncertain are those that are labeled $!$ or $?$.
-
-\item {\it The effective decision boundary.}
-
-The true decision boundary is the set $\{x: \eta(x) = 1/2\}$, but this doesn't tell us which points we can reasonably hope to correctly label with a finite sample.
-
-For any $p, \gamma > 0$, let $\X_{p, \gamma}^+$ consist of all points $x \in \supp(\mu)$ such that:
-\begin{itemize}
-\item $\eta(x) \geq 1/2 + \gamma$
-\item $\eta(B) \geq 1/2$ for all $B \in \B_{x,p}$
-\item $\eta(B) \geq 1/2 + \gamma$ for some $B \in \B_{x,p}$ with $\mu(B) \geq c_o p$
-\end{itemize}
-We can define $\X_{p,\gamma}^-$ similarly, and $\X_{p,\gamma} = \X_{p,\gamma}^+ \cup \X_{p,\gamma}^-$. The {\it effective boundary} then consists of all remaining points,
-$$ \partial_{p,\gamma} = \X \setminus \X_{p,\gamma} .$$
-
-\end{enumerate}
-
-{\bf Yoav:} Might be possible to prove an upper bound on the number of queries if the probability of the boundary $\partial_{p,\gamma}$ is small.
-
-\subsection{One round of sampling}
-
-The algorithm proceeds in rounds, each with an equal amount of {\it background} sampling and {\it focused} sampling. At the beginning of round $t$:
-\begin{itemize}
-\item $Q_t \subseteq X$ is the set of points queried so far.
-
-Each point $x \in Q_t$ has accompanying information that tells us, for any $B \in \B$, whether $x$ can be considered a random sample from $B$.
-
-\item $I_t \subseteq X$ is the set of points whose labels were inferred.
-
-\item $U_t \subseteq X$ is the set of points deemed ``uncertain'' according to some criterion.
-\end{itemize}
-Based on these, the focused sampling region for round $t$ is
-$$ S_t = \bigcup_{x \in U_t} \bigcup_{B \in \B_{x, p_t}} B .$$
-Here $(p_t)$ is a predefined sequence with $p_t \downarrow 0$.
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{6.3in}
-{\bf function Extend}($k$, $p$, $\delta$):
-
-\vspace{.1in}
-Global variables: $Q, I, U$.
-
-\vspace{.1in}
-Setup:
-\begin{itemize}
-\item Sampling region:
-$$ S =  \left( \bigcup_{x \in U} \bigcup_{B \in \B_{x,p}} B \right) \cap X $$
-\item Choose number of samples:
-  $$ m = \frac{|S|}{|X|} \cdot \frac{k + c_1 \sqrt{k}}{c_o p} $$
-\end{itemize}
-
-\vspace{.1in}
-Querying:
-\begin{itemize}
-\item Let $M$ consist of $m$ points chosen uniformly at random from $S$
-\item Query the labels of all $x \in M \setminus Q$ and add these points to $Q$
-\end{itemize}
-
-\vspace{.1in}
-Infer labels using $M$:
-\begin{itemize}
-\item For all $B \in \{B \in \B_{x,p}: x \in U\}$: Use $M \cap B$ to set $\yh(B)$
-\item For all $x \in U \setminus Q$:
-\begin{itemize}
-\item Use $\{\yh(B): B \in \B_{x,p}\}$ to set $\yh_{p}(x)$
-\item If $\yh_{p}(x) \in \{+,-\}$ and $\mbox{\tt oracle-check}(x,p)$: Add $x$ to $I$ with the given label
-\end{itemize}
-\end{itemize}
-
-\vspace{.1in}
-Update uncertainty region: 
-\begin{itemize}
-\item $U = U \setminus (Q \cup I)$
-\end{itemize}
-
-\end{minipage}}
-\caption{Subroutine for active learning given {\tt oracle-check}.}
-\label{fig:extend}
-\end{figure}
-
-\subsection{Large deviation bounds}
-
-We will use large deviation bounds for the class $\B$ of regions. 
-\begin{lemma}
-Choose any $0 < \delta < 1$ and define $C_\delta = \ln (4 |\B|/\delta)$.
-There are universal constants $c',c''>0$ for which the following is true.
-If $m$ points are drawn i.i.d.
-from $\mu|_S$, the restriction of $\mu$ to a subset $S \subset \X$, and labeled according to $\eta$, then with probability at least $1-\delta$, the following properties hold for all $B \in \B$:
-\begin{enumerate}
-\item[(a)] If $k \geq C_\delta$ and $\mu|_S(B) \geq (k + c' \sqrt{k C_\delta})/m$, then $B$ gets at least $k$ points.
-\item[(b)] If $B \subseteq S$ gets at least $k$ points, then the empirical bias of the labels in $B$ is equal to $\eta(B) \pm c'' \sqrt{C_\delta/k}$.
-\end{enumerate}
-\label{lemma:bias-estimates}
-\end{lemma}
-\begin{proof}
-These properties follow immediately from Lemmas~\ref{lemma:points-in-balls-finite} and \ref{lemma:bias}.
-\end{proof}
-
-Let $G_t$ denote the good event that these conditions hold for the points drawn in round $t$ of sampling, and let $G_{\leq t} = G_1 \cap G_2 \cap \cdots \cap G_t$.
-
-In what follows, we will take $\delta > 0$ to be a constant and will assume that $k \geq C_\delta$. We also define $c_1 = c' \sqrt{C_\delta}$ and $c_2 = c'' \sqrt{C_\delta}$. Finally, let
-$$ \Delta_k = \frac{c_2}{k^{1/2}}.$$
-
-Our bias-estimates are computed as follows.
-On any round of the algorithm, let $S \subseteq \X$ denote the sampling region and $M$ the randomly chosen samples in this region.
-For any region $B \subseteq S$ under consideration,
-$$ \yh(B) = 
-\left\{
-\begin{array}{ll}
-+ & \mbox{if $|B \cap M| \geq k$ and empirical bias $\geq 1/2 + \Delta_k$} \\
-- & \mbox{if $|B \cap M| \geq k$ and empirical bias $\leq 1/2 - \Delta_k$} \\
-? & \mbox{otherwise}
-\end{array}
-\right.
-$$
-The key property we will need for these estimates is that they are $\Delta_k$-accurate in the following sense.
-\begin{defn}
-Pick any $\Delta > 0$. We say bias-estimates $\yh(B) \in \{+,-,?\}$ are $\Delta$-accurate if they satisfy the following properties:
-\begin{itemize}
-\item If $\eta(B) \geq 1/2 + 2\Delta$, then $\yh(B) = +$.
-\item If $\eta(B) \leq 1/2 - 2\Delta$, then $\yh(B) = -$.
-\item If $\yh(B) = +$ then $\eta(B) > 1/2$.
-\item If $\yh(B) = -$ then $\eta(B) < 1/2$.
-\end{itemize}
-\label{defn:delta-accurate}
-\end{defn}
-
-\begin{lemma}
-Consider round $t$ of sampling, with uncertainty set $U$ and parameters $k,p$.
-Suppose event $G_t$ holds.
-Then for all $x \in U$, the bias-estimates $\{\yh(B): B \in \B_{x,p}\}$ are $\Delta_k$-accurate.
-\label{lemma:accurate-bias}
-\end{lemma}
-\begin{proof}
-This follows immediately from Lemma~\ref{lemma:bias-estimates}.
-\end{proof}
-
-\begin{thm}
-Suppose that in round $t$, the starting uncertainty region is $U$. Assume event $G_{\leq t}$ occurs. Then for all $x \in U$ that aren't queried, the following two properties hold.
-\begin{enumerate}
-\item[(a)] If $x \in \X_{p,2\Delta_k}$, then by the end of the round, its label is inferred to be $\yh_{p}(x) = g^*(x)$.
-\item[(b)] If $\yh_{p}(x) \in \{+,-\}$ but is not equal to $g^*(x)$, then $x \not\in \X_{p,0}$.
-\end{enumerate}
-\end{thm}
-\begin{proof}
-Pick any $x \in U$ whose label isn't queried by the end of round $t$. Without loss of generality $\eta(x) \geq 1/2$.
-
-For part (a), if $x \in \X_{p,2\Delta_k}$, then every $B \in \B_{x,p}$ has $\eta(B) \geq 1/2$ and thus, by Lemma~\ref{lemma:accurate-bias}, gets $\yh(B) \in \{+,?\}$ in any round of sampling.
-Moreover, by assumption there is some $B' \in \B_{x,p}$ for which $\mu(B') \geq c_o p$ and $\eta(B') \geq 1/2 + 2\Delta_k$.
-Letting $S$ denote the sampling region in round $t$ and ignoring the small discrepancy between $\mu(S)$ and $|S|/|X|$,
-$$ \mu|_S(B') = \frac{\mu(B')}{\mu(S)} \geq \frac{c_o p}{\mu(S)} \geq \frac{k + c_1 \sqrt{k}}{m}$$
-and thus by Lemma~\ref{lemma:bias-estimates}(a) we have $|M \cap B'| \geq k$. Thus $\yh(B)$ would be assigned $+$ in round $t$, if not earlier, whereupon $\yh_p(x) = +$.
-
-For part (b), suppose $\yh_{p}(x) = -$. Then some $B \in \B_{x,p}$ has $\yh(B) = -$, implying by Lemma~\ref{lemma:accurate-bias} that $\eta(B) < 1/2$. This can only be happen if $x \not\in \X_{p,0}$. 
-\end{proof}
-
-The Algorithm in Figure~\ref{fig:extend} makes use of an external subroutine called {\tt oracle-check}.
-This has the following semantics: $\mbox{\tt oracle-check}(x,p)$ returns {\tt true} if $x \in \X_{p,\gamma}$ and {\tt false} otherwise.
-
-\begin{cor}
-\item Pick any round $t$, and say parameters $k,p$ are used in that round.
-Suppose that $G_{\leq t}$ holds and that $\gamma \leq 2\Delta_k$.
-Let $U$ denote the uncertainty set at the start of the round.
-During round $t$, all points in $U \cap \X_{p,2\Delta_k}$ will be placed in $I$.
-Moreover, assuming the correctness of the {\tt oracle-check} subroutine, any point placed in $I$ will be labeled correctly.
-\end{cor}
-
-\subsection{An active learning loop (using the oracle)}
-
-Here's a simple idea: keep $k$ fixed, with the goal of trying to correctly classify all $x \in \X_{p,\gamma}$ with $p > 1/n$ and $\gamma = \Delta_k$.
-
-\begin{itemize}
-\item Given: $X$, $\gamma$, $\delta$
-\item Choose $k$ so that $\Delta_k = \gamma$ (i.e., $k = 4c_2^2/\gamma^2$)
-\item $Q = I = \emptyset$
-\item for $t = 0, 1, \ldots, T$:
-\begin{itemize}
-\item $p_t = 1/c^t$ for some $c > 1$
-\item $\delta = \delta_o/(T + 1)$
-\item Extend($k$, $p_t$, $\delta$)
-\end{itemize}
-\end{itemize}
-
-Pieces of the analysis:
-\begin{enumerate}
-\item {\it One useful instantiation of $\B_x$.}
-
-Suppose we define $\B_x$ to be all balls centered at $x$, and $\B_{x,p}$ to be all such balls with radius at most $r_p(x)$.
-In our earlier work we found that in this setting, in order to get good labels for $\X_{p,\gamma}$,
-$$ n \sim \frac{1}{p \gamma^2}$$
-labels are necessary and sufficient. This is our point of reference.
-
-\item {\it Invariants after round $t$.}
-
-At the end of round $t$, every point in $\X_{p_t,\gamma}$ will either have been queried or will have had its label correctly inferred.
-What remains, at most, is then $\partial_{p_t,\gamma}$.
-The sampling region in the next round is then a subset of $R_{t+1} = \partial_{p_t, p_{t+1}, \gamma}$, where we define
-$$ \partial_{p,q,\gamma} = \bigcup_{x \in \partial_{p,\gamma}} \bigcup_{B \in \B_{x,q}} B .$$
-
-\item {\it Label complexity: high level insights.}
-
-The total number of queries is upper-bounded by, roughly,
-$$ \sum_{t=0}^T \mu(R_t) \frac{k_t}{p_t} .$$
-If we keep $k$ fixed at $1/\gamma^2$ and use geometrically decreasing $p_t$, this is at most $1/(p_T \gamma^2)$, even if $\mu(R_t)$ remains stuck at 1.
-This is exactly the label complexity of passive learning for $\X_{p_T, \gamma}$.
-So the benefit of active sampling comes from the shrinking of $\mu(R_t)$.
-
-\end{enumerate}
-
-\subsection{Label complexity under smoothness assumptions}
-
-\begin{enumerate}
-
-\item {\it A further assumption on $\B_{x,p}$.}
-
-In this section we will impose the further condition that $B \in \B_{x,p} \implies B \subseteq B(x, r_p(x))$. As a consequence,
-$$ \partial_{p,q,\gamma} \subseteq \bigcup_{x \in \partial_{p,\gamma}} B(x, r_q(x)) .$$
-
-\item {\it Smoothness and margin conditions.}
-
-\begin{itemize}
-\item Notion of margin: for any $t \in (0, 1/2)$, define the {\it $\tau$-width margin} to be
-$$ M(\tau) = \{x \in \X: |\eta(x) - 1/2| \leq \tau\} .$$
-\item Smoothness: we define a notion of $(\alpha,L)$-smoothness condition under which, for any $p,q > 0$,
-$$ \bigcup_{x \in \partial_p} B(x, r_q(x)) \subseteq M(\gamma + L(p^\alpha + q^\alpha)) .$$
-\item Tsybakov margin condition: there is $C > 0$ and $\beta > 0$ for which
-$$ \mu(M(\tau)) \leq C \tau^\beta.$$
-\end{itemize}
-
-\item {\it Bounding the size of the sampling region.}
-
-Under the smoothness condition, we have by Lemma~\ref{lemma:boundary-plus} that
-$$ R_{t+1} \subseteq M(\gamma + L(p_t^\alpha + (p_t/c)^\alpha)) .$$
-Define $L' = L(1 + c^{-\alpha})$. Under the margin condition, we then get
-$$ \mu(R_{t+1}) 
-\ \leq \ 
-C (\gamma + L' p_t^\alpha)^\beta 
-\ \leq \ 
-2C \cdot \max \left(\frac{1}{k^{1/2}}, \frac{L'}{c^{\alpha t}} \right)^\beta.
-$$
-Define $t_o = (1/(2 \alpha)) \log_c k$. 
-\begin{itemize}
-\item When $t \leq t_o$, the dominant term is $(1/c^{\alpha \beta})^t$, which is exponentially shrinking, as we'd like.
-\item When $t > t_o$, the dominant term becomes $1/k^{\beta/2}$, which is constant.
-\end{itemize}
-During the second regime, there is no benefit to active learning. So we should stop when $t = t_o$.
-
-\item {\it Label complexity: more detail.}
-
-Use $p_t = 1/2^t$ and stop after $T = (1/(2 \alpha)) \lg k$ rounds. Then the label complexity is, roughly,
-$$ \sum_{t=0}^T \mu(R_t) \frac{k}{p_t} = k \sum_{t=0}^T \frac{2^t}{2^{\alpha\beta t}} = k \sum_{t = 0}^T \left( \frac{1}{2^{\alpha\beta - 1}}\right)^t .$$
-Consider three cases for the geometric series $(1/2^{\alpha\beta - 1})^t$.
-\begin{itemize}
-\item $\alpha \beta > 1$. Then the geometric series has ratio $< 1$ and the label complexity is $k$.
-\item $\alpha \beta = 1$. The geometric series has ratio 1 and the label complexity is $k T \sim k \lg k$.
-\item $\alpha \beta < 1$. The geometric series has ratio $> 1$ and the label complexity is $k 2^{(1-\alpha\beta)T} = k^{(2\alpha + 1 - \alpha\beta)/(2 \alpha)}$.
-\end{itemize}
-This correctly classifies $\X_{p_T, \gamma}$, for which passive sampling would need roughly 
-$$ \frac{1}{p_T^2 \gamma^2} = k2^T = k^{(2\alpha + 1)/(2 \alpha)}$$
-samples. If we think of this as $n$, then we should set $k = n^{2\alpha/(2\alpha+1)}$.
-Putting it all together, we get the following label complexity bounds (compared to $n$):
-$$ 
-\begin{array}{ll}
-n^{2\alpha/(2\alpha+1)} & \mbox{if $\alpha \beta > 1$} \\
-n^{2\alpha/(2\alpha+1)} \log n & \mbox{if $\alpha \beta = 1$} \\
-n^{(2\alpha + 1 - \alpha\beta)/(2\alpha+1)} & \mbox{if $\alpha \beta < 1$}
-\end{array}
-$$
-
-\end{enumerate}
-
-\section{Deterministic characterizations}
-
-The goal of this section is to characterize the conditional distribution of the labels given a point using quantities that control the random quantities define earlier.
-The general plan is to use these quantities in guarantees regarding the query complexity of the algorithm.
-Note that the sample space $\X$ is fixed and know.
-The unknown quaantity is $\eta(x)$
-
-\begin{enumerate}
-
-\item {\it Sampling regions.}
-
-There is a set of ``balls'' $\B$, regions of different sizes that can be as small as a single point in $\X$. These are the atomic regions on which we assess label bias.
-
-\item {\it bias of balls} For each $B \in \B$ we define the bias $B$ as 
-$$ \eta(B) = \frac{1}{|B|} \sum_{x \in B}\eta(x) $$
-\item {\it thresholding}
-We define a provisional label for a ball:
-$$\yh(B) =  
-\begin{cases}
-+ & \mbox{ if } \eta(B)>\gamma \\
-? & \mbox{ if } |\eta(B)| \leq \gamma \\
-- & \mbox{ if } \eta(B)<-\gamma 
-\end{cases}
-$$
-
-\item {\it Probability radius.}
-
-For any $x \in \X$ and any $p > 0$, let $r_p(x)$ denote the smallest radius $r$ such that $\mu(B(x, r)) \geq p$. 
-In other words $\mu(B) = |B| / |\X|$
-
-\item {\it Regions at a given probability scale.}
-
-For any point $x \in \X$ and value $p > 0$, let $\B_{x,p}$ be a subset of $\B_x$ that is considered to be at ``probability-scale $p$''. For instance, we could define
-$$ \B_{x,p} = \{B | x \in B, p\leq \mu(B) \leq 2p\}$$
-
-\item {\it Provisional labels for data points.}
-
-given a scale $p$ and a conditional probability threshold $\gamma$, we  assign $x \in \X$ a provisional label
-$$ \yh_{p,\gamma}(x) = 
-\left\{
-\begin{array}{ll}
-? & \mbox{if all $B \in \B_{x,p}$ have provisional label $\yh(B) = \ ?$} \\
-+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
-- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
-! & \mbox{if there is both a $+$ ball and a $-$ ball}
-\end{array}
-\right.
-$$
-
-\item{\it mind changes in $[0,1]$}
-We fix $\gamma>0$ and $x$, $s(x,p) = \yh_{p,\gamma}(x)$ defines a maps $p \in [0,1]$ to $(!,-,+,?)$.
-$s$ partitions the segment $[0,1]$ to sub-segments with constant labels.
-The segment that contains $p=0$ is the convergence, or, no-more-mind-changes segments.
-The segment above it is the final $!$ segment.
-My hypothesis is that our algorithm relies on the uniform sampling half for all segments but the last 2, and it uses the active region sampling for the final $!$ sampling until it reaches convergence.
-(it does not know that it reached convergence, therefor it has to continue uniform sampling.)
-
-\item {\it The effective decision boundary.}
-The true decision boundary is the set $\{x: \eta(x) = 1/2\}$, but this doesn't tell us which points we can reasonably hope to correctly label with a finite sample.
-
-The effective decision boundary for a given level $\gamma$ and a radius $p$ is the subset of $x \in \X$ for which 
-$p$ is not in the final convergence segment.
-
-\end{enumerate}
-
-\section{Another way to write the algorithm}
-\label{sec:alg-2}
-
-\subsection{Key ideas}
-
-\begin{enumerate}
-
-\item {\it Pool of unlabeled points.}
-
-We begin with a pool $X$ of $n$ points drawn at random from an unknown underlying distribution $\mu$ on the instance space $\X$.
-We are allowed to query the label of any $x \in X$; this label follows the conditional probability distribution
-$$ \pr(Y = 1|X = x) = \eta(x)$$
-where $\eta$ is also unknown.
-
-\item {\it Poisson sampling.}
-
-We adopt a slightly unusual sampling process.
-\begin{itemize}
-\item Typical sampling scheme: Choose $m'$ points at random from a given sampling region $S$ of size $m$.
-\item What we do: Pick each $x \in S$ with probability $m'/m$, independently.
-\end{itemize}
-This permits a cleaner implementation and analysis, as we now outline.
-
-Every point $x \in X$ has its own random value $T_x$ chosen independently from the uniform distribution over $[0,1]$.
-If we merely wanted to query uniform-random points from $X$, we could achieve this by choosing points in increasing order of their $T_x$ values.
-This can be visualized as having a ``knob'' $\tau$ that begins at zero and is gradually increased; at any given time, the queried points are then precisely those with $T_x \leq \tau$.
-
-For active learning, will have a combination of
-\begin{itemize}
-\item Background random sampling of the entire data set $X$.
-\item Focused random sampling of subsets of $X$.
-\end{itemize}
-We will do this by having several knobs: a background sampling knob that works as described above, and other, higher-valued, knobs that apply to different subsets of $X$ where we want denser sampling.
-
-\item {\it Sampling regions.}
-
-There is a set of ``balls'' $\B$, regions of different sizes that can get arbitrarily small. These are the regions on which we assess label bias.
-
-\begin{enumerate}
-
-\item[(a)] {\it Random samples within a region.} The construction of $\B$ may or may not depend on the unlabeled points $X$, and this distinction determines which points can be considered {\it random draws} from a particular $B \in \B$.
-
-For instance, consider these two alternatives for defining $\B$:
-\begin{itemize}
-\item $\B$ consists of a pre-defined set of regions.
-\item $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
-\end{itemize}
-In the first case, all points $X \cap B$ are random draws from $\mu|_B$, the restriction of $\mu$ to region $B$.
-In the second case, this is also true except for the two points $x,x'$ that define the ball.
-
-Let $X_B$ denote the points in $X$ that can be considered random draws from $\mu|_B$; this will typically be almost all of $B \cap X$.
-
-\item[(b)] {\it Query-set for a region.} For each region $B$, we define its {\it query-set} as the $O(k)$ points in it with lowest $T_x$ values.
-More precisely, we have a threshold $\tau(B) \in [0,1]$  associated with $B$, and the query-set for $B$ consists of
-$$\{x \in X_B: T_x \leq \tau(B)\} .$$ 
-The threshold $\tau(B)$ will be roughly $O(k/|X_B|)$.
-
-\item[(c)] {\it Processed regions.} When the labels of $B$'s query-set have all been queried, we say {\it $B$ has been processed}.
-We then use the labels of these points to assign a label to the region as a whole:
-$$ \yh(B) = 
-\left\{
-\begin{array}{ll}
-+ & \mbox{significant $+$ bias} \\
-- & \mbox{significant $-$ bias} \\
-? & \mbox{otherwise}
-\end{array}
-\right.
-$$
-
-\item[(d)] {\it Regions at different sampling scales.} Regions $B$ that are large---in the sense of containing many points---have low thresholds $\tau(B)$ and are thus likely to get processed earlier, modulo differences in the density of focused sampling across $X$.
-
-We will group regions $B$ according to intervals of $[0,1]$ in which their thresholds $\tau(B)$ lie. Define a sequence 
-$$ 0 < \tau_0 < \tau_1 < \ldots < \tau_{\ell_o} = 1, $$
-for some positive integer $\ell_o$.
-Each $0 \leq \ell \leq \ell_o$ denotes a ``level of sampling''.
-A region $B$ belongs to the smallest level $\ell$ such that $\tau(B) \leq \tau_\ell$; let $\B_\ell$ be the collection of all such regions.
-
-Thus $\B_0, \B_1, \ldots$ is a partition of $\B$ by sampling level, with $\B_0$ consisting of very large regions and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller regions that may not need to be considered if they are far from the decision boundary.
-
-\end{enumerate}
-
-\item {\it Balls used to evaluate a given point.}
-
-For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that may be used in determining $x$'s label. This could be, for instance,
-\begin{itemize}
-\item all balls containing $x$, or
-\item all balls centered at $x$.
-\end{itemize}
-As before, we can partition these regions by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.
-
-\begin{enumerate}
-\item[(a)] {\it Provisional label at a given level.} The provisional label of $x$ at a given sampling level $\ell$ is a function of the labels $\yh(B)$ of regions $B \in \B_\ell(x)$.
-It is defined as follows:
-\begin{equation}
-\yh_\ell(x) = 
-\left\{
-\begin{array}{ll}
-? & \mbox{if all $B \in \B_\ell(x)$ have $\yh(B) = \ ?$} \\
-+ & \mbox{if there is a $+$ ball and the rest are $+$ or $?$} \\ 
-- & \mbox{if there is a $-$ ball and the rest are $-$ or $?$} \\ 
-! & \mbox{if there is both a $+$ ball and a $-$ ball}
-\end{array}
-\right.
-\label{eq:provisional-label}
-\end{equation}
-Points labeled $!$ are ``known unknowns'': they are near the decision boundary and need further investigation. Points labeled $?$ are also uncertain, and will be treated as such.
-
-\item[(b)] {\it Point processed at level $\ell$.} When $\yh_\ell(x)$ becomes known, we say ``$x$ has been processed at level $\ell$''.
-In the case $\yh_\ell(x) = \ !$, this could happen before some of the individual regions $B \in \B_\ell(x)$ have been processed.
-
-\item[(c)] Let $L(x) \in \{0,1,\ldots,\ell_o+1\}$ be the first level at which $x$ has {\it not} yet been processed. Initially, all $L(x) = 0$.
-
-At any given time, the density of sampling will vary across $X$.
-It might be high at points known to be near the decision boundary, low near points believed to be far from the boundary, and intermediate at points which were once believed to be far from the boundary but have recently be found to possibly be close to the boundary.
-
-\end{enumerate}
-
-\item {\it Uncertainty region.} At any given time, the focused sampling takes place in the neighborhood of an ``uncertainty region'' $U \subset X$.
-We can partition this region by level: $U_\ell = \{x \in U: L(x) = \ell\}$.
-
-Initially, $U = X$.
-
-\end{enumerate}
-
-\subsection{One round of sampling}
-
-The algorithm proceeds in rounds, each with an equal amount of {\it background} sampling and {\it focused} sampling. At the beginning of a round:
-\begin{itemize}
-\item $Q \subseteq X$ is the set of points queried so far.
-
-\item $U \subseteq X \setminus Q$ is the ``uncertain'' region. We can partition it by level: $U_\ell = \{x \in U: L(x) = \ell\}$.
-\end{itemize}
-The focused sampling for the round occurs around the {\it lowest-numbered non-empty uncertainty set} $U_\ell$.
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{6.4in}
-{\bf function One-Round}
-
-\vspace{.1in}
-Global variables: 
-\begin{itemize}
-\item $Q$ - points queried so far.
-\item $U$ - uncertain region. Can be partitioned by level: $U_\ell = \{x \in U: L(x) = \ell\}$
-\end{itemize}
-
-\vspace{.1in}
-{\it Focused sampling.}
-\begin{itemize}
-\item Find the smallest $0 \leq \ell \leq \ell_o$ such that $U_\ell \neq \emptyset$. If there is no such $\ell$, move to the next step.
-\item Query region:
-$$ S =  \bigcup_{x \in U_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau(B)\} $$
-\item Query the next point in $S \setminus Q$, ordered by $T_z$ values, and add to $Q$
-\end{itemize}
-
-\vspace{.1in}
-{\it Background sampling.}
-\begin{itemize}
-\item Query the next point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
-\end{itemize}
-
-\vspace{.1in}
-{\it Processing points and regions.}
-\begin{itemize}
-\item For any region $B$ that is now processed: compute $\yh(B)$
-\item For any point $x$ that is now processed at level $L(x)$:
-\begin{itemize}
-\item Determine $\yh_\ell(x)$ using (\ref{eq:provisional-label})
-\item If $\yh_\ell(x) \in \{+,-\}$: set $U = U \setminus \{x\}$
-\item If $\yh_\ell(x) \in \{!,?\}$: set $U = U \cup \{x\}$
-\item $L(x) = L(x) + 1$
-\end{itemize}
-\end{itemize}
-
-\end{minipage}}
-\caption{One round of active learning.}
-\label{fig:one-round}
-\end{figure}
-
-\subsection{Accuracy of bias estimates}
-
-Our algorithm will rely upon a uniform bound on the bias estimates of all $B \in \B$, under Poisson sampling of the data.
-
-Recall that the pool of data points $X$ is sampled from an underlying distribution $\mu$, and that each such point $x$ is labeled according to the conditional probability $\pr(Y = 1|X = x) = \eta(x)$.
-Let $\eta(B)$ be the average $\eta$-value in $B$, that is,
-$$ \eta(B) = \frac{1}{\mu(B)} \int_B \eta(x) \ d\mu .$$
-The specific property we need is encapsulated in the following definition.
-\begin{defn}
-For any region $B \in \B$ and $\Delta > 0$, we say bias estimate $\yh(B) \in \{+,-,?\}$ is {\it $\Delta$-accurate} if the following holds:
-\begin{itemize}
-\item If $\eta(B) \geq 1/2 + 2 \Delta$, then $\yh(B) = +$.
-\item If $\yh(B) = +$ then $\eta(B) > 1/2$.
-\item If $\eta(B) \leq 1/2 - 2 \Delta$, then $\yh(B) = -$.
-\item If $\yh(B) = -$ then $\eta(B) < 1/2$.
-\end{itemize}
-\label{def:accurate-bias-estimate}
-\end{defn}
-
-For any region $B \in \B$, let $X_B$ be the points in $X$ that are random draws from $\mu|_B$.
-The number of such points, $n(B) = |X_B|$, determines the ``level'' in which $B$ lies.
-We now make this precise.
-
-In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
-\begin{equation}
-c_1 = \sqrt{2 \ln \frac{2|\B|}{\delta}} .
-\label{eq:c1}
-\end{equation}
-
-\begin{defn}
-For any region $B \in \B$, set 
-\begin{equation}
-\tau(B) = \frac{k + c_1 \sqrt{2k}}{n(B)} .
-\label{eq:ball-threshold}
-\end{equation}
-We define the {\it query-set} for $B$ to be 
-$$ \{x \in X_B: T_x \leq \tau(B)\}.$$
-\label{def:query-set}
-\end{defn}
-
-Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-single-ball} in the appendix.
-\begin{thm}
-With probability at least $1-\delta$, the following is true for all $B \in \B$:
-\begin{enumerate}
-\item[(a)] The query-set for $B$ has size at least $k$.
-\item[(b)] The average label on the query-set, call it $\widehat{\eta}(B)$, satisfies
-$$ \left| \widehat{\eta}(B) - \eta(B) \right| \leq \frac{c_1}{2\sqrt{k}}.$$
-\end{enumerate}
-\label{thm:large-deviation-bounds}
-\end{thm}
-\begin{proof}
-These follow immediately from Lemma~\ref{lemma:large-deviation-single-ball}.
-Note that $X_B$ consists of $n(B)$ random draws from $\mu|_B$; picking those with $T_x \leq \tau(B)$ is tantamount to picking each of these points with probability $\tau(B)$.
-\end{proof}
-
-\begin{cor}
-For $c_1$ as in (\ref{eq:c1}), define
-$$ \Delta = \frac{c_1}{2 \sqrt{k}} .$$
-For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set, as above, and define the bias-estimate $\yh(B)$ as follows:
-$$ \yh(B)
-= 
-\left\{
-\begin{array}{ll}
-+ & \mbox{if $\widehat{\eta}(B) > 1/2 + \Delta$} \\
-- & \mbox{if $\widehat{\eta}(B) < 1/2 - \Delta$} \\
-? & \mbox{otherwise}
-\end{array}
-\right.
-$$
-With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\Delta$-accurate.
-\label{cor:accurate-bias-estimates}
-\end{cor}
-In what follows, we will assume that this $(1-\delta)$-probability event holds.
-
-\subsection{Algorithmic guarantees}
-
-Pick any $x \in \X$ with $\eta(x) > 1/2$. We say $x$ has {\it significant bias at level $\ell$} if the following is true:
-\begin{itemize}
-\item $\eta(B) > 1/2$ for all $B \in \B_\ell(x)$
-\item $\eta(B) \geq 1/2 + 2\Delta$ for some $B \in \B_\ell(x)$
-\end{itemize}
-We say $x$ is {\it resolved at level $\ell$} if it has significant bias at level $\ell$ and moreover $\eta(B) > 1/2$ for all $B \in \B_{\ell'}(x), \ell' > \ell$.
-
-We can make a similar definition for $x$ with $\eta(x) < 1/2$.
-
-\begin{thm}
-Suppose all bias estimates $\yh(B)$ are $\Delta$-accurate as defined above. Suppose $x$ has $\eta(x) > 1/2$ and is resolved at level $\ell$. Then:
-\begin{enumerate}
-\item[(a)] $\yh_\ell(x) = +$ and 
-\item[(b)] for any level $\ell' > \ell$, we have $\yh_{\ell'}(x) \in \{+, ?\}$ and $x \not\in U_{\ell'}$.
-\end{enumerate}
-Likewise if $\eta(x) < 1/2$.
-\end{thm}
-
-\begin{proof}
-Suppose $\eta(x) > 1/2$ and $x$ is resolved at level $\ell$.
-
-For any $\ell' \geq \ell$ and $B \in \B_{\ell'}(x)$, we have $\eta(B) > 1/2$ and thus $\yh(B) \in \{+,?\}$.
-Moreover, since $x$ has significant bias at level $\ell$, there is some $B_o \in \B_\ell(x)$ with $\eta(B_o) > 1/2 + 2\Delta$ and thus $\yh(B_o) = +$.
-
-These imply that $\yh_\ell(x) = +$ and $\yh_{\ell'}(x) \in \{+,?\}$ for all $\ell' > \ell$.
-Thus, if $x \in U_\ell$, it will be removed from $U_\ell$ as soon as it is processed at level $\ell$.
-It will never be added to $U$ thereafter.
-\end{proof}
-
-\pagebreak
-\appendix
-
-\section{Technicalities}
-
-\subsection{Large deviation bounds}
-
-We need large deviation bounds that will guarantee two properties: (i) regions of interest receive at least $k$ samples, and (ii) the estimated label bias within these regions is close to the true bias.
-
-For both results, we have an instance space $\X$ with underlying distribution $\mu$, and label space $\Y = \{-1,+1\}$ with conditional probability distribution $\eta$.
-The first bound is comes from [BBL05].
-\begin{lemma}
-Let $\B$ consist of finitely many subsets of $\X$.
-Pick any $0 < \delta < 1$.
-Suppose $n$ points $S$ are drawn i.i.d.\ from a distribution $\mu$ on $\X$.
-Then with probability at least $1-\delta$ over the random sample, for all $B \in \B$, we have
-$$ \mu(B) \geq \frac{k}{n} + \frac{5 \sqrt{k \ln (4|\B|/\delta)}}{n} \ \implies \ |S \cap B| \geq k,$$
-provided $k \geq \ln (4|\B|/\delta)$.
-\label{lemma:points-in-balls-finite}
-\end{lemma}
-\begin{proof}
-From [BBL05, Thm 5.1], we have that with probability $> 1-\delta$, for all $B \in \B$, 
-$$ \frac{|S \cap B|}{n} \geq \mu(B) - 2 \sqrt{\frac{\mu(B) \ln (4|\B|/\delta)}{n}} .$$ 
-If $k \geq \ln (4|\B|/\delta)$ and $\mu(B)$ is written in the form
-$$ \mu(B) = \frac{k + \alpha\sqrt{k \ln (4|\B|/\delta)}}{n} $$
-for some $\alpha > 0$, then by algebraic manipulation we get $|S \cap B| \geq k$ as long as $\alpha \geq 5$.
-\end{proof}
-
-If $\B$ is a class of subsets of $\X$ of VC dimension $d_o$, then the same bound holds with $|\B|$ replaced by $d_o \ln n$.
-
-\iffalse
-Here is a weaker bound that is a bit cleaner.
-\begin{lemma}[CD10, Lemma 7]
-There is a universal constant $c_1$ such that the following holds.
-Let $\B$ be a class of measurable subsets of $\X$ of VC dimension $d_o$.
-Pick any $0 < \delta < 1$.
-Then with probability at least $1-\delta$, if a set of $n$ points $S$ is drawn i.i.d.\ from $\mu$, then for any integer $k$,
-$$ \mu(B) \geq \frac{k}{n} + \frac{c_1}{n} \max \left( k, d_o \log \frac{n}{\delta} \right)
-\ \implies \ 
-|S \cap B| \geq k.$$ 
-\label{lemma:points-in-balls-vc}
-\end{lemma}
-\fi
-
-The next lemma, bounding empirical biases, is from [BDFM19].
-\begin{lemma}[BDFM19, Lemma 7]
-There is a universal constant $c_2$ for which the following holds.
-Let $\mathcal{C}$ be a class of subsets of $\X$ with VC dimension $d_o$.
-Suppose $n$ points $(x_1, y_1), \ldots, (x_n, y_n)$ are drawn i.i.d.\ from the underlying distribution on $\X \times \Y$ given by $\mu$ and $\eta$, and are used to determine empirical biases
-$$ \eta_n(C) = \frac{|\{i \in [n]: x_i \in C, y_i = 1\}|}{n}.$$
-Pick any $0 < \delta < 1$. Then with probability at least $1-\delta$ over the sample, for all $C \in \mathcal{C}$,
-$$ \left| \eta_n(C) - \eta(C) \right| \ \leq \ \Delta(n, \#_n(C), \delta) $$
-where $\#_n(C) = |\{i: x_i \in C\}|$ is the number of points in $C$ and 
-\begin{equation}
-\Delta(n,k,\delta) = c_2 \sqrt{\frac{d_o \log n + \log (1/\delta)}{k}} .
-\label{eq:delta-defn}
-\end{equation}
-\label{lemma:bias}
-\end{lemma}
-
-\subsection{Smoothness condition}
-
-The work of [CD14] introduced a nearest neighbor-inspired variant of the Holder condition on $\eta$: for some constants $\alpha, L > 0$, for all $x \in \supp(\mu)$ and all $r > 0$,
-$$ |\eta(B(x,r)) - \eta(x)| \leq L \mu(B^o(x,r))^\alpha .$$
-It was shown in [CD14, Lemma 18] that under this condition, for any $p, \gamma$,
-$$ \partial_{p,\gamma} \cap \supp(\mu) \subset M(\gamma + Lp^\alpha) ,$$
-where $M(\tau)$ is a shorthand for $\{x \in \X: |\eta(x) - 1/2| \leq \tau\}$.
-
-In this paper, we consider a stronger condition.
-\begin{defn}
-We say $\eta$ is $(\alpha, L)$-smooth if for all $x \in \supp(\mu)$ and all $p > 0$,
-$$ |\eta(x') - \eta(x)| \leq L p^\alpha \ \ \mbox{for all $x' \in B(x,r_p(x))$}.$$
-\end{defn}
-
-\begin{lemma}
-Suppose $\eta$ is $(\alpha, L)$-smooth. Then for any $p,q > 0$,
-$$ \bigcup_{x \in \partial_{p, \gamma}} B(x, r_q(x)) \ \subset \ M(\gamma + L(p^\alpha + q^\alpha)).$$
-\label{lemma:boundary-plus}
-\end{lemma}
-\begin{proof}
-Pick any $x \in \partial_{p, \gamma}$ and $x' \in B(x,r_q(x))$. Then
-$$
-|\eta(x') - 1/2| 
-\ \leq \ 
-|\eta(x) - 1/2| + |\eta(x') - \eta(x)|
-\ \leq \ 
-(\gamma + Lp^\alpha) + Lq^\alpha,
-$$
-using the smoothness of $\eta$ along with Lemma 18 of [CD14].
-\end{proof}
-
-A more common assumption is that $\X = \R^d$ and $\eta$ is $\alpha_H$-Holder continuous for some $\alpha_H > 0$, that is,
-$$ |\eta(x) - \eta(x')| \leq C \|x-x'\|^{\alpha_H}$$
-for some $C > 0$. Typically such analysis also requires $\mu$ to be bounded below. Here we relate this to our assumption.
-\begin{lemma}
-Suppose that $\X \subseteq \R^d$, that $\eta$ is $\alpha_H$-Holder continuous, and that for some $\mu_{\scriptsize\rm min} > 0$ it is the case that 
-$$\mu(B(x,r)) \geq \min(\mu_{\scriptsize\rm min} r^d,1)$$ 
-for all $x \in \X$ and $r > 0$. Then there is a constant $L > 0$ such that $\eta$ is $(\alpha_H/d, L)$-smooth.
-\end{lemma}
-\begin{proof}
-Suppose $\eta$ satisfies the $\alpha_H$-Holder condition: for some $C > 0$,
-$$ |\eta(x) - \eta(x')| \leq C \|x-x'\|^{\alpha_H}.$$
-Now pick any $x \in \X$, any $p > 0$, and a point $x' \in B(x, r_p(x))$. Then
-$$ |\eta(x) - \eta(x')| 
-\leq C \|x-x'\|^{\alpha_H}
-\leq C \left( \frac{\mu(B(x, \|x-x'\|))}{\mu_{\scriptsize\rm min}} \right)^{\alpha_H/d}
-\leq C \left( \frac{p}{\mu_{\scriptsize\rm min}} \right)^{\alpha_H/d}.
-$$
-The result follows by taking $L = C \mu_{\scriptsize\rm min}^{-\alpha_H/d}$.
-\end{proof}
-
 \end{document}

### 2024-07-31 old/active-neighborhood-oct.tex
Initial Overleaf Import

--- 2022-12-28 older stuff/main.tex
@@
 \def\PL{{\rm PL}}
-\def\sign{{\rm sign}}
 
@@
 
-\title{Active learning using the near-neighborhoods algorithm}
+\title{Active learning using the near-Neighborhoods algorithm}
 
@@
 \maketitle
-
 \section{Setup}
 
-\begin{itemize}
-\item {\it Underlying model of data.}
+\begin{enumerate}
+\item {\it Pool-based active learning.}
 
 \begin{itemize}
-\item Instances lie in a metric space $(\X, d)$. 
-\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to an unknown conditional probability function $\Pr(Y|X)$, with
-$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x) = 2 \pr(Y=1|X=x) - 1 .$$
-\item The {\it Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
+\item We have a finite dataset $X$ of $n$ points, and we are able to query the labels of any of them.
+\item We want a procedure that chooses the next point, or batch of points, to query. This process might be stopped at any time, and then the remaining labels will need to be inferred.
+\item The goal is for the inferred labels to converge to their ``gold-standard'' values (to be defined) as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
 \end{itemize}
 
-\item {\it Task: Finite population labeling.}
+\item {\it Model of data.}
 
 \begin{itemize}
-\item We are given a finite dataset of $n$ points, and have the
-  ability to query the labels of any of them. The response to the
-  query $x$ is +1 or -1, chosen according the conditional probability
-  function $\eta(x)$. When two or more queries are made on the
-  instance $x$, the responses are chosen independently at random
-  according to $\eta(x)$, in other words, the labels are not persistant.
-
-\item We are given parameters $0 < \gamma, \delta < 1$.
-\item We want a procedure that chooses the next point, or batch of points, to query.
-This process will be applied repeatedly and forever.
-At any point the procedure can be asked to make a prediction on any point in $X$.
-The prediction is $\yh(x) \in \{-1,0,+1\}$.
-We consider the prediction correct if $\yh(x)=+1$ and $\eta(x)=+1$, if $\yh(x)=-1$ and $\eta(x)=-1$ or $\yh(x)=0$ and $|\eta(x)| < \gamma$.
-\item 
-Choosing the queries indepndently at random is guaranteed to reach the correct prediction after querying each example $O(1/\gamma^2)$ times for a total of $O(|X|/\gamma^2)$ queries.
-Our goal is to reach convergence after much fewer queries.
-\item 
-Our bounds are pointwise, i.e.
-we say that the query complexity of the point $x$ is $q(x)$ if for any number of queries larger than $q(x)$ the prediction on $x$ is correct.
-The hope is that for most $x$, $q(x) \ll |X|/\gamma^2)$.
-\item 
-Our bounds should be a function of the advantage, as defined in the adaptive KNN paper: $p \gamma^2$
-\iffalse
-We would like the inferred labels $\yh(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
-\item Specifically, we will then be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
-$$ \sum_{x \in X} {\bf 1}(\yh(x) \neq g^*(x),\ |\eta(x)| > \gamma) .$$
-The procedure is allowed to fail with probability $\delta$, to account for sampling error.
-\fi
+\item The dataset $X$ consists of $n$ independent samples from a fixed but unknown distribution $\mu$ on a metric space $(\X, d)$. 
+\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$, also unknown.
+\item The Bayes-optimal classifier on $\X$ is $g^*(x) = \mbox{sign}(2 \eta(x) - 1)$. This is the gold-standard for inferred labels.
 \end{itemize}
 
+\end{enumerate}
+
+s\section{Deterministic algorithm}
+
+\begin{enumerate}
+\item Define the highest level to be $l_{max} = \log_2 n$.
+\item We have a collection of balls $\B$, organized into levels $\B_\ell$, where $\ell = 0, 1, 2, \ldots$.
+$\B_\ell$ contains all balls whose probability (wrt the uniform distribution over the examples) satisfies $\frac{1}{2^l} \leq p(B) < \frac{1}{2^{l-1}}$
+\item We define $\eta(B) = \pr(Y=1|X \in B)$,
+%We will use the shorthand
+%$$ \B_{\leq \ell} = \bigcup_{\ell' \leq \ell} \B_{\ell'} $$
+%to denote balls upto and including level $\ell$.
+
+\item Let $\B(x) \subset \B$ denote the balls that may be consulted in determining the label of point $x$.
+These could, for instance, be all balls that contain $x$.
+Or they could be all balls centered at some $x \in X$
+
+\item The algorithm can query an oracle for a (of labels) for a set of balls $Q_\ell \subset \B_\ell$.
+The oracle responds with a set of data points such that there are at least $k$ independent samples in each ball $B \in Q_\ell$.
+
+\item
+Using these samples we can label each ball in $Q_\ell$ as $y(B)=+1/0/-1$ such that, with probability of $Q_\ell$ such that with probability $1-\delta$ and for some $\epsilon>0$ the following holds for each $B \in Q_\ell$:
+\begin{itemize}    
+\item $y(B) = +1 \implies \eta(B) \geq 1-\epsilon$
+\item $y(B) = -1 \implies \eta(B) \leq \epsilon$
+\item $y(B) = 0 \implies |\eta(B)-1/2| \leq \epsilon$
 \end{itemize}
+
+\item Once we have queried all balls in $\B_{\ell}(x)$, we determine the set of possible labels of $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$. It is defined as follows:
+\begin{itemize}
+\item $\PL_\ell(x)$ contains $+1$ iff there exists $B \in \B_{\ell}(x)$ with $y(B) = +1$ and there is  no $B' \in \B_{ \ell}(x)$ has both $B' \subset B$ and $y(B') = -1$.
+\item A similar condition determines whether $\PL_\ell(x)$ contains $-1$.
+\item
+Given this set of possible labels of $x$, we assign a label
+$$ y_\ell(x)
+=
+\left\{
+\begin{array}{cl}
++1 & \mbox{if $+1 \in \PL_\ell(x)$  and $-1 \notin \PL_\ell(x)$}\\
+-1 & \mbox{if $-1 \in \PL_\ell(x)$  and $+1 \notin \PL_\ell(x)$}\\
+0 & \mbox{if $\PL_\ell(x) = \{0\}$} \\
+! & \mbox{if $+1 \in \PL_\ell(x)$  and $-1 \in \PL_\ell(x)$}
+\end{array}
+\right.
+$$
+\item Define $U_\ell = \{x \in X: y_{\ell-1}(x) = !\}$
+\item Define the set of query balls 
+$$Q_l= \left\{B \in \B_l  |   B \cap U_{l-1} \neq \emptyset \right\}$$
+And define the focus set to be the union of the query balls:
+$$F_l = \bigcup_{B \in Q_l} B$$
+\end{itemize}
+
+\item Implementing the oracle. Let $P_\ell = \Pr[F_\ell]$
+Our hypothesis is that, if the set of balls has $d$, if we we select $M_\ell = C (\log d \log \delta^{-1})(n k 2^\ell) P_\ell$ independent points in $F_\ell$ then with probability at least $1-\delta$ each ball contains at least $k$ samples.
+
+\item Deterministic algorithm\\
+Set $F_1 = X$  \\
+For $l=1,\ldots, l_{max}$
+\begin{itemize}
+\item {\em Focused Query:} Query $M_\ell$ from $F_\ell$
+\item {\em Uniform query:} Query $M_\ell$ from $X$
+\item {\em Back-filling:} Consider a "hindsight known unknown" to be a set 
+$U_m \neq \emptyset$ for $m<\ell$ that was not identified in iteration $\ell-1$. Then focused sampling is added for levels $m+1$ till $\ell$. (needs clearer explanation)
+\end{itemize}
+\end{enumerate}
+
+\section{Sketch of analysis}
+
+Consider the number of focused queries made at level $\ell$: $M_\ell = C (\log d \log \delta^{-1})(n k 2^\ell) P_\ell$.
+Denote $N_\ell = 2C (\log d \log \delta^{-1})(n k 2^\ell)$. Then $M_{\ell} = N_\ell P_\ell$. is the number of samples needed to guarantee that the balls in $Q_\ell$ all contain at least $k$ samples. 
+If we replace $P_\ell$ with $1$ we get $N_l$: the number of uniform queries needed to get reliable labels for all 
+of the balls in $\B_\ell$.
+
+Fix a point $x\in X$, define the history $H(x)$ to be as binary sequence of length $l_{max}$ where each element is either 
+$C$ for {\em consistent} or $I$ for {\em Inconsistent}.  We use $H_i(x)$ to indicate the $i$'th element in $H(x)$.
+$H_i(x)=I$ if and only if $x \in F_\ell$.
+
+We now characterize the convergence of the predictions on $x$.
+We say that $c(x)$ is the {\em convergence point for $x$} if $j=c(x)$ is the smallest integer such that for all $i\geq j$, $H_i(x)=C$.
+we say that $d(x)$ is the {\em final disagreement point} if $j=d(x)$ is the smallest integer such that for all $j \geq i > c(x)$, $H_j(x)=I$
+
+Our analysis is based on the following intuition. For levels $1,\ldots,d(\ell)-1$ we need to use uniform sampling while for levels $d(\ell),\ldots c(\ell)-1$ we use focused sampling.
+
+As a result the number of samples needed is 
+$$ 2\left( \sum_{\ell=1}^{d(x)-1} N_{\ell} + \sum_{\ell=d(x)}^{c(x)-1} N_\ell P_\ell \right) $$ while the number of queries needed  using standard uniform queries is  
+$$ \sum_{\ell=1}^{c(x)-1} N_{\ell} $$
+Easy examples are ones where the $d(x)$ and $c(x)$ are small and $P_\ell$ is small for $d(x)\geq \ell \geq c(x)$.
+Note that the sequence $P_1,P_2$ is global and does not depend on $x$.
+The dependence on $x$ comes in through $d(x)$ and $c(x)$
+\pagebreak
+\section{1D example}
+I believe that for each point $x$ there are two critical levels. In the following I say that balls are "in agreement" if there is a sign $s \in \pm 1$ such that each ball is labeled either $s$ or $?$
+\begin{itemize}
+    \item The convergence level: all balls at this or higher level are in agreement with the polarity same $s$.
+    \item The last unknown-unknown level: the highest level in which there is agreement and the label is wrong.
+\end{itemize}
+Basically, we rely on uniform sampling until we pass the last unknown-unknown level. We rely on focused sampling from that level until convergence.
+
+In the figure above I sketch a simple case where there are two thresholds and $\eta$ is either $1/2-\gamma$ or $1/2+\gamma$.
+In this case the convergence radius is $d/2$ and the last UU is radii above $2W$
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=4in]{1Dbump.jpg}
+\end{center}
+\label{fig:cases}
+\end{figure}
 
@@
 
-\begin{figure}[ht!]
+\subsection{The algorithm}
+
+\begin{figure}[h!]
 \framebox{
 \begin{minipage}[t]{6.4in}
+\begin{itemize}
+\item Initialize uncertainty regions at all levels:
+\begin{itemize}
+\item $U_0 = X$
+\item $U_\ell = \emptyset$ for $\ell \geq 1$
+\end{itemize}
+
 \vspace{.1in}
-\emph{Variables:}
+\item Repeat:
 \begin{itemize}
-\item $Q$ (queries made so far)
+\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
+\begin{itemize}
+\item Let $\ell'$ be the smallest such level
+\item {\tt Focused-query}($\ell', U_{\ell'}$)
 \end{itemize}
+\item {\tt Background-query}
+\item Update $\yh_{\ell}$ values, $\ell \geq 0$, as per equation (\ref{eq:provisional-label})
+\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
+\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
+\end{itemize}
+
+\end{itemize}
+
+\end{minipage}}
+\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
+\label{alg:main}
+\end{figure}
+
+\begin{figure}[h!]
+\framebox{
+\begin{minipage}[t]{6.4in}
 
@@
 \begin{itemize}
-\item Set $Q := \emptyset$ (points queried so far)
+\item Set $Q = \emptyset$ (points queried so far)
 \item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
@@
 \end{itemize}
-\yoav{I am not sure whether the set from which to pick the query
-  should exclude $Q$: the queries that have been made o far. if a set
-  consists of one point and that point has a weak bias, we need to
-  sample the point several times.}
+
 \end{minipage}}
@@
 \label{alg:sampling}
-\end{figure}
-
-\begin{figure}[ht!]
-\framebox{
-  \begin{minipage}[t]{6.4in}
-    \vspace{.1in}
-\emph{Variables:}
-\begin{itemize}
-\item $U_\ell$ The set of points level $\ell$ that are ``known unknowns''
-\end{itemize}
-
-\vspace{.2in}
-{\bf Active Balls}
-\begin{itemize}
-\item Initialize uncertainty regions at all levels:
-\begin{itemize}
-\item $U_0 = X$
-\item $U_\ell = \emptyset$ for $\ell \geq 1$
-\end{itemize}
-
-\vspace{.1in}
-\item Repeat:
-\begin{itemize}
-\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
-\begin{itemize}
-\item Let $\ell'$ be the smallest such level
-\item {\tt Focused-query}($\ell', U_{\ell'}$)
-\end{itemize}
-\item {\tt Background-query}
-\item Update $\yh_{\ell}$ values, $\ell \geq 0$, as per equation (\ref{eq:provisional-label})
-\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = ! \}$
-%\yoav{The last run before convergence should contain only disagreements: $!$, not zeros $0$}
-\end{itemize}
-
-\end{itemize}
-
-\end{minipage}}
-\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
-\label{alg:main}
 \end{figure}
@@
 Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-discrete} in the appendix.
-
 \begin{thm}
@@
 
-\subsection{Mind changes}
-\iffalse
-\yoav{We say that a prediction {\bf has converged} at level $L_f$ if one of the following holds:
-\begin{itemize}
-\item For all $\ell\geq L_f$, $\yh_\ell(x) = +1 \mbox{ and } \eta(x) > 0$
-\item For all $\ell\geq L_f$, $\yh_\ell(x) = -1 \mbox{ and } \eta(x) < 0$
-\item For all $\ell\geq L_f$, $\yh_\ell(x) = 0  \mbox{ and } |\eta(x)| < \gamma$
-\end{itemize}
-
-}
-\fi
-
-As the level $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_{\ell}(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.
-
-%\yoav{There seems to be a  gap: $\yh_\ell(x)$ is a random variable, but $L_f(x),L_s(x)$ are not.}
-
-We distinguish two critical levels for each $x$: the level $L_s(x)$ beyond which this label no longer flips between $+1$ and $-1$ (although it may be $0$ or $!$), and the level $L_f(x)$ beyond which the label will be correct.
-See Figure~\ref{fig:two-levels} for a pictorial view.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{home-stretch.pdf}
-\end{center}
-\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively higher levels.
-Between levels $L_s(x)$ and $L_f(x)$ there are contradicting labels on $x$, i.e.
-$\yh_\ell(x)=!$.
-Beyond level $L_f(x)$, the correct label will be assigned.}
-\label{fig:two-levels}
-\end{figure}
-
-We begin with $L_s$.
-
-\begin{defn}
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We say $x$ is \emph{obscured} at level $\ell \geq 0$ if the following two conditions hold.
-\begin{itemize}
-\item No $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) > b_2$.
-\item Some $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) < -b_1$.
-\end{itemize}
-Let $L_s(x)$ denote the smallest integer such that $x$ is not obscured at any level $\ell \geq L_s(x)$. If there is no such level, we define $L_s(x) = \infty$.
-\label{defn:Ls}
-\end{defn}
-In words, for levels $\ell < L_s(x)$, there is no ball $B \in \B_\ell(x)$ which has a strong bias towards the correct (Bayes-optimal) label, and there are balls with at least a mild bias towards the wrong label.
-
-Next, we define the level by which the label of a point ought to be apparent.
-\begin{defn}
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label.
-We say $x$ is \emph{resolved} at level $\ell \geq 0$ if the following two conditions hold for all $\ell' \geq \ell$.
-\begin{itemize}
-\item All $B \in \B_{\ell'}(x)$ have $s \cdot \eta_X(B) > -b_1$.
-\item Some $B \in \B_{\ell'}(x)$ has $s \cdot \eta_X(B) > b_2$.
-\end{itemize}
-Let $R_\ell$ denote the points that are resolved by level $\ell$. Let $L_f(x)$ denote the smallest level at which $x$ is resolved, or $\infty$ is there is no such level.
-\label{defn:Lf}
-\end{defn}
-In words, for levels $\ell \geq L_f(x)$, no ball $B \in \B_\ell(x)$ has a strong bias towards the wrong label, and at least one ball has a strong bias towards the correct label.
-
-Thus $L_s(x) \leq L_f(x)$, and there might be a large gap between the two.
-Before the background sampling reaches level $L_s(x)$, it is quite possible that we will predict the wrong label for $x$.
-However, once all balls in $\B_{L_s(x)}(x)$ have been adequately sampled, point $x$ will either (i) be assigned label $0$ or $!$, and put in the uncertainty set, in which case it will benefit from focused sampling, or (ii) will get the correct label.
-Once the balls in $\B_{L_f(x)}(x)$ have been sampled, condition (ii) will hold.
-
-We define the set $F_\ell$ or include all of the points $x \in X$ such that $L_s(x) \leq \ell \leq L_f(x)$.
-The probability of $F_\ell$ determines the sample complexity of the active learning algorithm
-
-\subsection{A possible characterization of $L_f(x),L_s(x)$}
-
-\renewcommand{\P}{{\cal P}}
-We are given a distribution over $\P$ over $\X \times \Y$ and a point $x \in X$.
-We sample a set of unlabeled examples $X \subset \X$, and then run our active learning algorithm.
-
-Our goal is to upper bound the number of samples $|\X|$ and the number
-of queries the algorithm needs before the prediction of the algorithm
-on $x$ converges to the correct label (and will not change with additional samples).
-
-As we are dealing with the true distribution, we redifine $\B_\ell(x)$
-to be all balls containing $x$ whose probability is in the range
-$\left[ \frac{1}{2^{\ell-1}}, \frac{1}{2^\ell}\right]$.
-
-Given the distribution $\P$, an edge parameter $\gamma>0$, a
-prediction point $x$ and a level $\ell$ we partition the balls in $\B_{\ell}(x)$ into three 
-subsets:
-$$\B_{\ell}^+ = \{ B \in \B_\ell(x) | \eta(B)  > \gamma\},\;\;
-\B_{\ell}^- = \{ B \in \B_\ell(x) | \eta(B)  < -\gamma\},\;\;
-\B_{\ell}^0 = \{ B \in \B_\ell(x) | -\gamma \leq \eta(B) \leq \gamma\}$$
-
-let the ``correct label'' of x be $s$, i.e. $s = \sign(\eta(x))$.
-We define the ``true'' values of  $L_f(x),L_s(x)$ as follows:
-\begin{itemize}
-\item $L_s(x)$, is the smallest integer $m$ such that for all $\ell\geq
-  m$, $\B_\ell^s \neq \emptyset$ while $\B_\ell^{-s} = \emptyset$.
-\item $L_f(x)$ is the smallest integer $n \leq L_s(x)$ such that for
-  all $n\leq \ell < L_s(x)$, $\B_\ell^s \neq \emptyset$ and $\B_\ell^{-s} \neq \emptyset$.
-\end{itemize}
-
-In other words the level $L_s(x)$ is the minimal level below which the
-prediction is always correct. $L_f(x)$ is the minimal level above $L_s(x)$
-such that the prediction on $x$ for the levels between  $L_f(x)$ and
-$L_s(x)$ is always $!$.
-
-\subsection{Another definition of two critical levels}
-\yoav{I prefer to work with this definition}
-
-For each point $x \in X$, we define two critical levels $L_0(x)$ and $L_1(x)$.
-
-Let $s = \mbox{sign}(\eta(x)) \in \{-1, +1\}$. Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ after balls in $\B_\ell(x)$ have been queried.
-
-We will define $L_0(x)$ and $L_1(x)$ so that: $s$ is guaranteed to be in $\PL_\ell(x)$ for all $\ell \geq L_0(x)$ and $-s$ is guaranteed to \emph{not} be in $\PL_\ell(x)$ for all $\ell \geq L_1(x)$.
-More precisely:
-\begin{itemize}
-\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$,
-\begin{itemize}
-\item there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$, and
-\item any ball $B' \in \B_{\ell'}(x)$ with $B' \subset B$ has $s \cdot \eta_X(B') > -b_1$.
-\end{itemize}
-\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, any ball $B \in \B_{\ell'}(x)$ either has $s \cdot \eta_X(B) > -b_1$ or contains some $B' \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B') > b_2$.
-\item \yoav{I would leave the definition of $L_0$ unchanged, but define $L_1$ as follows} $L_1(x)$ is the smallest level $\ell$ such that for all $L_0(x) > \ell' \geq \ell$, 
-there exists two balls $B^+,B^- \in \B_{\ell'}(x)$ such that $\eta_X(B^+) > b_2$ and $\eta_X(B^-) < -b_2$.
-\end{itemize}
-
-Without containment:
-\begin{itemize}
-\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$.
-\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is no ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
-\end{itemize}
-
-\section{Characterizing $S_L,S_F$  for Euclidean balls}
-Suppose the input
-distribution is uniform on a ball $B$ of radius one  around the
-origin. Denote the ``positive'' set to be $C$. Let $\gamma>0$ be the
-edge such that
-\[
-  P(y=1|X=x) =\begin{cases}
-    1/2+\gamma & \mbox{ if } x \in C \\
-    1/2-\gamma & \mbox{ if } x \notin C
-    \end{cases}
-  \]
-
-  $L_f(x)$ corresponds to half the distance between $x$ and the boundary
-
-  $L_s(x)$ can be bounded independently of $x$ if $r$ is the largest
-  radius such that both $C$ and $C^c$ can be expressed as a union of
-  balls of radius $r$.
-
-  We can define a bondary of bounded convexity in this way.
-
- For other definitions of ``ball'' we can use a similar
- characterization: $L_s(x)$ corresponds to the smallest ball
- containing both $x$ and a point with the opposite label. $L_f(x)$
- corresponds to the largest radius such that both $C$ and $C^c$ can be
- written as a union of balls.
-
-\section{Examples}
-\begin{itemize}
-    \item 1D
-    \begin{itemize}
-        \item Single threshold with jump in $\eta$ ($-c$ on one side, $+c$ on the other side)
-        \item Single threshold with more general $\eta$, e.g. Tsybakov condition
-        \item Multiple thresholds with $-c, +c$
-    \end{itemize}
-    \item 2D all Euclidean balls
-    \begin{itemize}
-        \item Linear separator with $-c, +c$, e.g. uniform distribution over $x$\\
-        I this case $L_s(x)$ is lower bounded the the probability of a
-        ball whose radius is half the distance between $x$ and the
-        linear separator. $L_f(x)$ is close to 0. Two large balls, one
-        shifted towards the positive, the other shifted towards the
-        negative, will suffice to create a disagreement.
-
-        \item Separator of bounded curvature on both sides
-    \end{itemize}
-    \item Dyadic tree: instead of limiting attention to boxes that are cells of the tree, we look at all boxes of all sizes that supported on the underlying gridding of space.
-We get better bounds for two reasons:
-    \begin{itemize}
-        \item We look at large boxes, not just leaves of the tree.
-        \item We use all boxes, not just those that are nested like the tree; this makes the boundary apparent earlier using !.
-    \end{itemize}
-    \item Arbitrary dimension, but Holder-smoothness and Tsybakov margin conditions (usual setting in nonparametric estimation).
-The rates obtainable using passive random sampling were determined by Audibert-Tsybakov.
-For active, can look at Hanneke.
-Will find out.
-    \item A one-dimensional example, e.g. with data points in [0,1] and with
-$\eta > 3/4$ or $\eta < 1/4$ everywhere. In this case, the balls could
-be all intervals, or perhaps a finite collection of them based on some
-gridding of the line. The Bayes-optimal classifier would have a finite
-number of sign-changes, and we could quantify which of these we'd pick
-up by each level of sampling.
-
-\item A high-dimensional Euclidean setting in which we use balls
-corresponding to the cells of a tree (e.g. RP tree).
-
-\item A high-dimensional metric space setting in which we have a
-collection of metric balls and we give label complexity bounds based
-on smoothness conditions (Holder-smoothness of the $\eta$ function)
-and boundary conditions (like Tsybakov margin condition).
-
-\end{itemize}
+\pagebreak
 
@@
 
-\caption{\yoav{Picture should be changed to include circles containing $x$ but not centered at $x$.} This figure provides some examples for the concept of mind changes.
+\caption{This figure provides some examples for the concept of mind changes.
 Here we consider true probabilities, rather than estimates.
@@
 
-\pagebreak
+\iffalse
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in]{IMG_0840.jpg}
+
+%\includegraphics[width=1.5in]{boundary3.pdf}
+%\hskip1in
+%\includegraphics[width=1.75in]{boundary4.pdf}
+\end{center}
+\caption{\emph{Left:} In this picture, it could be the case that every ball $B \in \B(x)$ has $\eta(B) \geq 1/2$, in which case $L_s(x) = 0$.
+But these balls may not have significant positive bias until a higher level $L_f(x)$ (which includes the shown ball).
+\emph{Right:} This is a situation in which, for small $\ell$, balls $B \in \B_\ell(x)$ might have $\eta(B) < 1/2 - \gamma$.
+But for $\ell \geq L_s(x)$, we get balls (such as the outer ball) with a positive bias, and for $\ell \geq L_f(x)$ (such as the inner ball), a significant positive bias.}
+\label{fig:critical-interval}
+\end{figure}
+\fi
 

### 2024-09-02 AISTAT25/aistats.tex
introduced simplified algorithm, put into aistats format, seperatedd main from supplement

--- 2024-07-31 COLT/colt.tex
@@
-\documentclass[anon,12pt]{colt2024}
-
+\documentclass[twoside]{article}
+
+\usepackage{aistats2024}
 \usepackage{enumitem}
 \usepackage{multicol}
+\usepackage{times}
+\usepackage{amssymb}
+\usepackage{graphicx}
 
@@
 \newtheorem{thm}{Theorem}
-%\newtheorem{lemma}[thm]{Lemma}
+\newtheorem{lemma}[thm]{Lemma}
 \newtheorem{cor}[thm]{Corollary}
@@
 \newtheorem{assump}{Assumption}
-%\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}
+\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}
 
@@
 
-\title[Active learning using region-based sampling]{Active learning using region-based sampling}
-
-\usepackage{times}
-
-\coltauthor{
-\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
-\Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
-\addr University of California, San Diego}
+% If your paper is accepted, change the options for the package
+% aistats2024 as follows:
+%
+%\usepackage[accepted]{aistats2024}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+
+% If you set papersize explicitly, activate the following three lines:
+%\special{papersize = 8.5in, 11in}
+%\setlength{\pdfpageheight}{11in}
+%\setlength{\pdfpagewidth}{8.5in}
+
+% If you use natbib package, activate the following three lines:
+%\usepackage[round]{natbib}
+%\renewcommand{\bibname}{References}
+%\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+
+% If you use BibTeX in apalike style, activate the following line:
+%\bibliographystyle{apalike}
 
@@
 
-\maketitle
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
+\twocolumn[
+
+\aistatstitle{Non-parametric active learning without smoothness}
+
+\aistatsauthor{ Sanjoy Dasgupta \And Yoav Freund}
+
+\aistatsaddress{University of California, San Diego} ]
 
 \begin{abstract}%
-We present a general-purpose active learning scheme that maintains a collection of neighborhoods of different sizes and uses label queries to identify those with a strong bias towards one particular label.
-When two such neighborhoods intersect and have different labels, the region of overlap is treated as a ``known unknown'' and is targeted in future active queries.
-We give label complexity bounds for this method that do not rely on assumptions about the data, and we instantiate them in several cases of interest.
+  We present a general-purpose active learning scheme that maintains a
+  collection of balls of different sizes and uses label
+  queries to identify those with a strong bias towards one particular
+  label. When two such balls intersect and have different
+  labels, the region of overlap is treated as a ``known unknown'' and
+  is targeted in future active queries. We give label complexity
+  bounds for this method that do not rely on assumptions about the
+  data, and we instantiate them in several cases of interest.
 \end{abstract}
 
-\begin{keywords}
-Active learning, nonparametric estimation.
-\end{keywords}
-
 \section{Introduction}
 
-In \emph{active learning}, the starting point is a data set whose labels are hidden but can be obtained individually for a price.
-The goal is to label the data set, or to find a good classifier, at low cost.
-
-We consider a formulation in which we have an arbitrary collection of points $X = \{x_1, \ldots, x_n\}$ from some space $\X$.
-We can request the label of any of these points $x$, in which case we get a value $y \in \{-1, +1\}$ with conditional distribution
-$ \eta(x) = \E[y | x] = \pr(y=1|x) - \pr(y=-1|x) $
-for some unknown function $\eta: \X \to [-1,1]$.
-The \emph{Bayes-optimal} label for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and $+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.
-%$$ g^*(x) = \left\{
-%\begin{array}{cl}
-%-1 & \mbox{if $\eta(x) < 0$} \\
-%+1 & \mbox{if $\eta(x) > 0$} \\
-%\mbox{either} & \mbox{if $\eta(x)= 0$}
-%\end{array} \right.
-%$$
-We wish to find the Bayes-optimal labels for the given data points $X$.
-
-More precisely, at the outset we have the data $X$ and parameters $0 < \gamma, \delta < 1$.
-We want a procedure that chooses the next point, or batch of points, to query.
-This procedure will be applied repeatedly and can be stopped at any time, whereupon labels $\yh(x)$ must be provided for all $x \in X$, including those that were queried.
-Ideally, these will be the Bayes-optimal labels $g^*(x)$; however, we will only be judged on points with non-negligible bias, $|\eta(x)| \geq \gamma$.
-The number of mistakes is:
+In \emph{non-parametric active learning}, the starting point is a data
+set whose labels are hidden but can be obtained individually for a
+price. The goal is to label the data set, or to find a good
+classifier, at low cost.
+
+We consider a formulation in which we have an arbitrary collection of
+points $X = \{x_1, \ldots, x_n\}$ from some metric space
+$\X$.~\footnote{Our approach works for any topological space
+  that defines a sufficiently rich set of neighborhoods. For the same of concreteness, we
+  restrict ourselves to metric spaces where the neighborhoods are balls.}  We can
+query the label of any of these points $x$, in which case we get a
+value $y \in \{-1, +1\}$ with conditional expectation
+$ \eta(x) = \E[y | x] = \pr(y=1|x) - \pr(y=-1|x) $ for some arbitrary
+unknown function $\eta: \X \to [-1,1]$. The \emph{Bayes-optimal} label
+for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and
+$+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.  We
+wish to find an algorithm that approaches the Bayes-optimal labels for
+the points in $X$ using only a small number of queries, called the
+{\em query complexity} of the algorithm.
+
+More precisely, at the outset we have the data $X$ and parameters
+$0 < \gamma, \delta < 1$. We want a procedure that chooses the next
+point, or batch of points, to query. This procedure will be applied
+repeatedly and can be stopped at any time, whereupon labels $\yh(x)$
+must be provided for all $x \in X$, including those that were
+queried. Ideally, these will be the Bayes-optimal labels $g^*(x)$;
+however, we will only be judged on points with non-negligible bias,
+$|\eta(x)| \geq \gamma$. The number of mistakes is:
 $$ \sum_{x \in X} {\bf 1} \big(\yh(x) \neq g^*(x),\ |\eta(x)| \geq \gamma \big) .$$
@@
 
-\subsection{Approach and challenges}
-
-In our setting there is no underlying assumption about $g^*$, for instance that it follows a linear model. Thus we adopt a nonparametric approach.
-
-We begin with a large collection of \emph{neighborhoods}, $\B$, of varying sizes (Fig.~\ref{fig:challenges}(a)); we will refer to these as \emph{balls}.
-We use label queries to assign \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning with the largest balls.
-A bias of $+1$ means that the average $\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is significantly negative; and bias $0$ means the average $\eta$ is close to 0.
-As time goes on, we get biases for progressively smaller balls, as needed.
-The final label assigned to any point $x$ is based on the biases of balls containing $x$.
+Prior work on this problem~(see Section~\ref{sec:relatedwork})
+requires making various assumptions, typically on the smoothness of
+$\eta(x)$.  In contrast, we make no assumptions on either the set $X$
+or on the conditional expectation $\eta(x)$. Our algorithm is
+universal and works for any $X$ and $\eta(x)$. The {\em query
+  complexity} of our algorithm depends on $X$ and $\eta$, but it does
+not depend on smoothness of $\eta(x)$. Instead it depends on
+properties of $\eta(x)$ close to the decision boundary where $\eta(x)$
+changes sign. In particular, the algorithm is effective when $\eta(x)$
+is discontinuous and hence not smooth.
+
+\subsection{Approach}
+
+We begin with a large collection of \emph{balls}, $\B$, of varying
+sizes (Fig.~\ref{fig:challenges}(a)). We use label queries to assign
+\emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning
+with the largest balls. A bias of $+1$ means that the average
+$\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is
+significantly negative; and bias $0$ means the average $\eta$ is close
+to 0. As time goes on, we get biases for progressively smaller balls,
+as needed. The final label assigned to any point $x$ is based on the
+biases of balls containing just $x$.
 
@@
 \begin{center}
-\includegraphics[width=5in]{ideas.pdf}
+\includegraphics[width=3in]{ideas.pdf}
 \end{center}
@@
 
-%, me and to use label queries to estimate the signs associated with these regions.
-Ultimately we hope to cover $\X$ by a patchwork of neighborhoods with $\eta(x)$ values of uniform sign.
-Without strong smoothness conditions, these neighborhoods could be of very different sizes.
-The most beneficial for active learning are neighborhoods that contain a lot of $X$, but to get good coverage we might have to include smaller neighborhoods.
-To address this heterogeneity, we group the regions $\B$ by size.
-We begin by estimating the signs of the largest of them, and then move on to progressively smaller neighborhoods as the need arises.
-When the label budget runs out, the process is stopped, and provisional labels are assigned to the individual points in $X$.
-
 There are three key challenges to be addressed.
-The first is deciding \emph{where to query}.
-Initial random querying tells us the biases of the largest balls, which gives some information about the decision boundary.
-In what parts of $\X$ might it be helpful to know biases of smaller balls, warranting further label queries?
-Our approach is simple: if a ball $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the intersection $B \cap B'$ is a ``known unknown'' and is targeted for active querying  (Fig.~\ref{fig:challenges}(b)).
-
-This differs from the more conventional strategy of actively querying neighborhoods $B$ whose average $\eta$-value is close to zero.
-Such $B$ fall into two categories: (1) the $\eta$ values are close to zero throughout $B$, and (2) $B$ consists of two sub-regions, one strongly positive and the other strongly negative.
-In the first case, there is little merit in querying further.
-But in the second case, there is a lot to be gained.
-Our approach allows these two cases to be distinguished.
-
-%To distinguish these two cases, we use a collection of neighborhoods that are \emph{overlapping}.
+
+\begin{enumerate}
+\item \label{challange:wheretoquery}
+  {\bf Where to query?} \\
+  Initial random querying tells us the biases of the largest balls,
+  which gives some information about the decision boundary. In what
+  parts of $\X$ might it be helpful to know biases of smaller balls,
+  warranting further label queries? Our approach is simple: if a ball
+  $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the
+  intersection $B \cap B'$ is a ``known unknown'' and is targeted for
+  active querying (Fig.~\ref{fig:challenges}(b)).
+
+  This differs from the more conventional strategy of actively
+  querying balls $B$ whose average $\eta$-value is close to zero. Such
+  $B$ fall into two categories: (1) the $\eta$ values are close to
+  zero throughout $B$, and (2) $B$ consists of two sub-regions, one
+  strongly positive and the other strongly negative. In the first
+  case, there is little merit in querying further. But in the second
+  case, there is a lot to be gained. Our approach allows these two
+  cases to be distinguished.
+
+%To distinguish these two cases, we use a collection of balls that are \emph{overlapping}.
 For instance, we might take $\B$ to be \emph{all} balls in $\X$, which is effectively a finite collection once the given data points $X$ are taken into account.
-Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a neighborhood $B$ that is strongly positive as well as being in a neighborhood $B'$ that is strongly negative.
+Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a ball $B$ that is strongly positive as well as being in a ball $B'$ that is strongly negative.
 Such points are ``known unknowns'', and these are the targets of our active querying.
 
-A second challenge is that the sign of $\eta(\cdot)$ can change in an arbitrarily small neighborhood of any given point: the biases of balls containing $x$, say $B_1 \supset B_2 \supset \cdots$, might change sign a few times.
-At first, given only the bias of $B_1$, we might think $x$ has label $+1$.
-When we later estimate the bias of $B_2$, we could change our mind to $-1$.
-And then $+1$, and so on (Fig.~\ref{fig:challenges}(c)).
-
-What this means is that we can \emph{never} be sure of having correctly determined the label of $x$.
-Thus in addition to focused (active) querying, we also do background sampling of the entire space to pick up on possible mind changes.
-For simplicity, we make one background query per focused query.
-
-The third challenge is \emph{managing sampling} of overlapping balls.
-We are interested in detecting points, and thus neighborhoods, whose $\eta$-values are either $> \gamma$ or $<-\gamma$.
-This suggests querying $k \approx 1/\gamma^2$ points at random from each ball.
-Now, suppose we have queried this many points from ball $B$ and later want to query points from a different ball $B'$ that overlaps $B$.
-How can we reuse queries already made in $B \cap B'$?
-
-\emph{Poisson sampling} provides a clean solution.
-Rather than choosing $k$ points at random from $B$, we pick each point in $B$ with probability (roughly) $k/|B \cap X|$, independently.
-We implement this by assigning each $x \in X$ a uniform-random value $T_x \in [0,1]$, at the outset.
-When sampling $B$, we choose to query $x$ if $T_x \leq k/|B \cap X|$.
-And when it comes time to sample a different $B'$ that also contains $x$, we choose $x$ if $T_x \leq k/|B' \cap X|$; if its label has already been obtained, we are able to reuse it.
-In this way, the random querying of overlapping balls is seamlessly managed.
+\item \label{challange:nosmoothness}
+  {\bf No smootheness assumptions}\\
+  As no assumptions of smoothness are made, the sign of $\eta(\cdot)$
+  can change in an arbitrarily small ball of any given point: the
+  biases of balls containing $x$, say
+  $B_1 \supset B_2 \supset \cdots$, might change sign a few times. At
+  first, given only the bias of $B_1$, we might think $x$ has label
+  $+1$. When we later estimate the bias of $B_2$, we could change our
+  mind to $-1$. And then $+1$, and so on
+  (Fig.~\ref{fig:challenges}(c)).
+
+\item \label{challange:managingsampling}
+  {\bf Managing sampling} \\
+  The third challenge is \emph{managing sampling} of overlapping
+  balls. We are interested in detecting points, and thus balls, whose
+  $\eta$-values are either $> \gamma$ or $<-\gamma$. This suggests
+  querying $k \approx 1/\gamma^2$ points at random from each
+  ball. Now, suppose we have queried this many points from ball $B$
+  and later want to query points from a different ball $B'$ that
+  overlaps $B$. How can we reuse queries already made in $B \cap B'$?
+
+\end{enumerate}
 
 These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$.
-By tackling all three of them, we are able to give a general-purpose active learning scheme.
-
+By tackling all three of them, we are able to give a
+universal active learning scheme.
+
+The rest of the paper is organized as follows. Section~\ref{sec:relatedwork} describes related work.
+Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function.
+We identify two \emph{critical levels} for any $x \in X$: two scales (of ball sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).
+We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.
+
+Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$.
+We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.
+
+\iffalse
 \subsection{Results}
@@
 
-Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function.
-We identify two \emph{critical levels} for any $x \in X$: two scales (of neighborhood sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).
-We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.
-
-Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$.
-We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.
-
-\subsection{Related work}
-
-Existing works on nonparametric active learning can mostly be grouped by overall principle: either (1) they seek to obtain the Bayes-optimal labels of a few well-positioned points, and then propagate these to the rest of the space \citep{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or negative) of entire regions at a time \citep{DH08,M12}.
-In this paper, we follow the second strategy because the only reliable and general-purpose way to assess the bias of an individual point, $\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent smoothness assumptions, the sign can change abruptly in an arbitrarily small neighborhood around $x$.
-On the other hand, the bias of a region $B$---the sign of the average $\eta$ value in $B$---is easy to determine given a reasonable number of points from that region.
+\fi
+
+\subsection{Related work} \label{sec:relatedwork}
+
+Existing work on non-parametric active learning yields algorithms that are guaranteed to work when the underlying conditional expectation $\eta(x)$ is smooth in some way.
+Additional assumptions are often needed.
+
+This work can mostly be grouped by overall principle: either (1) they
+seek to obtain the Bayes-optimal labels of a few well-positioned
+points, and then propagate these to the rest of the space
+\cite{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or
+negative) of entire regions at a time \cite{DH08,M12}. In this paper,
+we follow the second strategy because the only reliable and
+general-purpose way to assess the bias of an individual point,
+$\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent
+smoothness assumptions, the sign can change abruptly in an arbitrarily
+small ball around $x$. On the other hand, the bias of a region
+$B$---the sign of the average $\eta$ value in $B$---is easy to
+determine given a reasonable number of points from that region.
 
@@
 
-\section{The active learning algorithm}
-\label{sec:algorithm}
-
-Recall that we have a collection of points $X$ from a space $\X$.
-The label of any $x \in X$ can be requested and the value returned will be either $-1$ or $+1$, according to the conditional probability function $\eta(x) = \E[y|x]$.
-We wish to assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.
-
-\subsection{A collection of sampling regions}
-
-In our active learning algorithm, sampling is organized around a collection $\B$ of subsets of $\X$.
-These are the atomic sets on which we assess label bias and we call them \emph{balls} or \emph{neighborhoods}.
-There are no formal requirements on $\B$---for instance, the so-called balls can be of arbitrary shape---but the intention is that each $x \in X$ is contained in balls of multiple sizes, including one that is so small as to exclude the rest of $X$.
-
-For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the data points that lie in it.
-%\footnote{Since $X_B$ is used to assess the bias of $B$, we need to exclude points that cannot be considered random samples from $B$.
-For instance, consider these two alternatives: (1) $\B$ consists of a pre-defined set of balls.
-(2) $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
-In the first case, all points $X \cap B$ are random draws from $B$.
-In the second case, we need to exclude the two points $x,x'$ that define the ball.}
-We group balls into levels by the number of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
+Recall that the algorithm can query any $x \in X$ and receive $-1$ or $+1$ at random,
+selected according to the conditional expectation function $\eta(x) = \E[y|x]$.
+
+The goal of the algorithm is to make a small number of queries and then
+assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.
+
+The goal of our analysis is to characterize conditions under which the
+number of queries is significantly smaller than
+$\frac{|X|}{\gamma^2}\log \frac{1}{\gamma^2}$ which is the expected
+  number of number of queries made by the naive algorithm that queries
+  random elements of $X$ until each $x \in X$ has been sampled
+  $\frac{1}{\gamma^2}$ times.
+
+\section{Definitions}
+
+The two basic entities in our algorithm are the collection of points
+$X$ and a collection $\B$ of balls. Sampling is organized around a the
+collection $\B$. Balls are the atomic sets on which we assess label
+bias.
+
+For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the
+data points that lie in it.  We group balls into levels by the number
+of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
 \begin{equation}
-\frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .
+\frac{|X|}{2^{\ell + 1}} \leq |X_B| < \frac{|X|}{2^\ell} .
 \label{eq:sampling-level}
 \end{equation}
-Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$.
-Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of highly-populated balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls.
-We will use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater, and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.
-
-Balls in lower levels contain more points, and thus their biases (average $\eta$ values) are easier to estimate. Our sampling algorithm goes from easier to harder levels.
-
-For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that contain $x$ and can thus be used in determining $x$'s label.
-We again partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.
-%We will also refer to $\B_\ell(x)$ as the {\em level $\ell$ neighborhood of $x$}.
-
-\subsection{Estimating the biases of balls}
-
-The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
-$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
-We estimate these biases using label queries and assign each ball a \emph{qualitative} bias estimate,
-$$ \yh(B) = 
+Let $\B_\ell$ consist of all balls in $\B$ that are at level
+$\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$
+consisting of highly-populated balls and subsequent
+$\B_1, \B_2, \ldots$ consisting of successively smaller balls. We will
+use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater,
+and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.
+
+Balls in lower levels contain more points, and thus their biases
+(average $\eta$ values) are easier to estimate. The main loop of the
+simplified learning algorithm iterates over the levels $\ell=0,1,2,\ldots$
+
+For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of
+balls that contain $x$ and can thus be used in determining $x$'s
+label. We again partition these balls by sampling-level, so that
+$\B_\ell(x) = \B(x) \cap \B_\ell$.
+
+\section{A general-purpose active learning algorithm}
+Our algorithm has several intricacies that might make it hard to
+follow at a first read.  We therefor start with a description of a
+simplified version of the algorithm (Fig~\ref{alg:siple}) that is not
+completely correct but serves to explain the main components of the
+detailed algorithm. The full description of the algorithm is given in
+subsection~\ref{sec:full description} and appendix~\ref{sec:detailedalgorithm}
+
+\begin{figure*}[t]
+\framebox{
+\begin{minipage}[t]{6in}  % Adjusted minipage width slightly
+\begin{center}
+\vspace{.1in}
+\textbf{Repeat for $\ell=0,1,2,\ldots,\log |X|-1$:}
+\begin{enumerate}[leftmargin=0.6cm]
+\item \label{step:Uell} {\bf Compute uncertainty set:}\\ If $\ell=0$, set $U_0 = X$; otherwise set $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$t
+\item \label{step:QS} {\bf Compute Focused Query set:}\\ Compute query ball set be $R=  \bigcup_{x \in U} \B_\ell(x)$
+and focused query set $S=\bigcup_{B \in R} B$
+\item \label{step:sample} {\bf Make queries:}\\
+\textbf{Repeat until} number of queries in each ball $B \in R$ is at least $C/\gamma^2$
+  \begin{enumerate}[leftmargin=0.6cm]
+  \item Select a {\tt query} uniformly at random from $S$
+  \end{enumerate}
+
+\item \label{step:balls} {\bf Update Ball labels:} For each ball $B \in R$, let $k(B)$ be
+  the number of queries (of any kind) made inside $B$ and let
+  $\widehat{\eta}(B)$ be the average of the labels returned by
+  these queries.
+
+  $$ \yh(B)
+  =
   \left\{
-\begin{array}{cl}
-+1 & \mbox{significant $+$ bias} \\
--1 & \mbox{significant $-$ bias} \\
-0 & \mbox{no significant bias} \\
-\bot & \mbox{not yet available}
+    \begin{array}{ll}
+%%%      \bot & \mbox{if $k(B) \leq \frac{C}{\gamma^2}$} \\
+%%%      \sign(\widehat{\eta}(B)) & \mbox{if  $k(B) > \frac{C}{\gamma^2}$ and $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+      \sign(\widehat{\eta}(B)) & \mbox{if  $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+      0 & \mbox{otherwise}
     \end{array}
@@
 
+\item \label{step:examples} {\bf Update example labels:} For each $x \in U_\ell$:
+  \begin{align}
+    \PL_\ell(x) &= \{s \in \{-1, +1\}: \ \text{there exists } B \in B_\ell(x) \text{ such that } \yh(B) = s \} \\ 
+    \yh_\ell(x) &= 
+                  \begin{cases}
+                    +1 & \text{if } \PL_\ell(x) = \{+1\} \\
+                    -1 & \text{if } \PL_\ell(x) = \{-1\} \\
+                    0  & \text{if } \PL_\ell(x) = \{\} \\
+                    !  & \text{if } \PL_\ell(x) = \{-1,+1\}
+                  \end{cases}
+                  \label{eq:provisional-label}
+  \end{align}
+\end{enumerate}
+\end{center}
+
+\end{minipage}}
+\caption{A simplified version of the active learning algorithm.}
+\label{alg:simple}
+\end{figure*}
+
+\subsection{Simplified algorithm}
+
+We now describe the simplified algorithm here referred to simply as
+the algorithm. The algorithm loops over increasing levels.  At each
+level it identifies the set of ``known unknowns'' which approximates
+the boundary. The algorithm then makes queries in the vicinity of the
+boundary to refine it and proceeds to the next finer level.
+
+What follows is a detailed explanation of  Figure~\ref{alg:simple}.
+
+The outer loop of the algorithm iterates over the
+levels, starting with $\ell=0$ that corresponds to balls that contain
+at least half of the points in $X$, and ending with $\ell=\log |X|-1$,
+corresponding to balls that contain a single point.
+
+The loop consists of five steps. The first computes the uncertainty set, the second computes the focused query
+set, the third defines the sampling process, the fourth and fifth
+update the labels of the balls and the points respectively.
+
+In step~\ref{step:Uell} the {\em uncertainty region} $U_\ell$ is
+calculated. Intuitively $U_l$ is the set of points whose label
+according to level $\ell-1$ is ! or ``known unknown'' and the label
+according to level $\ell$ is $\bot$ or ``not enough information''.
+These are the points for which we need better estimates.
+
+In step~\ref{step:QS} the algorithm identifies the balls that needs to
+be queried in order to get more information about $U_\ell$. These are
+the balls in $\B_l$ that contain at least one point of $U_\ell$.  The
+set $S$ is the union of these balls and defines the ``focus area''
+which is queried in order to reduce the known unknowns.
+
+In step~\ref{step:sample} the algorithm queries points inside the
+focus area repeatedly at random.  This step terminates once there are
+$C/\gamma^2$ samples in each of the balls in the focus area.
+
+In step ~\ref{step:balls} associates a label with each ball. The
+label is one of $\{-1,+1,0\}$.  The labels $+1$ and $-1$ correspond to
+a confident statement that $\eta(B)>0$ and $\eta(B)<0$
+respectively. Finally, the label $0$ indicates that the sign of $B$
+is uncertain given the answers to the queries.
+
+In step~\ref{step:examples} these labels are propagated to the
+examples.  This is done in two sub-steps. First the {\em Possible
+  Labels} for $x$ denoted $\PL_\ell(x)$ are calculated. $\PL_\ell(x)$
+contains $+1$ if there is a ball containing $x$ whose label is $+1$,
+symmetrically for $-1$. The label $\yh_\ell(x)$ is determined by the
+possible labels set $\PL_\ell(x)$ as described in
+step~\ref{step:examples}. The most significant label is ``!'', which
+indicates balls with conflicting polarity containing $x$. Points
+labeled ``!'' determine the uncertainty region and the focused query
+region in the following level and the process repeats.
+
+\iffalse
+\subsection{Estimating the biases of balls}
+
+The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
+$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
+We estimate these biases using label queries and assign each ball a
+\emph{qualitative} bias estimate $\yh(B)$,
+
+Suppose that $k'$ queries have been made in the ball $B$ and let
+$\widehat{\eta}(B)$ denote the mean of the answers to these queries.
+
+step \ref{step:Uell} and \ref{step:QS} and \ref{step:balls} and \ref{step:sample} and \ref{step:examples}
+
+We set the qualitative bias estimate as follows:
+$$ \yh(B)
+= 
+\left\{
+  \begin{array}{ll}
+    \bot & \mbox{if $k \leq \frac{1}{\gamma^2}$} \\
+\sign(\widehat{\eta}(B)) & \mbox{if  $k > \frac{1}{\gamma^2}$ and $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+0 & \mbox{otherwise}
+\end{array}
+\right.
+$$
+
 The option $\bot$ is used until enough points in $X_B$ have been queried: the required number is $k = O(1/\gamma^2)$ since $\gamma$ is the smallest bias that needs to be detected.
 Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.
+
 These bias estimates will with high probability be seen to satisfy the following guarantee.
@@
 
-The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\citep{R11}.
+The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\cite{R11}.
 Our active learning algorithm makes its \emph{focused queries} at level $\ell+1$ in the vicinity of such points $x$, that is, in $\B_{\ell+1}(x)$.
@@
 
-\subsection{The algorithm}
-
-The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, has two types of query.
-\emph{Background queries} are random draws from $X$ and correspond to passive learning.
-\emph{Focused queries} are made in the vicinity of ``uncertain'' points and correspond to active learning.
-The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$.
-Focused queries are drawn at random from balls in $\B_\ell$ containing such points.
-
-On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
-
-Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$.
-Then the next iteration begins.
+\fi
+
+\subsection{The detailed algorithm}
+\label{sec:detailedalgorithm}
+
+The simplified algorithm does addresses the first challange of ``where
+to query'' (see challanges in the introduction). It does not address
+the other two challanges, which the detailed algorithm does address.
+
+\begin{enumerate}
+\item {\bf No smoothness assumptions:}\\
+  As described in the introduction, one can \emph{never} be sure of having
+  correctly determined the label of $x$. Thus in addition to focused
+  (active) querying, we also do background sampling of the entire
+  space to pick up on possible mind changes. For simplicity, we make
+  one background query per focused query.
+
+  A related problem is that having the outer loop iterate over levels
+  will cause the algorithm to miss some uncertainty region if they are
+  appear at a lower level. Therefor the full algorithm does not loop
+  over levels. Instead, at each iteration it finds the lowest level
+  where the uncertainty set is not empty and makes two queries, one
+  focused and one background.
+
+\item {\bf Managing Sampling:}\\
+  As described in the introduction, we would like to use queries that
+  fall in the intersection of two balls for estimating the bias in
+  both balls. However, selecting examples IID at random from the balls
+  can create an undesired dependence between the estimates. This is solved using Poisson sampling, as described in section~\ref{sec:poisson}.
+\end{enumerate}
+
+\iffalse
+We now describe the detailed active learning algorithm as shown in
+Figs~\ref{alg:main}, \ref{alg:sampling} and~\ref{alg:bias-estimate}.
+The algorithm makes two types of query. \emph{Background queries} are random draws
+from $X$ and correspond to passive learning. \emph{Focused queries}
+are made in the vicinity of ``uncertain'' points and correspond to
+active learning.
+
+Many of the elements of the detailed algorithm have been introduced in the simplified version.
+Specifically $U_\ell,S,\PL_\ell(x),\yh_\ell(x)$, are as defined there.
+
+However, in the detailed algorithm all of the levels are considered at each iteration.
+On each iteration of the main loop (at most) one focused
+query is made as well as a background query. The focused query comes
+from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
+
+Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
+updated using the Poisson method,\\
+$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
+accordingly  and the next iteration begins.
 
@@
 \end{itemize}
-\item {\tt Background-query} // Fig.~\ref{alg:sampling}
+\item Run {\tt Background-query} // Fig.~\ref{alg:sampling}
 \item Update labels:
@@
 
-\subsection{Sampling}
+\subsubsection{Poisson Sampling}
+\label{sec:poisson}
 
@@
 \end{figure}
+
+\fi
 
@@
 
-\begin{thm}[Local label complexity]
+\begin{thm}[Local query complexity]
 Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$.
@@
 \includegraphics[width=2.5in]{oned-monotonic.pdf}
-\hspace{.5in}
+%\hspace{.5in}
 \includegraphics[width=2.5in]{oned-massart.pdf}
@@
 
-\subsection{A sufficiently-rich collection of neighborhoods}
+\subsection{A sufficiently-rich collection of balls}
 
@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=6in]{COLT/curvature.pdf}
-\end{center}
-\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=3in]{curvature.pdf}
+%\end{center}
+%\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
 The orange and light blue indicate the boundary sets for negative and positive respectively.
@@
 All of these circles are monochrome.}
-\label{fig:curvature}
-\end{figure}
+%\label{fig:curvature}
+%\end{figure}
 
@@
 We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.
-By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\citep{BC03}.
+By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\cite{BC03}.
 
@@
 The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$.
-It is a common notion of curvature~\citep{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
+It is a common notion of curvature~\cite{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
 A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
 \begin{thm}
-Under the two conditions above, there are constants $c_2, c_3$ for which the following holds.
-Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
-If the algorithm of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$ queries, it assigns Bayes-optimal labels to all but $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
+  Under the two conditions above, there are constants $c_2, c_3$ for
+  which the following holds. Pick $0 < \delta < 1$ and take
+  $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm
+  of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$
+  queries, it assigns Bayes-optimal labels to all but
+  $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
 \label{thm:massart-dist}
 \end{thm}
-Here the constant $c_2$ includes a term $(1/r_o)^d$.
-The $\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over the usual $1/m^{1/d}$ rate for random querying and was also observed earlier, in a more restricted setting, using specialized algorithms~\citep{CN08}.
-
-\newpage
+
+ Here the constant $c_2$ includes a term $(1/r_o)^d$. The
+$\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over
+the usual $1/m^{1/d}$ rate for random querying and was also observed
+earlier, in a more restricted setting, using specialized
+algorithms~\cite{CN08}.
 
@@
 \bibliography{refs}
-%\bibliographystyle{plain}
-
-\newpage
-\appendix
-
-\section{Technicalities: discrete setting}
-
-\subsection{Large deviation bounds for the discrete setting}
-
-\begin{lemma}
-Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.
-
-Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
-$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
-Now consider the following estimator $Z$ of this quantity:
-\begin{itemize}
-\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
-\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
-\end{itemize}
-If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
-\begin{enumerate}
-\item[(a)] $N \geq k$, and 
-\item[(b)] $| Z - \eta_o| < \sqrt{(48/k) \ln (4/\delta)}$.
-\end{enumerate}
-\label{lemma:large-deviation-discrete}
-\end{lemma}
-\begin{proof}
-Let's start with (a).
-Define $c = \sqrt{3 \ln (4/\delta)}$.
-We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c \sqrt{2k}$ since this is the worst case.
-By assumption, $k \geq 2c^2$ and thus $qm \leq 2k$.
-
-Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
-\begin{align*}
-\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
-\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
-\end{align*}
-Take $\epsilon = c/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c$, we see that with probability at least $1-\delta/2$, we get 
-$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
-The lower bound implies $N > qm(1-\epsilon) = qm - c\sqrt{qm} = k + c \sqrt{2k} - c \sqrt{qm} \geq k$.
-
-For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$.
-The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$.
-Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$.
-Thus their sum $S$ has expectation
-$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
-and variance
-$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
-We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
-$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + t/3)} \right) .$$
-Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.
-
-Combining with the high-probability bound on $N$ above, we get
-$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
-whereupon (recalling $Z = S/N$)
-$$ 
-\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
-Since $|\eta_o| \leq 1$,
-$$ |Z - \eta_o | < \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
-\leq 
-4 \epsilon,$$
-as claimed.  
-\end{proof}
-
-\subsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}
-
-We will use Lemma~\ref{lemma:large-deviation-discrete} to obtain a uniform guarantee on the bias estimates for all balls $B \in \B$.
-
-Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
-$$ \eta(x) = \E[y | x].$$ 
-For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value of the points in $B$, that is,
-$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
-
-For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$, where the sampling threshold $\tau_\ell$ for level $\ell$ is defined as:
-\begin{equation}
-\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
-\end{equation}
-for some $k$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.
-
-For what follows, define
-\begin{equation*}
-c = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
-\end{equation*}
-
-\begin{lemma}
-Suppose $k \geq 2c^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
-\begin{enumerate}
-\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
-\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
-$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c}{\sqrt{k}}.$$
-\end{enumerate}
-\label{lemma:large-deviation-bounds}
-\end{lemma}
-\begin{proof}
-Pick $B \in \B$. There are two cases to consider.
-
-Case 1: $|X_B| \geq 2k$.
-The choice of $\tau_\ell$ then ensures $|X_B| \tau_\ell \geq 2k$, so that $|\widehat{\eta}(B) - \eta_X(B)|$ can be bounded by applying Lemma~\ref{lemma:large-deviation-discrete} to the points $X_B$ with sampling probability $q = \tau_\ell$.
-
-Case 2: $k \leq |X_B| < 2k$.
-In this case, $B$ lies at a level $\ell$ for which $\tau_\ell = 1$.
-Lemma~\ref{lemma:large-deviation-discrete} does not apply directly, but its conclusion still holds.
-In particular, the query-set $\Gamma(B)$ is all of $X_B$, and the same Bernstein bound from the proof of Lemma~\ref{lemma:large-deviation-discrete} can be again be applied.
-
-To complete the proof, we take a union bound over all $B \in \B$.
-\end{proof}
-
-The following corollary of Lemma~\ref{lemma:large-deviation-bounds} is immediate.
-\begin{cor}
-Suppose that $k \geq (8c/\gamma)^2$. For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
-$$ \yh(B)
-= 
-\left\{
-\begin{array}{ll}
-\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
-0 & \mbox{otherwise}
-\end{array}
-\right.
-$$
-With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\gamma$-accurate.
-\label{cor:accurate-bias-estimates}
-\end{cor}
-\begin{proof}
-By the choice of $k$, we have from Lemma~\ref{lemma:large-deviation-bounds} that 
-$| \widehat{\eta}(B) - \eta_X(B) | < \gamma/2$
-for all $B \in \B$.
-\end{proof}
-
-\subsection{Critical levels: Proof of Lemma~\ref{lemma:boundary}}
-
-For part (a), note that some $B_o \in \B_{L_1(x)}(x)$ has significant bias (that is, bias $\geq \gamma$) towards the correct label $s(x)$, as does any ball $B \in \B(x)$ contained within it.
-For $\ell \geq L_1(x)$, the set $\{B \in \B_{\leq \ell}(x): X_{B} \subseteq X_{B_o}\}$ is nonempty (it contains $B_o$); pick any minimal ball $B$ within it.
-Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimates}) and thus $s(x) \in \PL_\ell(x)$.
-
-For (b), take any $\ell \geq L_2(x)$.
-Consider any $B \in \B_{\leq \ell}(x)$ for which $s(x) \cdot \eta_X(B) < 0$.
-By definition of $L_2(x)$, this $B$ must lie in $B_{< L_2(x)}(x)$, and moreover there must exist $B' \in \B_{\leq L_2(x)}(x) \subset \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
-Thus, any minimal $B \in \B_{\leq \ell}(x)$ has bias $\geq 0$ towards the correct label $s(x)$, whereupon $\yh(B) \in \{0, s(x)\}$ by the $\gamma$-accuracy of bias estimates.
-Therefore, $-s(x) \not\in \PL_\ell(x)$.
-
-\subsection{The region of focused sampling: Proof of Lemma~\ref{lemma:focused}}
-
-Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$.
-From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty set at level $\ell$.
-In short,
-\begin{equation}
-\overline{U}_{\ell} \subset \{x \in X: L_2(x) \geq \ell\}.
-\label{eq:uncertainty}
-\end{equation}
-
-We see from the {\tt Focused-query} subroutine (Figure~\ref{alg:sampling}) that all focused samples at level $\ell$ lie in
-$$
-\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
-\ 
-\subset 
-\ 
-\bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
-\ 
-=
-\ 
-\{z \in \Delta_\ell: T_z \leq \tau_\ell \}.
-$$
-
-\subsection{A generic label complexity bound}
-
-We start by showing that various subsets of interest contain roughly the expected number of points at each level.
-\begin{lemma}
-With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/2k)$.
-\begin{enumerate}
-\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| < 2 n \tau_\ell$.
-\item[(b)] If $\Delta_\ell \neq \emptyset$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| < 2 |\Delta_\ell| \tau_\ell$.
-\end{enumerate}
-\label{lemma:level-distribution}
-\end{lemma}
-\begin{proof} 
-Pick any subset $S \subset X$ and let $m = |S|$.
-Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$.
-The probability that it is greater than or equal to twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m \tau_{\ell}/3}$, which is $\leq e^{-k/3}$ as long as $m \tau_\ell \geq k$.
-
-Both parts follow from this principle; and we take a union bound over all $\lg (n/k)$ levels.
-For (b), we need to check that $|\Delta_\ell| \tau_\ell \geq k$.
-To see this, observe from the definition (\ref{eq:sampling-region}) of $\Delta_\ell$ that if it is non-empty, then it contains $X_B$ for at least one ball $B \in \B_\ell$, and every such ball has at least $n/2^{\ell+1}$ points.
-Combining this with the definition $\tau_\ell = \min(2^{\ell+2}k/n, 1)$ yields $|\Delta_\ell| \tau_\ell \geq k$.
-\end{proof}
-
-Theorem~\ref{thm:label-complexity} is a restated version of the following.
-\begin{thm}
-Suppose the active learning algorithm makes $0 < m \leq n$ queries. Then all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
-$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
-and $\ell_2$ is the largest integer $\leq \lg (n/2k)$ such that
-$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} |\Delta_{\ell}| \, \tau_{\ell} \ < \ \frac{m}{8} .$$
-\label{thm:label-complexity-0}
-\end{thm}
-
-\begin{proof}
-Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}.
-We will analyze the effect of background sampling in phase one and focused sampling in phase two.
-We start with the former.
-
-Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.
-
-Now, for $\ell_1$ as defined, we have that $\tau_{\ell_1} \leq m/8n$ and thus from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy $T_x \leq \tau_{\ell_1}$.
-Therefore all such points are queried in phase one, and all label-estimates $\{\yh_{\ell_1}(x): x \in X\}$ are set.
-
-It then follows from Lemma~\ref{lemma:boundary}(a) that the following hold for any $x \in X$ with $L_1(x) \leq \ell_1$:
-\begin{enumerate}
-\item[(a)] By the end of phase one, $\yh_{\ell_1}(x) \in \{g^*(x), !\}$.
-\item[(b)] For any $\ell > \ell_1$, when $\yh_\ell(x)$ becomes available, it lies in $\{g^*(x), !\}$.
-\item[(c)] If $x$ ever leaves the combined uncertainty region $U = \cup_\ell U_\ell$ during phase two, then its final label as defined in (\ref{eq:final-label}) is henceforth always $\yh(x) = g^*(x)$.
-\end{enumerate}
-
-Now let's move on to phase two.
-Let $A = \{x \in X: L_1(x) \leq \ell_1, L_2(x) \leq \ell_2\}$.
-We will show that every point in $A$ must leave the uncertainty region $U$ at some time during phase two.
-From (c), we can conclude that all these points have their final labels set correctly, once and for all.
-
-We will break the argument into two cases.
-
-{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.
-
-{\it Case 2:} A full $m/4$ focused queries are made in phase two.
-By the analysis of phase one, none of these queries can be at level $\leq \ell_1$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_1+1$ through $\ell_2$ inclusive is at most
-$$\sum_{\ell=\ell_1+1}^{\ell_2} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
-\ \leq \ 
-\sum_{\ell=\ell_1+1}^{\ell_2} 2 |\Delta_{\ell}| \tau_\ell
-\ < \ 
-\frac{m}{4} ,
-$$
-where the first inequality is from Lemma~\ref{lemma:level-distribution}(b).
-Thus at least one query in phase two must be at level $\ell_2 +1$.
-When this query is made, every $U_\ell$ with $\ell \leq \ell_2$ must be empty and thus all of $A$ must have left the uncertainty region; recall from (\ref{eq:uncertainty}) that no $x \in A$ can be part of $U_\ell$ for $\ell > \ell_2 \geq L_2(x)$.
-
-Thus every $x \in A$ must leave the uncertainty region at some point in phase two, and their final labels are subsequently set correctly.
-\end{proof}
-
-\subsection{One-dimensional monotonic $\eta$: Proof of Theorem~\ref{thm:oned-monotonic}}
-
-We begin by bounding the critical levels $L_1$ and $L_2$ for points in $X$.
-\begin{lemma}
-Pick any $x \in [0, 1]$.
-\begin{enumerate}
-\item[(a)] Define $n^- = |[0,\lambda_L] \cap X|$ and $n^+ = |[\lambda_R,1] \cap X|$. Then
-$$
-L_1(x)
-\leq
-\left\{
-\begin{array}{ll}
-\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
-\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
-\end{array}
-\right.
-$$
-\item[(b)] Let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$, counting $x$ itself.
-That is, $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$.
-Then $L_2(x) \leq \lg (n/r(x))$.
-\end{enumerate}
-\label{lemma:oned-monotonic-L}
-\end{lemma}
-\begin{proof}
-For (a), take any $x \geq \lambda_R$ (the other case is similar).
-The interval $B = [\lambda,1]$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta_X(B) \geq \gamma$.
-Furthermore, any $B' \subset B$ also has $\eta_X(B') \geq \gamma$.
-
-For (b), take $x \geq \lambda_R$ and $\ell \geq \lg (n/r(x))$.
-Any $B \in \B_\ell(x)$ contains $< n/2^\ell \leq r(x)$ points and thus cannot possibly extend to the other side of the boundary.
-It follows that every $B \in \B_{\geq \ell}(x)$ has $\eta_X(B) \geq 0$.
-Moreover, any interval $B'$ that does extend to the other side of the boundary contains $B' \in \B_{\ell}(x)$ that is entirely on the same side as $x$.
-\end{proof}
-
-We can now bound the size of the query region at each level and find that it shrinks exponentially with $\ell$.
-\begin{lemma}
-For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq 4n/2^\ell$.
-\label{lemma:oned-monotonic-query-region}
-\end{lemma}
-
-\begin{proof}
-We have
-\begin{align*}
-\Delta_\ell 
-&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $n/2^{\ell-1}$ points from $X$ on either side of $\lambda$. \end{proof}
-
-We are now ready for the proof of Theorem~\ref{thm:oned-monotonic}.
-
-Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}.
-Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.
-
-First observe that for any $\ell_1 \leq \ell_2$, we have from Lemma~\ref{lemma:oned-monotonic-query-region} that
-$$ \sum_{\ell=\ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell 
-\ \leq \ \sum_{\ell=\ell_1+1}^{\ell_2} \frac{4n}{2^\ell} \cdot \frac{2^{\ell+2} k}{n} 
-\ = \ 16k(\ell_2-\ell_1).$$
-Now, let's define
-$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor, 
-\ \ \ \ell_2 = \min \left( \ell_1 + \frac{m}{128k}, \ \ \lg \frac{n}{2k} \right).$$
-Then for any $x \in [0, \lambda_L] \cup [\lambda_R,1]$, we have
-$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor 
-\ \geq \ 
-\lg \frac{m}{64k}
-\ \geq \ 
-\lg \frac{n}{\min(n^+,n^-)}
-\ \geq \ 
-L_1(x)$$
-and, if $\min(r^+, r^-) \geq 2k$,
-\begin{align*}
-\ell_2 
-\ = \ \ell_1 + \frac{m}{128k} 
-&\geq \lg \frac{m}{64k} + \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \\
-&\geq \ \lg \frac{m}{64k} + \lg \frac{64kn/m}{\min(r^+,r^-)} 
-\ = \ \lg \frac{n}{\min(r^+,r^-)}
-\ \geq \  
-L_2(x).
-\end{align*}
-We get the algorithmic guarantee by applying Theorem~\ref{thm:label-complexity}.
-
-\subsection{One-dimensional data with Massart noise}
-\label{sec:oned-massart}
-
-We now turn to another one-dimensional setting.
-Once again, $X$ consists of $n$ arbitrarily-placed points in $\X = [0,1]$.
-This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
-\begin{itemize}
-\item There are $p$ disjoint open intervals $I_1, \ldots, I_p$, such that $\X$ is (the closure of) their union, and
-\item for each $j$, either $\eta(x) > \gamma$ for all $x \in I_j$ or $\eta(x) < -\gamma$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
-\end{itemize}
-Here $0 < \gamma < 1$ is some constant. See Figure~\ref{fig:oned}(b) for an illustrative example. For concreteness, the intervals $I_j$ can be written in the form $(\lambda_{j-1}, \lambda_j)$, where
-$0 = \lambda_0 < \lambda_1 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$
-Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
-
-We will take $\B$ to consist of all open intervals of $[0,1]$, with $\B(x)$ denoting intervals that contain point $x$.
-
-We begin with bounds on the $L_1$ and $L_2$ levels for each point.
-\begin{lemma}
-Pick any $x \in \X$; suppose $x \in I_j$.
-\begin{enumerate}
-\item[(a)] Let $n_j = |X \cap I_j|$. Then $L_1(x) \leq \lg (n/n_j)$.
-\item[(b)] Let $r(x)$ be the minimum number of data points that lie between $x$ and a boundary point, counting $x$ as well; this is either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller.
-Then $L_2(x) \leq \lg (n/r(x))$.
-\end{enumerate}
-\label{lemma:oned-massart-L}
-\end{lemma}
-\begin{proof}
-For (a), notice first that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$.
-Moreover, $s(I_j) \cdot \eta_X(I_j) > \gamma$.
-Thus $I_j$ belongs to $\B_\ell(x)$ and is strongly biased towards the correct label.
-This strong bias also holds for any subset of $I_j$.
-
-For (b), consider any $\ell \geq \lg (n/r(x))$.
-Any $B \in \B(x)$ with $\mbox{sign}(\eta_X(B)) \neq s(I_j)$ must contain either the entire left-interval $(\lambda_{j-1},x]$ or the entire right-interval $[x, \lambda_j)$, and thus has at least $r(x)$ points, which means that it is too large to be in $\B_\ell$.
-Thus all intervals $B \in \B_{\geq \ell}(x)$ have $s(I_j) \cdot \eta_X(B) > 0$.
-Also, for any interval $B \in \B_{<\ell}(x)$ there is some $B' \in \B_{\ell}(x)$ that is strictly contained within it.
-\end{proof}
-
-With $L_1(x)$ and $L_2(x)$ under control, it is easy to bound the size of the focused query region $\Delta_\ell$ at each level.
-\begin{lemma}
-For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq (p-1)n/2^{\ell-2}$.
-\label{lemma:oned-massart-query-region}
-\end{lemma}
-
-\begin{proof}
-We have
-\begin{align*}
-\Delta_\ell 
-&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $n/2^{\ell-1}$ points from $X$ on either side of each boundary point $\lambda_j$.
-\end{proof}
-
-Notice that $|\Delta_\ell|$ shrinks exponentially with $\ell$.
-With $L_1$, $L_2$, and $|\Delta_\ell|$ values in place, Theorem~\ref{thm:label-complexity} can be applied directly to give the following label complexity bound.
-
-\begin{thm}
-Pick any $0 < \epsilon, \delta < 1$. Suppose we run the algorithm of Figure~\ref{alg:main} with $k = O((1/\gamma^2) \ln (n/\delta))$. With probability at least $1-\delta$, after making
-$$ O \left( \frac{p-1}{\gamma^2} \ln \frac{p-1}{\epsilon} \ln \frac{n}{\delta} \right) $$
-queries, the algorithm will assign the correct label $g^*(x)$ to at least $1-\epsilon$ fraction of $X$, except possibly the $2k$ points of either side of each boundary point.
-\label{thm:oned-massart}
-\end{thm}
-
-\begin{proof}
-Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}.
-Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.
-
-Next, using Lemma~\ref{lemma:oned-massart-query-region}, we have that for any integers $0 \leq \ell_1 < \ell_2$,
-$$ \sum_{\ell = \ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell
-\ \leq \ 
-\sum_{\ell = \ell_1+1}^{\ell_2} \frac{(p-1) n}{2^{\ell-2}} \cdot \frac{2^{\ell+2}k}{n}
-\ = \ 
-16(p-1)k (\ell_2 - \ell_1).
-$$
-We can then apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$, for
-\begin{align*}
-\ell_1
-&= 
-\left\lfloor \lg \frac{m}{32k} \right\rfloor \\
-\ell_2
-&=
-\min \left( \ell_1 + \frac{m}{128 k(p-1)}, \ \lg \frac{n}{2k} \right)
-\end{align*}
-Using Lemma~\ref{lemma:oned-massart-L}, we have that in every target interval $I_j$ with $n_j/n = \Omega(k/m)$, all but $n \cdot 2^{-\ell_2-1}$ points will be correctly classified.
-For large enough $m$, this means that the fraction of misclassified or unclassified points in $X$ will be at most $\epsilon$ after $O(k(p-1) \log ((p-1)/\epsilon))$ queries, apart from the $2k$ points nearest the boundaries, which will remain unclassified.
-\end{proof}
-
-The $2k$ points nearest each boundary cannot be labeled by our algorithm with any certainty because they lie in intervals with a strongly positive bias as well as in intervals with a strongly negative bias.
-This qualification would be removed if were allowed to make multiple queries to each point, because in that case we would include $O(k)$ copies of each point, as explained earlier.
-
-\section{Technicalities: distributional setting}
-
-In the distributional setting, $X$ is drawn i.i.d.\ from a distribution $\mu$ on a topological space $\X$. We have a collection of regions $\B$.
-
-\subsection{Sampling level and probability mass}
-
-We start by showing how the fidelity of $\B$ (Definition~\ref{defn:conformity}) can be established.
-\begin{lemma}
-With probability at least $1-\delta$,
-\begin{enumerate}
-\item[(a)] For all $B \in \B$ with $n \mu(B) \geq 12 \ln (2|\B|/\delta)$, we have
-$$ \frac{\mu(B)}{2} \leq \frac{|X_B|}{n} \leq 2 \mu(B) .$$
-\item[(b)] For all $B \in \B$ with $n \mu(B) < 12 \ln (2|\B|/\delta)$, we have $|X_B| < 24 \ln (2 |\B|/\delta)$.
-\end{enumerate}
-\label{lemma:ball-size-bounds}
-\end{lemma}
-\begin{proof}
-It is an immediate consequence of the multiplicative Chernoff bound that with probability at least $1-\delta$, for all $B \in \B$,
-$$ \frac{|X_B|}{n} = \mu(B) \left(1 \pm \sqrt{\frac{3}{n \mu(B)} \ln \frac{2|\B|}{\delta}} \right) .$$
-\end{proof}
-
-In particular, taking $k > 24 \ln (2 |\B|/\delta)$ ensures that condition (a) hold will hold for all $B \in \B$ with $|X_B| \geq k$.
-
-Next, we will see that balls of probability mass $p$ belong to level $\ell \approx \lg (1/p)$.
-\begin{lemma}
-For any level $\ell \geq 0$ and any $B \in \B_{\ell}$ with $|X_B| \geq k$,
-$$ \left\lceil \lg \frac{1}{\mu(B)} \right\rceil -2 \leq \ell \leq  \left\lceil \lg \frac{1}{\mu(B)} \right\rceil.$$
-\label{lemma:mass-vs-level}
-\end{lemma}
-\begin{proof}
-Recall the definition of level $\ell$:
-$$ B \in \B_\ell 
-\ \Longleftrightarrow \ \frac{n}{2^{\ell+1}} \leq |X_B| < \frac{n}{2^\ell} 
-\ \Longleftrightarrow \ 2^\ell < \frac{n}{|X_B|} \leq 2^{\ell + 1}.$$
-By Lemma~\ref{lemma:ball-size-bounds}, 
-$$ \frac{1}{2 \mu(B)} \leq \frac{n}{|X_B|} \leq \frac{2}{\mu(B)}.$$
-Thus $2^{\ell} < 2/\mu(B)$ and $1/(2 \mu(B)) \leq 2^{\ell+1}$. These translate into the stated bounds on $\ell$.
-\end{proof}
-
-\subsection{Bounding $L_2$ using probabilistic distance: Proof of Lemma~\ref{lemma:L2-bound}}
-
-Let $p = \dist(x, \X^{-s(x)}) \geq 2k/n$ and let $\ell = \lceil \lg (1/p) \rceil + 1 \leq \lg (2n/k)$.
-
-For any $B \in \B_{\geq \ell}(x)$, we have $\mu(B) \leq 2|X_B|/n < 2/2^\ell \leq p$, using fidelity (Definition~\ref{defn:conformity}), the definition of sampling levels, and the definition of $\ell$, in that order.
-It follows that $B$ does not intersect $\X^{-s(x)}$, whereupon $s(x) \cdot \eta_X(B) \geq 0$.
-
-Next, pick any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$; thus $B \in \B_{< \ell}(x)$.
-By the richness of $\B$, there exists $B' \in \B_\ell(x)$ with $X_{B'} \subset X_B$; and so $s(x) \cdot \eta_X(B') \geq 0$.
-% Thus $B$ must intersect $\X^{-s(x)}$ and has probability mass $\geq p$.
-We will show that there exists $B' \in \B_{\leq \ell}(x)$ such that $B' \subset B$ and $B'$ does not intersect $\X^{-s(x)}$; whereupon $s(x) \cdot \eta_X(B') \geq 0$.
-Indeed, take $B' \in \B(x)$ to be a subset of $B$ of $\mu$-mass $p-\epsilon$ for some very small $\epsilon$; we can do this because of the absolute continuity of $\mu$.
-Then $B'$ does not touch $\X^{-s(x)}$ and by Lemma~\ref{lemma:mass-vs-level}, for small enough $\epsilon$, it lies at level $\leq \ell$.
-
-\subsection{Uncertainty region in the distributional setting: Proof of Lemma~\ref{lemma:delta-continuous}}
-
-Recall from (\ref{eq:sampling-region}) that
-$$ \Delta_\ell = \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .$$
-Consider any $z \in \Delta_\ell$.
-Then there exists $x \in \X$ with $L_2(x) \geq \ell$ and $B \in \B_\ell(x)$ such that $z \in B$.
-Now, $B \in \B_\ell(x)$ implies $|X_B|/n < 1/2^\ell$ and so (by the fidelity property) $\mu(B) < 2/2^\ell$.
-We will show that $x \in \partial_{4/2^\ell}$ and thus $z \in \partial_{4/2^\ell, 2/2^{\ell}}$.
-
-There are two cases for $x$. If $\dist(x, \X^{-s(x)}) < 2k/n$ then we immediately have $x \in \partial_{4/2^\ell}$ since $4/2^\ell \geq 2k/n$.
-
-On the other hand, if $\dist(x, \X^{-s(x)}) \geq 2k/n$ then, since $L_2(x) \geq \ell$, we can apply Lemma~\ref{lemma:L2-bound} to get 
-$$ \ell 
-\ \leq \ \left\lceil \lg \frac{1}{\dist(x,\X^{-s(x)})} \right\rceil + 1 
-\ \leq \ 
-\lg \frac{1}{\dist(x,\X^{-s(x)})} + 2 .
-$$
-Thus $\dist(x, \X^{-s(x)}) \leq 1/2^{\ell-2}$ and $x \in \partial_{4/2^\ell}$.
-
-\subsection{A generic label complexity bound in the distributional setting}
-
-We will need to relate the size of the query region $\Delta_\ell$ to the probability mass of the corresponding second-order boundary.
-For this, we provide an analog of Lemma~\ref{lemma:level-distribution} for the distributional setting.
-\begin{lemma}
-With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/4k)$.
-\begin{enumerate}
-\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
-\item[(b)] $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 n \mu(\partial_{4/2^\ell, 2/2^\ell}) \tau_\ell$.
-\end{enumerate}
-\label{lemma:level-distribution-dist}
-\end{lemma}
-\begin{proof} 
-Part (a) is as in Lemma~\ref{lemma:level-distribution}.
-
-For (b), pick any subset $S \subset \X$.
-If $X$ consists of $n$ independent draws from $\mu$, then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(n, \mu(S) \tau_\ell)$ distribution with expectation $n \mu(S) \tau_\ell$.
-The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-n \mu(S) \tau_\ell/3}$.
-We will apply this to the various sets $S = \partial_{4/2^\ell, 2/2^\ell}$ and take a union bound over them.
-In each application, we will also see that $n \mu(S) \tau_\ell \geq k$.
-
-Pick any $\ell \leq \lg (n/4k)$.
-If $\partial_{4/2^\ell, 2/2^\ell} = \emptyset$, then the statement in (b) is trivially true given Lemma~\ref{lemma:delta-continuous}.
-So assume this is not the case.
-Writing $p = 4/2^\ell$, we need to check that $\mu(\partial_{p, p/2}) \tau_\ell \geq k/n$, or equivalently, $\mu(\partial_{p, p/2}) \geq \max(1/2^{\ell+2}, k/n) = 1/2^{\ell+2}$.
-Now, $\partial_{p,p/2} \neq \emptyset \Longrightarrow \partial_p \neq \emptyset$.
-Pick any $x \in \partial_p$.
-We have by the richness of $\B$ (Definition~\ref{defn:conformity}) that there exists $B \in \B_\ell(x)$.
-By the definition of level, $p/8 \leq |X_B|/n < p/4$ and thus, by the fidelity property, $p/16 \leq \mu(B) < p/2$.
-It follows that $B \subset \partial_{p,p/2}$, whereupon and $\mu(\partial_{p,p/2}) \geq \mu(B) \geq p/16 = 1/2^{\ell+2}$.
-%By absolute continuity of $\mu$, we can grow a ball around $x$ of probability mass arbitrarily close to $p/2$, so that this ball is contained within $\partial_{p,p/2}$.
-Thus $\mu(\partial_{p,p/2}) \geq p/2-\epsilon$ for any $\epsilon > 0$.
-The required conditions then follow from the value of $p$.
-\end{proof}
-
-\begin{thm}
-Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$ and that the active learning algorithm makes $0 < m \leq n$ queries.
-Then with probability at least $1-3\delta$, all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
-$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
-and $\ell_2$ is the largest integer $\leq \lg (n/4k)$ such that
-$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell}) \ < \ \frac{m}{32k} .$$
-\label{thm:label-complexity-dist}
-\end{thm}
-
-\begin{proof}
-The proof is identical to that of Theorem~\ref{thm:label-complexity-0}; the only change is to use Lemma~\ref{lemma:level-distribution-dist}(b) in place of Lemma~\ref{lemma:level-distribution}(b).
-\end{proof}
-
-\subsection{Rates of convergence under distributional conditions}
-
-We now give label complexity bounds under three assumptions.
-Let $\X_\gamma = \{x \in \X: |\eta(x)| \geq \gamma\}$; thus $X \cap \X_\gamma$ is the set of points whose assigned labels will be judged.
-We further divide this set by label: for $s \in \{-1,+1\}$, let $\X^s_\gamma = \{x \in \X: s \cdot \eta(x) \geq \gamma\}$.
-
-The first assumption is on the curvature of the decision regions. It allows us to bound $L_1(x)$.
-\begin{enumerate}
-\item[(A1)] There is an absolute constant $p_o > 0$ for which the following holds: for any $x \in \X_\gamma$, there exists $B \in \B(x)$ such that $B \subset \X^{s(x)}_\gamma$ and $\mu(B) \geq p_o$.
-\end{enumerate}
-\begin{lemma}
-Under (A1), every $x \in \X_\gamma$ has $L_1(x) \leq \lceil \lg (1/p_o) \rceil$, provided $n > k/2p_o$.
-\label{lemma:L1-bound}
-\end{lemma} 
-\begin{proof}
-Pick any $x \in X_\gamma$; apply (A1) to get $B \in \B(x)$ for which $\mu(B) \geq p_o$ and $B \subset \X^{s(x)}_\gamma$.
-By Lemma~\ref{lemma:mass-vs-level}, $B \in \B_\ell$ for $\ell \leq \lceil \lg (1/p_o) \rceil$.
-Now, $s(x) \cdot \eta_X(B) \geq \gamma$; moreover, for any $B' \in \B(x)$ with $X_{B'} \subset X_B$ we have $X_{B'} \subset \X^{s(x)}_\gamma$ and thus $s(x) \cdot \eta_X(B') \geq \gamma$ as well.
-\end{proof}
-
-The second assumption is a variant of the Tsybakov margin condition and bounds $|\Delta_\ell|$.
-\begin{enumerate}
-\item[(A2)] There are absolute constants $C > 0$ and $0 < \sigma < 1$ for which the following holds: for any $p > 0$, we have $\mu(\partial_{p,p}) \leq C p^\sigma$.
-\end{enumerate}
-If, for instance, $\mu$ were the uniform distribution over $[0,1]^d$, we would expect $\sigma \sim 1/d$.
-
-The third assumption bounds the fraction of points with large $L_2$ values.
-It asks, what fraction of the $p$-boundary touches $\X_\gamma$?
-We consider two options: under (A3), the fraction is zero for $p$ small enough; under (A3'), the fraction is $p^\xi$.
-\begin{enumerate}
-\item[(A3)] There is an absolute constant $p_1 > 0$ such that $\mu(\partial_p \cap \X_\gamma) = 0$ for any $p < p_1$.
-\item[(A3')] There are constants $C' > 0$ and $0 < \xi < 1$ such that $\mu(\partial_{p} \cap \X_\gamma) \leq C' p^\xi$ for any $p > 0$.
-\end{enumerate}
-
-Under these assumptions, we get rates of convergence as follows.
-\begin{thm}
-Assume (A1), (A2) and either (A3) or (A3').
-There is an absolute constant $C''$ for which the following holds.
-Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
-If the algorithm of Fig.~\ref{alg:main} makes $m$ queries, for
-$$ \frac{128k}{p_o} \ \leq \ m \ \leq \ \frac{64C \cdot 8^\sigma}{1-\sigma} \, n^{1-\sigma} k^\sigma,$$
-then with probability at least $1-\delta$:
-\begin{itemize}
-\item Under (A3), all of $X \cap \X_\gamma$ get Bayes-optimal labels for $m > (512C/(1-\sigma)) \cdot (1/p_1)^{1-\sigma} \cdot k$.
-\item Under (A3'),  Bayes-optimal labels get assigned to all but $ C'' (k/m)^{\xi/(1-\sigma)} + (2/n) \log (4/\delta)$
-%$$ C'' \left(\frac{k}{m}\right)^{\xi/(1-\sigma)} + \frac{2 \log (4/\delta)}{n} $$
-fraction of $X \cap \X_\gamma$.
-\end{itemize}
-\label{thm:label-complexity-specific}
-\end{thm}
-This theorem yields the result of Section~\ref{sec:Massart}. Another example is in Section~\ref{sec:Tsybakov}, where bounds are given under smoothness and Tsybakov margin conditions.
-
-\begin{proof}
-In this case, $\B$ is infinite, but by standard VC-dimension arguments there are only $O(n^{d+1})$ balls with distinct sets $X_B$. This governs the setting of $k$.
-
-First, define $\ell_1 = \lfloor \lg (m/32k) \rfloor$ and observe that by Lemma~\ref{lemma:L1-bound}, all points in $\X_\gamma$ have 
-$$ L_1(x) 
-\ \leq \ 
-\left\lceil \lg \frac{1}{p_o} \right\rceil
-\ \leq \ 
-\lg \frac{2}{p_o} 
-\ \leq \ 
-\lg \frac{m}{64k}
-\ \leq \ 
-\ell_1 .
-$$
-Next, pick  
-$$ \ell_2 = \left\lfloor \frac{1}{1-\sigma} \left( \lg \frac{m}{32k} + \lg \frac{1-\sigma}{C \cdot 4^{1+\sigma}} \right) \right\rfloor.$$
-The upper bound on $m$ ensures that this is at most $\lg (n/2k)$. Then, using (A2),
-$$
-\sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell})
-\ \leq \ 
-C \sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \left( \frac{4}{2^\ell} \right)^\sigma 
-\ \leq \ 
-C \cdot 4^\sigma \cdot \frac{2}{2^{1-\sigma}-1} \cdot 2^{(1-\sigma)\ell_2}
-\ \leq \ 
-\frac{m}{32k} .
-$$
-Applying Theorem~\ref{thm:label-complexity-dist}, we then see that all points in $\X_\gamma$ with $L_2(x) \leq \ell_2$ will be correctly classified.
-Any remaining point $x \in \X_\gamma$ has $L_2(x) > \ell_2$ and thus (by Lemma~\ref{lemma:L2-bound})
-$$ \left\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \right\rceil + 1 > \ell_2 
-\ \Longrightarrow \ 
-\dist(x, \X^{-s(x)}) < \frac{4}{2^{\ell_2}} \leq  \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} . 
-$$
-Call this quantity $p$; thus any such $x$ lies in $\partial_p$.
-
-Under (A3), $\partial_p \cap \X_\gamma$ has zero probability mass for $p < p_1$, that is, if 
-$$ \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} < p_1
-\ \Longleftrightarrow \ 
-m > \frac{512C}{1-\sigma} \cdot \frac{1}{p_1^{1-\sigma}} \cdot k .
-$$
-
-Under (A3'), $\mu(\partial_p \cap \X_\gamma) \leq C' p^{\xi}$; we can then apply a Bernstein bound to assert that with probability at least $1-\delta$, 
-$$ |X \cap (\partial_p \cap \X_\gamma)| \ \leq \ \frac{3}{2} C' n p^{\xi} + 2 \log \frac{1}{\delta},$$
-from which the bound in the theorem follows by defining $C''$ appropriately.
-\end{proof}
-
-\subsection{Label complexity under curvature and Massart noise: Proof of Theorem~\ref{thm:massart-dist}}
-
-Let (C1) denote the strong density condition and (C2) the Massart noise and boundary condition.
-Theorem~\ref{thm:massart-dist} follows immediately from the more general result of Theorem~\ref{thm:label-complexity-specific}, once we establish how (C1) and (C2) relate to assumptions (A1), (A2), and (A3').
-
-\begin{lemma}
-Conditions (C1) and (C2) yield assumption (A1), with $p_o = r_o^d \cdot c_o \cdot v_d$, where $v_d$ is the volume of the unit ball in $\R^d$.
-\label{lemma:A1}
-\end{lemma}
-
-\begin{proof}
-Pick any $x \in \X_\gamma^+$ (the negative case is similar), and let $r = \inf_{z \in \X^0} \|x - z\|$.
-If $r \geq r_o$, then $B = B(x,r_o)$ is entirely in $\X_\gamma^+$.
-Otherwise, the reach condition (C2) implies the existence of a ball $B \subset \X_\gamma^+$ that contains $x$ and has radius $r_o$.
-Either way, $\mu(B) \geq c_o \vol(B) = c_o v_d r_o^d$ by (C1).
-\end{proof}
-
-\begin{lemma}
-Conditions (C1) and (C2) yield assumptions (A2) and (A3') with $\sigma = \xi = 1/d$. 
-\end{lemma}
-
-\begin{proof}
-Under (C1), any ball of probability mass $\leq p$ has volume $\leq p/c_o$ and radius $\leq (p/(c_o v_d))^{1/d}$.
-Let's call this latter quantity $r$.
-Thus, any point in $\partial_p$ lies within distance $2r$ of the boundary, while a point in $\partial_{p,p}$ lies within distance $4r$.
-
-To bound the volume of $\partial_{p,p}$, we can associate each point in this region with its nearest neighbor in $\X^0$; by condition (C1), this projection map is uniquely defined for $r < r_o/4$.
-The volume of the region is thus $O(r)$ and under (C1), has probability mass $O(r) = O(p^{1/d})$.
-\end{proof}
-
-\subsection{Label complexity under smoothness and Tsybakov noise}
-\label{sec:Tsybakov}
-
-Now we specialize Theorem~\ref{thm:label-complexity-specific} to another situation, where the data distribution satisfies Holder smoothness and Tsybakov margin conditions.
-
-Recall that $\dist(x,S)$ denotes the probability-distance from $x$ and set $S$. We will overload notation so that for $x,x' \in \X$,
-$$ \dist(x,x') = \dist(x, \{x'\}) = \inf\{\mu(B): B \in \B(x) \cap \B(x')\},$$
-that is, the probability mass of the smallest ball containing both $x$ and $x'$. We will impose the following conditions on the data.
-\begin{enumerate}
-\item[(C1)] [Strong density condition] This was introduced in the earlier (Massart) example.
-\item[(C2')] [Holder-smoothness of conditional probability function] There exist constants $L, \alpha$ such that
-$$ |\eta(x) - \eta(x')| \ \leq \ L \cdot \dist(x,x')^\alpha$$
-for all $x,x' \in \X$.
-\item[(C3')] [Tsybakov margin condition] There exists constants $M, \beta$ such that
-$$ \mu(\{x \in \X: |\eta(x)| \leq \tau\}) \leq M \tau^\beta$$
-for all $\tau \in (0,1)$.
-\item[(C4')] [Bounded curvature] The boundaries $\{x \in \X: \eta(x) = \gamma\}$ and $\{x \in \X: \eta(x) = -\gamma\}$ are $(d-1)$-dimensional Riemannian manifolds of reach $r_o > 0$.
-\end{enumerate}
-
-Lemma~\ref{lemma:A1} continues to hold, with condition (C4') doing the job of (C2). This yields assumption (A1). For the remaining assumptions, we first obtain a consequence of the Holder condition.
-
-\begin{lemma}
-Under (C2'), for any $p, q > 0$,
-\begin{enumerate}
-\item[(a)] $x \in \partial_p \implies |\eta(x)| \leq L p^\alpha$.
-\item[(b)] $x \in \partial_{p,q} \implies |\eta(x)| \leq L(p^\alpha + q^\alpha)$.
-\end{enumerate}
-\label{lemma:holder-consequence}
-\end{lemma}
-\begin{proof}
-Pick any $x \in \partial_p$.
-Let $s(x) = \mbox{sign}(\eta(x))$.
-By definition of the $p$-boundary set, for any $\epsilon > 0$, there exists $x' \in \X^{-s(x)}$ such that $\dist(x,x') < p+\epsilon$.
-By the Holder condition, $|\eta(x) - \eta(x')| < L(p+\epsilon)^\alpha$ and thus $|\eta(x)| < L(p+\epsilon)^\alpha$.
-Since this holds for any $\epsilon > 0$, we get part (a).
-
-For (b), pick $x \in \partial_{p,q}$ and $\epsilon > 0$.
-Then there exists $x' \in \partial_p$ with $\dist(x,x') < q+\epsilon$.
-As before, we use the Holder condition to conclude that $|\eta(x)| < |\eta(x')| + L(q+\epsilon)^\alpha$ and then invoke (a).
-\end{proof}
-
-\begin{lemma}
-Conditions (C2') and (C3') yield assumptions (A2) and (A3) with $C = (2L)^\beta M$, $\sigma = \alpha \beta$, and $p_1 = (\gamma/L)^{1/\alpha}$.
-\label{lemma:boundary-constants} 
-\end{lemma}
-
-\begin{proof}
-By Lemma~\ref{lemma:holder-consequence}, $\partial_{p,p} \subset \{x \in \X: |\eta(x)| \leq 2Lp^\alpha\}$; the probability mass of this set can be bounded by (C3').
-
-Also by Lemma~\ref{lemma:holder-consequence}, $\partial_p$ is entirely contained in $\{x \in \X: |\eta(x)| \leq L p^\alpha\}$. For $p < p_1$, this does not intersect $\X_\gamma$.
-\end{proof}
-
-\begin{thm}
-Assume conditions (C1) and (C2')--(C4').
-There are constants $c_2, c_3$ for which the following holds.
-Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
-Suppose the algorithm of Fig.~\ref{alg:main} makes $m$ queries, where
-$$ \frac{c_2k}{p_1^{1-\sigma}} \ \leq \ m \ \leq \ c_3 n^{1-\sigma} ,$$
-with $p_1, \sigma$ as given in Lemma~\ref{lemma:boundary-constants}. Then with probability at least $1-\delta$, all of $X \cap \X_\gamma$ is assigned Bayes-optimal labels.
-\label{thm:smoothness-dist}
-\end{thm}
-
-In this setting, regular $k$-nearest neighbor classification would also correctly classify all of $X \cap X_\gamma$ if given $O(k/p_1)$ random labeled points.
-The benefit of active learning is to thus reduce the label complexity to $O(k/p_1^{1-\alpha\beta})$.
+\bibliographystyle{plain}
+
 \end{document}

### 2024-09-02 AISTAT25/supplement.tex
introduced simplified algorithm, put into aistats format, seperatedd main from supplement

--- 2024-07-31 COLT/colt.tex
@@
 
-%\setlength{\leftmargini}{0.5cm}
-%\setlength{\leftmarginii}{0.25cm}
-%\setlength{\leftmarginiii}{0.25cm}
-
-\title[Active learning using region-based sampling]{Active learning using region-based sampling}
-
-\usepackage{times}
-
-\coltauthor{
-\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
-\Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
-\addr University of California, San Diego}
+% If your paper is accepted, change the options for the package
+% aistats2022 as follows:
+%
+%\usepackage[accepted]{aistats2024}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+
+% If you set papersize explicitly, activate the following three lines:
+%\special{papersize = 8.5in, 11in}
+%\setlength{\pdfpageheight}{11in}
+%\setlength{\pdfpagewidth}{8.5in}
+
+% If you use natbib package, activate the following three lines:
+%\usepackage[round]{natbib}
+%\renewcommand{\bibname}{References}
+%\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+
+% If you use BibTeX in apalike style, activate the following line:
+%\bibliographystyle{apalike}
 
@@
 
-\maketitle
-
-\begin{abstract}%
-We present a general-purpose active learning scheme that maintains a collection of neighborhoods of different sizes and uses label queries to identify those with a strong bias towards one particular label.
-When two such neighborhoods intersect and have different labels, the region of overlap is treated as a ``known unknown'' and is targeted in future active queries.
-We give label complexity bounds for this method that do not rely on assumptions about the data, and we instantiate them in several cases of interest.
-\end{abstract}
-
-\begin{keywords}
-Active learning, nonparametric estimation.
-\end{keywords}
-
-\section{Introduction}
-
-In \emph{active learning}, the starting point is a data set whose labels are hidden but can be obtained individually for a price.
-The goal is to label the data set, or to find a good classifier, at low cost.
-
-We consider a formulation in which we have an arbitrary collection of points $X = \{x_1, \ldots, x_n\}$ from some space $\X$.
-We can request the label of any of these points $x$, in which case we get a value $y \in \{-1, +1\}$ with conditional distribution
-$ \eta(x) = \E[y | x] = \pr(y=1|x) - \pr(y=-1|x) $
-for some unknown function $\eta: \X \to [-1,1]$.
-The \emph{Bayes-optimal} label for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and $+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.
-%$$ g^*(x) = \left\{
-%\begin{array}{cl}
-%-1 & \mbox{if $\eta(x) < 0$} \\
-%+1 & \mbox{if $\eta(x) > 0$} \\
-%\mbox{either} & \mbox{if $\eta(x)= 0$}
-%\end{array} \right.
-%$$
-We wish to find the Bayes-optimal labels for the given data points $X$.
-
-More precisely, at the outset we have the data $X$ and parameters $0 < \gamma, \delta < 1$.
-We want a procedure that chooses the next point, or batch of points, to query.
-This procedure will be applied repeatedly and can be stopped at any time, whereupon labels $\yh(x)$ must be provided for all $x \in X$, including those that were queried.
-Ideally, these will be the Bayes-optimal labels $g^*(x)$; however, we will only be judged on points with non-negligible bias, $|\eta(x)| \geq \gamma$.
-The number of mistakes is:
-$$ \sum_{x \in X} {\bf 1} \big(\yh(x) \neq g^*(x),\ |\eta(x)| \geq \gamma \big) .$$
-The overall procedure is allowed to fail with probability $\delta$, to account for sampling error.
-
-\subsection{Approach and challenges}
-
-In our setting there is no underlying assumption about $g^*$, for instance that it follows a linear model. Thus we adopt a nonparametric approach.
-
-We begin with a large collection of \emph{neighborhoods}, $\B$, of varying sizes (Fig.~\ref{fig:challenges}(a)); we will refer to these as \emph{balls}.
-We use label queries to assign \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning with the largest balls.
-A bias of $+1$ means that the average $\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is significantly negative; and bias $0$ means the average $\eta$ is close to 0.
-As time goes on, we get biases for progressively smaller balls, as needed.
-The final label assigned to any point $x$ is based on the biases of balls containing $x$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=5in]{ideas.pdf}
-\end{center}
-\caption{(a) A collection of balls $\B$.
-Ball $B$ has bias $\yh(B) \in \{-1,0,+1\}$.
-(b) If one ball has bias $+1$ and the other $-1$, points in their intersection are targeted for active querying.
-(c) Biases of progressively smaller balls containing $x$ might flip sign, leading to \emph{mind changes} about the label of $x$.}
-\label{fig:challenges}
-\end{figure}
-
-%, me and to use label queries to estimate the signs associated with these regions.
-Ultimately we hope to cover $\X$ by a patchwork of neighborhoods with $\eta(x)$ values of uniform sign.
-Without strong smoothness conditions, these neighborhoods could be of very different sizes.
-The most beneficial for active learning are neighborhoods that contain a lot of $X$, but to get good coverage we might have to include smaller neighborhoods.
-To address this heterogeneity, we group the regions $\B$ by size.
-We begin by estimating the signs of the largest of them, and then move on to progressively smaller neighborhoods as the need arises.
-When the label budget runs out, the process is stopped, and provisional labels are assigned to the individual points in $X$.
-
-There are three key challenges to be addressed.
-The first is deciding \emph{where to query}.
-Initial random querying tells us the biases of the largest balls, which gives some information about the decision boundary.
-In what parts of $\X$ might it be helpful to know biases of smaller balls, warranting further label queries?
-Our approach is simple: if a ball $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the intersection $B \cap B'$ is a ``known unknown'' and is targeted for active querying  (Fig.~\ref{fig:challenges}(b)).
-
-This differs from the more conventional strategy of actively querying neighborhoods $B$ whose average $\eta$-value is close to zero.
-Such $B$ fall into two categories: (1) the $\eta$ values are close to zero throughout $B$, and (2) $B$ consists of two sub-regions, one strongly positive and the other strongly negative.
-In the first case, there is little merit in querying further.
-But in the second case, there is a lot to be gained.
-Our approach allows these two cases to be distinguished.
-
-%To distinguish these two cases, we use a collection of neighborhoods that are \emph{overlapping}.
-For instance, we might take $\B$ to be \emph{all} balls in $\X$, which is effectively a finite collection once the given data points $X$ are taken into account.
-Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a neighborhood $B$ that is strongly positive as well as being in a neighborhood $B'$ that is strongly negative.
-Such points are ``known unknowns'', and these are the targets of our active querying.
-
-A second challenge is that the sign of $\eta(\cdot)$ can change in an arbitrarily small neighborhood of any given point: the biases of balls containing $x$, say $B_1 \supset B_2 \supset \cdots$, might change sign a few times.
-At first, given only the bias of $B_1$, we might think $x$ has label $+1$.
-When we later estimate the bias of $B_2$, we could change our mind to $-1$.
-And then $+1$, and so on (Fig.~\ref{fig:challenges}(c)).
-
-What this means is that we can \emph{never} be sure of having correctly determined the label of $x$.
-Thus in addition to focused (active) querying, we also do background sampling of the entire space to pick up on possible mind changes.
-For simplicity, we make one background query per focused query.
-
-The third challenge is \emph{managing sampling} of overlapping balls.
-We are interested in detecting points, and thus neighborhoods, whose $\eta$-values are either $> \gamma$ or $<-\gamma$.
-This suggests querying $k \approx 1/\gamma^2$ points at random from each ball.
-Now, suppose we have queried this many points from ball $B$ and later want to query points from a different ball $B'$ that overlaps $B$.
-How can we reuse queries already made in $B \cap B'$?
-
-\emph{Poisson sampling} provides a clean solution.
-Rather than choosing $k$ points at random from $B$, we pick each point in $B$ with probability (roughly) $k/|B \cap X|$, independently.
-We implement this by assigning each $x \in X$ a uniform-random value $T_x \in [0,1]$, at the outset.
-When sampling $B$, we choose to query $x$ if $T_x \leq k/|B \cap X|$.
-And when it comes time to sample a different $B'$ that also contains $x$, we choose $x$ if $T_x \leq k/|B' \cap X|$; if its label has already been obtained, we are able to reuse it.
-In this way, the random querying of overlapping balls is seamlessly managed.
-
-These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$.
-By tackling all three of them, we are able to give a general-purpose active learning scheme.
-
-\subsection{Results}
-
-%Our algorithm does not require any assumption on the input distribution.
-In the worst case the number of queries is twice the number of queries of a passive algorithm.
-We show that under several natural conditions the query complexity of our algorithm is logarithmic in the error.
-
-Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function.
-We identify two \emph{critical levels} for any $x \in X$: two scales (of neighborhood sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).
-We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.
-
-Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$.
-We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.
-
-\subsection{Related work}
-
-Existing works on nonparametric active learning can mostly be grouped by overall principle: either (1) they seek to obtain the Bayes-optimal labels of a few well-positioned points, and then propagate these to the rest of the space \citep{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or negative) of entire regions at a time \citep{DH08,M12}.
-In this paper, we follow the second strategy because the only reliable and general-purpose way to assess the bias of an individual point, $\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent smoothness assumptions, the sign can change abruptly in an arbitrarily small neighborhood around $x$.
-On the other hand, the bias of a region $B$---the sign of the average $\eta$ value in $B$---is easy to determine given a reasonable number of points from that region.
-
-Early results of \cite{CN08} established upper and lower bounds on label complexity in situations where the Bayes-optimal boundary is of a simple form: a single threshold for one-dimensional data or a ``smooth boundary fragment'' in higher dimension.
-
-An algorithm for active learning based on hierarchical sampling was given by \cite{DH08} and was analyzed under smoothness conditions by \cite{KUB15}.
-The idea is to begin with a hierarchical clustering of $X$, and to then use queries to discover a pruning of this tree whose leaf-clusters are almost-pure in their labels.
-The method is not well-suited to situations with significant noise levels.
-Another approach using dyadic partitions was given by \cite{M12} and analyzed under commonly-used smoothness, margin, and density assumptions---namely, that $\eta$ is Holder-smooth, the fraction of points with $|\eta(x)| \leq t$ is some polynomial in $t$, and the marginal density is close to uniform---along with an additional smoothness requirement on $\eta$.
-
-A different strategy using nearest neighbors was explored by \cite{ASU18}.
-Their idea was to choose an appropriate scale $s$, find an $s$-cover of $X$, estimate the Bayes-optimal label for each point in this cover by querying its neighbors, and then use these cleanly-labeled points for 1-nearest neighbor classification.
-A somewhat more general approach was given by \cite{H17} and studied under the usual smoothness, margin, and density conditions, with resulting rates of convergence comparable to those found by \cite{M12}.
-
-Finally, \cite{DNZ15} suggested a graph-based method for active learning based on adaptively looking for the cut in the graph corresponding to the correct decision boundary.
-Their assumptions are based on properties of this cut and are not easily comparable with earlier work.
-
-The idea of predicting with balls of multiple sizes is a central theme of \cite{BDFM19}, which predicts at a point $x$ by choosing a ball centered at $x$ that is as small as possible while having a clear bias towards one label.
-That work does not consider active learning, which demands considerable further innovation.
-For instance, to assess the uncertainty at a point $x$, we need to look at balls containing $x$, not just those centered at $x$ (Fig.~\ref{fig:challenges}(b)).
-
-\section{The active learning algorithm}
-\label{sec:algorithm}
-
-Recall that we have a collection of points $X$ from a space $\X$.
-The label of any $x \in X$ can be requested and the value returned will be either $-1$ or $+1$, according to the conditional probability function $\eta(x) = \E[y|x]$.
-We wish to assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.
-
-\subsection{A collection of sampling regions}
-
-In our active learning algorithm, sampling is organized around a collection $\B$ of subsets of $\X$.
-These are the atomic sets on which we assess label bias and we call them \emph{balls} or \emph{neighborhoods}.
-There are no formal requirements on $\B$---for instance, the so-called balls can be of arbitrary shape---but the intention is that each $x \in X$ is contained in balls of multiple sizes, including one that is so small as to exclude the rest of $X$.
-
-For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the data points that lie in it.
-%\footnote{Since $X_B$ is used to assess the bias of $B$, we need to exclude points that cannot be considered random samples from $B$.
-For instance, consider these two alternatives: (1) $\B$ consists of a pre-defined set of balls.
-(2) $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$.
-In the first case, all points $X \cap B$ are random draws from $B$.
-In the second case, we need to exclude the two points $x,x'$ that define the ball.}
-We group balls into levels by the number of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
-\begin{equation}
-\frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .
-\label{eq:sampling-level}
-\end{equation}
-Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$.
-Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of highly-populated balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls.
-We will use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater, and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.
-
-Balls in lower levels contain more points, and thus their biases (average $\eta$ values) are easier to estimate. Our sampling algorithm goes from easier to harder levels.
-
-For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that contain $x$ and can thus be used in determining $x$'s label.
-We again partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$.
-%We will also refer to $\B_\ell(x)$ as the {\em level $\ell$ neighborhood of $x$}.
-
-\subsection{Estimating the biases of balls}
-
-The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
-$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
-We estimate these biases using label queries and assign each ball a \emph{qualitative} bias estimate,
-$$ \yh(B) = 
-\left\{
-\begin{array}{cl}
-+1 & \mbox{significant $+$ bias} \\
--1 & \mbox{significant $-$ bias} \\
-0 & \mbox{no significant bias} \\
-\bot & \mbox{not yet available}
-\end{array}
-\right.
-$$
-
-The option $\bot$ is used until enough points in $X_B$ have been queried: the required number is $k = O(1/\gamma^2)$ since $\gamma$ is the smallest bias that needs to be detected.
-Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.
-These bias estimates will with high probability be seen to satisfy the following guarantee.
-\begin{defn}
-For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
-\begin{itemize}
-\item $\yh(B) = +1 \implies \eta_X(B) > 0$
-\item $\yh(B) = -1 \implies \eta_X(B) < 0$
-\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
-\end{itemize}
-\label{def:accurate-bias-estimate}
-\end{defn}
-
-\subsection{Assigning labels to points}
-
-The label of any point is inferred from the biases of balls containing it.
-
-Pick any $x \in X$ and any level $\ell$.
-Once qualitative bias estimates $\yh(B)$ are available for all balls at levels $\leq \ell$ that contain $x$ (that is, $\B_{\leq \ell}(x)$), we use these to infer a level-$\ell$ label for $x$.
-To begin with, the set of \emph{possible} labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$, is defined thus:
-\begin{itemize}[leftmargin=0.5cm]
-\item $\PL_\ell(x)$ includes $+1$ if there exists a \emph{minimal} ball $B \in \B_{\leq \ell}(x)$ (with no other $B' \in \B_{\leq \ell}(x)$ strictly contained in it) for which $\yh(B) = +1$.
-\item $\PL_\ell(x)$ includes $-1$ under a symmetrical condition.
-\end{itemize}
-This is made precise in Eqn~(\ref{eq:PL}) in Fig.~\ref{alg:main}.
-The label-estimate for $x$ at level $\ell$, denoted $\yh_\ell(x)$, is $+1$ if $\PL_\ell(x)=\{+1\}$, $-1$ if $\PL_\ell(x) = \{-1\}$, $0$ if $\PL_\ell(x) = \{\}$, and $!$ if $\PL_\ell(x) = \{-1,+1\}$ (Eqn~(\ref{eq:provisional-label})).
-%The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.
-
-The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\citep{R11}.
-Our active learning algorithm makes its \emph{focused queries} at level $\ell+1$ in the vicinity of such points $x$, that is, in $\B_{\ell+1}(x)$.
-
-% to be $-1$ if $\PL_\ell(x) = %\{-1\}$, or $+1$ if $\PL_\ell(x) = \{+1\}$, or $0$ if $\PL_\ell(x) = \{\}$, or $!$ %if $\PL_\ell(x) = \{-1,+1\}$, %and remains fixed thereafter.
-Points labeled $!$ are ``known unknowns'': they are %near the decision boundary and %need further investigation.
-Points labeled $0$ show no significant bias.
-
-\subsection{The algorithm}
-
-The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, has two types of query.
-\emph{Background queries} are random draws from $X$ and correspond to passive learning.
-\emph{Focused queries} are made in the vicinity of ``uncertain'' points and correspond to active learning.
-The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$.
-Focused queries are drawn at random from balls in $\B_\ell$ containing such points.
-
-On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
-
-Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$.
-Then the next iteration begins.
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
+% Supplementary material: To improve readability, you must use a single-column format for the supplementary material.
+\onecolumn
+\aistatstitle{Non-parametric active learning without smoothness\\
+Supplementary Materials}
+
+\section{Detailed algorithm}
+
+We now describe the detailed active learning algorithm as shown in
+Figs~\ref{alg:main}, \ref{alg:sampling} and~\ref{alg:bias-estimate}.
+The algorithm makes two types of query. \emph{Background queries} are random draws
+from $X$ and correspond to passive learning. \emph{Focused queries}
+are made in the vicinity of ``uncertain'' points and correspond to
+active learning.
+
+Many of the elements of the detailed algorithm have been introduced in the simplified version.
+Specifically $U_\ell,S,\PL_\ell(x),\yh_\ell(x)$, are as defined there.
+
+However, in the detailed algorithm all of the levels are considered at each iteration.
+On each iteration of the main loop (at most) one focused
+query is made as well as a background query. The focused query comes
+from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
+
+Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
+updated using the Poisson method,\\
+$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
+accordingly  and the next iteration begins.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[t]
-\framebox{
-\begin{minipage}[t]{6in}
-\begin{center}
-\begin{minipage}[t]{6in}
-\vspace{.05in}
+  \begin{framebox}
+%    \begin{minipage}
+% [t]{6in}
+%\begin{center}
+%\begin{minipage}[t]{6in}
+%\vspace{.05in}
 
@@
 
-\vspace{.1in}
+%\vspace{.1in}
 
@@
 
-\end{minipage}
-\end{center}
-\end{minipage}}
+%\end{minipage}
+%\end{center}
+%\end{minipage}
+\end{framebox}
 \caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
@@
 
-\subsection{Sampling}
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Poisson Sampling}
+\label{sec:poisson}
 
@@
 
-\section{Analysis: finite population setting}
-\label{sec:discrete}
-
-We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on the manner in which $X$ is generated.
-
-\subsection{Accuracy of bias estimates}
-
-Fix the set of balls $\B$ and let $0 < \delta < 1$ be a
-predefined confidence parameter. We start with a uniform guarantee on the
-bias estimates for all balls $B \in \B$.
-
-In Fig.~\ref{alg:sampling}, we see that the points queried in a ball $B$ at level $\ell$ are $\{z \in X_B: T_z \leq \tau_\ell\}$, which from the definition of $\tau_\ell$ has size $O(k)$.
-We use these labels to estimate the bias of $B$.
-Since we need to detect biases of magnitude greater than $\gamma$, we set $k \propto 1/\gamma^2$.
-%This intuition is borne out by the following result, proved in the appendix. 
-\begin{lemma}
-Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. 
-Let $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}.
-Then with probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are $\gamma$-accurate in the sense of Definition~\ref{def:accurate-bias-estimate}.
-\label{lemma:accurate-bias-estimates}
-\end{lemma}
-Henceforth assume $\yh(B)$ is $\gamma$-accurate for all  $B \in \B$.
-
-\subsection{Critical levels}
-
-Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes the possible labels for $x$ given information from balls at levels $\leq \ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$.
-Despite randomness in querying, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ reliably contains the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ reliably omits the wrong label.
-
-\begin{defn}[Critical levels $L_1,L_2$]
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$. Define $L_1(x)$ to be the smallest level $\ell$ such that there exists $B_o \in \B_\ell(x)$ for which
-\begin{itemize}[leftmargin=0.5cm]
-\item $s(x) \cdot \eta_X(B_o) \geq \gamma$, and
-\item $s(x) \cdot \eta_X(B) \geq \gamma$ for any $B \in \B(x)$ with $X_{B} \subset X_{B_o}$.
-\end{itemize}
-Define $L_2(x)$ to be the smallest level $\ell$ such that:
-\begin{itemize}[leftmargin=0.5cm]
-\item For all $B \in \B_{\geq \ell}(x)$, we have $s(x) \cdot \eta_X(B) \geq 0$.
-\item For any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$, there exists $B' \in \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
-\end{itemize}
-Take $L_1(x)$ or $L_2(x)$ to be $\infty$ if no level meets the requirements, or if $\eta(x) = 0$. 
-\label{defn:L12}
-\end{defn}
-
-Once $\ell$ reaches a critical level, $\PL_\ell(x)$ and thus $\yh_\ell(x)$ behave more predictably, as follows.
-\begin{lemma}
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) \in \{+1,-1\}$ denote its Bayes-optimal label. Then for any level $\ell$ and any time at which $\yh_\ell(x) \neq \bot$:
-\begin{enumerate}
-\item[(a)] If $\ell \geq L_1(x)$, then $s(x) \in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), !\}$. 
-\item[(b)] If $\ell \geq L_2(x)$, then $-s(x) \not\in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), 0\}$.
-\end{enumerate}
-\label{lemma:boundary}
-\end{lemma}
-
-\subsection{Boundary sets and label complexity}
-
-A common intuition about active learning is that successive queries gradually constrain the possible locations of the decision boundary.
-Let's consider the state of affairs when all balls at level $\leq \ell - 1$ have been sampled.
-The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points must have $L_2(x) \geq \ell$.
-Focused sampling at level $\ell$ is restricted to balls that contain these points.
-We will think of this region as the \emph{boundary set} at level $\ell$.
-\begin{defn}[Boundary set $\bf \Delta_\ell$]
-For any level $\ell$, define the \emph{boundary set} at level $\ell$ to be
-\begin{equation}
-\Delta_\ell 
-= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B 
-.
-\label{eq:sampling-region}
-\end{equation}
-\end{defn}
-%It can be shown that all focused samples at level $\ell$ must lie in this set.
-\begin{lemma}
-All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
-\label{lemma:focused}
-\end{lemma}
-
-We can now give generic label complexity bounds in terms of $L_1$, $L_2$, and $\Delta_\ell$.
-These take two equivalent forms: a \emph{global} version that specifies what parts of $X$ are correctly labeled after $m$ queries and a \emph{local} version that specifies the number of queries after which a particular $x$ is correctly labeled.
-The global version (Theorem~\ref{thm:label-complexity-0}) is in the Appendix; here is the local version.
-
-\begin{thm}[Local label complexity]
-Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$.
-With probability at least $1-2\delta$, the following holds simultaneously for all $x \in X$.
-Let $L_1(x)$ and $L_2(x)$ be the critical levels for $x$, as in Definition~\ref{defn:L12}.
-If $L_2(x) \leq \lg (n/2k)$, let
-\[
-m_o(x) = 32k \cdot \max\bigg( 2^{L_1(x)}, \ \frac{1}{n} \sum_{\ell=L_1(x)+1}^{L_2(x)} |\Delta_\ell| \, 2^\ell\bigg) .
-\]
-Once the active learner has made $m_o(x)$ queries, $\yh(x)$ is fixed at the Bayes-optimal label $g^*(x)$.
-\label{thm:label-complexity}
-\end{thm}
-Let's interpret this bound.
-For a specific $x \in X$, there is no telling how $\yh_\ell(x)$ will behave at levels $\ell < L_1(x)$.
-But for $\ell \geq L_1(x)$, either $\yh_\ell(x) = \, !$, in which case $x$ is in the uncertainty region, or the label of $x$ is correctly set once and for all, which occurs by level $L_2(x)$.
-
-We will say that $x$ \emph{reaches level $\ell$} when bias estimates have been obtained for all balls in $\B_{\leq \ell}(x)$ and thus $\yh_\ell(x)$ is set.
-Recalling that the algorithm alternates between background and focused queries, we show that $m_o(x)/4$ \emph{background} queries suffice for $x$ to reach level $L_1(x)$ and an additional $m_o(x)/4$ \emph{focused} queries take it to level $L_2(x)$, by which time its label is correct.
-
-In particular, $O(k \cdot 2^\ell)$ background queries give us the biases of all balls in $\B_{\leq \ell}$; this is the first term in $m_o(x)$.
-And there are $O(k \cdot 2^\ell \cdot |\Delta_\ell|/n)$ focused queries at any level $\ell$: the second term.
-
-A footnote: Our algorithm queries points at most once.
-However, in some applications, a point $x$ can be queried repeatedly, each time producing an independent draw from $\eta(x)$.
-If \emph{repeat queries} are permissible, $O(1/\gamma^2)$ copies should be made of each point in $X$ before the algorithm is applied; and in this case, points with $|\eta(x)| \geq \gamma$ will all be correctly labeled, eventually.
-
-\section{One-dimensional examples}
-
-To apply Theorem~\ref{thm:label-complexity}, we need bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each $x \in X$, and on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
-We now derive these in some canonical settings.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=2.5in]{oned-monotonic.pdf}
-\hspace{.5in}
-\includegraphics[width=2.5in]{oned-massart.pdf}
-\end{center}
-\caption{(a) The conditional probability function $\eta$ is monotonically increasing; $\eta(x) = -\gamma, 0, \gamma$ at $x = \lambda_L, \lambda, \lambda_R$, respectively.
-(b) On each subinterval, $\eta$ is either $> \gamma$ or $<-\gamma$.}
-\label{fig:oned}
-\end{figure}
-
-\subsection{Example: monotonic $\eta$}
-
-Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing (Fig.~\ref{fig:oned}(a)).
-Let $\lambda \in (0,1)$ be the point for which $\eta(\lambda) = 0$ and let $\lambda_L, \lambda_R$ be the points for which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$.
-Thus we are not required to label points in the interval $(\lambda_L, \lambda_R)$.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=3in]{oned-monotonic.pdf}
-%\end{center}
-%\caption{The conditional probability function $\eta$ is monotonically increasing. The values $x = %\lambda_L, \lambda, \lambda_R$ have $\eta(x) = -\gamma, 0, \gamma$ respectively.}
-%\label{fig:oned-monotonic}
-%\end{figure}
-
-Let $\B$ consist of all closed intervals. Although there are infinitely many of these, we need only consider those whose endpoints lie in $X$, so effectively $|\B| = O(n^2)$.
-
-The critical levels are easy to bound in this setting.
-Let $n^- = |[0,\lambda_L] \cap X|$  be the number of points to the left of $\lambda_L$ and $n^+ = |[\lambda_R,1] \cap X|$ the number of points to the right of $\lambda_R$.
-Then
-%\begin{align*}
-%L_1(x)
-%&\leq
-%\left\{
-%\begin{array}{ll}
-%\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
-%\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
-%\end{array}
-%\right.
-%\\
-%L_2(x)
-%&\leq
-%\lg (n/r(x))
-%\end{align*}
-$$
-L_1(x)
-\leq
-\left\{
-\begin{array}{ll}
-\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
-\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
-\end{array}
-\right.
-\mbox{\ \ \ \ and \ \ \ }
-L_2(x)
-\leq
-\lg (n/r(x))
-$$
-where $r(x)$ is the number of data points between $x$ and $\lambda$, counting $x$ as well.
-%\begin{itemize}
-%\item $L_1(x) \leq \lg (n/n^+)$ if $x \geq \lambda_R$ and $L_1(x) \leq \lg (n/n^-)$ if %$x \leq \lambda_L$, and
-%\item $L_2(x) \leq \lg (n/r(x))$, 
-%\end{itemize}
-
-A counting argument then shows that the boundary region shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
-\begin{thm}
-Define $r^+ = \min \{r(x): x \in X \cap [\lambda_R,1]\}$ and $r^- = \min \{r(x): x \in X \cap [0,\lambda_L]\}$.
-Pick any $0 < \delta < 1$.
-Suppose we run the algorithm of Fig.~\ref{alg:main} with $k = O((\log (n/\delta))/\gamma^2)$ and that $\min(r^+, r^-) \geq 2k$.
-Take any
-$$ m \geq 64k \cdot \max \left( \frac{n}{\min(n^+,n^-)}, \ 2 \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \right) .$$
-With probability $\geq 1-\delta$, after $m$ queries the algorithm correctly labels all $x \in X$ with $|\eta(x)| \geq \gamma$.
-\label{thm:oned-monotonic}
-\end{thm}
-If $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$, then the label complexity is $O(\log^2 n)$.
-If in addition $r^+, r^- = \Theta(n)$, that is, a constant fraction of the points lie between $\lambda_L, \lambda_R$ and the decision boundary, the label complexity is $O(\log n)$.
-
-\subsection{Example: Massart noise}
-
-In Section~\ref{sec:oned-massart}, we develop another one-dimensional example, in which $\X$ consists of several pieces, such that on each piece $\eta$ is either entirely above $\gamma$ or entirely below $-\gamma$ (Fig.~\ref{fig:oned}(b)).
-Once again, a logarithmic label complexity is obtained.
-
-\section{Analysis: distributional setting}
-\label{sec:continuous}
-
-We now turn to a setting where the points $X$ are drawn from some distribution $\mu$ on a topological space $\X$.
-For the same algorithm, we bound label complexity in terms of properties of $\mu$ and $\eta$.
-
-\subsection{A sufficiently-rich collection of neighborhoods}
-
-Let $\B$ be a finite collection of subsets of $\X$; we spell out a basic property we need it to possess.
-
-\begin{defn}
-We say $\B$ is \emph{$k$-conforming} with respect to $X$ if 
-\begin{itemize}
-\item ({\bf Richness}) For any $x \in X$, there exists $B \in \B_0(x)$.
-Moreover, for any level $0 \leq \ell \leq \lg (n/k)$ and any $B \in \B_\ell(x)$, there exists $B' \in \B_{\ell+1}(x)$ such that $X_{B'} \subset X_B$.
-\item ({\bf Fidelity}) $\mu(B)/2 \leq |X_B|/n \leq 2\mu(B)$ for all $B \in \B$ with $|X_B| \geq k$.
-\end{itemize}
-\label{defn:conformity}
-\end{defn}
-When $\B$ consists of Euclidean balls in $\R^d$, made finite by restricting to $X$, the first property is immediate and the second holds with high probability (Lemma~\ref{lemma:ball-size-bounds}).
-We also envisage scenarios where $\B$ is constructed using $X$, for instance from the cells of multiple spatial partition trees.
-
-Henceforth assume $\B$ satisfies this property.  We ignore $B$ with $|X_B| < k$, thereby restricting attention to levels $0 \leq \ell < \lg (n/k)$.
-
-\subsection{A probabilistic notion of distance}
-
-A crucial step in the analysis is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
-\begin{defn}
-For any $x \in \X$ and $S \subset \X$, define
-$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
-\label{defn:prob-dist}
-\end{defn}
-We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
-\begin{lemma}
-For any $x \in \X$, let $s(x) = \sign(\eta(x))$. If $\dist(x, \X^{-s(x)}) \geq 2k/n$, then
-$L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1$.
-\label{lemma:L2-bound}
-\end{lemma}
-We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
-\begin{defn}
-For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
-\label{defn:boundary}
-\end{defn}
-This includes all $x$ with $\eta(x) = 0$. Next, the $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
-\begin{defn}
-For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
-\label{defn:boundary2}
-\end{defn}
-
-We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
-\begin{lemma}
-For any $0 \leq \ell \leq \lg (2n/k)$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
-\label{lemma:delta-continuous}
-\end{lemma}
-Theorem~\ref{thm:label-complexity} thus still holds with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.
-
-Using these notions, we are able to specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under suitable margin conditions; this is given in Theorem~\ref{thm:label-complexity-specific} in the appendix.
-Here we just give a concrete scenario handled by the latter result; another such scenario, with smoothness and Tsybakov noise conditions, appears in the appendix.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=6in]{COLT/curvature.pdf}
-\end{center}
-\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
-The orange and light blue indicate the boundary sets for negative and positive respectively.
-{\em Right:} examples of the $L_1,L_2$ balls for different points.
-The green $L_1$ circles are the largest monochrome circle that contains the point.
-The purple $L_2$ circles are the union of all circles of half the radius or smaller.
-All of these circles are monochrome.}
-\label{fig:curvature}
-\end{figure}
-
-\subsection{Example: Curvature and Massart noise}
-\label{sec:Massart}
-
-By way of example, consider a case where $\X \subset \R^d$ and $\mu$ admits a density.
-We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.
-By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\citep{BC03}.
-
-We make two key assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
-\begin{itemize}[leftmargin=0.5cm]
-\item (Strong density condition)
-There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
-we have $c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) $, where $\vol(\cdot)$ is $d$-dimensional volume.
-\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
-\end{itemize}
-The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$.
-It is a common notion of curvature~\citep{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
-A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
-\begin{thm}
-Under the two conditions above, there are constants $c_2, c_3$ for which the following holds.
-Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$.
-If the algorithm of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$ queries, it assigns Bayes-optimal labels to all but $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
-\label{thm:massart-dist}
-\end{thm}
-Here the constant $c_2$ includes a term $(1/r_o)^d$.
-The $\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over the usual $1/m^{1/d}$ rate for random querying and was also observed earlier, in a more restricted setting, using specialized algorithms~\citep{CN08}.
-
-\newpage
-
-%\input{active-knn.bbl}
-\bibliography{refs}
-%\bibliographystyle{plain}
-
-\newpage
-\appendix
-
-\section{Technicalities: discrete setting}
-
-\subsection{Large deviation bounds for the discrete setting}
+%%%%%%%%%%%%
+\section{Analysis Details: Discrete setting}
+
+\subsection{Technicalities: discrete setting}
+
+\subsubsection{Large deviation bounds for the discrete setting}
 
@@
 
-\subsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}
+\subsubsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}
 
@@
 
-\section{Technicalities: distributional setting}
+\section{Analysis Details: distributional setting}
 
@@
 
+\section{FORMATTING INSTRUCTIONS}
+
+To prepare a supplementary pdf file, we ask the authors to use \texttt{aistats2024.sty} as a style file and to follow the same formatting instructions as in the main paper.
+The only difference is that the supplementary material must be in a \emph{single-column} format.
+You can use \texttt{supplement.tex} in our starter pack as a starting point, or append the supplementary content to the main paper and split the final PDF into two separate files.
+
+Note that reviewers are under no obligation to examine your supplementary material.
+
+\section{MISSING PROOFS}
+
+The supplementary materials may contain detailed proofs of the results that are missing in the main paper.
+
+\subsection{Proof of Lemma 3}
+
+\textit{In this section, we present the detailed proof of Lemma 3 and then [ ... ]}
+
+\section{ADDITIONAL EXPERIMENTS}
+
+If you have additional experimental results, you may include them in the supplementary materials.
+
+\subsection{The Effect of Regularization Parameter}
+
+\textit{Our algorithm depends on the regularization parameter $\lambda$. Figure 1 below illustrates the effect of this parameter on the performance of our algorithm. As we can see, [ ... ]}
+
+\vfill
+
+\end{document}
+
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: t
+%%% TeX-master: "aistats"
 %%% End:

### 2024-09-05 AISTAT25/aistats.tex
added reference to Poisson sampling

--- 2024-07-31 COLT/colt.tex
@@
 
-\subsection{The algorithm}
-
-The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, has two types of query.
-\emph{Background queries} are random draws from $X$ and correspond to passive learning.
-\emph{Focused queries} are made in the vicinity of ``uncertain'' points and correspond to active learning.
-The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$.
-Focused queries are drawn at random from balls in $\B_\ell$ containing such points.
-
-On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
-
-Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$.
-Then the next iteration begins.
-
-The querying process can be stopped at any time, whereupon labels are assigned as follows:
-\begin{equation}
-\yh(x) = 
-\left\{
-\begin{array}{cl}
-\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
-0 & \mbox{``don't know'', otherwise}
-\end{array}
-\right.
-\label{eq:final-label}
-\end{equation}
-
-\begin{figure}[t]
-\framebox{
-\begin{minipage}[t]{6in}
-\begin{center}
-\begin{minipage}[t]{6in}
-\vspace{.05in}
-
-Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
-
-\vspace{.1in}
-
-Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
-
-\vspace{.1in}
-
-Repeat:
-\begin{itemize}[leftmargin=0.5cm]
-\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
-\begin{itemize}[leftmargin=0.25cm]
-\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
-\end{itemize}
-\item {\tt Background-query} // Fig.~\ref{alg:sampling}
-\item Update labels:
-\begin{itemize}[leftmargin=0.25cm]
-\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
-\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
-\begin{align}
-\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
-             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
-\yh_\ell(x) &= 
-\left\{
-\begin{array}{cl}
-+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
--1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
-0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
-! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
-\end{array}
-\right.
-\label{eq:provisional-label}
-\end{align}
-
-\end{itemize}
-\item Update uncertainty regions:
-\begin{itemize}[leftmargin=0.25cm]
-\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
-\end{itemize}
-\end{itemize}
-
-\end{minipage}
-\end{center}
-\end{minipage}}
-\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
-\label{alg:main}
-\end{figure}
-
-\subsection{Sampling}
-
-The uncertainty region at level $\ell$ consists of the known-unknowns from level $\ell-1$, that is, points $x \in X$ with $\yh_{\ell-1}(x) = \, !$.
-This region is revealed to us piecemeal, a few points at a time.
-To get good label complexity we need to actively query the part of it we know.
-
-We manage this through \emph{Poisson sampling} (Fig.~\ref{alg:sampling}).
-To infer the bias of a ball $B$, we need labels for $k = \tilde{O}(1/\gamma^2)$ random points in it.
-Balls at level $\ell$ have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$ fraction of the ball to be queried.
-Instead, we query every point in the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, \emph{independently}.
-Specifically, we query points $x$ in the ball with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
-
-Once a point is queried, it is never queried again.
-
-\begin{figure}[t]
-\framebox{
-\begin{minipage}[t]{6in}
-
-\vspace{.05in}
-\emph{Initialization:}
-\begin{itemize}[leftmargin=0.5cm]
-\item Set $Q = \emptyset$ (points queried so far)
-\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
-\end{itemize}
-
-\vspace{.05in}
-{\bf Background-query}
-
-\begin{itemize}[leftmargin=0.5cm]
-\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
-\end{itemize}
-
-\vspace{.05in}
-{\bf Focused-query}($\ell, U$)
-
-\begin{itemize}[leftmargin=0.5cm]
-\item Define querying region:
-$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
-\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
-\end{itemize}
-
-\end{minipage}}
-\caption{The two sampling procedures.
-Each $x \in X$ has an associated r.v.
-$T_x$ drawn uniformly from $[0,1]$.
-The smaller $T_x$, the earlier $x$ is likely to be queried.
-\emph{Background queries} are drawn at random from $X$.
-\emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.}
-\label{alg:sampling}
-\end{figure}
-
-\begin{figure}
-\framebox{
-\begin{minipage}[t]{6in}
-\vspace{.05in}
-\begin{itemize}[leftmargin=0.25cm]
-\item Initially $\yh(B) = \bot$
-\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
-$$ \yh(B)
-= 
-\left\{
-\begin{array}{ll}
-\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
-0 & \mbox{otherwise}
-\end{array}
-\right.
-$$
-\end{itemize}
-\end{minipage}}
-\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
-\label{alg:bias-estimate}
-\end{figure}
+\fi
+
+\subsection{The detailed algorithm}
+\label{sec:detailedalgorithm}
+
+The simplified algorithm does addresses the first challange of ``where
+to query'' (see challanges in the introduction). It does not address
+the other two challanges, which the detailed algorithm does address.
+
+\begin{enumerate}
+\item {\bf No smoothness assumptions:}\\
+  As described in the introduction, one can \emph{never} be sure of having
+  correctly determined the label of $x$. Thus in addition to focused
+  (active) querying, we also do background sampling of the entire
+  space to pick up on possible mind changes. For simplicity, we make
+  one background query per focused query.
+
+  A related problem is that having the outer loop iterate over levels
+  will cause the algorithm to miss some uncertainty region if they are
+  appear at a lower level. Therefor the full algorithm does not loop
+  over levels. Instead, at each iteration it finds the lowest level
+  where the uncertainty set is not empty and makes two queries, one
+  focused and one background.
+
+\item {\bf Managing Sampling:}\\
+  As described in the introduction, we would like to use queries that
+  fall in the intersection of two balls for estimating the bias in
+  both balls. However, selecting examples IID at random from the balls
+  can create an undesired dependence between the estimates. This is solved using Poisson sampling, as described in section~\ref{sec:poisson}.
+\end{enumerate}
 
@@
 \bibliography{refs}
-%\bibliographystyle{plain}
-
-\newpage
+\bibliographystyle{plain}
+
+\iffalse
+\documentclass[twoside]{article}
+
+\usepackage{aistats2024}
+\usepackage{enumitem}
+\usepackage{multicol}
+\usepackage{times}
+\usepackage{amssymb}
+\usepackage{graphicx}
+
+\def\R{{\mathbb{R}}}
+\def\pr{{\rm Pr}}
+\def\E{{\mathbb E}}
+\def\X{{\mathcal X}}
+\def\Y{{\mathcal Y}}
+\def\H{{\mathcal H}}
+\def\G{{\mathcal G}}
+\def\B{{\mathcal B}}
+\def\yh{{\widehat{y}}}
+\def\bias{{\rm bias}}
+\def\supp{{\rm supp}}
+\def\dist{{\rm dist}}
+\def\sign{{\rm sign}}
+\def\vol{{\rm vol}}
+\def\PL{{\mbox{\rm PL}}}
+
+\newtheorem{thm}{Theorem}
+\newtheorem{lemma}[thm]{Lemma}
+\newtheorem{cor}[thm]{Corollary}
+%\newtheorem{claim}[thm]{Claim}
+\newtheorem{defn}[thm]{Definition}
+\newtheorem{assump}{Assumption}
+\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}
+
+\newcommand{\comment}[2]{ {\bf #1 :} #2}
+\newcommand{\yoav}[1]{\comment{Yoav}{\em #1}}
+\newcommand{\sanjoy}[1]{\comment{blue}{Sanjoy}{#1}}
+
+\DeclareMathOperator*{\argmax}{arg\,max}
+
+% If your paper is accepted, change the options for the package
+% aistats2022 as follows:
+%
+%\usepackage[accepted]{aistats2024}
+%
+% This option will print headings for the title of your paper and
+% headings for the authors names, plus a copyright note at the end of
+% the first column of the first page.
+
+% If you set papersize explicitly, activate the following three lines:
+%\special{papersize = 8.5in, 11in}
+%\setlength{\pdfpageheight}{11in}
+%\setlength{\pdfpagewidth}{8.5in}
+
+% If you use natbib package, activate the following three lines:
+%\usepackage[round]{natbib}
+%\renewcommand{\bibname}{References}
+%\renewcommand{\bibsection}{\subsubsection*{\bibname}}
+
+% If you use BibTeX in apalike style, activate the following line:
+%\bibliographystyle{apalike}
+
+\begin{document}
+
+% If your paper is accepted and the title of your paper is very long,
+% the style will print as headings an error message. Use the following
+% command to supply a shorter title of your paper so that it can be
+% used as headings.
+%
+%\runningtitle{I use this title instead because the last one was very long}
+
+% If your paper is accepted and the number of authors is large, the
+% style will print as headings an error message. Use the following
+% command to supply a shorter version of the authors names so that
+% they can be used as headings (for example, use only the surnames)
+%
+%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}
+
+% Supplementary material: To improve readability, you must use a single-column format for the supplementary material.
+\fi
+
 \appendix
-
-\section{Technicalities: discrete setting}
-
-\subsection{Large deviation bounds for the discrete setting}
+\onecolumn
+\begin{center}
+\LARGE{Non-parametric active learning without smoothness\\
+Supplementary Materials}
+\end{center}
+\section{Detailed algorithm}
+
+We now describe the detailed active learning algorithm as shown in
+Figs~\ref{alg:main}, \ref{alg:sampling} and~\ref{alg:bias-estimate}.
+The algorithm makes two types of query. \emph{Background queries} are random draws
+from $X$ and correspond to passive learning. \emph{Focused queries}
+are made in the vicinity of ``uncertain'' points and correspond to
+active learning.
+
+Many of the elements of the detailed algorithm have been introduced in the simplified version.
+Specifically $U_\ell,S,\PL_\ell(x),\yh_\ell(x)$, are as defined there.
+
+However, in the detailed algorithm all of the levels are considered at each iteration.
+On each iteration of the main loop (at most) one focused
+query is made as well as a background query. The focused query comes
+from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
+
+Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
+updated using the Poisson method,\\
+$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
+accordingly  and the next iteration begins.
+
+The querying process can be stopped at any time, whereupon labels are assigned as follows:
+\begin{equation}
+\yh(x) = 
+\left\{
+\begin{array}{cl}
+\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
+0 & \mbox{``don't know'', otherwise}
+\end{array}
+\right.
+\label{eq:final-label}
+\end{equation}
+
+\subsubsection{Poisson Sampling}
+\label{sec:poisson}
+
+The uncertainty region at level $\ell$ consists of the known-unknowns
+from level $\ell-1$, that is, points $x \in X$ with
+$\yh_{\ell-1}(x) = \, !$. This region is revealed to us piecemeal, a
+few points at a time. To get good label complexity we need to actively
+query the part of it we know.
+
+We manage this through \emph{Poisson
+  sampling}~\cite{ghosh2002sampling} (Fig.~\ref{alg:sampling}) which
+works as follows. Each instance $x \in X$ is assocciated with a random
+draw from the uniform distribution over $[0,1]$ denoted $T_x$.
+Pooints are sampled in the order of increasing $T_x$ under some
+conditions described below.  Note that restricting the sampled points
+to be from set $A$ and having $T_x \leq \tau$ is the same, in
+expectation to sampling each point in $A$ independently at random with
+probability $\tau$.
+
+To infer the bias of a ball $B$, we need labels for
+$k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$
+have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$
+fraction of the ball to be queried. Instead, we query every point in
+the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$,
+\emph{independently}. Specifically, we query points $x$ in the ball
+with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a
+uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
+
+Once a point is queried, it is never queried again.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h]
+  \begin{framebox}
+%    \begin{minipage}
+% [t]{6in}
+%\begin{center}
+%\begin{minipage}[t]{6in}
+%\vspace{.05in}
+
+Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
+
+%\vspace{.1in}
+
+Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
+
+%\vspace{.1in}
+
+Repeat:
+\begin{itemize}[leftmargin=0.5cm]
+\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
+\begin{itemize}[leftmargin=0.25cm]
+\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
+\end{itemize}
+\item Run {\tt Background-query} // Fig.~\ref{alg:sampling}
+\item Update labels:
+\begin{itemize}[leftmargin=0.25cm]
+\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
+\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
+\begin{align}
+\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
+             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
+\yh_\ell(x) &= 
+\left\{
+\begin{array}{cl}
++1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
+-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
+0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
+! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
+\end{array}
+\right.
+\label{eq:provisional-label}
+\end{align}
+
+\end{itemize}
+\item Update uncertainty regions:
+\begin{itemize}[leftmargin=0.25cm]
+\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
+\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
+\end{itemize}
+\end{itemize}
+
+%\end{minipage}
+%\end{center}
+%\end{minipage}
+\end{framebox}
+\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
+\label{alg:main}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[t]
+\framebox{
+\begin{minipage}[t]{6in}
+
+\vspace{.05in}
+\emph{Initialization:}
+\begin{itemize}[leftmargin=0.5cm]
+\item Set $Q = \emptyset$ (points queried so far)
+\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Background-query}
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Focused-query}($\ell, U$)
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Define querying region:
+$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
+\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\end{minipage}}
+\caption{The two sampling procedures.
+Each $x \in X$ has an associated r.v.
+$T_x$ drawn uniformly from $[0,1]$.
+The smaller $T_x$, the earlier $x$ is likely to be queried.
+\emph{Background queries} are drawn at random from $X$.
+\emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.}
+\label{alg:sampling}
+\end{figure}
+
+\begin{figure}
+\framebox{
+\begin{minipage}[t]{6in}
+\vspace{.05in}
+\begin{itemize}[leftmargin=0.25cm]
+\item Initially $\yh(B) = \bot$
+\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
+$$ \yh(B)
+= 
+\left\{
+\begin{array}{ll}
+\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+0 & \mbox{otherwise}
+\end{array}
+\right.
+$$
+\end{itemize}
+\end{minipage}}
+\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
+\label{alg:bias-estimate}
+\end{figure}
+
+%%%%%%%%%%%%
+\section{Analysis Details: Discrete setting}
+
+\subsection{Technicalities: discrete setting}
+
+\subsubsection{Large deviation bounds for the discrete setting}
 
@@
 
+\section{FORMATTING INSTRUCTIONS}
+
+To prepare a supplementary pdf file, we ask the authors to use \texttt{aistats2024.sty} as a style file and to follow the same formatting instructions as in the main paper.
+The only difference is that the supplementary material must be in a \emph{single-column} format.
+You can use \texttt{supplement.tex} in our starter pack as a starting point, or append the supplementary content to the main paper and split the final PDF into two separate files.
+
+Note that reviewers are under no obligation to examine your supplementary material.
+
+\section{MISSING PROOFS}
+
+The supplementary materials may contain detailed proofs of the results that are missing in the main paper.
+
+\subsection{Proof of Lemma 3}
+
+\textit{In this section, we present the detailed proof of Lemma 3 and then [ ... ]}
+
+\section{ADDITIONAL EXPERIMENTS}
+
+If you have additional experimental results, you may include them in the supplementary materials.
+
+\subsection{The Effect of Regularization Parameter}
+
+\textit{Our algorithm depends on the regularization parameter $\lambda$. Figure 1 below illustrates the effect of this parameter on the performance of our algorithm. As we can see, [ ... ]}
+
+\vfill
+
+\end{document}
+
+\end{document}
+
 %%% Local Variables:

### 2024-09-12 AISTAT25/question.tex
moved simulations out and added figures

--- 2022-08-15 COLT/colt.tex
@@
-\documentclass[anon,12pt]{colt2022} % Anonymized submission
-%\documentclass[final,12pt]{colt2022} % Include author names
+=== current ==
 
-% The following packages will be automatically loaded:
-% amsmath, amssymb, natbib, graphicx, url, algorithm2e
+\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
+\begin{align}
+\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
+             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\
 
-\usepackage{amsfonts}
+=== suggested ===
 
-\def\R{{\mathbb{R}}}
-\def\pr{{\textrm Pr}}
-\def\E{{\mathbb E}}
-\def\X{{\mathcal X}}
-\def\Y{{\mathcal Y}}
-\def\H{{\mathcal H}}
-\def\G{{\mathcal G}}
-\def\B{{\mathcal B}}
-\def\yh{{\widehat{y}}}
-\def\bias{{\rm bias}}
-\def\supp{{\rm supp}}
-\def\sign{{\rm sign}}
-\def\PL{{\mbox{\rm PL}}}
-\newcommand{\predict}{\mathbf{Predict}}
-\newcommand{\learn}{\mathbf{Learn}}
-\newcommand{\D}{{\cal D}}
-
-\newcommand\bigsubset[1][1.19]{%
-   \mathrel{\vcenter{\hbox{\scalebox{#1}{$\subset$}}}}}
-
-\newcommand{\comment}[3]{{\color{#1} {\bf #2 :} #3}}
-%\newcommand{\comment}[3]{}  % suppress comments
-\newcommand{\sanjoy}[1]{\comment{red}{Sanjoy}{#1}}
-\newcommand{\yoav}[1]{\comment{blue}{Yoav}{#1}}
-
-\DeclareMathOperator*{\argmax}{arg\,max}
-
-\title[Active learning using region-based sampling]{Active learning using region-based sampling}
-\usepackage{times}
-% Use \Name{Author Name} to specify the name.
-% If the surname contains spaces, enclose the surname
-% in braces, e.g. \Name{John {Smith Jones}} similarly
-% if the name has a "von" part, e.g \Name{Jane {de Winter}}.
-% If the first letter in the forenames is a diacritic
-% enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}
-
-% Two authors with the same address
-\coltauthor{\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
-\Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
-\addr UC San Diego}
-
-% Three or more authors with the same address:
-% \coltauthor{\Name{Author Name1} \Email{an1@sample.com}\\
-%  \Name{Author Name2} \Email{an2@sample.com}\\
-%  \Name{Author Name3} \Email{an3@sample.com}\\
-%  \addr Address}
-
-% Authors with different addresses:
-%\coltauthor{%
-% \Name{Author Name1} \Email{xyz@sample.com}\\
-% \addr Address 1
-% \AND
-% \Name{Author Name2} \Email{xyz@sample.com}\\
-% \addr Address 2%
-%}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}%
-Abstract goes here.
-\end{abstract}
-
-\begin{keywords}%
-List of keywords
-\end{keywords}
-
-\section{Introduction}
-
-\section{Setup}
-
-We assume that examples are drawn from a fixed but unknown
-distribution $\D$ over $\X \times \Y$, where $\X$ is a metric space and
-$\Y = \{-1,+1\}$~\footnote{In a later section we define the notion of
-  neighborhoods more generally to allow, for example, diadic
-  partitions}
-Given $x \in \X$, the label $y \in \Y$ is drawn
-according to the conditional probability function
-$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x).$$
-The \emph{Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
-
-The task we wish to solve is \emph{finite population annotation}. We
-are given a finite dataset $X \subset \X$ of $n$ points.
-We are given the ability to query the label of any point $x \in X$. The
-answer to the query: $a(x) \in \{-1,+1\}$ is a binary random variable
-drawn according to $\eta(x)$. In particular, multiple queries on the
-same point can recieve different answers.
-
-Our goal is to design an active learning algorithm that converges to the Bayes
-optimal $g^*(x)$ and does so with a much smaller number of queries
-than a passive algorithm.
-
-We want a procedure that chooses the next point to query.
-It will be applied repeatedly and then stopped at an arbitrary time, whereupon labels $y(x)$ must be provided for all $x \in X$, including those that were queried.
-We would like the inferred labels $y(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
-
-Specifically, we will be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
-$$ \sum_{x \in X} {\bf 1}(y(x) \neq g^*(x),\, |\eta(x)| > \gamma) .$$
-The procedure is allowed to fail with probability $\delta$, to account for sampling error.
-
-\section{Active learning algorithm}
-
-\subsection{Sampling regions}
-
-The sampling is organized around a collection $\B$ of measurable subsets of $\X$. These are the atomic sets on which we assess label bias and we call them ``balls''.
-
-\begin{enumerate}
-
-\item[(a)] \emph{Points within a ball.} For any $B \in \B$, let $X_B$ denote the points in $X$ that can be used to assess the bias of $B$.
-This is simply $X_B = X \cap B$ if $\B$ is a predefined set of balls.
-
-\item[(b)] \emph{Levels of sampling.} We group balls by the number of points that they contain: we put $B$ at {\it level} $\ell$ if
-$$ \frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .$$
-Here $\ell$ ranges from $0$ (consisting of balls that contain at least half the data points) to roughly $\lg n$ (containing a single point).
-
-Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$.
-Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of very large balls and $\B_1, \B_2, \ldots$ consisting of successively smaller balls.
-
-\item[(c)] \emph{The neighborhood of a point.} For any
-  $x \in \X$, and level $\ell$.
-  $N(\ell,x)$ is the collection of balls from $\B_\ell$ that
-  contain $x$.
-\item[(d)] We say that the neighborhood $N(x,\ell)$ is {\em covered}
-  if there are $k$ query points in each $B \in N(x,\ell)$.
-
-\end{enumerate}
-
-\subsection{Bias estimates and provisional labels}
-
-The active learning algorithm uses label-queries to assess the bias of balls $B \in \B$.
-In turn, these are used to estimate the labels of individual points.
-All these estimates are organized by sampling level.
-
-\begin{enumerate}
-
-\item[(a)] We take the \emph{bias of a ball} $B \in \B$ to be
-$\eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \}$.
-At any given time, the algorithm assigns $B$ a qualitative bias estimate,
-$$ \yh(B) = 
-\left\{
-\begin{array}{cl}
-+1 & \mbox{significant $+$ bias} \\
--1 & \mbox{significant $-$ bias} \\
-0 & \mbox{no significant bias} \\
-\bot & \mbox{not yet available}
-\end{array}
-\right.
-$$
-The option $\bot$ is used until sufficiently many points in $X_B$ have been labeled: the required number, $k$, is roughly $1/\gamma^2$ (recall that $\gamma$ is the smallest bias that needs to be detected).
-Once $k$ labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter.
-It will satisfy a basic correctness guarantee (Definition~\ref{def:accurate-bias-estimate}) based on large deviation theory.
-
-\item[(b)] \emph{Possible labels for a point.} Pick any $x$ and any level $\ell$.
-Once bias-estimates are available for all balls $B \in \B_\ell(x)$, the possible labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,0,+1\}$, are defined as follows:
-\begin{gather}
-+1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in
-  \B_\ell(x)$ with $\yh(B) = +1$} \notag \\
-0 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = 0$} \notag \\
--1 \in \PL_\ell(x) \Longleftrightarrow \mbox{there exists $B \in \B_\ell(x)$ with $\yh(B) = -1$}
-\label{eq:PL}
-\end{gather}
-The label of point $x$ at level $\ell$ is a value $\yh_\ell(x) \in \{-1,+1,0,!,\bot\}$.
-It is initially $\bot$, but once bias-estimates are available for all $B \in \B_\ell(x)$, it is set to a value in $\{-1,+1,0,!\}$ and remains fixed thereafter:
-\begin{equation}
-\yh_\ell(x) = 
-\left\{
-  \begin{array}{cl}
-+1   & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \notin PL_\ell(x)$} \\
--1   & \mbox{if $+1 \notin \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
-!    & \mbox{if $+1 \in \PL_\ell(x)$ and $-1 \in PL_\ell(x)$} \\
-0    & \mbox{if $\PL_\ell(x)=\{0\}$} \\
-\bot & \mbox{if $\PL_\ell(x)=\emptyset$}
-\end{array}
-\right.
-\label{eq:provisional-label}
-\end{equation}
-A point $x$ for which $\yh_\ell(x)=!$ is said to have a ``conflict at
-level $\ell$''.
-
-\end{enumerate}
-
-\subsection{The algorithm}
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
+  \item For each $x \in X$ and level $\ell \geq 0$:
   \begin{itemize}
-\item{\bf Initialize:} set $\ell=0$, cover all of the balls in $\B_0$.
-\item   $\predict(x)$
-  \begin{itemize}
-  \item Predict with $\yh_\ell(x)$
-  \end{itemize}
+  \item Set $C$ to be the set of balls $B \in \B_{\leq \ell(x)}\}$ for which $\yh(B)$ is available:
     \item
-  $\learn$  Increment sampling level\\
-  Repeat while there is an $x$ that has a conflict at level $0\leq i
-  \leq \ell$ and $N(x,i+1)$ is not fully covered. 
-  \begin{itemize}
-  \item Cover $N(x,i+1)$ using $m$ queries
-  \item Query $m$ points uniformly from the domain (background
-    sampling)
-  \item Update $\yh_{i+1}(y)$ For all $y$.
-  \end{itemize}
-  Let $\ell+1 \to \ell$
-\end{itemize}
-\end{minipage}}
-\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
-\label{alg:main}
-\end{figure}
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
-\begin{itemize}
-\item Initialize uncertainty regions at all levels:
-\begin{itemize}
-\item $U_0 = X$
-\item $U_\ell = \emptyset$ for $\ell \geq 1$
-\end{itemize}
-\item Initialize all neighborhoods to ``not yet sampled'':
-\begin{itemize}
-\item For all $B \in \B$: \ $\yh(B) = \bot$
-\end{itemize}
-\item Initialize all point-labels, at all levels, to ``not yet available'':
-\begin{itemize}
-\item For all $x \in X$ and $\ell \geq 0$: \ $\yh_\ell(x) = \bot$
-\end{itemize}
-%
-%\vspace{.1in}
-\item Repeat (main loop):
-\begin{itemize}
-\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
-\begin{itemize}
-\item Let $\ell$ be the smallest such level
-\item Set query region:
-$$ S =  \bigcup_{x \in U_\ell} \bigcup_{B \in \B_\ell(x)} X_B $$
-\item {\tt Query}($S$) \ (this queries one point from $S$)
-\end{itemize}
-\item Background sampling: {\tt Query}($X$)
-\item For any neighborhood $B$ that is now adequately sampled for the first time: 
-\begin{itemize}
-\item Compute bias estimate $\yh(B) \in \{-1,0,+1\}$
-\end{itemize}
-\item For any $\ell \geq 0$ and $x \in X$ for which $\B_\ell(x)$ is available for the first time:
-\begin{itemize}
-\item Set $\PL_\ell(x)$ and $\yh_{\ell}(x)$ as per equations (\ref{eq:PL}) and (\ref{eq:provisional-label})
-\end{itemize}
-\item Update uncertainty region at all levels:
-\begin{itemize}
-\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
-\end{itemize}
-\end{itemize}
-
-\end{itemize}
-
-\end{minipage}}
-\caption{The active learning algorithm. In each iteration, (at most) one focused query and one background query are made.}
-\label{alg:main}
-\end{figure}
-
-\begin{figure}[h!]
-\framebox{
-\begin{minipage}[t]{5.9in}
-
-\vspace{.1in}
-\emph{Initialization:}
-\begin{itemize}
-\item Set $Q = \emptyset$ (points queried so far)
-\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
-\end{itemize}
-
-\vspace{.2in}
-{\bf Query}($S, \ell$):
-\begin{itemize}
-\item Query all points in $S \setminus Q$ with $T_z \leq 2^\ell k/n$, and add to $Q$
-\end{itemize}
-
-\end{minipage}}
-\caption{The sampling procedure. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried.} 
-\label{alg:sampling}
-\end{figure}
-
-The algorithm of Figure~\ref{alg:main} can be run for any amount of time. Whenever it is stopped, the {\it final label} of any $x \in X$ is
-\begin{equation}
-\yh(x) = 
-\left\{
-\begin{array}{ll}
-\yh_\ell(x) & \mbox{for the largest $\ell < L[x]$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
-\mbox{as queried} & \mbox{if there is no such $\ell$ but $x \in Q$} \\
-? & \mbox{otherwise}
-\end{array}
-\right.
-\label{eq:final-label}
-\end{equation}
-
-\yoav{I don't think there is anything special about a point that has
-  been queried, a single query is not enough, we need $k$ queries to
-  determine the prediction.}
-
-\section{Analysis of algorithm}
-
-\subsection{Accuracy of bias estimates}
-
-We'll start with a uniform guarantee on the bias estimates for all balls $B \in \B$.
-
-Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution $\eta(x) = \E[Y | X=x]$.
-For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value in $B$, that is,
-$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$
-The specific property we need is encapsulated in the following definition.
-\begin{definition}
-For any $B \in \B$ and $0 < b_1 < b_2$, we say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$(b_1,b_2)$-accurate} if the following holds:
-\begin{itemize}
-\item $\yh(B) = +1 \implies \eta_X(B) > b_1$
-\item $\yh(B) = -1 \implies \eta_X(B) < -b_1$
-\item $\yh(B) = 0 \implies |\eta_X(B)| < b_2$
-\end{itemize}
-\label{def:accurate-bias-estimate}
-\end{definition}
-
-In what follows, take $0 < \delta < 1$ to be a predefined constant, and define
-\begin{equation}
-c_1 = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
-\label{eq:c1}
-\end{equation}
-\yoav{So $\B$ is defined relative to $X$?}
-
-For any level $\ell$, define threshold $\tau_\ell$ as follows:
-\begin{equation}
-\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
-\label{eq:threshold-ell}
-\end{equation}
-For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B)
-= \{z \in X_B: T_z \leq \tau_\ell\}$. We base our bias-estimate for
-$B$ on the labels of points in $\Gamma(B)$. The following is a
-consequence of Lemma~\ref{lemma:large-deviation-bounds} in the
-appendix.
-\yoav{I guess $T_z \leq \tau_\ell$ refers to the poisson sampling
-  trick, which is not discussed. I think that initially we should
-  define an oracle that provides samples with the desired properties,
-  and later we can (in an appendix?) describe how the oracle is implemented.}
-\begin{theorem}
-Suppose that $k \geq (8c_1/(b_2-b_1))^2$ where $c_1$ is given in (\ref{eq:c1}). 
-For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
-$$ \yh(B)
-= 
-\left\{
-\begin{array}{ll}
-\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq (b_1 + b_2)/2$} \\
-0 & \mbox{otherwise}
-\end{array}
-\right.
-$$
-With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $(b_1,b_2)$-accurate.
-\label{thm:accurate-bias-estimates}
-\end{theorem}
-In what follows, we will assume that this $(1-\delta)$-probability event holds.
-
-\subsection{Mind changes}
-
-As the level of sampling $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_\ell(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.
-
-Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ at level $\ell$.
-We distinguish two critical levels: $L_0(x)$, a level beyond which the Bayes-optimal label $s = \mbox{sign}(\eta(x))$ is \emph{always} in $\PL_\ell(x)$; and $L_1(x)$, a level beyond which the opposite label, $-s$, is \emph{never} in $\PL_\ell(x)$.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=3.5in]{home-stretch.pdf}
-%\end{center}
-%\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively finer levels.
-Beyond level $L_s(x)$, this $\yh_\ell(x)$ will never be the wrong label, but might be $0$ or $!$.
-Beyond level $L_f(x)$, the correct label will be assigned.}
-%\label{fig:two-levels}
-%\end{figure}
-
-\begin{definition}
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label.
-We define $L_0(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there exists $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$.
-We define $L_1(x)$ to be the smallest integer $\ell \geq 0$ such that for all $\ell' \geq \ell$, there is no $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
-\label{defn:L01}
-\end{definition}
-
-The key properties of these critical levels are spelled out in the following lemma.
-\begin{lemma}
-Pick $x \in X$ with $\eta(x) \neq 0$. Let $g^*(x) \in \{+1,-1\}$ be its Bayes-optimal label. Then:
-\begin{enumerate}
-\item[(a)] For all $\ell \geq L_0(x)$, we must have $g^*(x) \in \PL_\ell(x)$.
-\item[(b)] For all $\ell \geq L_1(x)$, we must have $-g^*(x) \not\in \PL_\ell(x)$.
-\item[(c)] For all $\ell \geq \max(L_0(x), L_1(x))$, we must have $\yh_\ell(x) = g^*(x)$.
-\end{enumerate}
-\label{lemma:boundary}
-\end{lemma}
-\begin{proof}
-This lemma follows from the $(b_1, b_2)$-accuracy of the bias estimates for balls.
-Pick $x \in X$ with $\eta(x) \neq 0$.
-If $\ell \geq L_0(x)$, there is necessarily a ball $B \in \B_\ell(x)$ with $g^*(x) \eta_X(B) > b_2$ and thus $\yh(B) = g^*(x)$; therefore $g^*(x) \in \PL_\ell(x)$.
-If $\ell \geq L_1(x)$, every $B \in B_\ell(x)$ has $g^*(x) \eta_X(B) \geq -b_1$ and thus $\yh(B) \in \{0,g^*(x)\}$; therefore $-g^*(x) \not\in \PL_\ell(x)$.
-Finally, when $\ell \geq \max(L_0(x), L_1(x))$, it follows from (a) and (b) that $\PL_\ell(x) = \{g^*(x)\}$ and thus $\yh_\ell(x) = g^*(x)$.
-\end{proof}
-
-\subsection{Bounding the amount of focused sampling}
-
-We can now characterize the points that are queried as part of focused sampling.
-\begin{lemma}
-For any $\ell \geq 0$, define
-\begin{equation}
-\Delta_{\ell} \ \ \ = \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .
-\label{eq:sampling-region}
-\end{equation}
-Then, all focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
-\label{lemma:focused}
-\end{lemma}
-\begin{proof}
-Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$.
-For any point $x$ with $L_1(x) < \ell$, we have from Lemma~\ref{lemma:boundary}(b) that $\PL_{\ell-1}(x)$ is either $\{\}$ or $\{g^*(x)\}$; either way, $x$ is not in the uncertainty set at level $\ell$.
-Thus $\overline{U}_\ell \subset \{x \in X: L_1(x) \geq \ell\}$.
-
-The focused samples at level $\ell$ lie in
-$$
-\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau(B) \}
-\ \ \ 
-\bigsubset[1.44]
-\ 
-\bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \} .
-$$
-This is precisely $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
-\end{proof}
-
-A final property we will need for label complexity bounds is that various subsets of interest contain roughly the expected number of points at each level.
-\begin{lemma}
-Pick any positive integer $\ell_f$. With probability at least $1-2(\ell_f+1)e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \ell_f$.
-\begin{enumerate}
-\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
-\item[(b)] If $|\Delta_\ell| \geq k$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 |\Delta_\ell| \tau_\ell$.
-\end{enumerate}
-\label{lemma:level-distribution}
-\end{lemma}
-\begin{proof} 
-Pick any subset $S \subset X$ and let $m = |S|$.
-Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$.
-The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m/3}$.
-
-Both parts follow from this principle; and we take a union bound over all $\ell_o + 1$ levels.
-\end{proof}
-
-\subsection{Label complexity bounds}
-
-\yoav{I think this is the main theorem, however, I find it very hard
-  to parse. In particular, I don't see any reference to the fraction
-  of examples that are in the focused group at each level. These
-  fractions have to be small in order for our sampling to be effective.}
-
-\begin{theorem}
-Pick any $0 < \delta < 1$. Suppose that $k \geq (8c_1/(b_2-b_1))^2$, where $c_1$ is defined in (\ref{eq:c1}), and that the active learning algorithm is allowed to make $m$ queries. Let
-$$ \ell_0 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
-and let $\ell_1$ be the largest integer such that
-$$ \sum_{\ell = \ell_0 + 1}^{\ell_1} |\Delta_{\ell}| \, \tau_\ell \ \leq \ \frac{m}{8} .$$
-Then with probability at least $1-2\delta$, any point $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$ will have final label $\yh(x) = g^*(x)$.
-\label{thm:label-complexity}
-\end{theorem}
-\begin{proof}
-Under the setting of $k$, the high-probability events of Theorem~\ref{thm:accurate-bias-estimates} and Lemma~\ref{lemma:level-distribution} hold with probability at least $1-2\delta$.
-We will henceforth take these for granted.
-
-Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}.
-We will analyze the effect of background sampling in phase one and focused sampling in phase two.
-We start with the former.
-
-Of the $m/2$ queries in phase one, at least $m/4$ will be background samples.
-Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.
-Now, for $\ell_0$ as defined, we have from (\ref{eq:threshold-ell}) that $\tau_{\ell_0} \leq m/8n$ and from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy
-$$ T_x \leq \tau_{\ell_0} \leq \frac{m}{8n}.$$
-Therefore all such points are queried in phase one, and thus all balls $B$ at level $\leq \ell_0$ are processed and get bias estimates $\yh(B)$.
-
-It then follows from Lemma~\ref{lemma:boundary} that the following holds for any $x \in X$ with $L_0(x) \leq \ell_0$:
-\begin{enumerate}
-\item[(a)] For any $\ell \geq \ell_0$, if $\yh_\ell(x)$ is defined then it is either $g^*(x)$ or $!$.
-\item[(b)] If $x$ ever leaves $U$ during phase two, then its final label as defined in (\ref{eq:final-label}) is $\yh(x) = g^*(x)$.
-\item[(c)] If $x$ is processed at level $L_1(x)$, then it will leave $U$ for good.
-\end{enumerate}
-
-Now let's move on to phase two.
-Let $A = \{x \in X: L_0(x) \leq \ell_0, L_1(x) \leq \ell_1\}$.
-We will show that every point in $A$ must leave the uncertainty set at some time during phase two.
-From (b), we can conclude that all these points have their final labels set correctly, once and for all.
-
-We will break the argument into three cases.
-
-{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.
-
-{\it Case 2:} A full $m/4$ focused queries are made in phase two, but at least one of them is at level $\ell_1 + 1$.
-In that round, $U$ cannot contain any $x \in A$, since by (c) all such points will leave the uncertainty set by level $\ell_1$.
-
-{\it Case 3:} A full $m/4$ focused queries are made in phase two, all at levels $\leq \ell_1$.
-By the analysis of phase one, none of these queries can be at level $\leq \ell_0$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_0+1$ through $\ell_1$ inclusive is at most
-$$\sum_{\ell=\ell_0+1}^{\ell_1} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
-\ \leq \ 
-\sum_{\ell=\ell_0+1}^{\ell_1} 2 |\Delta_{\ell}| \tau_\ell
-\ \leq \ 
-\frac{m}{4} ,
-$$
-where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus by the end of the phase two, every $x \in A$ will have been processed until it leaves $U$.
-\end{proof}
-
-\section{Label complexity bounds in various settings}
-
-In order to apply Theorem~\ref{thm:label-complexity}, we need two things:
-\begin{itemize}
-\item An upper bound on the critical level $L_1(x)$ for each $x \in X$, which in turn (via Lemma~\ref{lemma:focused}) will yield an upper bound on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
-\item An upper bound on $L_0(x)$ for points $x$ that we expect to be correctly labeled by the end.
-\end{itemize}
-We now illustrate how this works out in several canonical settings.
-
-\subsection{One-dimensional data with Massart noise}
-
-Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
-\begin{itemize}
-\item The Bayes-optimal labeling is determined by $p$ disjoint open intervals $I_1 = (\lambda_0, \lambda_1), I_2 = (\lambda_1, \lambda_2), \ldots, I_p = (\lambda_{p-1}, \lambda_p)$, where
-$$ 0 = \lambda_0 < \lambda_1 < \lambda_2 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$$
-Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.
-\item For each $j$, either $\eta(x) > c$ for all $x \in I_j$ or $\eta(x) < -c$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
-\end{itemize}
-Here $c > 0$ is some constant.
-
-We will take $\B$ to be all open intervals of $[0,1]$, and $\B(x)$ to be intervals containing the point $x$. The accuracy parameters $b_1, b_2$ will be set so that $0 \leq b_1 < b_2 \leq c$.
-\begin{lemma}
-Pick any $x \in I_j$, for $1 \leq j \leq p$.
-\begin{enumerate}
-\item[(a)] $L_0(x) \leq \left\lceil \lg (n/n_j) - 1 \right\rceil$, where $n_j$ is the number of data points in $I_j$.
-\item[(b)] $L_1(x) \leq \left\lceil \lg (n/r(x)) - 1 \right\rceil$, where $r(x)$ is the minimum number of data points between $x$ and a boundary point. 
-\end{enumerate}
-\label{lemma:critical-one-d-massart}
-\end{lemma}
-\begin{proof}
-For (a), define $n_j = |X \cap I_j|$ and notice that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$.
-Moreover, $s(I_j) \cdot \eta_X(I_j) > c \geq b_2$.
-For levels $\ell' > \ell$, we can make a similar argument using a sub-interval $I \subset I_j$ that lies in $\B_{\ell'}(x)$.
-
-For (b), we define $r(x)$ to be either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller.
-For $\ell \geq \lceil (\lg n/r(x)) - 1 \rceil$, any $B \in \B(x)$ will lie entirely within $I_j$ and must therefore have $s(I_j) \cdot \eta_X(B) > 0$.
-\end{proof}
-
-Using these bounds on the critical levels, we find that all but an $\epsilon$ fraction of $X$ will be correctly classified after making $m = O(\frac{(p-1)\log n}{c^2} \log \frac{p-1}{\epsilon})$ queries.
-\begin{theorem}
-In the one-dimensional scenario above, there are absolute constants $C_1, C_2, C_3$ for which the following holds.
-Pick any $0 < \delta < 1$ and take $k \geq (C_1/c^2) \ln (n/\delta)$.
-Then with probability at least $1-\delta$, after making $m$ queries, the active learning algorithm will correctly classify all but $n \cdot 2^{-C_2 m/(k(p-1))}$ points in any target interval $I_j$ with $n_j/n \geq C_3 k/m$.
-\label{thm:one-d-massart}
-\end{theorem}
-
-\begin{proof}
-Using Lemma~\ref{lemma:critical-one-d-massart}(b), we see that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
-\begin{align*}
-\Delta_\ell \ \ \ 
-&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $2n/2^{\ell}$ points on either side of each boundary point, so $|\Delta_\ell| \leq 4(p-1)n/2^{\ell}$.
-
-We then set $b_1 = 0$, $b_2 = c$ and apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
-\begin{align*}
-\ell_0
-&= 
-\lg \frac{m}{32k} \\
-\ell_1
-&=
-\ell_0 + \frac{m}{128 k(p-1)},
-\end{align*}
-and $k \geq (192/c^2)(2 \ln n + \ln (2/\delta))$; in setting $k$, we have used $|\B| = {n \choose 2}$. The theorem statement follows by invoking Lemma~\ref{lemma:critical-one-d-massart}.
-\end{proof}
-
-\subsection{One-dimensional data with monotonic $\eta$}
-
-For our next example, we once again take $X$ to consist of $n$ arbitrarily-placed points in $\X = [0,1]$.
-This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing.
-Let $\lambda \in (0,1)$ denote the point at which $\eta(\lambda) = 0$.
-
-As before, $\B$ will consist of all open intervals, and $\B(x)$ be will all intervals containing $x$.
-Given a threshold $\gamma > 0$ such that we do not need to label points with $|\eta(x)| < \gamma$, we can set accuracy parameters $b_2 = \gamma, b_1 = 0$.
-
-\begin{lemma}
-Let $\lambda_L, \lambda_R$ be the points at which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$.
-\begin{enumerate}
-\item[(a)] Define $n^- = |[0,\lambda_L) \cap X|$ and $n^+ = |(\lambda_R,1] \cap X|$. Then
-$$
-L_0(x)
-\leq
-\left\{
-\begin{array}{ll}
-\lg \lceil (n/n^+) - 1 \rceil & \mbox{if $x > \lambda_R$} \\
-\lg \lceil (n/n^-) -1 \rceil & \mbox{if $x < \lambda_L$}
-\end{array}
-\right.
-$$
-\item[(b)] For any $x \in X$, let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$. Then
-$L_1(x) \leq \left\lceil \lg (n/r(x)) -1 \right\rceil$.
-\end{enumerate}
-\label{lemma:critical-one-d-monotonic}
-\end{lemma}
-
-\begin{proof}
-For (a), take $x > \lambda_R$, for instance.
-The interval $B = (\lambda,1)$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta(B) > b_2$.
-For higher levels $\ell$, we can take a sub-interval of $(\lambda, 1)$ containing $x$.
-
-For (b), we have $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$.
-For any $\ell \geq \lceil \lg (n/r(x)) - 1 \rceil$, no $B \in \B_\ell(x)$ extends to the other side of the boundary; thus $\eta_X(B) \geq 0$.
-\end{proof}
-
-In this setting, we are not required to label points in the interval $[\lambda_L, \lambda_R]$, and thus we don't need $L_0(x)$ values for them.
-However, we do need to bound $L_1(x)$ for \emph{all} points in order to control the size of the sampling region.
-
-With Lemma~\ref{lemma:critical-one-d-monotonic}, the rest of the argument follows the pattern of the previous one-dimensional case.
-
-\begin{theorem}
-There are absolute constants $C_1, C_2, C_3$ for which the following holds.
-Pick any $0 < \delta < 1$ and take $k \geq (C_1/\lambda^2) \ln (n/\delta)$.
-Then with probability at least $1-\delta$, after making $m \geq C_2 k \max(n/n^+, n/n^-)$ queries:
-\begin{enumerate}
-\item[(a)] The active learning algorithm will correctly label all but $n \cdot 2^{-C_3 m/k}$ points in $[0,\lambda_L) \cup (\lambda_R,1]$. 
-\item[(b)] If $n_o = \min(|X \cap [\lambda_L,\lambda)|, |X \cap (\lambda,\lambda_R]|) > 0$, then all points in $[0,\lambda_L) \cup (\lambda_R,1]$ will be correctly labeled after $C_3 k \lg (n/n_o)$ queries.
-\end{enumerate}
-\label{thm:one-d-monotonic}
-\end{theorem}
-
-\begin{proof}
-From Lemma~\ref{lemma:critical-one-d-monotonic}(b), we have that the sampling region at level $\ell$, as defined in (\ref{eq:sampling-region}), is
-\begin{align*}
-\Delta_\ell \ \ \ 
-&= \bigcup_{x \in X: L_1(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
-&\bigsubset[1.44] \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
-.
-\end{align*}
-This includes at most $2n/2^{\ell}$ points on either side of the boundary, whereupon $|\Delta_\ell| \leq 4n/2^{\ell}$.
-
-By Theorem~\ref{thm:label-complexity}, making $m$ queries will correctly classify all points with $L_0(x) \leq \ell_0$ and $L_1(x) \leq \ell_1$, for
-\begin{align*}
-\ell_0
-&= 
-\lg \frac{m}{32k} \\
-\ell_1
-&=
-\ell_0 + \frac{m}{128 k},
-\end{align*}
-and $k \geq (192/\lambda^2)(2 \ln n + \ln (2/\delta))$. If $m \geq 32k \cdot \max(n/n^+, n/n^-)$, this includes all points in $[0,\lambda_L) \cup (\lambda_R,1]$ with $r(x) \geq n \cdot 2^{-m/128k}$. 
-\end{proof}
-
-\subsection{Two-dimensional data with linear boundary}
-
-\subsection{Higher dimension, with dyadic partitioning}
-
-\subsection{Higher dimension, with smoothness conditions}
-
-% Acknowledgments---Will not appear in anonymized version
-\acks{We thank a bunch of people and funding agency.}
-
-\bibliography{yourbibfile}
-
-\vfill
-
-\pagebreak
-
-\appendix
-
-\section{Technicalities}
-
-\subsection{Large deviation bounds for the discrete setting}
-
-\begin{lemma}
-Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.
-
-Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
-$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
-Now consider the following estimator $Z$ of this quantity:
-\begin{itemize}
-\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
-\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
-\end{itemize}
-If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
-\begin{enumerate}
-\item[(a)] $N \geq k$, and 
-\item[(b)] $| Z - \eta_o| \leq \sqrt{(48/k) \ln (4/\delta)}$.
-\end{enumerate}
-\label{lemma:large-deviation-discrete}
-\end{lemma}
-\begin{proof}
-Let's start with (a).
-Define $c_1 = \sqrt{3 \ln (4/\delta)}$.
-We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c_1 \sqrt{2k}$ since this is the worst case.
-By assumption, $k \geq 2c_1^2$ and thus $qm \leq 2k$.
-
-Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
-\begin{align*}
-\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
-\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
-\end{align*}
-Take $\epsilon = c_1/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c_1$, we see that with probability at least $1-\delta/2$, we get 
-$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
-The lower bound implies $N > qm(1-\epsilon) = qm - c_1\sqrt{qm} = k + c_1 \sqrt{2k} - c_1 \sqrt{qm} \geq k$.
-
-For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$.
-The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$.
-Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$.
-Thus their sum $S$ has expectation
-$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
-and variance
-$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
-We can bound the concentration of $S$ around its expected value using Bernstein's inequality,
-$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + 2t/3)} \right) .$$
-Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.
-
-Combining with the high-probability bound on $N$ above, we get
-$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
-whereupon (recalling $Z = S/N$)
-$$ 
-\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
-Since $|\eta_o| \leq 1$,
-$$ |Z - \eta_o | \leq \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
-\leq 
-4 \epsilon,$$
-as claimed.  
-\end{proof}
-
-An immediate consequence is the following large deviation bound.
-Recall that for any ball $B \in \B_\ell$, its query-set $\Gamma(B)$ is $\{z \in X_B: T_z \leq \tau_\ell\}$, where $\tau_\ell$ is defined in (\ref{eq:threshold-ell}).
-\begin{lemma}
-Suppose that the constant $c_1$ is set as in (\ref{eq:c1}) and that $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
-\begin{enumerate}
-\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
-\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
-$\left| \widehat{\eta}(B) - \eta_X(B) \right| \leq 4c_1/\sqrt{k}$.
-\end{enumerate}
-\label{lemma:large-deviation-bounds}
-\end{lemma}
-\begin{proof}
-These follow by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
-\end{proof}
-
-\end{document}
+    \begin{align}
+    \PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in C$ s.t. $\yh(B) = s$ and} \notag \\ 
+             &                       \hspace{1.5in}\mbox{no $B' \in C$ such that  $X_{B'} \subset X_B$ and $\yh(B')=-s$} \} \label{eq:PL} \\

### 2024-09-12 AISTAT25/aistats.tex
updated figures
Made some changes to experimental section

@@
 \begin{center}
-\includegraphics[width=3in]{ideas.pdf}
+\includegraphics[width=3in]{figures/ideas.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=2.5in]{oned-monotonic.pdf}
+  \includegraphics[width=2.5in]{figures/Monotone.png}
 %\hspace{.5in}
-\includegraphics[width=2.5in]{oned-massart.pdf}
+\includegraphics[width=2.5in]{figures/Massart.png}
 \end{center}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=3in]{figures/MassartExperiment.png}
+\end{center}
+\caption{{\bf Experimental evidence:} This figure compares the performance of passive sampling alone to that of the active learning algorithm.
+The X axis represents the number of queries made and the Y axis represents the difference between the error achieved and the optimal achievable error.
+}
+\label{fig:Experiments}
+\end{figure}
+
+\section{Experiments}
+
+We provide experimental evidence that our algorithm works.
+We chose a simple 1D problem, described in figure~\ref{}.
+Although this problem is simple, it is outside the scope of algorithms that assume that the conditional probability is smooth.
+
+\section{Future Directions}
+
 %\input{active-knn.bbl}

### 2024-09-12 AISTAT25/aistats.tex
Removed hyphen in title
experimental section

@@
 
-\aistatstitle{Non-parametric active learning without smoothness}
+\aistatstitle{Nonparametric active learning without smoothness}
 
@@
 \end{center}
-\caption{{\bf Experimental evidence:} This figure compares the performance of passive sampling alone to that of the active learning algorithm.
-The X axis represents the number of queries made and the Y axis represents the difference between the error achieved and the optimal achievable error.
-}
+\caption{{\bf Experimental evidence:} This figure compares the
+  performance of passive sampling alone to that of the active learning
+  algorithm. The X axis represents the number of queries made and the
+  Y axis represents the difference between the error achieved and the
+  optimal achievable error.}
 \label{fig:Experiments}
@@
 
-We provide experimental evidence that our algorithm works.
-We chose a simple 1D problem, described in figure~\ref{}.
-Although this problem is simple, it is outside the scope of algorithms that assume that the conditional probability is smooth.
+We provide some experimental evidence that our algorithm is effective.
+We chose a simple 1D problem, described in figure~\ref{fig:oned}(b)
+(Massard Noise).  Although this problem is simple, it is outside the
+scope of algorithms that assume that the conditional probability is a
+smooth function of $x$. We used a sample of 2000 points and ran our
+algorithm 10 times. In figure~\ref{fig:experiments} we present the
+average error as a function of the number of queries. We compare the
+performance of our active algorithm (red line) and the performance of the passive algorithm which is the
+same algorithm where focused queries are disabled and only backgroud samples are used (blue line).
+
+Observe that the error of the active algorithm decreases rapidly to
+until iteration 600, while it takes 1,500 queries for the passive
+algorithm to reach the same error. This almost three-fold improvement
+demonstrates the effectiveness of our algorithm.
 

### 2024-09-25 AISTAT25/aistats.tex
added new experiments and reorganized

@@
 
-\begin{figure*}[t]
+\begin{figure*}[t]{6in}
 \framebox{
-\begin{minipage}[t]{6in}  % Adjusted minipage width slightly
+\begin{minipage}[t]{6in}
 \begin{center}
@@
   \begin{align}
-    \PL_\ell(x) &= \{s \in \{-1, +1\}: \ \text{there exists } B \in B_\ell(x) \text{ such that } \yh(B) = s \} \\ 
+    \PL_\ell(x) &= \{s \in \{-1, +1\}: \\
+               & \text{s.t. there exists } B \in B_\ell(x) \text{ s.t. } \yh(B) = s \} \\ 
     \yh_\ell(x) &= 
@@
 \end{enumerate}
+
+\section{Detailed algorithm}
+
+We now describe the detailed active learning algorithm as shown in
+Figs~\ref{alg:main}, \ref{alg:sampling} and~\ref{alg:bias-estimate}.
+The algorithm makes two types of query. \emph{Background queries} are random draws
+from $X$ and correspond to passive learning. \emph{Focused queries}
+are made in the vicinity of ``uncertain'' points and correspond to
+active learning.
+
+Many of the elements of the detailed algorithm have been introduced in the simplified version.
+Specifically $U_\ell,S,\PL_\ell(x),\yh_\ell(x)$, are as defined there.
+
+However, in the detailed algorithm all of the levels are considered at each iteration.
+On each iteration of the main loop (at most) one focused
+query is made as well as a background query. The focused query comes
+from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
+
+Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
+updated using the Poisson method,\\
+$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
+accordingly  and the next iteration begins.
+
+The querying process can be stopped at any time, whereupon labels are assigned as follows:
+\begin{equation}
+\yh(x) = 
+\left\{
+\begin{array}{cl}
+  \yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$,1}\\
+  & \mbox{if such an $\ell$ exists} \\
+0 & \mbox{``don't know'', otherwise}
+\end{array}
+\right.
+\label{eq:final-label}
+\end{equation}
+
+\subsubsection{Poisson Sampling}
+\label{sec:poisson}
+
+The uncertainty region at level $\ell$ consists of the known-unknowns
+from level $\ell-1$, that is, points $x \in X$ with
+$\yh_{\ell-1}(x) = \, !$. This region is revealed to us piecemeal, a
+few points at a time. To get good label complexity we need to actively
+query the part of it we know.
+
+We manage this through \emph{Poisson
+  sampling}~\cite{ghosh2002sampling} (Fig.~\ref{alg:sampling}) which
+works as follows. Each instance $x \in X$ is assocciated with a random
+draw from the uniform distribution over $[0,1]$ denoted $T_x$.
+Pooints are sampled in the order of increasing $T_x$ under some
+conditions described below.  Note that restricting the sampled points
+to be from set $A$ and having $T_x \leq \tau$ is the same, in
+expectation to sampling each point in $A$ independently at random with
+probability $\tau$.
+
+To infer the bias of a ball $B$, we need labels for
+$k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$
+have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$
+fraction of the ball to be queried. Instead, we query every point in
+the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$,
+\emph{independently}. Specifically, we query points $x$ in the ball
+with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a
+uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
+
+Once a point is queried, it is never queried again.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h]
+\framebox{
+\begin{minipage}[t]{3in}
+
+Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
+
+%\vspace{.1in}
+
+Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
+
+%\vspace{.1in}
+
+Repeat:
+\begin{itemize}[leftmargin=0.5cm]
+\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
+\begin{itemize}[leftmargin=0.25cm]
+\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
+\end{itemize}
+\item Run {\tt Background-query} // Fig.~\ref{alg:sampling}
+\item Update labels:
+\begin{itemize}[leftmargin=0.25cm]
+\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
+\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
+\begin{align}
+  \PL_\ell(x) &= \{s \in \{-1, +1\}: \\
+             & \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
+             & \mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
+\yh_\ell(x) &= 
+\left\{
+\begin{array}{cl}
++1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
+-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
+0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
+! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
+\end{array}
+\right.
+\label{eq:provisional-label}
+\end{align}
+
+\end{itemize}
+\item Update uncertainty regions:
+\begin{itemize}[leftmargin=0.25cm]
+\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
+\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
+\end{itemize}
+\end{itemize}
+
+\end{minipage}
+%\end{center}
+%\end{minipage}
+}
+\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
+\label{alg:main}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[t]
+\framebox{
+\begin{minipage}[t]{3in}
+
+\vspace{.05in}
+\emph{Initialization:}
+\begin{itemize}[leftmargin=0.5cm]
+\item Set $Q = \emptyset$ (points queried so far)
+\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Background-query}
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\vspace{.05in}
+{\bf Focused-query}($\ell, U$)
+
+\begin{itemize}[leftmargin=0.5cm]
+\item Define querying region:
+$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
+\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
+\end{itemize}
+
+\end{minipage}}
+\caption{The two sampling procedures.
+Each $x \in X$ has an associated r.v.
+$T_x$ drawn uniformly from $[0,1]$.
+The smaller $T_x$, the earlier $x$ is likely to be queried.
+\emph{Background queries} are drawn at random from $X$.
+\emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.}
+\label{alg:sampling}
+\end{figure}
+
+\begin{figure}
+\framebox{
+\begin{minipage}[t]{3in}
+\vspace{.05in}
+\begin{itemize}[leftmargin=0.25cm]
+\item Initially $\yh(B) = \bot$
+\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
+$$ \yh(B)
+= 
+\left\{
+\begin{array}{ll}
+\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
+0 & \mbox{otherwise}
+\end{array}
+\right.
+$$
+\end{itemize}
+\end{minipage}}
+\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
+\label{alg:bias-estimate}
+\end{figure}
 
@@
 
-\section{Analysis: distributional setting}
-\label{sec:continuous}
-
-We now turn to a setting where the points $X$ are drawn from some distribution $\mu$ on a topological space $\X$.
-For the same algorithm, we bound label complexity in terms of properties of $\mu$ and $\eta$.
-
-\subsection{A sufficiently-rich collection of balls}
-
-Let $\B$ be a finite collection of subsets of $\X$; we spell out a basic property we need it to possess.
-
-\begin{defn}
-We say $\B$ is \emph{$k$-conforming} with respect to $X$ if 
-\begin{itemize}
-\item ({\bf Richness}) For any $x \in X$, there exists $B \in \B_0(x)$.
-Moreover, for any level $0 \leq \ell \leq \lg (n/k)$ and any $B \in \B_\ell(x)$, there exists $B' \in \B_{\ell+1}(x)$ such that $X_{B'} \subset X_B$.
-\item ({\bf Fidelity}) $\mu(B)/2 \leq |X_B|/n \leq 2\mu(B)$ for all $B \in \B$ with $|X_B| \geq k$.
-\end{itemize}
-\label{defn:conformity}
-\end{defn}
-When $\B$ consists of Euclidean balls in $\R^d$, made finite by restricting to $X$, the first property is immediate and the second holds with high probability (Lemma~\ref{lemma:ball-size-bounds}).
-We also envisage scenarios where $\B$ is constructed using $X$, for instance from the cells of multiple spatial partition trees.
-
-Henceforth assume $\B$ satisfies this property.  We ignore $B$ with $|X_B| < k$, thereby restricting attention to levels $0 \leq \ell < \lg (n/k)$.
-
-\subsection{A probabilistic notion of distance}
-
-A crucial step in the analysis is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
-\begin{defn}
-For any $x \in \X$ and $S \subset \X$, define
-$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
-\label{defn:prob-dist}
-\end{defn}
-We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
-\begin{lemma}
-For any $x \in \X$, let $s(x) = \sign(\eta(x))$. If $\dist(x, \X^{-s(x)}) \geq 2k/n$, then
-$L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1$.
-\label{lemma:L2-bound}
-\end{lemma}
-We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
-\begin{defn}
-For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
-\label{defn:boundary}
-\end{defn}
-This includes all $x$ with $\eta(x) = 0$. Next, the $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
-\begin{defn}
-For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
-\label{defn:boundary2}
-\end{defn}
-
-We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
-\begin{lemma}
-For any $0 \leq \ell \leq \lg (2n/k)$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
-\label{lemma:delta-continuous}
-\end{lemma}
-Theorem~\ref{thm:label-complexity} thus still holds with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.
-
-Using these notions, we are able to specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under suitable margin conditions; this is given in Theorem~\ref{thm:label-complexity-specific} in the appendix.
-Here we just give a concrete scenario handled by the latter result; another such scenario, with smoothness and Tsybakov noise conditions, appears in the appendix.
-
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=3in]{curvature.pdf}
-%\end{center}
-%\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
-The orange and light blue indicate the boundary sets for negative and positive respectively.
-{\em Right:} examples of the $L_1,L_2$ balls for different points.
-The green $L_1$ circles are the largest monochrome circle that contains the point.
-The purple $L_2$ circles are the union of all circles of half the radius or smaller.
-All of these circles are monochrome.}
-%\label{fig:curvature}
-%\end{figure}
-
-\subsection{Example: Curvature and Massart noise}
-\label{sec:Massart}
-
-By way of example, consider a case where $\X \subset \R^d$ and $\mu$ admits a density.
-We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.
-By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\cite{BC03}.
-
-We make two key assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
-\begin{itemize}[leftmargin=0.5cm]
-\item (Strong density condition)
-There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
-we have $c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) $, where $\vol(\cdot)$ is $d$-dimensional volume.
-\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
-\end{itemize}
-The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$.
-It is a common notion of curvature~\cite{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
-A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
-\begin{thm}
-  Under the two conditions above, there are constants $c_2, c_3$ for
-  which the following holds. Pick $0 < \delta < 1$ and take
-  $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm
-  of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$
-  queries, it assigns Bayes-optimal labels to all but
-  $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
-\label{thm:massart-dist}
-\end{thm}
-
- Here the constant $c_2$ includes a term $(1/r_o)^d$. The
-$\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over
-the usual $1/m^{1/d}$ rate for random querying and was also observed
-earlier, in a more restricted setting, using specialized
-algorithms~\cite{CN08}.
-
 \begin{figure}
 \begin{center}
-\includegraphics[width=3in]{figures/MassartExperiment.png}
+\includegraphics[width=3in]{figures/MassartExperiment5.png}
 \end{center}
@@
   performance of passive sampling alone to that of the active learning
-  algorithm. The X axis represents the number of queries made and the
+  algorithm. The X axis corresponds to the number of queries made and the
   Y axis represents the difference between the error achieved and the
@@
 
-We provide some experimental evidence that our algorithm is effective.
+In this section we provide experimental evidence that our algorithm has low query complexity.
 We chose a simple 1D problem, described in figure~\ref{fig:oned}(b)
@@
 scope of algorithms that assume that the conditional probability is a
-smooth function of $x$. We used a sample of 2000 points and ran our
-algorithm 10 times. In figure~\ref{fig:experiments} we present the
+smooth function of $x$. We used a sample of 10,000 points and ran our
+algorithm 20 times. In figure~\ref{fig:experiments} we present the
 average error as a function of the number of queries. We compare the
@@
 
-Observe that the error of the active algorithm decreases rapidly to
-until iteration 600, while it takes 1,500 queries for the passive
-algorithm to reach the same error. This almost three-fold improvement
-demonstrates the effectiveness of our algorithm.
+Number of queries as a function of the accuracy for passive and active algorithms.
+\begin{tabular}{|c|rr|c|}
+\hline
+err-opt &  Active&  Passive \\
+\hline
+0.02  &    320 &     517 \\
+0.01 &     429 &    1035 \\
+0.005 &678 &    2045 & \\
+\hline
+\end{tabular}
 
 \section{Future Directions}
+Our algorithm is generic, while we used euclidean balls for clarity,
+any sets can be used. While this generality is attractive, it leaves
+open the question of what are good ways to define the collection of
+regions. Another aspect of this generality is that the computational
+complexity of the algorithm is high. Future work will be directed
+towards characterizing region sets that give both high query
+efficiency and low computational complexity.
 
@@
 \end{center}
-\section{Detailed algorithm}
-
-We now describe the detailed active learning algorithm as shown in
-Figs~\ref{alg:main}, \ref{alg:sampling} and~\ref{alg:bias-estimate}.
-The algorithm makes two types of query. \emph{Background queries} are random draws
-from $X$ and correspond to passive learning. \emph{Focused queries}
-are made in the vicinity of ``uncertain'' points and correspond to
-active learning.
-
-Many of the elements of the detailed algorithm have been introduced in the simplified version.
-Specifically $U_\ell,S,\PL_\ell(x),\yh_\ell(x)$, are as defined there.
-
-However, in the detailed algorithm all of the levels are considered at each iteration.
-On each iteration of the main loop (at most) one focused
-query is made as well as a background query. The focused query comes
-from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.
-
-Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
-updated using the Poisson method,\\
-$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
-accordingly  and the next iteration begins.
-
-The querying process can be stopped at any time, whereupon labels are assigned as follows:
-\begin{equation}
-\yh(x) = 
-\left\{
-\begin{array}{cl}
-\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
-0 & \mbox{``don't know'', otherwise}
-\end{array}
-\right.
-\label{eq:final-label}
-\end{equation}
-
-\subsubsection{Poisson Sampling}
-\label{sec:poisson}
-
-The uncertainty region at level $\ell$ consists of the known-unknowns
-from level $\ell-1$, that is, points $x \in X$ with
-$\yh_{\ell-1}(x) = \, !$. This region is revealed to us piecemeal, a
-few points at a time. To get good label complexity we need to actively
-query the part of it we know.
-
-We manage this through \emph{Poisson
-  sampling}~\cite{ghosh2002sampling} (Fig.~\ref{alg:sampling}) which
-works as follows. Each instance $x \in X$ is assocciated with a random
-draw from the uniform distribution over $[0,1]$ denoted $T_x$.
-Pooints are sampled in the order of increasing $T_x$ under some
-conditions described below.  Note that restricting the sampled points
-to be from set $A$ and having $T_x \leq \tau$ is the same, in
-expectation to sampling each point in $A$ independently at random with
-probability $\tau$.
-
-To infer the bias of a ball $B$, we need labels for
-$k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$
-have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$
-fraction of the ball to be queried. Instead, we query every point in
-the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$,
-\emph{independently}. Specifically, we query points $x$ in the ball
-with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a
-uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
-
-Once a point is queried, it is never queried again.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h]
-  \begin{framebox}
-%    \begin{minipage}
-% [t]{6in}
-%\begin{center}
-%\begin{minipage}[t]{6in}
-%\vspace{.05in}
-
-Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
-
-%\vspace{.1in}
-
-Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
-
-%\vspace{.1in}
-
-Repeat:
-\begin{itemize}[leftmargin=0.5cm]
-\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
-\begin{itemize}[leftmargin=0.25cm]
-\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
-\end{itemize}
-\item Run {\tt Background-query} // Fig.~\ref{alg:sampling}
-\item Update labels:
-\begin{itemize}[leftmargin=0.25cm]
-\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
-\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
-\begin{align}
-\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
-             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
-\yh_\ell(x) &= 
-\left\{
-\begin{array}{cl}
-+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
--1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
-0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
-! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
-\end{array}
-\right.
-\label{eq:provisional-label}
-\end{align}
-
-\end{itemize}
-\item Update uncertainty regions:
-\begin{itemize}[leftmargin=0.25cm]
-\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
-\end{itemize}
-\end{itemize}
-
-%\end{minipage}
-%\end{center}
-%\end{minipage}
-\end{framebox}
-\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
-\label{alg:main}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[t]
-\framebox{
-\begin{minipage}[t]{6in}
-
-\vspace{.05in}
-\emph{Initialization:}
-\begin{itemize}[leftmargin=0.5cm]
-\item Set $Q = \emptyset$ (points queried so far)
-\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
-\end{itemize}
-
-\vspace{.05in}
-{\bf Background-query}
-
-\begin{itemize}[leftmargin=0.5cm]
-\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
-\end{itemize}
-
-\vspace{.05in}
-{\bf Focused-query}($\ell, U$)
-
-\begin{itemize}[leftmargin=0.5cm]
-\item Define querying region:
-$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
-\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
-\end{itemize}
-
-\end{minipage}}
-\caption{The two sampling procedures.
-Each $x \in X$ has an associated r.v.
-$T_x$ drawn uniformly from $[0,1]$.
-The smaller $T_x$, the earlier $x$ is likely to be queried.
-\emph{Background queries} are drawn at random from $X$.
-\emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.}
-\label{alg:sampling}
-\end{figure}
-
-\begin{figure}
-\framebox{
-\begin{minipage}[t]{6in}
-\vspace{.05in}
-\begin{itemize}[leftmargin=0.25cm]
-\item Initially $\yh(B) = \bot$
-\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
-$$ \yh(B)
-= 
-\left\{
-\begin{array}{ll}
-\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
-0 & \mbox{otherwise}
-\end{array}
-\right.
-$$
-\end{itemize}
-\end{minipage}}
-\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
-\label{alg:bias-estimate}
-\end{figure}
 
@@
 This qualification would be removed if were allowed to make multiple queries to each point, because in that case we would include $O(k)$ copies of each point, as explained earlier.
+
+\section{Analysis: distributional setting}
+\label{sec:continuous}
+
+We now turn to a setting where the points $X$ are drawn from some distribution $\mu$ on a topological space $\X$.
+For the same algorithm, we bound label complexity in terms of properties of $\mu$ and $\eta$.
+
+\subsection{A sufficiently-rich collection of balls}
+
+Let $\B$ be a finite collection of subsets of $\X$; we spell out a basic property we need it to possess.
+
+\begin{defn}
+We say $\B$ is \emph{$k$-conforming} with respect to $X$ if 
+\begin{itemize}
+\item ({\bf Richness}) For any $x \in X$, there exists $B \in \B_0(x)$.
+Moreover, for any level $0 \leq \ell \leq \lg (n/k)$ and any $B \in \B_\ell(x)$, there exists $B' \in \B_{\ell+1}(x)$ such that $X_{B'} \subset X_B$.
+\item ({\bf Fidelity}) $\mu(B)/2 \leq |X_B|/n \leq 2\mu(B)$ for all $B \in \B$ with $|X_B| \geq k$.
+\end{itemize}
+\label{defn:conformity}
+\end{defn}
+When $\B$ consists of Euclidean balls in $\R^d$, made finite by restricting to $X$, the first property is immediate and the second holds with high probability (Lemma~\ref{lemma:ball-size-bounds}).
+We also envisage scenarios where $\B$ is constructed using $X$, for instance from the cells of multiple spatial partition trees.
+
+Henceforth assume $\B$ satisfies this property.  We ignore $B$ with $|X_B| < k$, thereby restricting attention to levels $0 \leq \ell < \lg (n/k)$.
+
+\subsection{A probabilistic notion of distance}
+
+A crucial step in the analysis is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
+\begin{defn}
+For any $x \in \X$ and $S \subset \X$, define
+$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
+\label{defn:prob-dist}
+\end{defn}
+We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
+\begin{lemma}
+For any $x \in \X$, let $s(x) = \sign(\eta(x))$. If $\dist(x, \X^{-s(x)}) \geq 2k/n$, then
+$L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1$.
+\label{lemma:L2-bound}
+\end{lemma}
+We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
+\begin{defn}
+For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
+\label{defn:boundary}
+\end{defn}
+This includes all $x$ with $\eta(x) = 0$. Next, the $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
+\begin{defn}
+For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
+\label{defn:boundary2}
+\end{defn}
+
+We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
+\begin{lemma}
+For any $0 \leq \ell \leq \lg (2n/k)$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
+\label{lemma:delta-continuous}
+\end{lemma}
+Theorem~\ref{thm:label-complexity} thus still holds with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.
+
+Using these notions, we are able to specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under suitable margin conditions; this is given in Theorem~\ref{thm:label-complexity-specific} in the appendix.
+Here we just give a concrete scenario handled by the latter result; another such scenario, with smoothness and Tsybakov noise conditions, appears in the appendix.
+
+%\begin{figure}[t]
+%\begin{center}
+%\includegraphics[width=3in]{curvature.pdf}
+%\end{center}
+%\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
+The orange and light blue indicate the boundary sets for negative and positive respectively.
+{\em Right:} examples of the $L_1,L_2$ balls for different points.
+The green $L_1$ circles are the largest monochrome circle that contains the point.
+The purple $L_2$ circles are the union of all circles of half the radius or smaller.
+All of these circles are monochrome.}
+%\label{fig:curvature}
+%\end{figure}
+
+\subsection{Example: Curvature and Massart noise}
+\label{sec:Massart}
+
+By way of example, consider a case where $\X \subset \R^d$ and $\mu$ admits a density.
+We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$.
+By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\cite{BC03}.
+
+We make two key assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
+\begin{itemize}[leftmargin=0.5cm]
+\item (Strong density condition)
+There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
+we have $c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) $, where $\vol(\cdot)$ is $d$-dimensional volume.
+\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
+\end{itemize}
+The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$.
+It is a common notion of curvature~\cite{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$).
+A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
+\begin{thm}
+  Under the two conditions above, there are constants $c_2, c_3$ for
+  which the following holds. Pick $0 < \delta < 1$ and take
+  $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm
+  of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$
+  queries, it assigns Bayes-optimal labels to all but
+  $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
+\label{thm:massart-dist}
+\end{thm}
+
+ Here the constant $c_2$ includes a term $(1/r_o)^d$. The
+$\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over
+the usual $1/m^{1/d}$ rate for random querying and was also observed
+earlier, in a more restricted setting, using specialized
+algorithms~\cite{CN08}.
 

### 2024-09-26 AISTAT25/aistats.tex
Experimental section

@@
   $\frac{1}{\gamma^2}$ times.
+
+The rest of the paper is organized as follows. xxxxxxx
 
@@
 
-\begin{figure}
+\begin{figure}[t]
 \begin{center}
@@
   Y axis represents the difference between the error achieved and the
-  optimal achievable error.}
+  Bayes error.}
 \label{fig:Experiments}
@@
 
-In this section we provide experimental evidence that our algorithm has low query complexity.
-We chose a simple 1D problem, described in figure~\ref{fig:oned}(b)
-(Massard Noise).  Although this problem is simple, it is outside the
-scope of algorithms that assume that the conditional probability is a
-smooth function of $x$. We used a sample of 10,000 points and ran our
-algorithm 20 times. In figure~\ref{fig:experiments} we present the
-average error as a function of the number of queries. We compare the
-performance of our active algorithm (red line) and the performance of the passive algorithm which is the
-same algorithm where focused queries are disabled and only backgroud samples are used (blue line).
-
-Number of queries as a function of the accuracy for passive and active algorithms.
-\begin{tabular}{|c|rr|c|}
+We describe an experiment designed to evaluate our algorithm.
+We use a simple 1D problem, described in figure~\ref{fig:oned}(b)
+(Massard Noise).  Although this problem is simple, it is outside the scope of most non-parametric active learning algorithms, which assume smoothness of the conditional probability.
+
+We compare two algorithms, the "active" agorithm described here, and the "passive" algorithm which uses only "background queries".
+We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$.
+From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon})$.
+
+ For our experiment we used a sample of 10,000 unlabeled points, we repeated each of our experiments 20 times.
+In Figure~\ref{fig:Experiments} we show the average performance of these experiments.
+Note the horizontal line segments marked "active vs.
+passive ...
+", these compare the number of queries required by the active and the passive algorithms for different accuracies
+ These numbers are repeated in the following table.
+\begin{tabular}{|c|rr|}
 \hline
@@
 0.01 &     429 &    1035 \\
-0.005 &678 &    2045 & \\
+0.005 &678 &    2045 \\
 \hline
 \end{tabular}
+Considering the columns for the Active and Passive algorithms respectively, we see that the experimental query complexities concur with the theoretical predictions.
 

### 2024-09-26 AISTAT25/aistats.tex
Cleanup

@@
 
-\iffalse
-\subsection{Results}
-
-%Our algorithm does not require any assumption on the input distribution.
-In the worst case the number of queries is twice the number of queries of a passive algorithm.
-We show that under several natural conditions the query complexity of our algorithm is logarithmic in the error.
-
-\fi
-
 \subsection{Related work} \label{sec:relatedwork}
@@
   $\frac{1}{\gamma^2}$ times.
-
-The rest of the paper is organized as follows. xxxxxxx
 
@@
 \section{A general-purpose active learning algorithm}
+\label{sec:algorithm}
 Our algorithm has several intricacies that might make it hard to
 follow at a first read.  We therefor start with a description of a
-simplified version of the algorithm (Fig~\ref{alg:siple}) that is not
+simplified version of the algorithm (Fig~\ref{alg:simple}) that is not
 completely correct but serves to explain the main components of the
-detailed algorithm. The full description of the algorithm is given in
-subsection~\ref{sec:full description} and appendix~\ref{sec:detailedalgorithm}
+detailed algorithm. The detailed algorithm is described in
+subsection~\ref{sec:detailedalgorithm}.
 
@@
 
-In step ~\ref{step:balls} associates a label with each ball. The
+Step ~\ref{step:balls} associates a label with each ball. The
 label is one of $\{-1,+1,0\}$.  The labels $+1$ and $-1$ correspond to
@@
 These bias estimates will with high probability be seen to satisfy the following guarantee.
-\begin{defn}
-For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
-\begin{itemize}
-\item $\yh(B) = +1 \implies \eta_X(B) > 0$
-\item $\yh(B) = -1 \implies \eta_X(B) < 0$
-\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
-\end{itemize}
-\label{def:accurate-bias-estimate}
-\end{defn}
-
 \subsection{Assigning labels to points}
@@
 The label-estimate for $x$ at level $\ell$, denoted $\yh_\ell(x)$, is $+1$ if $\PL_\ell(x)=\{+1\}$, $-1$ if $\PL_\ell(x) = \{-1\}$, $0$ if $\PL_\ell(x) = \{\}$, and $!$ if $\PL_\ell(x) = \{-1,+1\}$ (Eqn~(\ref{eq:provisional-label})).
-%The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.
+The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.
 
@@
 \end{enumerate}
-
-\section{Detailed algorithm}
 
@@
 \right.
-\label{eq:provisional-label}
+\label{eq:provisional-label1}
 \end{align}
@@
 %This intuition is borne out by the following result, proved in the appendix.
-\begin{lemma}
-Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. 
-Let $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}.
-Then with probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are $\gamma$-accurate in the sense of Definition~\ref{def:accurate-bias-estimate}.
-\label{lemma:accurate-bias-estimates}
+
+\pagebreak
+In the next definition and lemma we establish that the bias estimates for all balls in $\B$ are sufficiently accurate. 
+\begin{defn}
+\label{def:accurate-bias-estimate}
+For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
+\begin{itemize}
+\item $\yh(B) = +1 \implies \eta_X(B) > 0$
+\item $\yh(B) = -1 \implies \eta_X(B) < 0$
+\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
+\end{itemize}
+\end{defn}
+
+\begin{lemma}
+\label{lemma:accurate-bias-estimate}
+  Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$.  Let
+  $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}.  Then with
+  probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are
+  $\gamma$-accurate in the sense of
+  Definition~\ref{def:accurate-bias-estimate}.
 \end{lemma}
@@
 Despite randomness in querying, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ reliably contains the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ reliably omits the wrong label.
+Intuitively, background sampling dominates until level $L_1(x)$ is eached while focused sampling dominates between $L_1(x)$ and convergence at $L_2(x)$.
+Therefor, roughly speaking, the query complexity for point $x$ is low when $L_2(x)-L_1(x)$ is large relative to $L_1(x)$. The precise statement is given in Theorem~\ref{thm:queryComplexity}.
 
@@
 
-\begin{thm}[Local query complexity]
+\begin{thm}[Local query complexity] \label{thm:queryComplexity}
 Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$.
@@
 \begin{center}
-  \includegraphics[width=2.5in]{figures/Monotone.png}
+  \includegraphics[width=2.5in,height=1.7in]{figures/Monotone.png}
 %\hspace{.5in}
-\includegraphics[width=2.5in]{figures/Massart.png}
+\includegraphics[width=2.5in,height=1.7in]{figures/Massart.png}
 \end{center}
@@
 \subsection{Example: Massart noise}
+\label{sec:example-massart}
 
@@
 Once again, a logarithmic label complexity is obtained.
+(Details of the proof are in Section~(\ref{sec:oned-massart})
 
@@
 We describe an experiment designed to evaluate our algorithm.
-We use a simple 1D problem, described in figure~\ref{fig:oned}(b)
-(Massard Noise).  Although this problem is simple, it is outside the scope of most non-parametric active learning algorithms, which assume smoothness of the conditional probability.
+We use a simple 1D problem, described in Section~\ref{sec:example-massart}.
+Although this problem is simple, it is outside the scope of most non-parametric active learning algorithms, which assume smoothness of the conditional probability.
 
@@
 We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$.
-From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon})$.
+From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}{\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon})$.
 
@@
 
-\subsubsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}
+\subsubsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimate}}
 
@@
 For $\ell \geq L_1(x)$, the set $\{B \in \B_{\leq \ell}(x): X_{B} \subseteq X_{B_o}\}$ is nonempty (it contains $B_o$); pick any minimal ball $B$ within it.
-Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimates}) and thus $s(x) \in \PL_\ell(x)$.
+Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimate}) and thus $s(x) \in \PL_\ell(x)$.
 
@@
 
-%\begin{figure}[t]
-%\begin{center}
-%\includegraphics[width=3in]{curvature.pdf}
-%\end{center}
-%\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=3in]{figures/curvature.pdf}
+\end{center}
+\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue.
 The orange and light blue indicate the boundary sets for negative and positive respectively.
@@
 All of these circles are monochrome.}
-%\label{fig:curvature}
-%\end{figure}
+\label{fig:curvature}
+\end{figure}
 

### 2024-09-27 AISTAT25/aistats.tex
cleaning

@@
 \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning
-with the largest balls. A bias of $+1$ means that the average
-$\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is
+with the largest balls. A bias of $+1$ means that the average of 
+$\eta(x)$ over $s \in B$ is significantly positive; bias $-1$ means it is
 significantly negative; and bias $0$ means the average $\eta$ is close
@@
 
-These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$.
-By tackling all three of them, we are able to give a
-universal active learning scheme.
-
-The rest of the paper is organized as follows. Section~\ref{sec:relatedwork} describes related work.
-Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function.
-We identify two \emph{critical levels} for any $x \in X$: two scales (of ball sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}).
-We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.
-
-Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$.
-We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.
+These three challenges go beyond earlier work in active learning,
+which was able to avoid problems like mind-changes by making
+smoothness assumptions on $\eta$. By tackling all three of them, we
+are able to give a universal active learning scheme. Our focus is on
+the query complexity of the algorithm, largely ignoring the
+computational complexity. However, we do provide some experimental results that demonstrate that the algorithm works on a toy example.
+
+The rest of the paper is organized as
+follows. Section~\ref{sec:relatedwork} describes related work.
+Section~\ref{sec:algorithm} presents the active learning algorithm and
+Section~\ref{sec:discrete} analyzes it in the finite population
+setting, taking $X$ to be an arbitrary point-set and allowing any
+$\eta$ function. We identify two \emph{critical levels} for any
+$x \in X$: two scales (of ball sizes) that control how many queries
+are sufficient for $x$ to be correctly labeled
+(Theorem~\ref{thm:label-complexity}). We instantiate these bounds in
+canonical one-dimensional settings (Theorems~\ref{thm:oned-massart}
+and \ref{thm:oned-monotonic}) to get label complexities logarithmic in
+$|X|$. In section~\ref{sec:experiment} we describe an experiment that supports the theoretical bound given in Theorems~\ref{thm:oned-massart}.
+
+The supplementary material in
+Section~\ref{sec:FinitePopulationProofs}, which contains the proofs of
+the theorems from Section~\ref{sec:discrete}.  Finally,
+Section~\ref{sec:continuous} considers the statistical setting where
+$X$ is drawn from an underlying distribution $\mu$ on $\X$. We give
+rates of convergence under distributional conditions
+(Theorem~\ref{thm:label-complexity-specific}), and instantiate them
+under common assumptions from learning theory and computational
+geometry.
 
@@
 $B$---the sign of the average $\eta$ value in $B$---is easy to
-determine given a reasonable number of points from that region.
+determine given a reasonable number of queries chosen at random from that region.
 
@@
 The two basic entities in our algorithm are the collection of points
-$X$ and a collection $\B$ of balls. Sampling is organized around a the
+$X$ and a collection of balls $\B$. Sampling is organized around a the
 collection $\B$. Balls are the atomic sets on which we assess label
@@
 \begin{enumerate}[leftmargin=0.6cm]
-\item \label{step:Uell} {\bf Compute uncertainty set:}\\ If $\ell=0$, set $U_0 = X$; otherwise set $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$t
-\item \label{step:QS} {\bf Compute Focused Query set:}\\ Compute query ball set be $R=  \bigcup_{x \in U} \B_\ell(x)$
-and focused query set $S=\bigcup_{B \in R} B$
+\item \label{step:Uell} {\bf Compute uncertainty region:}\\ If $\ell=0$, set $U_0 = X$; otherwise set $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$t
+\item \label{step:QS} {\bf Compute Focused Query region:}\\ Compute query region be $R=  \bigcup_{x \in U} \B_\ell(x)$
+and focused query region $S=\bigcup_{B \in R} B$
 \item \label{step:sample} {\bf Make queries:}\\
@@
 
-\subsection{Simplified algorithm}
+\subsection{The simplified algorithm}
 
@@
 the algorithm. The algorithm loops over increasing levels.  At each
-level it identifies the set of ``known unknowns'' which approximates
+level it identifies the region of ``known unknowns'' which approximates
 the boundary. The algorithm then makes queries in the vicinity of the
@@
 What follows is a detailed explanation of  Figure~\ref{alg:simple}.
-
 The outer loop of the algorithm iterates over the
@@
 at least half of the points in $X$, and ending with $\ell=\log |X|-1$,
-corresponding to balls that contain a single point.
-
-The loop consists of five steps. The first computes the uncertainty set, the second computes the focused query
-set, the third defines the sampling process, the fourth and fifth
-update the labels of the balls and the points respectively.
+corresponding to balls that contain a single point. Initially, all balls are labeled $\bot$, which indicates that there are not enough queries inside the ball to determine it's label.
+
+The loop consists of five steps. The first computes the {\em uncertainty
+region}, the second computes the {\em focused query region}, the third defines the
+sampling process, the fourth and fifth update the labels of the balls
+and the points respectively.
 
@@
 the balls in $\B_l$ that contain at least one point of $U_\ell$.  The
-set $S$ is the union of these balls and defines the ``focus area''
-which is queried in order to reduce the known unknowns.
+region $S$ is the union of these balls and defines the set
+from which queries are selected in order to reduce the known unknowns.
 
 In step~\ref{step:sample} the algorithm queries points inside the
-focus area repeatedly at random.  This step terminates once there are
-$C/\gamma^2$ samples in each of the balls in the focus area.
-
-Step ~\ref{step:balls} associates a label with each ball. The
+focus region repeatedly at random.  This step terminates once there are
+$C/\gamma^2$ samples in each of the balls in the focus region.
+
+Step ~\ref{step:balls} associates a label with each ball in the focus region. The
 label is one of $\{-1,+1,0\}$.  The labels $+1$ and $-1$ correspond to
@@
 possible labels set $\PL_\ell(x)$ as described in
-step~\ref{step:examples}. The most significant label is ``!'', which
+Equation~\ref{eq:provisional-label}. The most significant label is ``!'', which
 indicates balls with conflicting polarity containing $x$. Points
@@
   over levels. Instead, at each iteration it finds the lowest level
-  where the uncertainty set is not empty and makes two queries, one
+  where the uncertainty region is not empty and makes two queries, one
   focused and one background.
@@
 \begin{array}{cl}
-  \yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$,1}\\
+  \yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$}\\
   & \mbox{if such an $\ell$ exists} \\
@@
 \begin{itemize}[leftmargin=0.5cm]
-\item Define querying region:
+\item Define focus region:
 $$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
@@
 
-\pagebreak
 In the next definition and lemma we establish that the bias estimates for all balls in $\B$ are sufficiently accurate. 
@@
 Focused sampling at level $\ell$ is restricted to balls that contain these points.
-We will think of this region as the \emph{boundary set} at level $\ell$.
+We will think of this set as the \emph{boundary set} at level $\ell$.
 \begin{defn}[Boundary set $\bf \Delta_\ell$]
@@
 
-To apply Theorem~\ref{thm:label-complexity}, we need bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each $x \in X$, and on the size of the sampling region $\Delta_\ell$ at each level $\ell$.
+To apply Theorem~\ref{thm:label-complexity}, we need bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each $x \in X$, and on the size of the sampling set $\Delta_\ell$ at each level $\ell$.
 We now derive these in some canonical settings.
@@
 
-A counting argument then shows that the boundary region shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
+A counting argument then shows that the boundary set shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
 \begin{thm}
@@
 \section{Experiments}
-
+\label{sec:experiment}
 We describe an experiment designed to evaluate our algorithm.
@@
 Note the horizontal line segments marked "active vs.
-passive ...
-", these compare the number of queries required by the active and the passive algorithms for different accuracies
+passive <error>", these compare the number of queries required by the active and the passive algorithms for different accuracies
  These numbers are repeated in the following table.
+\begin{center}
 \begin{tabular}{|c|rr|}
@@
 \end{tabular}
+\end{center}
+
 Considering the columns for the Active and Passive algorithms respectively, we see that the experimental query complexities concur with the theoretical predictions.
@@
 complexity of the algorithm is high. Future work will be directed
-towards characterizing region sets that give both high query
-efficiency and low computational complexity.
+towards characterizing collections of regions that give both low query
+complexity and low computational complexity.
 
@@
 %%%%%%%%%%%%
-\section{Analysis Details: Discrete setting}
-
-\subsection{Technicalities: discrete setting}
-
-\subsubsection{Large deviation bounds for the discrete setting}
+\section{Analysis Details: Finite Population setting}
+\label{sec:FinitePopulationProofs}
+
+\subsection{Technicalities: finite population setting}
+
+\subsubsection{Large deviation bounds for the finite population setting}
 
@@
 
-Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$.
-From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty set at level $\ell$.
+Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty region at level $\ell$.
+From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty region at level $\ell$.
 In short,
@@
 
-{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.
+{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two.
+This can only happen if some round of sampling has an empty uncertainty region, meaning that all of $A$ has left $U$ at that point.
 

### 2024-10-05 AISTAT25/aistats.tex
Updates from Overleaf
Sanjoy's final version

@@
 
-\usepackage{aistats2024}
+\usepackage{aistats2025}
 \usepackage{enumitem}
@@
 
-In \emph{non-parametric active learning}, the starting point is a data
+In \emph{nonparametric active learning}, the starting point is a data
 set whose labels are hidden but can be obtained individually for a
-price. The goal is to label the data set, or to find a good
+price. The goal is to label the data set, or to find a corresponding
 classifier, at low cost.
@@
 $\X$.~\footnote{Our approach works for any topological space
-  that defines a sufficiently rich set of neighborhoods. For the same of concreteness, we
-  restrict ourselves to metric spaces where the neighborhoods are balls.}  We can
+  that defines a sufficiently rich set of neighborhoods. For concreteness, we
+  focus on metric spaces and take the neighborhoods to be balls.}  We can
 query the label of any of these points $x$, in which case we get a
@@
 or on the conditional expectation $\eta(x)$. Our algorithm is
-universal and works for any $X$ and $\eta(x)$. The {\em query
-  complexity} of our algorithm depends on $X$ and $\eta$, but it does
-not depend on smoothness of $\eta(x)$. Instead it depends on
-properties of $\eta(x)$ close to the decision boundary where $\eta(x)$
-changes sign. In particular, the algorithm is effective when $\eta(x)$
-is discontinuous and hence not smooth.
+universal and works for any $X$ and $\eta(x)$. Our {\em query
+  complexity} bound depends on $X$ and $\eta$, but has no explicit dependence on the smoothness of $\eta(x)$. Instead it depends on
+properties of $\eta(x)$ close to the decision boundary where $\eta(x)$ changes sign. In particular, the algorithm is effective even when $\eta(x)$
+is discontinuous.
 
@@
 with the largest balls. A bias of $+1$ means that the average of 
-$\eta(x)$ over $s \in B$ is significantly positive; bias $-1$ means it is
+$\eta(x)$ over $x \in B$ is significantly positive; bias $-1$ means it is
 significantly negative; and bias $0$ means the average $\eta$ is close
@@
 as needed. The final label assigned to any point $x$ is based on the
-biases of balls containing just $x$.
+biases of balls containing $x$.
 
@@
 \begin{center}
-\includegraphics[width=3in]{figures/ideas.pdf}
+\includegraphics[width=3.25in]{figures/ideas.pdf}
 \end{center}
@@
 
+\paragraph{1. Where to query?}\label{challange:wheretoquery}
+
+  Initial random querying tells us the biases of the largest balls,
+  which gives some information about the decision boundary. In what
+  parts of $\X$ might it be helpful to know biases of smaller balls,
+  warranting further label queries? Our approach is simple: if a ball
+  $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the
+  intersection $B \cap B'$ is a ``known unknown'' and is targeted for
+  active querying (Fig.~\ref{fig:challenges}(b)).
+
+  This differs from the more conventional strategy of actively
+  querying balls $B$ whose average $\eta$-value is close to zero. Such
+  $B$ fall into two categories: (1) the $\eta$ values are close to
+  zero throughout $B$, and (2) $B$ consists of two sub-regions, one
+  strongly positive and the other strongly negative. In the first
+  case, there is little merit in querying further. But in the second
+  case, there is a lot to be gained. Our approach allows these two
+  cases to be distinguished.
+
+\paragraph{2. No smoothness assumptions} \label{challange:nosmoothness}
+
+  As no assumptions of smoothness are made, the sign of $\eta(\cdot)$
+  can change in an arbitrarily small neighborhood of any given point: the
+  biases of balls containing $x$, say
+  $B_1 \supset B_2 \supset \cdots$, might change sign a few times. At
+  first, given only the bias of $B_1$, we might think $x$ has label
+  $+1$. When we later estimate the bias of $B_2$, we could change our
+  mind to $-1$. And then $+1$, and so on
+  (Fig.~\ref{fig:challenges}(c)).
+
+\paragraph{3. Managing sampling} \label{challange:managingsampling}
+
+  The third challenge is \emph{managing sampling} of overlapping
+  balls. We are interested in detecting points, and thus balls, whose
+  $\eta$-values are either $> \gamma$ or $<-\gamma$. This suggests
+  querying $k \approx 1/\gamma^2$ points at random from each
+  ball. Now, suppose we have queried this many points from ball $B$
+  and later want to query points from a different ball $B'$ that
+  overlaps $B$. How can we reuse queries already made in $B \cap B'$?
+
+\iffalse
 \begin{enumerate}
@@
 \item \label{challange:nosmoothness}
-  {\bf No smootheness assumptions}\\
+  {\bf No smoothness assumptions}\\
   As no assumptions of smoothness are made, the sign of $\eta(\cdot)$
@@
 \end{enumerate}
+\fi
 
@@
 and \ref{thm:oned-monotonic}) to get label complexities logarithmic in
-$|X|$. In section~\ref{sec:experiment} we describe an experiment that supports the theoretical bound given in Theorems~\ref{thm:oned-massart}.
+$|X|$. In section~\ref{sec:experiment} we describe an experiment that supports the theoretical bound given in Theorem~\ref{thm:oned-massart}.
 
 The supplementary material in
-Section~\ref{sec:FinitePopulationProofs}, which contains the proofs of
+Section~\ref{sec:FinitePopulationProofs} contains the proofs of
 the theorems from Section~\ref{sec:discrete}.  Finally,
@@
 
-Existing work on non-parametric active learning yields algorithms that are guaranteed to work when the underlying conditional expectation $\eta(x)$ is smooth in some way.
+Existing work on nonparametric active learning yields algorithms that are guaranteed to work when the underlying conditional expectation $\eta(x)$ is smooth in some way.
 Additional assumptions are often needed.
@@
 negative) of entire regions at a time \cite{DH08,M12}. In this paper,
-we follow the second strategy because the only reliable and
+we follow the second strategy because the bias of a region
+$B$---the sign of the average $\eta$ value in $B$---is easy to
+determine given a reasonable number of queries chosen at random from that region. 
+On the other hand, the only reliable and
 general-purpose way to assess the bias of an individual point,
@@
 smoothness assumptions, the sign can change abruptly in an arbitrarily
-small ball around $x$. On the other hand, the bias of a region
-$B$---the sign of the average $\eta$ value in $B$---is easy to
-determine given a reasonable number of queries chosen at random from that region.
+small ball around $x$.
 
@@
 The idea of predicting with balls of multiple sizes is a central theme of \cite{BDFM19}, which predicts at a point $x$ by choosing a ball centered at $x$ that is as small as possible while having a clear bias towards one label.
-That work does not consider active learning, which demands considerable further innovation.
-For instance, to assess the uncertainty at a point $x$, we need to look at balls containing $x$, not just those centered at $x$ (Fig.~\ref{fig:challenges}(b)).
-
+That work does not consider active learning, which demands further innovation.
+For instance, to assess the uncertainty at a point $x$, we look at balls \emph{containing} $x$, not just \emph{centered} at $x$ (Fig.~\ref{fig:challenges}(b)).
+
+\iffalse
 Recall that the algorithm can query any $x \in X$ and receive $-1$ or $+1$ at random,
@@
   $\frac{1}{\gamma^2}$ times.
+\fi
 
@@
 The two basic entities in our algorithm are the collection of points
-$X$ and a collection of balls $\B$. Sampling is organized around a the
-collection $\B$. Balls are the atomic sets on which we assess label
+$X$ and a collection of balls $\B$. Sampling is organized around $\B$:
+these balls are the atomic sets on which we assess label
 bias.
@@
 For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the
-data points that lie in it.  We group balls into levels by the number
+data points lying in it.  We group balls into levels by the number
 of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
@@
 \end{equation}
-Let $\B_\ell$ consist of all balls in $\B$ that are at level
+Let $\B_\ell$ consist of all balls in $\B$ at level
 $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$
@@
 Balls in lower levels contain more points, and thus their biases
-(average $\eta$ values) are easier to estimate. The main loop of the
-simplified learning algorithm iterates over the levels $\ell=0,1,2,\ldots$
-
-For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of
+(average $\eta$ values) are easier to estimate. The overall querying
+strategy begins with balls at the lowest levels.
+%main loop of the
+%simplified learning algorithm iterates over the levels $\ell=0,1,2,\ldots$
+
+For any point $x \in \X$, let $\B(x) \subset \B$ denote the collection of
 balls that contain $x$ and can thus be used in determining $x$'s
@@
 Our algorithm has several intricacies that might make it hard to
-follow at a first read.  We therefor start with a description of a
-simplified version of the algorithm (Fig~\ref{alg:simple}) that is not
-completely correct but serves to explain the main components of the
-detailed algorithm. The detailed algorithm is described in
-subsection~\ref{sec:detailedalgorithm}.
-
-\begin{figure*}[t]{6in}
+follow at a first read.  We therefore start with a
+simplified version of the algorithm (Fig~\ref{alg:simple}) that serves
+to explain the main ideas but is missing some details.  
+The detailed algorithm is described in
+Section~\ref{sec:detailedalgorithm}.
+
+\begin{figure*}[t]
 \framebox{
-\begin{minipage}[t]{6in}
-\begin{center}
+\begin{minipage}[t]{6.5in}
+%\begin{center}
 \vspace{.1in}
-\textbf{Repeat for $\ell=0,1,2,\ldots,\log |X|-1$:}
+\noindent\textbf{Repeat for $\ell=0,1,2,\ldots,$:}
 \begin{enumerate}[leftmargin=0.6cm]
-\item \label{step:Uell} {\bf Compute uncertainty region:}\\ If $\ell=0$, set $U_0 = X$; otherwise set $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$t
-\item \label{step:QS} {\bf Compute Focused Query region:}\\ Compute query region be $R=  \bigcup_{x \in U} \B_\ell(x)$
-and focused query region $S=\bigcup_{B \in R} B$
+\item \label{step:Uell} {\bf Compute uncertainty region:}\\ If $\ell=0$, set $U_0 = X$; otherwise set $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!\}$ %, \ \yh_\ell(x) = \bot\}$
+\item \label{step:QS} {\bf Compute focused query region:}\\ 
+The balls to be queried are $R=  \bigcup_{x \in U} \B_\ell(x)$
+and the focused query region is their union, $S=\bigcup_{B \in R} B$
 \item \label{step:sample} {\bf Make queries:}\\
 \textbf{Repeat until} number of queries in each ball $B \in R$ is at least $C/\gamma^2$
-  \begin{enumerate}[leftmargin=0.6cm]
-  \item Select a {\tt query} uniformly at random from $S$
-  \end{enumerate}
-
-\item \label{step:balls} {\bf Update Ball labels:} For each ball $B \in R$, let $k(B)$ be
-  the number of queries (of any kind) made inside $B$ and let
+  \begin{itemize}[leftmargin=0.6cm]
+  \item Select a {\tt query} point uniformly at random from $S$
+  \end{itemize}
+
+\item \label{step:balls} {\bf Update ball labels:} For each ball $B \in R$, let 
+    %$k(B)$ be the number of queries (of any kind) made inside $B$ and let
   $\widehat{\eta}(B)$ be the average of the labels returned by
-  these queries.
+  queries in $B$.
 
@@
 \end{enumerate}
-\end{center}
+%\end{center}
 
@@
 
-We now describe the simplified algorithm here referred to simply as
+We now describe the simplified algorithm, here referred to simply as
 the algorithm. The algorithm loops over increasing levels.  At each
@@
 What follows is a detailed explanation of  Figure~\ref{alg:simple}.
-The outer loop of the algorithm iterates over the
-levels, starting with $\ell=0$ that corresponds to balls that contain
-at least half of the points in $X$, and ending with $\ell=\log |X|-1$,
-corresponding to balls that contain a single point. Initially, all balls are labeled $\bot$, which indicates that there are not enough queries inside the ball to determine it's label.
+The outer loop of the algorithm iterates over
+levels, starting with $\ell=0$ that corresponds to balls with at least half of the points in $X$, then moving onto $\ell=1$ which has balls with between $1/4$ and $1/2$ of the points in $X$, and so on.
+%and ending with $\ell=\log |X|-1$,
+%corresponding to balls that contain a single point. %Initially, all balls are labeled $\bot$, which indicates that there are not enough queries inside the ball to determine it's label.
 
@@
 region}, the second computes the {\em focused query region}, the third defines the
-sampling process, the fourth and fifth update the labels of the balls
-and the points respectively.
+sampling process, the fourth and fifth update the labels of balls
+and points respectively.
 
@@
 calculated. Intuitively $U_l$ is the set of points whose label
-according to level $\ell-1$ is ! or ``known unknown'' and the label
-according to level $\ell$ is $\bot$ or ``not enough information''.
+according to level $\ell-1$ is ! or ``known unknown''.
+%and the label
+%according to level $\ell$ is $\bot$ or ``not enough information''.
 These are the points for which we need better estimates.
 
-In step~\ref{step:QS} the algorithm identifies the balls that needs to
-be queried in order to get more information about $U_\ell$. These are
-the balls in $\B_l$ that contain at least one point of $U_\ell$.  The
-region $S$ is the union of these balls and defines the set
-from which queries are selected in order to reduce the known unknowns.
+In step~\ref{step:QS} the algorithm identifies balls that need to
+be queried to get more information about $U_\ell$. These are
+the balls in $\B_\ell$ that contain at least one point of $U_\ell$. 
+Region $S$ is the union of these balls. It is the set
+from which queries are selected in order to reduce the known-unknowns.
 
@@
 
-Step ~\ref{step:balls} associates a label with each ball in the focus region. The
-label is one of $\{-1,+1,0\}$.  The labels $+1$ and $-1$ correspond to
-a confident statement that $\eta(B)>0$ and $\eta(B)<0$
-respectively. Finally, the label $0$ indicates that the sign of $B$
+Step ~\ref{step:balls} assigns a label $\yh(B)$ to each ball $B$ in the focus region. The
+label is one of $\{-1,+1,0\}$.  Labels $+1$ and $-1$ correspond to
+a confident statement that $\eta_X(B)>0$ and $\eta_X(B)<0$
+respectively; here $\eta_X(B)$ is the average value of $\eta(x)$ over points $x \in B \cap X$. The label $0$ indicates that the sign of $B$
 is uncertain given the answers to the queries.
@@
 examples.  This is done in two sub-steps. First the {\em Possible
-  Labels} for $x$ denoted $\PL_\ell(x)$ are calculated. $\PL_\ell(x)$
+  Labels} for $x$, denoted $\PL_\ell(x)$, are calculated. $\PL_\ell(x)$
 contains $+1$ if there is a ball containing $x$ whose label is $+1$,
@@
 
-The simplified algorithm does addresses the first challange of ``where
-to query'' (see challanges in the introduction). It does not address
-the other two challanges, which the detailed algorithm does address.
-
+The simplified algorithm addresses the first challenge of ``where
+to query'' (see challenges in the introduction). It does not address
+the other two challenges, which the detailed algorithm does address.
+
+\paragraph{No smoothness assumptions.} As discussed in the introduction, one
+is \emph{never} sure of having correctly determined the label of any 
+specific point $x$. Thus in addition to focused (active) querying, we also
+do background sampling of the entire space to pick up on possible mind
+changes. For the sake of simplicity, we make one background query per focused query.
+
+A related problem is that having the outer loop iterate over levels will
+cause the algorithm to miss uncertain regions that later appear in lower
+levels as a result of background sampling (the so-called mind changes).
+Instead, each iteration of the full algorithm looks for the lowest level 
+in which the uncertainty region is non-empty, and makes just two queries, 
+one focused and one background.
+
+\paragraph{Managing sampling.} As described in the introduction, we would
+like to use queries that fall in the intersection of two balls for estimating
+the bias in each of the balls. To facilitate query re-use, it is helpful to 
+use Poisson sampling, as described in Section~\ref{sec:poisson}.
+
+\iffalse
 \begin{enumerate}
@@
 Once these queries are made, bias estimates $\widehat{\eta}(B)$ are
-updated using the Poisson method,\\
-$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated
-accordingly  and the next iteration begins.
+updated, the values of 
+$\yh(B), \PL_\ell(x), \yh_\ell(x), U_\ell$ are updated,
+ and the next iteration begins.
 
@@
 query the part of it we know.
-
+As a result, we are repeatedly in situations where the current region of uncertainty---in particular, the balls $B$ whose labels $\yh(B)$ we wish to determine---already contains some labels from earlier queries, and we wish to reuse those labels in a manner that is statistically sound.
 We manage this through \emph{Poisson
-  sampling}~\cite{ghosh2002sampling} (Fig.~\ref{alg:sampling}) which
-works as follows. Each instance $x \in X$ is assocciated with a random
-draw from the uniform distribution over $[0,1]$ denoted $T_x$.
-Pooints are sampled in the order of increasing $T_x$ under some
-conditions described below.  Note that restricting the sampled points
-to be from set $A$ and having $T_x \leq \tau$ is the same, in
-expectation to sampling each point in $A$ independently at random with
-probability $\tau$.
-
-To infer the bias of a ball $B$, we need labels for
-$k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$
-have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$
-fraction of the ball to be queried. Instead, we query every point in
-the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$,
-\emph{independently}. Specifically, we query points $x$ in the ball
-with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a
-uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
-
-Once a point is queried, it is never queried again.
+  sampling}~\cite{ghosh2002sampling} (Fig.~\ref{alg:sampling}).
+
+As an example of what we are able to achieve, suppose we want to determine the label of a ball $B$ at level $\ell$ and have already queried points in several other balls, at levels $\leq \ell$, that intersect $B$.
+We will be able to apply \emph{all} these queries that fall in $B$ towards estimating $B$'s bias.
+
+Here is the first key idea.
+In order to infer the label $\yh(B)$ of a ball $B$, we need to estimate the average bias of the ball, $\eta_X(B)$, within $\gamma$, which requires querying roughly $k = \tilde{O}(1/\gamma^2)$ random points in the ball.
+Instead of picking $k$ points at random in $B$, we query every point in $B$ with probability $\approx k/|X_B|$, \emph{independently}.
+More precisely, if $B$ lies at level $\ell$ then it has roughly $n/2^{\ell+1}$ points, so we query every point in $B$ with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$.
+
+The second key idea is to manage this independent sampling in a manner that makes it easy to reuse earlier queries.
+At the outset of the algorithm, each instance $x \in X$ is assigned a random draw from the uniform distribution over $[0,1]$, denoted $T_x$.
+To query each point in a ball $B$ with probability $\tau_\ell$, we simply select all points in $x \in B$ that have $T_x \leq \tau_\ell$.
+If a point $x$ was queried earlier---because it also belongs to a ball $B'$ at level $\ell' \leq \ell$, say---then it will now get reused (because $T_x \leq \tau_{\ell'} \implies T_x \leq \tau_\ell$).
+
+%Points are sampled in the order of increasing $T_x$ under some
+%conditions described below.  Note that restricting the sampled points
+%to be from set $A$ and having $T_x \leq \tau$ is the same, in
+%expectation to sampling each point in $A$ independently at random with
+%probability $\tau$.
+
+%To infer the bias of a ball $B$, we need labels for
+%$k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$
+%have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$
+%fraction of the ball to be queried. Instead, we query every point in
+%the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$,
+%%%\emph{independently}. Specifically, we query points $x$ in the ball
+%with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a
+%uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.
+
+Once a point is queried, it is never queried again; but see the note at the end of Section~\ref{sec:discrete} about situations in which it is permissible to query a point multiple times.
 
@@
 
-Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
-
+Initialize uncertainty regions at all levels: 
+\begin{itemize}
+\item $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$
+\end{itemize}
 %\vspace{.1in}
 
-Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
+Initialize labels at all levels to ``unavailable'': 
+\begin{itemize}
+\item $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$
+\end{itemize}
 
@@
 \item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
-\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
+\item For all levels $\ell \geq 1$: \\ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
 \end{itemize}
@@
 
-We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on the manner in which $X$ is generated.
+We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on how $X$ is generated.
 
@@
 Fix the set of balls $\B$ and let $0 < \delta < 1$ be a
-predefined confidence parameter. We start with a uniform guarantee on the
-bias estimates for all balls $B \in \B$.
+predefined confidence parameter. We start with a uniform guarantee on the bias estimates for all balls $B \in \B$.
 
@@
 \label{def:accurate-bias-estimate}
-For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
+For any $B \in \B$, let $\eta_X(B)$ denote $\mbox{mean}(\{\eta(x): x \in X_B\})$. We say bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
 \begin{itemize}
@@
 
-Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes the possible labels for $x$ given information from balls at levels $\leq \ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$.
+Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes possible labels for $x$ given information from balls at levels $\leq \ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$.
 Despite randomness in querying, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ reliably contains the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ reliably omits the wrong label.
-Intuitively, background sampling dominates until level $L_1(x)$ is eached while focused sampling dominates between $L_1(x)$ and convergence at $L_2(x)$.
-Therefor, roughly speaking, the query complexity for point $x$ is low when $L_2(x)-L_1(x)$ is large relative to $L_1(x)$. The precise statement is given in Theorem~\ref{thm:queryComplexity}.
+Intuitively, background sampling dominates until level $L_1(x)$ is reached while focused sampling dominates between $L_1(x)$ and convergence at $L_2(x)$.
+Therefore, roughly speaking, the query complexity for point $x$ is low when $L_2(x)-L_1(x)$ is large relative to $L_1(x)$. The precise statement is in Theorem~\ref{thm:queryComplexity}.
 
 \begin{defn}[Critical levels $L_1,L_2$]
-Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$. Define $L_1(x)$ to be the smallest level $\ell$ such that there exists $B_o \in \B_\ell(x)$ for which
+Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$. Define $L_1(x)$ to be the smallest level $\ell$ containing some $B_o \in \B_\ell(x)$ for which
 \begin{itemize}[leftmargin=0.5cm]
@@
 
-A common intuition about active learning is that successive queries gradually constrain the possible locations of the decision boundary.
+A common intuition about active learning is that successive queries gradually constrain the location of the decision boundary.
 Let's consider the state of affairs when all balls at level $\leq \ell - 1$ have been sampled.
-The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points must have $L_2(x) \geq \ell$.
+The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points have $L_2(x) \geq \ell$.
 Focused sampling at level $\ell$ is restricted to balls that contain these points.
-We will think of this set as the \emph{boundary set} at level $\ell$.
+We will think of this as the \emph{boundary set} at level $\ell$.
 \begin{defn}[Boundary set $\bf \Delta_\ell$]
@@
 Let's interpret this bound.
+First, recall that levels $\ell$ range from $0$ to $\log (n/k)$.
+In the many scenarios we later consider, $L_1(x)$ is a constant, so the second term in the bound dominates.
+The benefit of active learning is captured by the critical level $L_2(x)$ and, most importantly of all, the size of the boundary, $|\Delta_\ell|$, at each level.
+If $|\Delta_\ell|$ is always $n$, we might end up querying all $n$ points.
+But in the cases we study, either $|\Delta_\ell| = 0(1)$, in which case we get label complexity polylogarithmic in $n$, or $|\Delta_\ell| = n/{2^{\alpha \ell}}$, $\alpha > 0$, which yields label complexity $\widetilde{O}(n^{1-\alpha})$.
+
+Now let's see how the bound arises.
 For a specific $x \in X$, there is no telling how $\yh_\ell(x)$ will behave at levels $\ell < L_1(x)$.
@@
 
-In particular, $O(k \cdot 2^\ell)$ background queries give us the biases of all balls in $\B_{\leq \ell}$; this is the first term in $m_o(x)$.
+In particular, $O(k \cdot 2^\ell)$ background queries give the biases of all balls in $\B_{\leq \ell}$; this is the first term in $m_o(x)$.
 And there are $O(k \cdot 2^\ell \cdot |\Delta_\ell|/n)$ focused queries at any level $\ell$: the second term.
@@
 %\end{align*}
-$$
+\begin{align*}
 L_1(x)
-\leq
+&\leq
 \left\{
@@
 \end{array}
-\right.
-\mbox{\ \ \ \ and \ \ \ }
+\right. \\
+%\mbox{\ \ \ \ and \ \ \ }
 L_2(x)
-\leq
+&\leq
 \lg (n/r(x))
-$$
+\end{align*}
 where $r(x)$ is the number of data points between $x$ and $\lambda$, counting $x$ as well.
@@
 \end{thm}
-If $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$, then the label complexity is $O(\log^2 n)$.
+If $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$, the label complexity is $O(\log^2 n)$.
 If in addition $r^+, r^- = \Theta(n)$, that is, a constant fraction of the points lie between $\lambda_L, \lambda_R$ and the decision boundary, the label complexity is $O(\log n)$.
@@
 Once again, a logarithmic label complexity is obtained.
-(Details of the proof are in Section~(\ref{sec:oned-massart})
+Details of the proof are in Section~\ref{sec:oned-massart}.
 
@@
 
+\section{The distributional setting}
+
+In Section~\ref{sec:continuous} (appendix), we consider the familiar scenario where $X$ is drawn from a distribution $\mu$ on $\X$.
+We show how Theorem~\ref{thm:label-complexity} implies label complexity bounds based on curvature and smoothness properties of the boundary (Theorem~\ref{thm:label-complexity-specific}).
+This recovers minimax bounds obtained earlier for active learning under smoothness~\cite{CN08}.
+
+A key component of the analysis is a new probabilistic notion of distance.
+We take the distance from a point $x$ to a set $S$ to be the probability mass of the smallest ball in $\B$ that contains $x$ and touches $S$, that is, $d(x,S) = \inf\{\mu(B): B \in \B(x), B \cap S \neq \emptyset\}$.
+We then find, for instance, that the second critical level $L_2(x)$ of any point $x$ is easily bounded in terms of its distance to the set of points with the opposite label (Lemma~\ref{lemma:L2-bound}).
+Other relevant quantities such as boundary size $|\Delta_\ell|$ also have simple expressions in terms of this distance.
+See the appendix for further details.
+
 \section{Experiments}
@@
 We use a simple 1D problem, described in Section~\ref{sec:example-massart}.
-Although this problem is simple, it is outside the scope of most non-parametric active learning algorithms, which assume smoothness of the conditional probability.
-
-We compare two algorithms, the "active" agorithm described here, and the "passive" algorithm which uses only "background queries".
+Although this problem is simple, it is outside the scope of nonparametric active learning algorithms that assume smoothness of the conditional probability.
+
+We compare two algorithms, the ``active'' agorithm described here, and the ``passive'' algorithm which uses only background queries.
 We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$.
-From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}{\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon})$.
-
- For our experiment we used a sample of 10,000 unlabeled points, we repeated each of our experiments 20 times.
+From the theory, we expect the query complexity of the passive algorithm to be $O(\frac{1}{\epsilon})$ and of the active algorithm to be $O(\log (\frac{1}{\epsilon}))$.
+
+ For our experiment we used a sample of 10,000 unlabeled points and repeated each of our experiments 20 times.
 In Figure~\ref{fig:Experiments} we show the average performance of these experiments.
-Note the horizontal line segments marked "active vs.
-passive <error>", these compare the number of queries required by the active and the passive algorithms for different accuracies
+The horizontal line segments marked ``active vs.
+passive [error]'' compare the number of queries required by the active and the passive algorithms for different accuracies.
 These numbers are repeated in the following table.
@@
 
-\section{Future Directions}
-Our algorithm is generic, while we used euclidean balls for clarity,
-any sets can be used. While this generality is attractive, it leaves
+\section{Future directions}
+Our algorithm is generic; while we used Euclidean balls for clarity,
+any neighborhoods can be used. Although this generality is attractive, it leaves
 open the question of what are good ways to define the collection of
 regions. Another aspect of this generality is that the computational
-complexity of the algorithm is high. Future work will be directed
+complexity of the algorithm could be high. Future work will be directed
 towards characterizing collections of regions that give both low query
@@
 \begin{center}
-\LARGE{Non-parametric active learning without smoothness\\
+\LARGE{Nonparametric active learning without smoothness\\
 Supplementary Materials}
@@
 %%%%%%%%%%%%
-\section{Analysis Details: Finite Population setting}
+\section{Analysis details: finite population setting}
 \label{sec:FinitePopulationProofs}
@@
 
-\section{Analysis Details: distributional setting}
+\section{Analysis details: distributional setting}
 

### 2024-10-06 AISTAT25/aistats.tex
Yoavs final

@@
 \usepackage{graphicx}
+
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{0pt}{0pt}
+\titlespacing*{\subsection}{0pt}{0pt}{0pt}
+\titlespacing*{\subsubsection}{0pt}{0pt}{0pt}
 
@@
   collection of balls of different sizes and uses label
-  queries to identify those with a strong bias towards one particular
+  queries to identify those with a strong bias towards a particular
   label. When two such balls intersect and have different
@@
   is targeted in future active queries. We give label complexity
-  bounds for this method that do not rely on assumptions about the
+  bounds for this method that do not rely on any assumptions about the
   data, and we instantiate them in several cases of interest.
@@
 
+As is well known~\cite{CH67, S80}, the $k$-NN method converges to
+Bayes optimal rule without any assumptions on the underlying
+distribution. Convergence rate is not uniform. It is slower at points
+close to the decision boundary and faster at points further from the
+boundary.  Our goal is to mimic the asymptotic performance of $k$-NN
+while achieving lower query complexity by reducing the number of
+queries far from the boundary.
+
 We begin with a large collection of \emph{balls}, $\B$, of varying
@@
 to 0. As time goes on, we get biases for progressively smaller balls,
-as needed. The final label assigned to any point $x$ is based on the
+as needed. The label $\yh(x)$ assigned to any point $x \in X$ is based on the
 biases of balls containing $x$.
@@
 The rest of the paper is organized as
-follows. Section~\ref{sec:relatedwork} describes related work.
-Section~\ref{sec:algorithm} presents the active learning algorithm and
-Section~\ref{sec:discrete} analyzes it in the finite population
-setting, taking $X$ to be an arbitrary point-set and allowing any
-$\eta$ function. We identify two \emph{critical levels} for any
-$x \in X$: two scales (of ball sizes) that control how many queries
-are sufficient for $x$ to be correctly labeled
+follows. Section~\ref{sec:relatedwork} describes related
+work. Section~\ref{sec:definitions} defines some terminology and
+notation.  Section~\ref{sec:algorithm} presents the active learning
+algorithm and Section~\ref{sec:discrete} analyzes it in the finite
+population setting, taking $X$ to be an arbitrary point-set and
+allowing any $\eta$ function. We identify two \emph{critical levels}
+for any $x \in X$: two scales (of ball sizes) that control how many
+queries are sufficient for $x$ to be correctly labeled
 (Theorem~\ref{thm:label-complexity}). We instantiate these bounds in
@@
 and \ref{thm:oned-monotonic}) to get label complexities logarithmic in
-$|X|$. In section~\ref{sec:experiment} we describe an experiment that supports the theoretical bound given in Theorem~\ref{thm:oned-massart}.
+$|X|$. In section~\ref{sec:experiment} we describe an experiment that
+supports the theoretical bound given in
+Theorem~\ref{thm:oned-massart}.
 
@@
 \section{Definitions}
-
+\label{sec:definitions}
 The two basic entities in our algorithm are the collection of points
@@
 
-We now describe the simplified algorithm, here referred to simply as
-the algorithm. The algorithm loops over increasing levels.  At each
-level it identifies the region of ``known unknowns'' which approximates
-the boundary. The algorithm then makes queries in the vicinity of the
-boundary to refine it and proceeds to the next finer level.
-
-What follows is a detailed explanation of  Figure~\ref{alg:simple}.
+We now describe the simplified algorithm. The algorithm loops over
+increasing levels.  At each level it identifies the region of ``known
+unknowns'' which approximates the boundary. The algorithm then makes
+queries in the vicinity of the boundary to refine it and proceeds to
+the next finer level.
+
+Figure~\ref{alg:simple} contains the pseudo-code of the simplified algorithm.
+What follows is a detailed explanation of this code.
 The outer loop of the algorithm iterates over
-levels, starting with $\ell=0$ that corresponds to balls with at least half of the points in $X$, then moving onto $\ell=1$ which has balls with between $1/4$ and $1/2$ of the points in $X$, and so on.
-%and ending with $\ell=\log |X|-1$,
-%corresponding to balls that contain a single point. %Initially, all balls are labeled $\bot$, which indicates that there are not enough queries inside the ball to determine it's label.
+levels, starting with $\ell=0$ that corresponds to balls with at least half of the points in $X$, then
+$\ell=1$ which has balls with between $1/4$ and $1/2$ of the points in $X$, and so on.
 
@@
 to query'' (see challenges in the introduction). It does not address
-the other two challenges, which the detailed algorithm does address.
+the other two challenges, addressed here with the detailed algorithm.
 
@@
 
-A related problem is that having the outer loop iterate over levels will
+A related problem is that having the outer loop iterate over levels might
 cause the algorithm to miss uncertain regions that later appear in lower
 levels as a result of background sampling (the so-called mind changes).
-Instead, each iteration of the full algorithm looks for the lowest level 
+Instead, each iteration of the detailed algorithm looks for the lowest level 
 in which the uncertainty region is non-empty, and makes just two queries, 
@@
 
-We compare two algorithms, the ``active'' agorithm described here, and the ``passive'' algorithm which uses only background queries.
+We compare two algorithms, the ``active'' algorithm described here, and the ``passive'' algorithm which uses only background queries.
 We denote the difference between the error of the generated classifier and that of the Bayes optimal classifier by $\epsilon$.
